### 2023-12-15 GSI_and_Fst.Rmd
Initial commit

---
title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('references.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
# inputs:
input_list <- list(
  final_baseline = "data/cali-chinook-baseline.rds",
  pop_labels = "inputs/reference-collection-names.csv"
)
# outputs:
output_list <- list(
  assig_table = "results/GSI_and_Fst/assignment-table-full-baseline.pdf",
  fst_table = "results/GSI_and_Fst/full-baseline-but-no-lfar-Fst-table.pdf",
  assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Introduction

Here we assess power of the baseline for population assignment

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rubias)
library(cowplot)
library(hierfstat)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_rds(input_list$final_baseline) %>%
  filter(collection != "SacNonW")

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything())

# finally, get the order of collections and repunits we want
collection_order <- unique(pop_labels$collection)
```

# Do the self-assignment

In this round, we go directly to assigning to repunit, and then to the max-likelihood
collection within that repunit.

```{r}
full_sa <- self_assign(full_base2, gen_start_col = 5)
```

```{r}
top_ass <- full_sa %>%
  group_by(indiv, collection, repunit, inferred_repunit) %>%
  mutate(repu_sclike = sum(scaled_likelihood)) %>%
  group_by(indiv) %>%
  filter(repu_sclike == max(repu_sclike)) %>%
  arrange(indiv, desc(scaled_likelihood)) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(coll_f = factor(collection, levels = collection_order))
```

And, once again, we want to investigate the ones that are in the wrong repunit, and look at their
scaled_repu_likelihoods, the others.
```{r}
source("R/colors.R")
ta_corr <- top_ass %>%
  filter(repunit == inferred_repunit)
ta_wrong <- top_ass %>%
  filter(repunit != inferred_repunit) %>%
  arrange(collection, repu_sclike) %>%
  group_by(collection) %>%
  mutate(y = 190 + (70/9) * 1:n())

ggplot() +
  geom_histogram(
    data = ta_corr, 
    mapping = aes(x = repu_sclike, fill = inferred_repunit)
  ) +
  geom_point(
    data = ta_wrong,
    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
    size = 3, 
    shape = 21
  ) +
  facet_wrap(~coll_f, ncol = 3) +
  scale_fill_manual(values = repunit_colors)
```
Which looks to me like there are really only 7 here that we are not getting right.
Those others we will filter out and make a table of them:

```{r}
tafilt <- top_ass %>%
  filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
```

Here we calculate the number of correctly and incorrectly assigned fish:
```{r}
tafilt %>%
  mutate(correct = repunit == inferred_repunit) %>%
  count(correct) %>%
  mutate(fract = n / sum(n))
```

And now we will make a table of those:
```{r}
# get the function
source("R/taf-prep.R")
source("R/table-as-figure.R")

# get the RC_groups
RC_groups <- pop_labels %>%
  select(-old_name) %>%
  distinct() %>%
  rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
  select(external_group, internal_group, label)

X2 <- tafilt %>%
  mutate(
    collection_f = factor(collection, levels = collection_order),
    inferred_coll_f = factor(inferred_collection, levels = collection_order)
  ) %>%
  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
  taf_prep()

# get the result
TAF2 <- table_as_figure(
  X = X2, 
  RC_groups = RC_groups,
  external_colors = run_time_colors,
  internal_colors = repunit_colors, 
  plot_margins = c(1, 0.1, 0.7, 0.1)
)

# and plot it
g <- plot_grid(
  TAF2$full_plot,
  plot_grid(
    get_legend(TAF2$for_external_legend),
    get_legend(TAF2$for_internal_legend),
    nrow = 1
  ),
  nrow = 2,
  rel_heights = c(7,3)
)

g
```

That is a considerable improvement. Let's write it out:
```{r}
ggsave(g, filename = output_list$assig_table, width = 8, height = 10)
```

# Calculating Fst between the collections from the microhaplotype markers

We use the 'hierfstat' package to get the pairwise
$F_\mathrm{ST}$ values (Weir & Cockerham method) between the collections from these microhaplotype data.

In order to do this we need to convert the data to a lumped format.
This means making integers of the alleles and then catenating them.
No locus has more than 9 alleles, so we can do single-digits for alleles.  We drop the
LFAR from this, because those were chosen specifically to be weird in the late fall.
```{r}
hfs_ref_long <- full_base2 %>% 
  select(-starts_with("NC_"), -starts_with("Ots")) %>%  # see what we get when we kep just the tag_id markers
  mutate(
    Pop = as.integer(factor(collection, levels = collection_order)),
    .before = collection
  ) %>%
  select(-repunit, -collection, -sample_type) %>%
  pivot_longer(
    cols = c(-indiv, -Pop),
    names_to = c("locus", "gene_copy"),
    names_pattern = "^(.*)_([12])$"
  ) %>%
  group_by(locus) %>%
  mutate(
    alle_int =  as.integer(factor(value)),
    alle_int = ifelse(is.na(alle_int), 0L, alle_int)
  ) %>%
  ungroup() %>%
  select(-value)

# then we can lump the alleles 
hfs_ref_wide <- hfs_ref_long %>%
  group_by(indiv, Pop, locus) %>%
  summarise(geno = str_c(alle_int[1], alle_int[2])) %>%
  ungroup() %>%
  mutate(geno = ifelse(geno == "00", NA_character_, geno)) %>%
  mutate(geno = as.integer(geno)) %>%
  pivot_wider(
    names_from = c(locus),
    values_from = geno
  ) %>%
  arrange(Pop, indiv)

hfstat_dat <- hfs_ref_wide %>% 
  select(-indiv) %>%
  as.data.frame()

FST <- pp.fst(dat = hfstat_dat, diploid = TRUE)

# now, get the values out of it and put them in the upper diagonal
# of a matrix for plotting with the WGS values.
fst_mat <- FST$fst.pp 
colnames(fst_mat) <- collection_order
```

Something is very, very weird here.  Look at that Fst matrix:
```{r}
tmp <- fst_mat
rownames(tmp) <- collection_order
tmp
```

For some reason, BCF, BCS, MDS, and BCF have quite low Fst values to EVERYONE, even Winter run and the
Klamath.  That just cannot be right!!  They have super low Fst values to FRHF and FRHF, but those
two pops have Fst of around 0.14 to winter-run, and BCS to winter run is only 0.0269.  WTF?!

```{r}

all_values_tib <- fst_mat %>%
  as_tibble() %>%
  mutate(row_label = collection_order, .before = SRW) %>%
  pivot_longer(
    -row_label, 
    names_to = "col_label",
    values_to = "fst"
  )

# to get it to the right format we make it look like rubias
# output and the pipe it into taf_prep()
upper_tri_fst <- all_values_tib %>%
  filter(!is.nan(fst)) %>%
  mutate(
    collection_f = factor(row_label, levels = collection_order),
    inferred_coll_f = factor(col_label, levels = collection_order)
  ) %>% 
  select(-row_label, -col_label) %>%
  mutate(n = sprintf("%0.3f", fst)) %>%
  select(-fst) %>%
  taf_prep()

```

Now that we have the `upper_tri_fst` values in the right format, we just need
to add the WGS values to the lower triangle part of the matrix.

# Adding the WGS Fst values and making a figure

```{r}
# for now I am just going to put the WGS Fst's into the lower diagonal, but I think it will be better, ultimately,
# to but the bootstrapped CI's in there.
wgs_mat <- read_rds("../anderson-et-al-chinook-ecotypic-genetic-architecture/outputs/001/full-wgs-fst-mat.rds")

# once again, make it looks like rubias output and taf_prep() it.
lower_tri_diag <- wgs_mat %>%
  mutate(
    collection_f = factor(pop1_f, levels = collection_order),
    inferred_coll_f = factor(pop2_f, levels = collection_order),
    n = fst
  ) %>%
  select(collection_f, inferred_coll_f, n) %>%
  complete(collection_f, inferred_coll_f) %>%
  filter(as.integer(collection_f) >= as.integer(inferred_coll_f)) %>%  # this gets lower tri + diagonal
  taf_prep()

# then we put those together and make a figure
full_fst <- bind_rows(
  upper_tri_fst,
  lower_tri_diag
) %>%
  arrange(row_label, col_label)

# get the result
TAF_FST <- table_as_figure(
  X = full_fst, 
  RC_groups = RC_groups,
  external_colors = run_time_colors,
  internal_colors = repunit_colors,
  Xs_on_diagonal = TRUE,
  plot_margins = c(0.7, 0.1, 0.1, 0.1)
)

# and plot it
TAF_FST$full_plot

```

And we can save that too:
```{r}
ggsave(TAF_FST$full_plot, file = output_list$fst_table, width = 6, height = 5.1)
```

# Add the LFAR markers to that

Anthony compiled the LFAR markers for the baseline fish.  Not every fish in the baseline
was typed at the LFAR markers.  So, let us see which ones that includes.
```{r}
lfar <- read_csv(input_list$lfar_base)
```

As we saw previously, the loci that we are interested in here are:
```
"NC_037130.1:1062935-1063235",
"NC_037130.1:828619-828919",
"NC_037130.1:864908-865208"
```

However, the 828619 gobbles up reads that don't all map to the right place and
turns out not to be super informative, so we will drop it.

Also, we can call these things as SNPs, not microhaps, which we will do.  The Canonical
Variation file in the snakemake workflow tells us which position is the target SNP in each:

```{r}
focal_loci <- c(
  "NC_037130.1:1062935-1063235",  # target SNP is the second one (position 151 in the amplicon)
  "NC_037130.1:864908-865208"     # target SNP is the first one (position 151 in the amplicon)
)

# so we can pick those out:
lfar_focal <- lfar[,c("Sample_ID", paste(rep(focal_loci, each = 2), c("_1", "_2"), sep = ""))]

# then we will pick out just the target SNPs, too
lfar_focal[, 2] <- str_sub(lfar_focal[[2]], 2, 2)
lfar_focal[, 3] <- str_sub(lfar_focal[[3]], 2, 2)
lfar_focal[, 4] <- str_sub(lfar_focal[[4]], 1, 1)
lfar_focal[, 5] <- str_sub(lfar_focal[[5]], 1, 1)
```

We will left join this onto the baseline by the NMFS_DNA_ID:
```{r}
full_base_lfar <- full_base2 %>%
  left_join(lfar_focal, by = join_by(indiv == Sample_ID))
```

Quickly, let's count up how many individuals we have typed at all these loci.
```{r}
lfar_check <- full_base_lfar[, c("indiv", "repunit", "collection", paste(focal_loci, 1, sep = "_"))] %>%
  mutate(Baseline = "Baseline") %>%
  pivot_longer(-(indiv:collection), names_to = "set") %>%
  mutate(HasData = !is.na(value))

lfar_check %>%
  group_by(repunit, collection, set) %>%
  summarise(
    num_fish_with_genos = sum(HasData)
  ) %>%
  pivot_wider(names_from = set, values_from = num_fish_with_genos)

```
OK! That is great.  We have genotyped pretty much everyone at these LFAR markers except for about 50 of the
FRHS and 10 of the MDF.  If need be would could toss fish that don't have genotypes at any LFAR markers
when we make the assignment table.  Let's get a tibble of baseline individuals telling us whether or not
they have data for at least one of the LFAR markers:
```{r}
at_least_one_lfar <- lfar_check %>%
  filter(set != "Baseline") %>%
  group_by(indiv) %>%
  summarise(at_least_one_lfar = any(HasData == TRUE))

# just see how many there are in total
at_least_one_lfar %>%
  count(at_least_one_lfar)
```
OK! We will compute the assignment table with and without those 66.

## Self-assignment

```{r}
sa_lfar_all <- self_assign(full_base_lfar, gen_start_col = 5)

top_ass_lf <- sa_lfar_all %>%
  group_by(indiv, collection, repunit, inferred_repunit) %>%
  mutate(repu_sclike = sum(scaled_likelihood)) %>%
  group_by(indiv) %>%
  filter(repu_sclike == max(repu_sclike)) %>%
  arrange(indiv, desc(scaled_likelihood)) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(coll_f = factor(collection, levels = collection_order))
```

## Checking for individuals sampled into the wrong collections

We investigate the ones that are in the wrong repunit, and look at their
scaled_repu_likelihoods, to the others.
```{r}
source("R/colors.R")
ta_corr_lf <- top_ass_lf %>%
  filter(repunit == inferred_repunit)
ta_wrong_lf <- top_ass_lf %>%
  filter(repunit != inferred_repunit) %>%
  arrange(collection, repu_sclike) %>%
  group_by(collection) %>%
  mutate(y = 190 + (70/9) * 1:n())

ggplot() +
  geom_histogram(
    data = ta_corr_lf, 
    mapping = aes(x = repu_sclike, fill = inferred_repunit)
  ) +
  geom_point(
    data = ta_wrong_lf,
    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
    size = 3, 
    shape = 21
  ) +
  facet_wrap(~coll_f, ncol = 3) +
  scale_fill_manual(values = gen_bkgrnd_colors)
```
This is much like it was before, however one CHLF fish that was inferred as a spring
run is now inferred as a CHLF.

We toss those 6 that are assigned quite strongly (> 0.98) to a repunit
that they were not sampled in. (Again, the winter-run are clearly true winter
run sampled in the fall or spring collections, and the same is probably true
for MDS-collected fish that are really fall run. We could follow the latter
up with the rosa markers.)

```{r}
ta_lf_filt <- top_ass_lf %>%
  filter(!(repunit != inferred_repunit & repu_sclike > 0.98))

# get the total fraction of assigned vs misassigned fish
ta_lf_filt %>%
  mutate(correct = repunit == inferred_repunit) %>%
  count(correct) %>%
  mutate(fract = n / sum(n))
```
Pretty good.

But, what we really want to see if the table for the collections, to see if we can distinguish
late-fall from fall run any better.

## Assignment table with all the fish

```{r}
# get the function
source("R/taf-prep.R")
source("R/table-as-figure.R")

# get the RC_groups
RC_groups <- pop_labels %>%
  select(-old_name) %>%
  distinct() %>%
  rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
  select(external_group, internal_group, label)

X2 <- ta_lf_filt %>%
  mutate(
    collection_f = factor(collection, levels = collection_order),
    inferred_coll_f = factor(inferred_collection, levels = collection_order)
  ) %>%
  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
  taf_prep()

# get the result
TAF2_lf <- table_as_figure(
  X = X2, 
  RC_groups = RC_groups,
  external_colors = run_time_colors,
  internal_colors = gen_bkgrnd_colors, 
  plot_margins = c(1, 0.1, 0.7, 0.1)
)

# and plot it
g <- plot_grid(
  TAF2_lf$full_plot,
  plot_grid(
    get_legend(TAF2_lf$for_external_legend),
    get_legend(TAF2_lf$for_internal_legend),
    nrow = 1
  ),
  nrow = 2,
  rel_heights = c(7,3)
)

g
```

```{r}
ggsave(g, filename = output_list$assig_table_lfar_all, width = 6, height = 8)
```

## Assignment table with only the fish that have at least one-non-missing genotype among the 3 LFAR genotypes.

We can just filter out fish that don't have a non-missing genotype amongst the three LFARs and call it good.

```{r}
X2_alolf <- ta_lf_filt %>%
  semi_join(at_least_one_lfar %>% filter(at_least_one_lfar), by = join_by(indiv)) %>%
  mutate(
    collection_f = factor(collection, levels = collection_order),
    inferred_coll_f = factor(inferred_collection, levels = collection_order)
  ) %>%
  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
  taf_prep()

# get the result
TAF2_alolf <- table_as_figure(
  X = X2_alolf, 
  RC_groups = RC_groups,
  external_colors = run_time_colors,
  internal_colors = gen_bkgrnd_colors, 
  plot_margins = c(1, 0.1, 0.7, 0.1)
)

# and plot it
g <- plot_grid(
  TAF2_alolf$full_plot,
  plot_grid(
    get_legend(TAF2_alolf$for_external_legend),
    get_legend(TAF2_alolf$for_internal_legend),
    nrow = 1
  ),
  nrow = 2,
  rel_heights = c(7,3)
)

g
```

```{r}
ggsave(g, filename = output_list$assig_table_lfar_no_miss, width = 6, height = 8)
```

OK, that is good to see.  The missassignments to or from CHLF are *not* because
they don't have data at the LFAR.  So, we can just use the full data set.

### 2023-12-15 R/colors.R

@@ -1,0 +1,22 @@
+
+# These are for the genetic-marker-resolvable genomic backgrounds
+repunit_colors <- c(
+  `Klamath-Trinity` = "#6e8b3d",
+  `Coastal` = "#79cdcd",
+  `CV-Winter` = "#CC99FF",            # a "cool-ish" color for winter
+  `CV-Spring` = "#99CC33",             # green like spring
+  `CV-Fall` = "#F06292",
+  `CV-Late-Fall` = "#cd9b1d"
+)
+
+
+# These are for the "nominal run timing groups" colored like they
+# are in the Science paper
+run_time_colors <- c(
+  `Winter run` = "#ffff99",
+  `Spring run` = "gold",
+  `Fall run` = "blue",
+  `Late-fall run` = "#a6cee3"
+)
+
+

### 2023-12-15 R/table-as-figure.R

@@ -1,0 +1,200 @@
+require(tidyverse)
+
+#' A function to create a table as a figure
+#'
+#' This is intended for non-symmetrical measures, so that the table
+#' is represented as a full matrix. These labels will
+#' go on both the right and left sides, as well as the top and bottom.
+#'
+#' The primary reason for doing this is to create a table as a figure
+#' with text labels and also colors in some of the cells, and also
+#' in the row and column labels.
+#' @param X a tibble with the columns:
+#'   - row_label: a factor, which row is should be in
+#'   - col_label: a factor, which column it should be in
+#'   - cell_label: a character vector of what goes in the cell
+#'   - cell_fill: a character vector that will be mapped to colors.
+#'   Should be "white" if it is to be left as white.
+#' @param RC_groups a tibble with two columns:
+#'   - external group: the group that each row_label belongs to, that it will be colored
+#'     by in the external labels.
+#'   - internal_group: the group that the row/column belongs to inside the table.  These
+#'     should be contiguous
+#'   - label: the label itself, as a character vector. The whole tibble
+#'     Should be ordered by label according to the row_label factor levels in X
+#' @param external_colors a named vector with the names being the groups
+#' found in RC_groups$external_group and the values being colors
+#' @param internal_colors a named vector with the names being the groups
+#' found in RC_groups$internal_group and the values being colors
+#' @param Xs_on_diagonal  If TRUE then line segments are drawn in a big X along
+#' the diagonal (i.e., for Fst values, etc.)
+table_as_figure <- function(
+    X,
+    RC_groups,
+    internal_colors,
+    external_colors,
+    internal_color_legend_name = "Genetic-marker-resolvable\ngenomic background",
+    external_color_legend_name = "Nominal group based\non run timing",
+    Xs_on_diagonal = FALSE,
+    skinny_line_width = 0.3,
+    thick_line_width = 1.3,
+    plot_margins = c(0, 0, 1.5, 0)
+) {
+
+  # first, expand X to include the xmin/max and ymin/ymax for each internal cell
+  N <- length(levels(X$row_label))
+  X2 <- X %>%
+    mutate(
[…]

### 2023-12-15 R/taf-prep.R

@@ -1,0 +1,36 @@
+require(tidyverse)
+
+#' simple function to take rubias output and prep it
+#' for the table_as_figure function
+taf_prep <- function(X) {
+  X %>% rename(
+    row_label = collection_f,
+    col_label = inferred_coll_f,
+  ) %>%
+    mutate(
+      cell_label = as.character(n)
+    ) %>%
+    select(-n) %>%
+    left_join(
+      pop_labels %>% select(collection, repunit),
+      by = c("row_label" = "collection"),
+      keep = TRUE
+    ) %>%
+    select(-collection) %>%
+    rename(row_rep = repunit) %>%
+    left_join(
+      pop_labels %>% select(collection, repunit),
+      by = c("col_label" = "collection"),
+      keep = TRUE
+    ) %>%
+    select(-collection) %>%
+    rename(col_rep = repunit) %>%
+    mutate(
+      cell_fill = case_when(
+        row_rep == col_rep ~ row_rep,
+        TRUE ~ NA_character_
+      )
+    ) %>%
+    select(-col_rep, -row_rep)
+
+}

### 2023-12-15 tex/abstract.tex
commit some tex stuff

%!TEX root = main.tex

We need to write an abstract.

### 2023-12-15 tex/author-title-etc.tex
commit some tex stuff

%!TEX root = main.tex

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{A multipurpose microhaplotype panel for analysis of California Chinook salmon}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{
$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,Santa Cruz, California, USA.
}

\renewcommand{\AuthorAddresses}{\myAffiliations}

\renewcommand{\KeyWords}{Genetic stock identfication, population assignment, parentage based tagging, amplicon sequencing}

\renewcommand{\CorrespondingAuthor}{eric.c.anderson@noaa.gov}

% the email address for the corresponding author
\newcommand{\myEmailAddress}{eric.c.anderson@noaa.gov}
\newcommand{\myEmailFootnote}{$^\S$}

% here you can put your very own copyright notice
\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}

% here you can put a running title (a short title that goes on the left of the
% even pages)
\newcommand{\myRunningTitle}{Microhaplotypes for California Chinook salmon}
\renewcommand{\RunningTitle}{\myRunningTitle}

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson et al.}

%%%% DONE WITH AUTHOR/TITLE/ETC INFORMATION DEFINITION

### 2023-12-15 tex/body-text-of-paper.tex
commit some tex stuff

%!TEX root = main.tex

\section*{Introduction}

Some verbiage here

\section*{Methods}

More verbiage

\subsection*{Population Assignment}

More Verbiage.

\subsection*{Fisher Information and Effective Sample Size}

\begin{figure*}
\newcommand{\esssimscap}{\footnotesize {\bf a)} An example figure Observed information calculated for simulated data summarized either as fully observed genotypes (purple)
or as genotype likelihoods (orange) computed from sequencing read data of different depths simulated from the genotypes. Fully observed genotype data is not affected by read depth, but an independent set of fully observed genotypes was simulated for each different value of read depth, and these are all shown in the figure.  {\bf b)} Effective sample sizes calculated for simulated genotype likelihood data.
In each figure the facet headers give the true population allele frequency, the $x$-axis gives the
average read depth in the simulations, and the distribution of quantities in the $y$ direction are summarized
as boxplots showing the median (dark line) the first and third quartiles (the edges of the boxes)  the largest (or smallest) value no further than $1.5 \times\mbox{}$ the interquartile range from the first (third) quartiles (the
whiskers) and outliers beyond the whiskers (individual points). All simulations had 100 individuals.}
\begin{center}
\includegraphics[width=\textwidth]{images/info_and_eff_size.pdf}
\end{center}
\caption[\esssimscap]{\esssimscap}
\label{fig:ess_sims}
\end{figure*}

### 2023-12-15 tex/data-accessibility.tex
commit some tex stuff

\section*{Data Accessibility}

{\sc WGSassign} is available as a Python package with these associated links:
\begin{itemize}
\item Development version and entire revision history on GitHub: \url{https://anonymous.4open.science/r/WGSassign-7155}
\item Zenodo archive of initial package release: UPDATE AFTER ANONYMOUS REVIEW
\item Online version of data and scripts used in paper: \url{https://anonymous.4open.science/r/WGSassign-manuscript-data-3072}
\item Data repository with full datasets used in paper. UPDATED WHEN JOURNAL PROVIDES DOI. \url{https://dryad.something.or.other}
\end{itemize}
\mbox{}

### 2023-12-15 tex/newcomms.tex
commit some tex stuff

\newcommand{\relat}{R}
\newcommand{\relatset}{\mathscr{R}}
\newcommand{\nA}{\mathcal{A}}
\newcommand{\nL}{\mathcal{L}}
\newcommand{\nG}{\mathcal{G}}
\newcommand{\MAF}{\mathrm{MAF}}

\newcommand{\Pe}{\mathrm{P}_\epsilon}
\newcommand{\PR}{\mathrm{P}_\relat}

\newcommand{\bkappa}{\bm{\kappa}}

\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\BY}{\mathbf{Y}}
\newcommand{\BX}{\mathbf{X}}
\newcommand{\BC}{\mathbf{C}}
\newcommand{\I}{\mathbb{1}}
\newcommand{\bomega}{\bm{\omega}}

\newcommand{\Exp}{\mathbb{E}}
\newcommand{\Var}{\mathrm{Var}}

\newcommand{\heta}{\hat{\theta}}
\newcommand{\bg}{\bm{g}}

\newcommand{\GG}{\mathscr{G}}

### 2023-12-15 tex/body-text-of-paper.tex
Working on some writing...

@@
 
-Some verbiage here
+There is a long and rich tradition of using genetic markers in the 
+management of Pacific salmon, and work on Pacific salmon has been
+a prominent driving force behind a number of advances in molecular ecology,
+both in data generation \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
+and in statistical analysis
+\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
+Two broadly applicable innovations that have been
+actively fostered by the Pacific salmon research community are genetic
+stock identification
+(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development}) and parentage based tagging
+(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
+
+ In the 1980's, electrophoretically
+detectable genetic variation, in the form of allozymes
+\citep{ayala1972allozymes,allendorf1981use}, was used to
+establish a program of GSI for Chinook salmon,
+{\em Oncorhynchus tshawytscha} \citep{milner1982genetic}.  Extensive sampling
+revealed that these allozyme markers
+exhibited different allele frequencies among major stocks of Chinook salmon on the West Coast.  These allele frequencies, in turn, implied different expected frequencies of 
+multilocus genotypes in the different stocks, such that the genotypes of fish caught in fisheries 
+could be used to estimate the fraction of fish originating from each stock.
+\citet{milner1985genetic} showed that proportions of stocks in
+the Washington state coastal troll fishery could be estimated by GSI.
+Since that time, with the development of novel molecular markers,
+and, now, the advent of the genomic sequencing era, the scope and scale of GSI
+has expanded considerably.
+
+With greater numbers of more variable markers it is now possible to accurately
+identify the population of origin of individual fish, and to resolve groups of
+more closely related populations.  Furthermore, reference data sets of allele 
+frequencies in hundreds of populations from throughout the range of multiple species of salmon and
+other anadromous species
+\citep{seeb2007development,gilbey2018microsatellite,barclay2019genetic} now exist, and are routinely used to assign fish caught thousands of
+kilometers from their natal streams to their stock of origin. Applications include estimating fishery
+composition \citep{satterthwaite2015stock}, providing real-time information for fishery openings and closings \citep{beacham2004dna}, assessing the spatial distribution of different stocks in the
+ocean \citep{urawa2009stock} and the temporal distribution in upstream migrations
+\citep{hess2014monitoring},  and monitoring  bycatch \citep{hasselman2016genetic} or illegal captures \citep{wilmot1999origins} in marine fisheries.
 
@@
 
-\subsection*{Population Assignment}
+\subsection*{Population Sampling}
 
@@
 
-\subsection*{Fisher Information and Effective Sample Size}
+\subsection*{Genetic Variation}
+
+\subsection*{Power for genetic stock identification and population assignment}
+
+\subsection*{Power for relationship inference}
 
 \begin{figure*}
-\newcommand{\esssimscap}{\footnotesize {\bf a)} An example figure Observed information calculated for simulated data summarized either as fully observed genotypes (purple)
-or as genotype likelihoods (orange) computed from sequencing read data of different depths simulated from the genotypes.
-Fully observed genotype data is not affected by read depth, but an independent set of fully observed genotypes was simulated for each different value of read depth, and these are all shown in the figure.
-{\bf b)} Effective sample sizes calculated for simulated genotype likelihood data.
-In each figure the facet headers give the true population allele frequency, the $x$-axis gives the
-average read depth in the simulations, and the distribution of quantities in the $y$ direction are summarized
-as boxplots showing the median (dark line) the first and third quartiles (the edges of the boxes)  the largest (or smallest) value no further than $1.5 \times\mbox{}$ the interquartile range from the first (third) quartiles (the
-whiskers) and outliers beyond the whiskers (individual points). All simulations had 100 individuals.}
+\newcommand{\examplecap}{\footnotesize This is just an example figure thrown
+in here for now to illustrate the formatting and how the captions are put in. It is not what
+we will use for the paper, but we will make another Joseph and the Magic Technicolor
+assignment table for this paper.  {\bf a)} Pairwise $F_\mathrm{ST}$ values.
+{\bf b)} Assignment rates.}
 \begin{center}
-\includegraphics[width=\textwidth]{images/info_and_eff_size.pdf}
+\includegraphics[width=0.35\textwidth]{images/assign_and_fst_table.pdf}
 \end{center}
-\caption[\esssimscap]{\esssimscap}
-\label{fig:ess_sims}
+\caption[\examplecap]{\examplecap}
+\label{fig:example}
 \end{figure*}

### 2023-12-17 tex/body-text-of-paper.tex
working through intro more

@@
 
-There is a long and rich tradition of using genetic markers in the 
-management of Pacific salmon, and work on Pacific salmon has been
-a prominent driving force behind a number of advances in molecular ecology,
-both in data generation \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
-and in statistical analysis
+Genetic markers have long been used in the 
+management of Pacific salmon, and salmon conservation and management has been
+a driving force behind a number of advances in molecular ecology,
+from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
+to statistical methodology
 \citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
-Two broadly applicable innovations that have been
-actively fostered by the Pacific salmon research community are genetic
+One broadly applicable innovation that has been
+actively fostered by the Pacific salmon research community is genetic
 stock identification
-(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development}) and parentage based tagging
-(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
+(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
 
@@
 revealed that these allozyme markers
-exhibited different allele frequencies among major stocks of Chinook salmon on the West Coast.  These allele frequencies, in turn, implied different expected frequencies of 
-multilocus genotypes in the different stocks, such that the genotypes of fish caught in fisheries 
-could be used to estimate the fraction of fish originating from each stock.
-\citet{milner1985genetic} showed that proportions of stocks in
+exhibited different allele frequencies among major stocks of Chinook salmon on the West Coast.
+These allele frequency differences make GSI possible, and
+\citet{milner1985genetic} soon showed that proportions of stocks in
 the Washington state coastal troll fishery could be estimated by GSI.
-Since that time, with the development of novel molecular markers,
-and, now, the advent of the genomic sequencing era, the scope and scale of GSI
+Since that time, with the development of novel molecular markers, and, now
+with increasing capacity to sequence genomic material, the scope and scale of GSI
 has expanded considerably.
 
-With greater numbers of more variable markers it is now possible to accurately
-identify the population of origin of individual fish, and to resolve groups of
-more closely related populations.  Furthermore, reference data sets of allele 
-frequencies in hundreds of populations from throughout the range of multiple species of salmon and
+By using greater numbers of more variable markers than the allozymes available
+in the 1980s it is possible to accurately
+identify the population of origin of individual fish, rather than simply estimating
+aggregated stock proportions.  It is also possible to resolve populations of fish that
+are much more closely related than before.  Furthermore, reference data sets with genotypes
+from hundreds of populations throughout the range of multiple species of salmon and
 other anadromous species
@@
 kilometers from their natal streams to their stock of origin. Applications include estimating fishery
-composition \citep{satterthwaite2015stock}, providing real-time information for fishery openings and closings \citep{beacham2004dna}, assessing the spatial distribution of different stocks in the
-ocean \citep{urawa2009stock} and the temporal distribution in upstream migrations
+composition \citep{satterthwaite2015stock}, providing real-time information for genetics-informed fishery closures \citep{beacham2004dna}, assessing the spatial distribution of different stocks in the
+ocean \citep{urawa2009stock} and their temporal distribution in upstream migrations
 \citep{hess2014monitoring},  and monitoring  bycatch \citep{hasselman2016genetic} or illegal captures \citep{wilmot1999origins} in marine fisheries.
+
+Although sequencing costs continue to decline, they are high enough
+that there remains a tradeoff
+between reference baselines that include information from a large number of
+populations across a broad scale, and those that have been tailored
+to distinguish between closely related populations on a smaller, regional scale.
+Because of cost considerations, reference baselines that include populations across a broad 
+spatial scale may include only a few populations from each sub-region.  Furthermore,
+baselines tailored to a specific region often assemble markers that specifically
+show allele frequency differences between the closely related---and hence difficult
+to resolve---stocks within the region.  Consequently, regionally targeted baselines typically
+outperform broad-scale baselines in resolving populations within the region.
+
+Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
+structure within the Central Valley of California. Chinook salmon of the two main river basins---the 
+Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
+within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
+(Fall-, Late-Fall-, Winter-, and Spring-run), adult Chinook salmon can be found migrating or residing
+in freshwater every month of the year in California.
+
+ and parentage based tagging
+(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
 
@@
 \end{figure*}
+
+\section*{Discussion}
+
+When, and if, sequencing costs drop sufficiently, it may  be possible to include
+all different populations of a single species within a single standardized reference baseline
+that performs equally well at broad and regional scales. A statistical approach for population
+assignment from low coverage whole genome sequences appeared recently
+\citep{desaixINPRESSpopulation}, and its authors noted that such a data type could
+be ideal for producing reference baselines simultaneously applicable to broad scale
+coverage and high resolution within sub-regions.   For the present, however,
+baselines tailored to specific regions are essential for regional management questions.
+It should be noted that recent

### 2023-12-18 GSI_and_Fst.Rmd
Further updates

@@
 This means making integers of the alleles and then catenating them.
-No locus has more than 9 alleles, so we can do single-digits for alleles.  We drop the
+One locus has 11 allele, so we should use two digits for each allele.  We drop the
 LFAR from this, because those were chosen specifically to be weird in the late fall.
 ```{r}
-hfs_ref_long <- full_base2 %>% 
-  select(-starts_with("NC_"), -starts_with("Ots")) %>%  # see what we get when we kep just the tag_id markers
+full_base_long <- full_base2 %>% 
+  select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
   mutate(
@@
     names_pattern = "^(.*)_([12])$"
-  ) %>%
+  )
+
+hfs_ref_long <- full_base_long %>%
   group_by(locus) %>%
   mutate(
-    alle_int =  as.integer(factor(value)),
-    alle_int = ifelse(is.na(alle_int), 0L, alle_int)
+    alle_int =  sprintf("%02d", as.integer(factor(value)))
   ) %>%
@@
   ungroup() %>%
-  mutate(geno = ifelse(geno == "00", NA_character_, geno)) %>%
+  mutate(geno = ifelse(str_detect(geno, "NA"), NA_character_, geno)) %>%
   mutate(geno = as.integer(geno)) %>%
@@
 
+# let's also bootstrap that:
+FST_boot <- boot.ppfst(
+  dat = hfstat_dat,
+  nboot = 1000,
+  quant = c(0.025,0.975),
+  diploid = TRUE
+)
+
 # now, get the values out of it and put them in the upper diagonal
@@
 ```
-
-Something is very, very weird here.  Look at that Fst matrix:
-```{r}
-tmp <- fst_mat
-rownames(tmp) <- collection_order
-tmp
-```
-
-For some reason, BCF, BCS, MDS, and BCF have quite low Fst values to EVERYONE, even Winter run and the
-Klamath.  That just cannot be right!!  They have super low Fst values to FRHF and FRHF, but those
-two pops have Fst of around 0.14 to winter-run, and BCS to winter run is only 0.0269.  WTF?!
 

### 2023-12-18 tex/body-text-of-paper.tex
Further updates

@@
 \citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
-One broadly applicable innovation that has been
-actively fostered by the Pacific salmon research community is genetic
+Two broadly applicable innovations that has been
+actively fostered by the Pacific salmon research community are genetic
 stock identification
 (GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
+and parentage based tagging
+(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
 
@@
 
+A further consideration in the development of a reference baseline for GSI is that
+the same panel of loci used for GSI, if properly designed, can also serve as a
+markers for parentage based tagging (PBT).
+
 Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
@@
 in freshwater every month of the year in California.
-
- and parentage based tagging
-(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
 

### 2023-12-19 tex/body-text-of-paper.tex
Further updates

--- 2023-12-17 tex/body-text-of-paper.tex
@@
 \citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
-One broadly applicable innovation that has been
-actively fostered by the Pacific salmon research community is genetic
+Two broadly applicable innovations that have been
+actively fostered by the Pacific salmon research community are genetic
 stock identification
 (GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
+and parentage based tagging
+(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
 
@@
 
+The ascendance over the last decade of parentage based tagging (PBT) as a management
+tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
+Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
+programme \citep{nandor2010overview}
+with PBT, it has been noted that one of the major advantages of a genetic program for
+PBT is that the genetic markers used for PBT could also be useful for GSI (and
+vice versa).  This
+
 Here, we present a reference baseline for Chinook salmon, targeted to the complex population 

### 2023-12-20 map-of-samples.Rmd
Working on the map

---
title: "Make a Map for our Paper"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('references.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r}
# inputs:
input_list <- list(
)
# outputs:
output_list <- list(
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

```{r}
library(terra)
library(sf)
library(tidyverse)
library(ggspatial)
#library(ggsn)
library(cowplot)
library(plotly)
```

# Download spatial data

## Raster

```{r}
if(!file.exists("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/raster/HYP_HR_SR_OB_DR.zip",
    destfile = "geo-spatial/HYP_HR_SR_OB_DR.zip"
  )

  unzip(
    zipfile = "geo-spatial/HYP_HR_SR_OB_DR.zip", 
    exdir = "geo-spatial/HYP_HR_SR_OB_DR"
  )

  file.remove("geo-spatial/HYP_HR_SR_OB_DR.zip")
}
```

## State lines

```{r}
if(!file.exists("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_1_states_provinces_lines.zip",
    destfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip", 
    exdir = "geo-spatial/ne_10m_admin_1_states_provinces_lines"
  )

  file.remove("geo-spatial/ne_10m_admin_1_states_provinces_lines.zip")
}
```

## Coastline

```{r}
if(!file.exists("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/physical/ne_10m_coastline.zip",
    destfile = "geo-spatial/ne_10m_coastline.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_coastline.zip", 
    exdir = "geo-spatial/ne_10m_coastline"
  )

  file.remove("geo-spatial/ne_10m_coastline.zip")
}
```

## California Major Rivers and Creeks

This is from California Nat Res Agency.  Find it at: https://data.cnra.ca.gov/dataset/national-hydrography-dataset-nhd

```{r}
# note.  Once when I did this, download.file did not see capable of downloading this file on my Mac
# (the download is super sloooooow), and that time I 
# ended up downloading it with Chrome.  But it seems those connection problems
# had been fixed when I did this again.
if(!file.exists("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://data.cnra.ca.gov/dataset/511528b2-f7d3-4d86-8902-cc9befeeeed5/resource/7d1e7e44-81b1-43fe-95f6-1862eea6ac24/download/nhd_major_rivers_and_creeks.zip",
    destfile = "geo-spatial/nhd_major_rivers_and_creeks.zip"
  )

  unzip(
    zipfile = "geo-spatial/nhd_major_rivers_and_creeks.zip", 
    exdir = "geo-spatial"
  )

  file.remove("geo-spatial/nhd_major_rivers_and_creeks.zip")
}
```

# Read in the spatial data

```{r}
nat.earth <- rast("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") #
state_prov <- st_read("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")
coastline <- st_read("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")
all_rivers <- st_read("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp") %>%
  st_zm() %>%
  st_transform(., st_crs(state_prov))
```

## Subset the rivers

We plot this with plotly so we can see the names of the rivers easily, and whittle
it down to the ones that we need.

```{r}
# I don't know the gnis_id for mill creek, and there are a lot of Mill Creeks,
# so lets get it:
mill_area <- c(
  xmin = -121.4,
  xmax = -122.3,
  ymin = 40,
  ymax = 40.3
)
mill_candi <- all_rivers %>%
  filter(str_detect(gnis_name, "Mill Creek")) %>%
  st_crop(mill_area)

our_rivers <- all_rivers %>%
  filter(gnis_name %in% c(
    "Sacramento River",
    "San Joaquin River",
    "Feather River",
    "Russian River",
    "Trinity River"
    #"Deer Creek",
    #"Battle Creek"
  ) | 
    gnis_id %in% c(
      "00220293", "00237425", "00263498", "00266522",  # these are butte creek
      "00233775", "01655075",    # these are deer
      "00218740", "00229640", "00234966"   # these are battle creek
    ) 
  ) %>%
  bind_rows(mill_candi)

# get candidate positions for notations
labels <- read_tsv("inputs/map-notations.tsv")

g <- ggplot() + 
  geom_sf(data = our_rivers, aes(colour = gnis_name)) +
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)

ggplotly(g)
```

That looks good.  So let us proceed.

## Crop
```{r}
# important to put them in this order and named like this
domain <- c(
  xmin = -128,
  xmax = -116,
  ymin = 36,
  ymax = 42.5
)
nat_crop <- crop(nat.earth, y = ext(domain))
state_subset <- st_crop(state_prov, domain)
coastline_cropped <- st_crop(coastline, domain)

```

# Make the base map
```{r base-map-with-sampling-sites}
base_map <- ggplot() +
  ggspatial::layer_spatial(nat_crop) +
  geom_sf(data = state_subset, color = "gray30", fill = NA) +
  geom_sf(data = coastline_cropped, color = "gray30", fill = NA) +
  geom_sf(data = our_rivers, colour = "blue")

base_map
```

Now look add stuff to that:

```{r}
base_map + 
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)
```

# Grab data for sampling sites

```{r}
fish_sites <- read_csv("./data/100-RoSA_sample_sites.csv")
```

```{r}

  geom_text(data = fish_sites, aes(x = name_long, y = name_lat, label = pop, family = "serif", fontface = font_type), size = 3.5) +
  geom_segment(data = fish_sites %>% filter(!pop %in% c("San Joaquin", "River (F)")), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.05, yend = latitude), size = 0.5) +
  geom_segment(data = fish_sites %>% filter(pop == "San Joaquin"), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.55, yend = latitude), size = 0.5) +
  geom_point(data = fish_sites, mapping = aes(x = longitude, y = latitude), fill = fish_sites$color2, colour = fish_sites$color2) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("Longitude") +
  scale_y_continuous("Latitude", expand = c(0, 0)) +
  coord_sf(xlim = domain[1:2], ylim = domain[3:4]) +
  scalebar(
    x.min = domain[1], x.max = domain[2], y.min = domain[3], y.max = domain[4],
    location = "bottomright", model = "WGS84", dist = 250, anchor = c(x = -123, y = 36), st.size = 2.5, transform = TRUE, dist_unit = "km"
  ) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", size = 1),
    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8, family = "serif"),
    axis.title.y = element_text(family = "serif", size = 10),
    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
    legend.position = "none"
  )

# base_map
```

## add north arrow to map.
```{r adding-arrow-to-base-map}
arrow_map <- base_map +
  north(x.min = domain[-1], x.max = domain[2], y.min = domain[3], y.max = domain[4], location = "bottomright", anchor = c(x = -120.25, y = 36.5))

# arrow_map
```

## Now, work on the world-scale map with the inset:
```{r inset-map}
wrld <- map_data("state")
domain_df <- data_frame(point = 1:length(domain), long = rep(domain[1:2], each = 2), lat = c(domain[3:4], rev(domain[3:4])))
inset_world <- ggplot() +
  geom_path(data = wrld, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3) +
  coord_map("ortho", orientation = c(41, -132, 0)) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, -1, -1), "mm")
  )
# inset_world
```

## Use cowplot to make the inset appear on the bigger map

```{r cowplot-to-make-inset}
final_map <- ggdraw() +
  draw_plot(arrow_map) +
  draw_plot(inset_world, x = 0.7, y = 0.725, width = 0.25, height = 0.2)

ggsave(final_map, filename = "./outputs/100/RoSA_figure1_map_with_inset.pdf", width = 4.75, height = 4.25)
# and then plot it in the notebook (albeit at different size) too
final_map
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2023-12-20 R/colors.R

@@ -3,10 +3,10 @@
 repunit_colors <- c(
   `Klamath-Trinity` = "#6e8b3d",
   `Coastal` = "#79cdcd",
-  `CV-Winter` = "#CC99FF",            # a "cool-ish" color for winter
+  `CV-Winter` = "#ffff99",            # a "cool-ish" color for winter
   `CV-Spring` = "#99CC33",             # green like spring
   `CV-Fall` = "#F06292",
-  `CV-Late-Fall` = "#cd9b1d"
+  `CV-Late-Fall` = "#a6cee3"
 )
 
 

### 2023-12-20 R/table-as-figure.R

@@ -33,8 +33,8 @@ table_as_figure <- function(
     RC_groups,
     internal_colors,
     external_colors,
-    internal_color_legend_name = "Genetic-marker-resolvable\ngenomic background",
-    external_color_legend_name = "Nominal group based\non run timing",
+    internal_color_legend_name = "Reporting Units",
+    external_color_legend_name = "Run Timing Groups",
     Xs_on_diagonal = FALSE,
     skinny_line_width = 0.3,
     thick_line_width = 1.3,

### 2023-12-20 map-of-samples.Rmd
Working through stuff

@@
 output_list <- list(
+  map = "results/map_of_samples/map.pdf"
 )
@@
   geom_sf(data = coastline_cropped, color = "gray30", fill = NA) +
-  geom_sf(data = our_rivers, colour = "blue")
+  geom_sf(data = our_rivers, colour = "blue", linejoin = "round", lineend = "round")
 
@@
 ```{r}
-base_map + 
-  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
-  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)
+bit <- 0.0
+kick <- 0.4
+source("R/colors.R")
+
+mm <- base_map + 
+  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", size = 0.4) +
+  geom_point(data = labels, aes(x = label_long + bit + kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
+  scale_fill_manual(values = run_time_colors) +
+  theme_bw() +
+  theme(
+    panel.border = element_rect(colour = "black", linewidth = 1),
+    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
+    axis.text.y = element_text(size = 8, family = "serif"),
+    axis.title.y = element_text(family = "serif", size = 10),
+    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
+    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
+    legend.position = "none"
+  ) +
+  xlab("Longitude") +
+  ylab("Latitude")
+
+ggsave(mm, filename = output_list$map, width = 5, height = 3.5)
 ```

### 2023-12-21 map-of-samples.Rmd
Finish the map and insert it into the TeX

--- 2023-12-20 map-of-samples.Rmd
@@
     "Russian River",
-    "Trinity River"
+    "Trinity River",
+    "Klamath River"
     #"Deer Creek",
@@
 domain <- c(
-  xmin = -128,
-  xmax = -116,
+  xmin = -127.9,
+  xmax = -117.4,
   ymin = 36,
-  ymax = 42.5
+  ymax = 42.6
 )
+sf_use_s2(FALSE)
 nat_crop <- crop(nat.earth, y = ext(domain))
@@
   geom_sf(data = state_subset, color = "gray30", fill = NA) +
-  geom_sf(data = coastline_cropped, color = "gray30", fill = NA) +
-  geom_sf(data = our_rivers, colour = "blue")
+  geom_sf(data = coastline_cropped, color = "gray30", fill = NA, linewidth = 0.15) +
+  geom_sf(data = our_rivers, colour = "blue", linejoin = "round", lineend = "round")
 
-base_map
 ```
@@
 ```{r}
-base_map + 
-  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
-  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)
-```
+bit <- 0.0
+kick <- 0.4
+source("R/colors.R")
 
-# Grab data for sampling sites
+labels_1line <- labels %>%
+  group_by(name_text) %>%
+  filter(rank == max(rank)) %>%
+  ungroup()
 
-```{r}
-fish_sites <- read_csv("./data/100-RoSA_sample_sites.csv")
-```
-
-```{r}
-
-  geom_text(data = fish_sites, aes(x = name_long, y = name_lat, label = pop, family = "serif", fontface = font_type), size = 3.5) +
-  geom_segment(data = fish_sites %>% filter(!pop %in% c("San Joaquin", "River (F)")), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.05, yend = latitude), size = 0.5) +
-  geom_segment(data = fish_sites %>% filter(pop == "San Joaquin"), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.55, yend = latitude), size = 0.5) +
-  geom_point(data = fish_sites, mapping = aes(x = longitude, y = latitude), fill = fish_sites$color2, colour = fish_sites$color2) +
-  scale_x_continuous(expand = c(0, 0)) +
-  xlab("Longitude") +
-  scale_y_continuous("Latitude", expand = c(0, 0)) +
-  coord_sf(xlim = domain[1:2], ylim = domain[3:4]) +
-  scalebar(
-    x.min = domain[1], x.max = domain[2], y.min = domain[3], y.max = domain[4],
-    location = "bottomright", model = "WGS84", dist = 250, anchor = c(x = -123, y = 36), st.size = 2.5, transform = TRUE, dist_unit = "km"
+mm <- base_map + 
+  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", size = 0.4) +
+  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
+  geom_label(
+    data = labels_1line,
+    aes(
+      x = label_long + bit + (1 - hjust * 2) * kick * ( 0.65 + rank - 1),
+      y = label_lat, 
+      label = name_text,
+      hjust = hjust
+    ),
+    size = 2.2,
+    label.padding = unit(0.09, "lines")
   ) +
+  scale_fill_manual(values = run_time_colors) +
   theme_bw() +
   theme(
-    panel.border = element_rect(colour = "black", size = 1),
+    panel.border = element_rect(colour = "black", linewidth = 1),
     axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
@@
     legend.position = "none"
-  )
+  ) +
+  xlab("Longitude") +
+  ylab("Latitude") +
+  coord_sf(
+    expand = FALSE,
+  ) +
+  guides(fill = guide_legend(title = "Run Timing:", nrow = 2)) +
+  ggspatial::annotation_north_arrow(location = "br", height = unit(10, "mm"), width = unit(5, "mm"), style = north_arrow_fancy_orienteering()) +
+  ggspatial::annotation_scale(location = "br", height = unit(0.1, "cm"))
 
-# base_map
-```
-
-## add north arrow to map.
-```{r adding-arrow-to-base-map}
-arrow_map <- base_map +
-  north(x.min = domain[-1], x.max = domain[2], y.min = domain[3], y.max = domain[4], location = "bottomright", anchor = c(x = -120.25, y = 36.5))
-
-# arrow_map
 ```
@@
 ```{r inset-map}
-wrld <- map_data("state")
-domain_df <- data_frame(point = 1:length(domain), long = rep(domain[1:2], each = 2), lat = c(domain[3:4], rev(domain[3:4])))
+wrld <- map_data("world")
+states <- map_data("state")
+domain_df <- tibble(point = 1:length(domain), long = rep(domain[1:2], each = 2), lat = c(domain[3:4], rev(domain[3:4])))
+
 inset_world <- ggplot() +
-  geom_path(data = wrld, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
-  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3) +
-  coord_map("ortho", orientation = c(41, -132, 0)) +
+  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", size = 0.1) +
+  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
+  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3, linewidth = 0.1) +
+  coord_map("ortho", orientation = c(30, -114, 0), xlim = c(-150, -75), ylim = c(28, 60)) +
   theme_bw() +
@@
   )
-# inset_world
+inset_world
 ```
@@
 final_map <- ggdraw() +
-  draw_plot(arrow_map) +
-  draw_plot(inset_world, x = 0.7, y = 0.725, width = 0.25, height = 0.2)
+  draw_plot(mm) +
+  draw_plot(inset_world, x = 0.181, y = 0.139, width = 0.20, height = 0.17)
 
-ggsave(final_map, filename = "./outputs/100/RoSA_figure1_map_with_inset.pdf", width = 4.75, height = 4.25)
-# and then plot it in the notebook (albeit at different size) too
-final_map
+ggsave(final_map, filename = output_list$map, width = 5, height = 3.5)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$map, collapse = " ")
+system(CALL)
+
+```
+
+# Make colored ball for inclusion into the caption
+
+We can do this and just kick them directly into the tex/images folder.  
+```{r}
+rc_col_tib <- enframe(run_time_colors) %>%
+  mutate(
+    nospace = str_replace_all(name, " ", "-"),
+    file = str_c("tex/images/", nospace, "-ball.pdf")
+  )
+
+for(r in 1:nrow(rc_col_tib)) {
+  tmp <- rc_col_tib[r,]
+  g <- ggplot(tmp, aes(x = 1, y = 1)) +
+    geom_point(shape = 21, fill = tmp$value, size = 3.5) +
+    theme_void()
+  ggsave(g, filename = tmp$file)
+
+  CALL <- str_c("pdfcrop ", tmp$file)
+  system(CALL)
+  file.remove(tmp$file)
+}
 ```

### 2023-12-21 tex/body-text-of-paper.tex
Update body-text-of-paper.tex
Finish the map and insert it into the TeX

@@
 (Fall-, Late-Fall-, Winter-, and Spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California.
+in freshwater every month of the year in California. NEED MORE HERE.
 
@@
 
-More Verbiage.
+We sampled fish for the reference baselines from 8 locations on the across the two
+main tributaries (the Sacramento River and the San Joaquin River) within
+California's Central Valley. The baseline also includes samples collected from
+one location in the Klamath River, and three rivers of the California coast
+Fig.~\ref{fig:map}
+%%%%%%%%%%%%%
+\begin{figure}
+\newcommand{\mapcap}{\footnotesize  A map of sampling locations within
+California represented in the reference baseline.  The colored circles aside
+each location name represent the different run-timing groups of Chinook salmon
+included the samples from the location.  Color codes are as follows: \Wball~=~Winter run,
+\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run.}
+\begin{center}
+\includegraphics[width=\columnwidth]{images/map-crop.pdf}
+\end{center}
+\caption[\mapcap]{\mapcap}
+\label{fig:map}
+\end{figure}
+%%%%%%%%%%%%%%%
 

### 2023-12-21 tex/newcomms.tex
Finish the map and insert it into the TeX

@@
 \newcommand{\GG}{\mathscr{G}}
+
+\newcommand{\LFball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Late-fall-run-ball-crop.pdf}}}
+\newcommand{\Fball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Fall-run-ball-crop.pdf}}}
+\newcommand{\Wball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Winter-run-ball-crop.pdf}}}
+\newcommand{\Sball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Spring-run-ball-crop.pdf}}}

### 2023-12-21 GSI_and_Fst.Rmd
Additional adds

@@
   final_baseline = "data/cali-chinook-baseline.rds",
-  pop_labels = "inputs/reference-collection-names.csv"
+  pop_labels = "inputs/reference-collection-names.csv",
+  map_notations = "inputs/map-notations.tsv"
 )
@@
 collection_order <- unique(pop_labels$collection)
+```
+
+# Make a table summary of all the samples
+
+I would like to break things down by general sampling location, run type, number of samples,
+month-range of sampling, and range of years.
+```{r}
+map_notations <- read_tsv(input_list$map_notations, comment = "#")
+full_base2 %>%
+  count(repunit, collection) %>%
+  left_join
 ```

### 2023-12-21 map-of-samples.Rmd
Additional adds

@@
     toc_float: true
-bibliography: "`r here::here('references.bib')`"
+bibliography: "`r here::here('tex/citation.bib')`"
 ---
@@
       "00220293", "00237425", "00263498", "00266522",  # these are butte creek
-      "00233775", "01655075",    # these are deer
-      "00218740", "00229640", "00234966"   # these are battle creek
+      "00233775", "01655075"    # these are deer
+     # "00218740", "00229640", "00234966"   # these are battle creek
     ) 
@@
 # get candidate positions for notations
-labels <- read_tsv("inputs/map-notations.tsv")
+labels <- read_tsv("inputs/map-notations.tsv", comment = "#")
 

### 2023-12-21 tex/body-text-of-paper.tex
Additional adds

@@
 
-We sampled fish for the reference baselines from 8 locations on the across the two
+We sampled fish for the reference baselines from 7 locations on the across the two
 main tributaries (the Sacramento River and the San Joaquin River) within
@@
 \begin{figure}
-\newcommand{\mapcap}{\footnotesize  A map of sampling locations within
-California represented in the reference baseline.  The colored circles aside
+\newcommand{\mapcap}{\footnotesize  A map of sampling locations represented in the reference 
+baseline.  The colored circles aside
 each location name represent the different run-timing groups of Chinook salmon
@@
 %%%%%%%%%%%%%%%
+
+We are going to stick a table in here:
+\begin{table}
+\caption{Gonna make a caption. UKTR: Upper Klamath and Trinity Rivers ESU;
+CVFR: Central Valley Fall Run ESU; CVSR: Central Valley Spring Run ESU; CVWR:
+Central Valley Winter Run ESU
+}
+\end{table}
 

### 2024-01-07 GSI_and_Fst.Rmd
A whole bunch of updates

@@
   pop_labels = "inputs/reference-collection-names.csv",
-  map_notations = "inputs/map-notations.tsv"
+  map_notations = "inputs/map-notations.tsv",
+  lfar_base = "data/metagenos_hap2col_lfar.csv"
 )
@@
   assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
-  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf"
+  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf",
+  tex_table = "tex/inputs/samples-table.tex",
+  gsi_fst_fig = "results/GSI_and_Fst/gsi_and_fst_fig.pdf",
+  gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
+  gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf"
 )
@@
 library(hierfstat)
+library(lubridate)
 
@@
 ```{r}
-map_notations <- read_tsv(input_list$map_notations, comment = "#")
-full_base2 %>%
-  count(repunit, collection) %>%
-  left_join
+meta <- read_csv("data/baseline_repository_meta_complete.csv.gz") %>%
+  rename(indiv = NMFS_DNA_ID...1) %>%
+  select(
+    indiv, REPORTED_LIFE_STAGE, PHENOTYPE, COLLECTION_DATE, ESTIMATED_DATE, WATERSHED, TRIB_1,
+    TRIB_2, WATER_NAME, REACH_SITE, HATCHERY, LATITUDE_F,
+    LONGITUDE_F, LOCATION_COMMENTS_F
+  ) %>%
+  mutate(
+    collection_date = mdy(COLLECTION_DATE),
+    month = month(collection_date),
+    year = year(collection_date)
+  )
+
+# get the notations on the map, and, in particular get the distinct values of
+# abbrv and name_text
+map_notations <- read_tsv(input_list$map_notations)
+
+map_notations_dist1 <- map_notations %>%
+  distinct(abbrv, repository_name, map_text, run_timing, run_timing_for_table)
+
+# join the baseline to the meta data and standardize names and nest everything on abbrv and repository name
+fbmet <- full_base2 %>%
+  select(1:4) %>%
+  left_join(meta, by = join_by(indiv)) %>%
+  rename(abbrv = collection) %>%
+  mutate(
+    repository_name = case_when(
+      !is.na(HATCHERY) ~ HATCHERY,
+      TRUE ~ WATER_NAME
+    )
+  ) %>%
+  left_join(map_notations_dist1, by = join_by(abbrv, repository_name)) %>%
+  group_by(abbrv, repository_name, run_timing_for_table, map_text, WATERSHED) %>%
+  nest()
+
+# then extract bits of information from the data and set sparkline paths
+fbmet2 <- fbmet %>%
+  mutate(
+    N = map_int(.x = data, .f = function(x) nrow(x)),
+    `Run Timing` = map_chr(.x = data, .f = function(x) x$run_timing[1]),
+    min_year = map_int(.x = data, .f = function(x) base::min(x$year, na.rm = TRUE)),
+    max_year = map_int(.x = data, .f = function(x) base::max(x$year, na.rm = TRUE)),
+    sparkline_path = str_c("tex/images/months-", abbrv, "_", str_replace_all(repository_name, " ", "-"), ".pdf")
+  )
+
+# then make the images.  Put them right into the tex directory
+source("R/month_sample_sparkline.R")
+
+dir.create("tex/images", recursive = TRUE, showWarnings = FALSE)
+
+fbmet3 <- fbmet2 %>%
+  mutate(
+    ggplot = map2(.x = data, .y = sparkline_path, .f = function(x, y) month_sample_sparkline(tib = x, path = y))
+  )
+
+# now, add all the columns we might need.
+fbmet4 <- fbmet3 %>%
+  left_join(map_notations %>% select(abbrv, repository_name, map_text, basin, ESU, `Mainstem River` ))
+
+# and now do some formatting
+fbmet5 <- fbmet4 %>%
+  ungroup() %>%
+  mutate(
+    `Sampling Months` = sprintf("\\raisebox{-0.12 em}{\\includegraphics[height=1.02em]{../%s}}", sparkline_path),
+    `Year Range` = sprintf("%d--%d", min_year, max_year)
+  ) %>%
+  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, ESU, N, `Sampling Months`, `Year Range`) %>%
+  rename(
+    `Mainstem River$^1$` = `Mainstem River`,
+    Code = abbrv,
+    `Location Name$^2$` = map_text,
+    Run = run_timing_for_table,
+    `ESU$^3$` = ESU,
+    `$N$` = N
+  )
+
+# a hack to get an hline under the heading:
+fbmet5$`Mainstem River$^1$`[1] <- str_c("\\hline ", fbmet5$`Mainstem River$^1$`[1])
+
+# Finally write out the table
+write_delim(fbmet5, delim = "&", eol = "\\tabularnewline\n", file = output_list$tex_table)
+
 ```
@@
 
+# Put them together into a single figure for the paper
+
+```{r}
+gsi_fst_fig_for_paper <- plot_grid(
+  TAF2$full_plot,
+  plot_grid(
+    get_legend(TAF2$for_external_legend),
+    get_legend(TAF2$for_internal_legend),
+    nrow = 1
+  ),
+  TAF_FST$full_plot,
+  nrow = 3,
+  rel_heights = c(7,3,7)
+)
+
+# and then save that
+ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 7, height = 17)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$gsi_fst_fig, collapse = " ")
+system(CALL)
+
+file.rename(from = output_list$gsi_fst_fig, to = output_list$gsi_fst_fig_tex)
+```
 # Add the LFAR markers to that

### 2024-01-07 map-of-samples.Rmd
A whole bunch of updates

@@
 output_list <- list(
-  map = "results/map_of_samples/map.pdf"
+  map = "results/map_of_samples/map.pdf",
+  cropmap = "results/map_of_samples/map-crop.pdf",
+  texmap = "tex/images/map-crop.pdf"
 )
@@
 labels_1line <- labels %>%
-  group_by(name_text) %>%
+  group_by(map_text) %>%
   filter(rank == max(rank)) %>%
@@
       y = label_lat, 
-      label = name_text,
+      label = map_text,
       hjust = hjust
@@
     size = 2.2,
-    label.padding = unit(0.09, "lines")
+    label.padding = unit(0.09, "lines"),
+    lineheight = 0.85
   ) +
@@
 
+file.rename(from = output_list$cropmap, to = output_list$texmap)
+
 ```
 
-# Make colored ball for inclusion into the caption
+# Make colored balls for inclusion into the caption
 

### 2024-01-07 tex/body-text-of-paper.tex
A whole bunch of updates

@@
 each location name represent the different run-timing groups of Chinook salmon
-included the samples from the location.  Color codes are as follows: \Wball~=~Winter run,
-\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run.}
+included in the collections from the location.  Color codes are as follows: \Wball~=~Winter run,
+\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run. Abbreviations used in location names as folows: R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
 \begin{center}
@@
 We are going to stick a table in here:
-\begin{table}
-\caption{Gonna make a caption. UKTR: Upper Klamath and Trinity Rivers ESU;
-CVFR: Central Valley Fall Run ESU; CVSR: Central Valley Spring Run ESU; CVWR:
-Central Valley Winter Run ESU
+%%%%%%%%%%%%%%%%
+\begin{table*}
+\caption{Collections of fish in the reference baseline.  The Sampling Months column shows
+a pictorial representation of the proportion of each collection sampled in each month.  Colors in
+this column correspond to run-timing groups as described in the caption of Figure~\protect\ref{fig:map}.}
+{\small 
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} lllllrcr}
+\hline\hline\vspace*{0.4ex}
+\input{inputs/samples-table.tex}
+\end{tabular*}
 }
-\end{table}
+\rule{3cm}{0.3pt}
+
+{\footnotesize
+$^1$R. = River.\\
+$^2$Ck. = Creek; NFH = National Fish Hatchery; R. = River; H. = Hatchery.\\
+$^3$ESU = Evolutionarily Significant Unit; CVFR = Central Valley Fall Run ESU; CVSR = Central Valley Spring Run ESU; CVWR = 
+Central Valley Winter Run ESU; CC = California Coastal ESU; UKTR = Upper Klamath and Trinity Rivers ESU.
+}
+\hrule\vspace*{0.3ex}\hrule
+\end{table*}
+%%%%%%%%%%%%%%%%%%%%%
+
+\subsection*{Genetic Markers}
+
+We will say that we took the SNPs from Clemento et al. and made amplicon based
+assays of them
+
+Then we must describe the process by which we got more microhaps.
+
+Then we can talk about assays for the subRoSA markers.
+
+Finally we mention that we have a whole section dedicated to finding the LFAR markers.
+
+\subsection*{Identifying and Designing markers from the Late-Fall Associated Region}
 
@@
 
-\begin{figure*}
-\newcommand{\examplecap}{\footnotesize This is just an example figure thrown
-in here for now to illustrate the formatting and how the captions are put in. It is not what
-we will use for the paper, but we will make another Joseph and the Magic Technicolor
-assignment table for this paper.  {\bf a)} Pairwise $F_\mathrm{ST}$ values.
-{\bf b)} Assignment rates.}
+\begin{figure}
+\newcommand{\examplecap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
+$F_\mathrm{ST}$ values ({\bf b)} for the collections in the baseline.  The codes of the populations
+appear on the outer margins of the table atop cells colored according to the run-timing group
+of the collection.  Interior cells in which the column and row correspond to the same reporting
+unit are colored according to the reporting unit.  For self-assignment tallies, the row corresponds
+to the true source collection and the column corresponds to the assigned collection. For $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
+data and the lower triangle was calculated with the whole genome sequence data (where available) from \citet{thompson2020complex}}
 \begin{center}
-\includegraphics[width=0.35\textwidth]{images/assign_and_fst_table.pdf}
+\includegraphics[width=\columnwidth]{images/gsi_and_fst_fig-crop.pdf}
 \end{center}
@@
 \label{fig:example}
-\end{figure*}
+\end{figure}
 

### 2024-01-07 tex/inputs/samples-table.tex
A whole bunch of updates

Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&Sampling Months&Year Range\tabularnewline
\hline Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
Sacramento R.&MDF&Deer Ck.&Fall&CVFR&29&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
Sacramento R.&MDF&Mill Ck.&Fall&CVFR&53&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&MDS&Mill Ck.&Spring&CVSR&51&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
Russian R.&RRF&Warm Springs H.&Fall&CC&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline

### 2024-01-07 R/month_sample_sparkline.R

@@ -1,0 +1,42 @@
+
+
+#' make a plot of months that will be suitable for sparklining
+#' @param tib a tibble like what you have in the data column of fbmet2
+#' @param path the path to write the pdf file to
+month_sample_sparkline <- function(tib, path) {
+
+  source("R/colors.R")
+
+  # first letters of each month
+  first_letters <- tibble(
+    letter = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"),
+    pos = 1:12
+  )
+
+  t2 <- tib %>%
+    filter(!is.na(month)) %>%
+    mutate(month = as.integer(month)) %>%
+    count(month, run_timing) %>%
+    mutate(fract = n/sum(n))
+
+  g <- ggplot() +
+    geom_rect(data = t2, mapping = aes(xmin = month - 0.45, xmax = month + 0.45, ymin = 0, ymax = fract, fill = run_timing, colour = run_timing)) +
+    scale_fill_manual(values = run_time_colors) +
+    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
+    geom_hline(yintercept = c(0, 1)) +
+    scale_x_continuous(limits = c(0, 13)) +
+    theme_void() +
+    theme(legend.position = "none") +
+    geom_text(
+      data = first_letters,
+      mapping = aes(x = pos, y = 0.5, label = letter),
+      hjust = 0.5, vjust = 0.5,
+      size = 10,
+      colour = "gray60"
+    )
+
+  ggsave(g, filename = path, width = 5, height = 0.6)
+
+  return(g)
+
+}

### 2024-01-08 GSI_and_Fst.Rmd
Lots more refinement and additions

@@
   pop_labels = "inputs/reference-collection-names.csv",
-  map_notations = "inputs/map-notations.tsv",
-  lfar_base = "data/metagenos_hap2col_lfar.csv"
+  map_notations = "inputs/map-notations.tsv"
 )
@@
   nrow = 3,
-  rel_heights = c(7,3,7)
+  rel_heights = c(6,1.5,6)
 )
@@
 # and then save that
-ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 7, height = 17)
+ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 7, height = 14)
 
@@
 ```
-# Add the LFAR markers to that
-
-Anthony compiled the LFAR markers for the baseline fish.  Not every fish in the baseline
-was typed at the LFAR markers.  So, let us see which ones that includes.
-```{r}
-lfar <- read_csv(input_list$lfar_base)
-```
-
-As we saw previously, the loci that we are interested in here are:
-```
-"NC_037130.1:1062935-1063235",
-"NC_037130.1:828619-828919",
-"NC_037130.1:864908-865208"
-```
-
-However, the 828619 gobbles up reads that don't all map to the right place and
-turns out not to be super informative, so we will drop it.
-
-Also, we can call these things as SNPs, not microhaps, which we will do.  The Canonical
-Variation file in the snakemake workflow tells us which position is the target SNP in each:
-
-```{r}
-focal_loci <- c(
-  "NC_037130.1:1062935-1063235",  # target SNP is the second one (position 151 in the amplicon)
-  "NC_037130.1:864908-865208"     # target SNP is the first one (position 151 in the amplicon)
-)
-
-# so we can pick those out:
-lfar_focal <- lfar[,c("Sample_ID", paste(rep(focal_loci, each = 2), c("_1", "_2"), sep = ""))]
-
-# then we will pick out just the target SNPs, too
-lfar_focal[, 2] <- str_sub(lfar_focal[[2]], 2, 2)
-lfar_focal[, 3] <- str_sub(lfar_focal[[3]], 2, 2)
-lfar_focal[, 4] <- str_sub(lfar_focal[[4]], 1, 1)
-lfar_focal[, 5] <- str_sub(lfar_focal[[5]], 1, 1)
-```
-
-We will left join this onto the baseline by the NMFS_DNA_ID:
-```{r}
-full_base_lfar <- full_base2 %>%
-  left_join(lfar_focal, by = join_by(indiv == Sample_ID))
-```
-
-Quickly, let's count up how many individuals we have typed at all these loci.
-```{r}
-lfar_check <- full_base_lfar[, c("indiv", "repunit", "collection", paste(focal_loci, 1, sep = "_"))] %>%
-  mutate(Baseline = "Baseline") %>%
-  pivot_longer(-(indiv:collection), names_to = "set") %>%
-  mutate(HasData = !is.na(value))
-
-lfar_check %>%
-  group_by(repunit, collection, set) %>%
-  summarise(
-    num_fish_with_genos = sum(HasData)
-  ) %>%
-  pivot_wider(names_from = set, values_from = num_fish_with_genos)
-
-```
-OK! That is great.  We have genotyped pretty much everyone at these LFAR markers except for about 50 of the
-FRHS and 10 of the MDF.  If need be would could toss fish that don't have genotypes at any LFAR markers
-when we make the assignment table.  Let's get a tibble of baseline individuals telling us whether or not
-they have data for at least one of the LFAR markers:
-```{r}
-at_least_one_lfar <- lfar_check %>%
-  filter(set != "Baseline") %>%
-  group_by(indiv) %>%
-  summarise(at_least_one_lfar = any(HasData == TRUE))
-
-# just see how many there are in total
-at_least_one_lfar %>%
-  count(at_least_one_lfar)
-```
-OK! We will compute the assignment table with and without those 66.
-
-## Self-assignment
-
-```{r}
-sa_lfar_all <- self_assign(full_base_lfar, gen_start_col = 5)
-
-top_ass_lf <- sa_lfar_all %>%
-  group_by(indiv, collection, repunit, inferred_repunit) %>%
-  mutate(repu_sclike = sum(scaled_likelihood)) %>%
-  group_by(indiv) %>%
-  filter(repu_sclike == max(repu_sclike)) %>%
-  arrange(indiv, desc(scaled_likelihood)) %>%
-  slice(1) %>%
-  ungroup() %>%
-  mutate(coll_f = factor(collection, levels = collection_order))
-```
-
-## Checking for individuals sampled into the wrong collections
-
-We investigate the ones that are in the wrong repunit, and look at their
-scaled_repu_likelihoods, to the others.
-```{r}
-source("R/colors.R")
-ta_corr_lf <- top_ass_lf %>%
-  filter(repunit == inferred_repunit)
-ta_wrong_lf <- top_ass_lf %>%
-  filter(repunit != inferred_repunit) %>%
-  arrange(collection, repu_sclike) %>%
-  group_by(collection) %>%
-  mutate(y = 190 + (70/9) * 1:n())
-
-ggplot() +
-  geom_histogram(
-    data = ta_corr_lf, 
-    mapping = aes(x = repu_sclike, fill = inferred_repunit)
-  ) +
-  geom_point(
-    data = ta_wrong_lf,
-    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
-    size = 3, 
-    shape = 21
-  ) +
-  facet_wrap(~coll_f, ncol = 3) +
-  scale_fill_manual(values = gen_bkgrnd_colors)
-```
-This is much like it was before, however one CHLF fish that was inferred as a spring
-run is now inferred as a CHLF.
-
-We toss those 6 that are assigned quite strongly (> 0.98) to a repunit
-that they were not sampled in. (Again, the winter-run are clearly true winter
-run sampled in the fall or spring collections, and the same is probably true
-for MDS-collected fish that are really fall run. We could follow the latter
-up with the rosa markers.)
-
-```{r}
-ta_lf_filt <- top_ass_lf %>%
-  filter(!(repunit != inferred_repunit & repu_sclike > 0.98))
-
-# get the total fraction of assigned vs misassigned fish
-ta_lf_filt %>%
-  mutate(correct = repunit == inferred_repunit) %>%
-  count(correct) %>%
-  mutate(fract = n / sum(n))
-```
-Pretty good.
-
-But, what we really want to see if the table for the collections, to see if we can distinguish
-late-fall from fall run any better.
-
-## Assignment table with all the fish
-
-```{r}
-# get the function
-source("R/taf-prep.R")
-source("R/table-as-figure.R")
-
-# get the RC_groups
-RC_groups <- pop_labels %>%
-  select(-old_name) %>%
-  distinct() %>%
-  rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
-  select(external_group, internal_group, label)
-
-X2 <- ta_lf_filt %>%
-  mutate(
-    collection_f = factor(collection, levels = collection_order),
-    inferred_coll_f = factor(inferred_collection, levels = collection_order)
-  ) %>%
-  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
-  taf_prep()
-
-# get the result
-TAF2_lf <- table_as_figure(
-  X = X2, 
-  RC_groups = RC_groups,
-  external_colors = run_time_colors,
-  internal_colors = gen_bkgrnd_colors, 
-  plot_margins = c(1, 0.1, 0.7, 0.1)
-)
-
-# and plot it
-g <- plot_grid(
-  TAF2_lf$full_plot,
-  plot_grid(
-    get_legend(TAF2_lf$for_external_legend),
-    get_legend(TAF2_lf$for_internal_legend),
-    nrow = 1
-  ),
-  nrow = 2,
-  rel_heights = c(7,3)
-)
-
-g
-```
-
-```{r}
-ggsave(g, filename = output_list$assig_table_lfar_all, width = 6, height = 8)
-```
-
-## Assignment table with only the fish that have at least one-non-missing genotype among the 3 LFAR genotypes.
-
-We can just filter out fish that don't have a non-missing genotype amongst the three LFARs and call it good.
-
-```{r}
-X2_alolf <- ta_lf_filt %>%
-  semi_join(at_least_one_lfar %>% filter(at_least_one_lfar), by = join_by(indiv)) %>%
-  mutate(
-    collection_f = factor(collection, levels = collection_order),
-    inferred_coll_f = factor(inferred_collection, levels = collection_order)
-  ) %>%
-  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
-  taf_prep()
-
-# get the result
-TAF2_alolf <- table_as_figure(
-  X = X2_alolf, 
-  RC_groups = RC_groups,
-  external_colors = run_time_colors,
-  internal_colors = gen_bkgrnd_colors, 
-  plot_margins = c(1, 0.1, 0.7, 0.1)
-)
-
-# and plot it
-g <- plot_grid(
-  TAF2_alolf$full_plot,
-  plot_grid(
-    get_legend(TAF2_alolf$for_external_legend),
-    get_legend(TAF2_alolf$for_internal_legend),
-    nrow = 1
-  ),
-  nrow = 2,
-  rel_heights = c(7,3)
-)
-
-g
-```
-
-```{r}
-ggsave(g, filename = output_list$assig_table_lfar_no_miss, width = 6, height = 8)
-```
-
-OK, that is good to see.  The missassignments to or from CHLF are *not* because
-they don't have data at the LFAR.  So, we can just use the full data set.

### 2024-01-08 lfar-allele-freqs.Rmd
Lots more refinement and additions

---
title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('references.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
# inputs:
input_list <- list(
  final_baseline = "data/cali-chinook-baseline.rds",
  pop_labels = "inputs/reference-collection-names.csv"
)
# outputs:
output_list <- list(
  lfar_tex_table = "tex/inputs/lfar-freqs.tex"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Introduction

The goal here is simply to get the allele frequenccies in different reporting units
at the LFAR markers into a table format.

First, get the baseline and attach the collection and repunit names that we want.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_rds(input_list$final_baseline) %>%
  filter(collection != "SacNonW")

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything())

# finally, get the order of collections and repunits we want
repunit_order <- rev(unique(pop_labels$repunit))
```

Now, whittle that down to just the LFAR and pivot it longer:
```{r}
lfar_base <- full_base2 %>%
  select(indiv:collection, starts_with("NC_037130.1")) %>%
  pivot_longer(
    cols = starts_with("NC_037130.1"), 
    names_to = c("locus", "gene_copy"),
    names_pattern = "(NC_037130.1.*)_([12])",
    values_to = "allele"
  )
```

And now we count up the alleles within repunits
```{r}
lfar_cnts <- lfar_base %>%
  filter(!is.na(allele)) %>%
  count(repunit, locus, allele) %>%
  complete(repunit, locus, allele, fill = list(n = 0L)) %>%
  group_by(repunit, locus) %>%
  mutate(
    tot_n = sum(n),
    fract = n / tot_n
  ) %>%
  ungroup()

lfar_cnts
```

Now, we just want to focus on the alleles that are most frequent in the late-fall.
```{r}
focal <- lfar_cnts %>%
  filter(repunit == "CV-Late-Fall") %>%
  group_by(locus) %>%
  filter(n == max(n))

alle_tib <- lfar_cnts %>%
  semi_join(focal, by = join_by(locus, allele)) %>%
  mutate(
    repunit_f = factor(repunit, levels = repunit_order),
    f = sprintf("%.3f", fract)
  ) %>%
  select(-repunit, -allele, -n, -fract) %>%
  pivot_wider(names_from = locus, values_from = c(f, tot_n)) %>%
  arrange(repunit_f) %>%
  select(repunit_f, `f_NC_037130.1:864908-865208`, `tot_n_NC_037130.1:864908-865208`, `f_NC_037130.1:1062935-1063235`, `tot_n_NC_037130.1:1062935-1063235`)

alle_tib
```

Now, we will print this out without any column names so that we can use it in a TeX table.
```{r}
write_delim(alle_tib, delim = "&", eol = "\\tabularnewline\n", file = output_list$lfar_tex_table, col_names = FALSE)
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2024-01-08 tex/body-text-of-paper.tex
Lots more refinement and additions

@@
 within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
-(Fall-, Late-Fall-, Winter-, and Spring-run), adult Chinook salmon can be found migrating or residing
+(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
 in freshwater every month of the year in California. NEED MORE HERE.
@@
 
-We sampled fish for the reference baselines from 7 locations on the across the two
+We sampled fish for the reference baselines from XX locations across California.
+These locations included sites in the two
 main tributaries (the Sacramento River and the San Joaquin River) within
-California's Central Valley. The baseline also includes samples collected from
-one location in the Klamath River, and three rivers of the California coast
-Fig.~\ref{fig:map}
+California's Central Valley, the Klamath-Trinity basin in Northern California, and
+YY rivers of the California coast
+(Fig.~\ref{fig:map}).
 %%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%
-
-We are going to stick a table in here:
+Collections in each location were separate into fish from the different run-timing ecotypes according
+to a variety of criteria.  (We need to talk about how these were split into spring run, fall-run, winter 
+run, etc.
+In total, the samples in the baseline represent fish from five different Evolutionarily Significant 
+Units (ESUs) of Chinook salmon \citep{waples1991pacific} A summary of the different collections 
 %%%%%%%%%%%%%%%%
 \begin{table*}
-\caption{Collections of fish in the reference baseline.  The Sampling Months column shows
+\caption{\footnotesize Collections of fish in the reference baseline.  The Sampling Months column shows
 a pictorial representation of the proportion of each collection sampled in each month.  Colors in
-this column correspond to run-timing groups as described in the caption of Figure~\protect\ref{fig:map}.}
+this column correspond to run-timing groups as described in the caption of Fig.~\protect\ref{fig:map}.}
 {\small 
@@
 
-We will say that we took the SNPs from Clemento et al. and made amplicon based
-assays of them
+To assemble the markers for this reference baseline, we drew upon several different
+sources, as detailed in the separate sections below.
 
-Then we must describe the process by which we got more microhaps.
+\subsubsection*{Conversion of existing SNPtype assays}
 
-Then we can talk about assays for the subRoSA markers.
+A number of SNP markers from earlier discovery efforts \citep{clemento2011discovery} had already 
+been validated as useful for GSI
+and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
+using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (CITE).
+In order to type these markers using amplicon sequencing, we\ldots
 
-Finally we mention that we have a whole section dedicated to finding the LFAR markers.
+\subsubsection*{Additional microhaplotype discovery}
 
-\subsection*{Identifying and Designing markers from the Late-Fall Associated Region}
+We undertook a new discovery process to develop additional multiallelic
+microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
+NEED TO DESCRIBE.
 
+\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}
+
+We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
+between spring-run and fall-run fish described in \citet{thompson2020complex}.
+
+Need more here.
+
+\subsubsection*{Identifying and Designing markers from the Late-Fall Associated Region}
+
+To identify loci with large allele frequency differences between late-fall and fall-run Chinook salmon, 
+we first conducted a simple case-control GWAS using the low-coverage whole genome sequencing 
+data of \citet{thompson2020complex}.  We conducted two separate association-study comparisons.  
+The first involved the 16 late-fall run fish as cases with 16 fall-run fish from the Feather River Hatchery as controls. 
+The second involved the same 16 late-fall fish as cases, but used 16 fall-run fish from the San 
+Joaquin River as controls.  The association studies were performed for each chromosome with 
+ANGSD \citep{pmid21663684,korneliussen_angsd_2014}, using the following command: {\footnotesize\tt angsd -yBin \{ybin\}  -r \{chrom\} 
+-minMapQ 30 -minQ 20 -minInd 12 -doAsso 1 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP\_pval 1e-6  
+-out \{prefix\}  -bam \{bamlist\}}, where {\tt \{ybin\} } is the path to a file indicating the cases and 
+controls, 
+{\tt \{chrom\} } indicates the chromosome name, {\tt \{prefix\} } is the prefix to use for output files, and 
+{\tt \{bamlist\} } is the path to the text file holding paths to the BAM files, one for each individual, ordered as in {\tt \{ybin\} }.
+
+There was only one peak in association scores that was shared between both comparisons (see 
+Results). From this peak we identified candidates for followup genotyping by including the 10 
+SNPs with the lowest association p-values, and supplementing those with additional SNPs that 
+showed:
+\begin{itemize}
+\item An estimated absolute allele frequency difference (between late-fall and all 32 fall-run fish)  $>0.75$,  with an approximate lower confidence interval of that figure of 0.5,  or
+\item An estimated absolute frequency difference of >0.5 with a lower confidence interval >0.25 and 
+with either fall-run or late-fall run being nearly fixed for one of the alleles, or
+\item An annotation from snpEff \citep{cingolani2012program} of “HIGH” or “MODERATE”, or
+\item An absolute frequency difference $>0.48$ and a lower confidence interval of $>0.25$ and a 
+genomic  coordinate $> 2.2$~Mb.
+\end{itemize}
+The final condition was implemented to gather several SNPs within the flanking region to the right of 
+the main peak of association.
+
+For the candidate markers found by the above process, we designed amplification primers using 
+Primer3 \citep{koressaar2007enhancements,untergasser2012primer3} and tested amplification in a 
+subset of Chinook salmon 
+samples. The primer pairs that amplified sequence that reliably mapped back to the association- 
+peak region were then typed
+in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring 
+run, and 188 fall/late-fall run. We divided the fish identified as fall/late-fall via the 125-locus genetic 
+stock identification data set into fall run and late fall run according to their sampling date at Keswick: 
+99 fish arriving before Dec 26 were classified as fall run and 89 fish arriving after Dec 26 were 
+classified as late-fall run.   The allele frequencies of the target SNPs at the remaining amplicons 
+were 
+estimated for each of these four groups: winter, spring, fall, and late-fall, and we retained
+amplicons showing large allele frequency  differences between late-fall fish and all other runs.
+
+\section*{Results}
+
+\subsection*{Genetic Markers}
+
+\subsubsection*{Late-fall Associated Region Markers}
+
+The case-control association test revealed a single peak on chromosome 34 with large allele frequency differences
+between Central Valley late-fall and fall-run fish.  This peak was evident both in the comparison of
+late-fall fish to fall-run fish in the Feather River Hatchery and of late-fall fish to the fall-run fish in
+the San Joaquin River (Fig.~\ref{fig:lfar-assoc}).
+%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\newcommand{\lfarcap}{\footnotesize Negative log base 10 of association $p$-values for
+individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
+with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
+to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
+upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
+panel is the comparison of late-fall to San Joaquin River fall. }
+\includegraphics[width=\textwidth]{images/lfar-assoc-faceted.jpg}
+\caption[\lfarcap]{\lfarcap}
+\label{fig:lfar-assoc}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
+The Manhattan plot of Fig.~\ref{fig:lfar-assoc} shows sporadic SNPs with low association
+$p$-values.  These are not shared between the two different late-fall to fall comparisons, and
+sites that we investigated were concentrated in areas of poor mapping suggesting that their
+$p$-values were artifactual.  By contrast, a large number of sites near the prominent peak on
+chromosome 34 had large allele frequency differences between late-fall and fall run.  Taking the
+top 10 most associated SNPs, and including additional ones from the filtering criteria given in the
+Methods, yielded 49 candidate markers for which we designed primers for amplication.  Only 8
+of these 49 primer pairs yielded reliable amplification and mapping.   Of these 8, two of the SNPs 
+showed inconsistent genotyping and were removed from consideration.  Of the remaining 6, only 3 
+showed marked allele frequency differences ($> 0.7$) between late-fall and fall run.  These three 
+SNPs (Chr34:828,768,  Chr34:865,057, and Chr34:1,063,084) also had pronounced differences in 
+allele frequency between 
+late-fall and spring or winter run.
+
+Analysis of the mapping of amplicons for Chr34:828,768 
+showed that many ($\approx$89\%) of the reads were off-target (aligning to other chromosomes, 
+etc.), making it costly (in terms of sequencing effort) to include the marker in the baseline panel. 
+Consequently it was dropped.  The frequency of the different alleles at the two remaining loci
+within different reporting units in the baseline are shown in Table~\ref{tab:lfar-freqs}.
+%%%%%%%%%%%%%%
+\begin{table}
+\caption{\footnotesize Allele frequencies across reporting units of the two late-fall associated
+markers.  Frequencies are given for the alleles most common amongst late-fall Chinook: nucleotide 
+{\tt G} at locus Chr34:865,057 and {\tt A} at Chr34:1,063,084.  $N$ is the total number of gene 
+copies available for estimating the allele frequency.}
+\label{tab:lfar-freqs}
+{\small
+\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrr}
+\hline\hline
+& \multicolumn{2}{c}{\underline{Chr34:865,057}} & \multicolumn{2}{c}{\underline{Chr34:1,063,084}} \\
+Reporting Unit & freq. & $N$ & freq. & $N$ \\ \hline
+\input{inputs/lfar-freqs.tex}
+\end{tabular*}
+}
+\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\end{table}
+%%%%%%%%%%%%%%%
 \subsection*{Genetic Variation}
@@
 to the true source collection and the column corresponds to the assigned collection. For $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
-data and the lower triangle was calculated with the whole genome sequence data (where available) from \citet{thompson2020complex}}
+data and the lower triangle was calculated with the whole genome sequence data (where available) 
+from \citet{thompson2020complex}.  NOTE: I THINK WE SHOULD MAKE THE LOWER 
+DIAGONAL THE P-VALUE FROM A TEST OF DIFFERENTIATION OR SOMETHING LIKE THAT.}
 \begin{center}

### 2024-01-08 tex/data-accessibility.tex
Lots more refinement and additions

@@
 
-{\sc WGSassign} is available as a Python package with these associated links:
+All data and code needed to reproduce the results here are available online.
 \begin{itemize}
-\item Development version and entire revision history on GitHub: \url{https://anonymous.4open.science/r/WGSassign-7155}
-\item Zenodo archive of initial package release: UPDATE AFTER ANONYMOUS REVIEW
-\item Online version of data and scripts used in paper: \url{https://anonymous.4open.science/r/WGSassign-manuscript-data-3072}
-\item Data repository with full datasets used in paper. UPDATED WHEN JOURNAL PROVIDES DOI. \url{https://dryad.something.or.other}
+\item Online version of data and scripts used in paper: \url{https://github.com/eriqande/california-chinook-microhaps}
+\item Archived version of data and scripts used in paper: \url{zenodo link when made}.
+\item Data repository with full baseline data set. UPDATED WHEN JOURNAL PROVIDES DOI. \url{https://dryad.something.or.other}
 \end{itemize}

### 2024-01-08 tex/author-title-etc.tex
Lots more refinement and additions
spacing in author affiliations.

@@
 % All caps tends to look better
-\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
-\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
+\newcommand{\myAuthors}{Eric C.
+Anderson$^{*,\S}$, Anthony J.
+Clemento$^{*}$, Matthew A.
+Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
+\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
 
@@
 \newcommand{\myAffiliations}{
-$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,Santa Cruz, California, USA.
+$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA. $^\dagger$Current address: Mac's home address in AK.
 }

### 2024-01-10 tex/body-text-of-paper.tex
rivers added

@@
 California's Central Valley, the Klamath-Trinity basin in Northern California, and
-YY rivers of the California coast
+3 rivers of the California coast (Russian, Eel, etc)
 (Fig.~\ref{fig:map}).

### 2024-01-10 tex/body-text-of-paper.tex
Added some edits to SNPtype conversion
Added Yoshiyama historical paper citation
quick edit

--- 2024-01-08 tex/body-text-of-paper.tex
@@
 subset of Chinook salmon 
-samples. The primer pairs that amplified sequence that reliably mapped back to the association- 
-peak region were then typed
+samples. The primer pairs that amplified sequence that reliably mapped back to the association-peak region were then typed
 in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring 

### 2024-01-10 tex/body-text-of-paper.tex

@@
 (fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California. NEED MORE HERE.
+in freshwater every month of the year in California. NEED MORE HERE. Certainly we should cite
+\citet{yoshiyama1998historical}.
 
@@
 using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (CITE).
-In order to type these markers using amplicon sequencing, we\ldots
+In order to type these markers using amplicon sequencing, we added Illumina sequencing
+primers following the protocol in \citep{campbell2015genotyping}.
 

### 2024-01-11 tex/body-text-of-paper.tex
Rough draft of SNPtype conversion and mapping to genome (stuck at end of microhaplotype discovery section for now).

@@
 using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (CITE).
-In order to type these markers using amplicon sequencing, we added Illumina sequencing
-primers following the protocol in \citep{campbell2015genotyping}.
+In order to type these markers using amplicon sequencing, we followed the methods in
+\citet{campbell2015genotyping} and added the Illumina small RNA sequencing primer and 
+read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus 
+Specific Primers respectively.  We then dropped any loci that produced genotypes that were 
+frequently inconsistent with genotypes produced with the SNPtype assays and any loci that 
+did not amplify well or took over PCRs when run with the other loci in the panel.  We then 
+used Freebayes \citep{garrison2012haplotype} to identify non mulltinucleotide or complex variants
+in addition the the SNPs targeted by the original TaqMan assays.
 
@@
 NEED TO DESCRIBE.
+
+After we designed these loci, we compiled the reference sequences for these and the
+other loci and mapped them to the Chinook v1.0 genome assembly \citep{christensen2018chinook}
+with both the Burrows-Wheeler Aligner (CITE) and the BLAST-like Alignment tool \citep{kent2002blat}.
+Genome locations found with both aligners were assumed to be the most likely locations
+of these target amplicons in the genome.
 

### 2024-01-25 tex/body-text-of-paper.tex
edits to Methods

@@
 %%%%%%%%%%%%%%%
-Collections in each location were separate into fish from the different run-timing ecotypes according
-to a variety of criteria.  (We need to talk about how these were split into spring run, fall-run, winter 
+Collections in each location were separated into fish from the different run-timing ecotypes according
+to a variety of criteria. The last vestiges of winter-run Chinook salmon are maintained at the Livingston
+Stone National Fish Hatchery for supplementation (released as juveniles into the river) and as a captive broodstock;
+sampling was performed by hatchery staff during spawning.   
+(We need to talk about how these were split into spring run, fall-run, winter 
 run, etc.

### 2024-01-25 tex/body-text-of-paper.tex
more Methods

@@
 Stone National Fish Hatchery for supplementation (released as juveniles into the river) and as a captive broodstock;
-sampling was performed by hatchery staff during spawning.   
+sampling of this population was performed by hatchery staff during spawning. In Central Valley Rivers with both 
+spring and fall ecotypes (i.e. Mill, Deer and Butte Creeks), previous research has documented distinct/disparate/unique
+spawning times (Julian date) for each ecotype (REF?). As such, sampling targeted carcasses found in each of these 
+rivers at times representative of their run-timing designation, which we were able to further verify using their genotypes
+at the RoSA markers. At the time of sampling, only fall-run Chinook were found in the San Joaquin River. Run-timing 
+designation at the Feather River Hatchery is more complicated but is managed at a fine-scale by hatchery practices, 
+where early returning fish in May and April (expressing the Spring-run life history) are visibly tagged and re-released 
+to the river. Subsequently n the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
+as Spring-run broodstock. Individual-based analysis
+
 (We need to talk about how these were split into spring run, fall-run, winter 

### 2024-01-26 tex/body-text-of-paper.tex
working...

@@
 where early returning fish in May and April (expressing the Spring-run life history) are visibly tagged and re-released 
-to the river. Subsequently n the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
-as Spring-run broodstock. Individual-based analysis
+to the river. Subsequently in the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
+as Spring-run broodstock. Individual-based analyses of RoSA genotypes have demonstrated
 

### 2024-02-13 002-GSI_and_Fst.Rmd
remove bibliiography line, update gitignore

--- 2024-01-08 GSI_and_Fst.Rmd
@@
     toc_float: true
-bibliography: "`r here::here('references.bib')`"
+
 ---
@@
 LFAR from this, because those were chosen specifically to be weird in the late fall.
-```{r}
+```{r, cache=TRUE}
 full_base_long <- full_base2 %>% 

### 2024-02-14 002-GSI_and_Fst.Rmd
Working through the CKMRsim stuff.

@@
   select(-repunit) %>%
-  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
+  left_join(pop_labels %>% select(select(old_name, collection, repunit)), by = join_by(old_name)) %>%
   select(-old_name) %>%

### 2024-02-14 ckmr-sim-evaluation.Rmd
Working through the CKMRsim stuff.

---
title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/cali-chinook-baseline.rds",
    pop_labels = "inputs/reference-collection-names.csv",
    locus_info = "inputs/ChinookFullPanelLociTable.csv",
    chrom_lengths = "inputs/Otsh_v1.0_chrom_lengths.tsv"
  )
}
```

## Wrangle the data

```{r}
library(tidyverse)
library(CKMRsim)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_rds(input_list$final_baseline) %>%
  filter(collection != "SacNonW")

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(old_name, collection, repunit, ckmr_sim_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, ckmr_sim_group, everything())

# after that, we can nest things into separate ckmr_sim_groups
ckmr_sim_nests <- full_base2 %>%
  group_by(ckmr_sim_group) %>%
  nest()

# now, we also want to get the genome positions of things.
# we take its position to be the start position of the amplicon
gen_coords <- read_csv(input_list$locus_info) %>%
  select(Locus, AmpliconGenomeV1Location) %>%
  separate(AmpliconGenomeV1Location, into = c("Chrom", "pos_str"), sep = ":") %>%
  separate(pos_str, into = c("Start", "Stop"), sep = "-", convert = TRUE) %>%
  rename(Pos = Start) %>%
  select(-Stop) %>%
  arrange(Chrom, Pos)

# now we have to do some funny stuff to deal with things on unplaced scaffolds.
# We will just sprinkle scaffolds into the existing chromosome lengths from the
# Otsh_v1.0 genome
chrom_lengths <- read_tsv(input_list$chrom_lengths)

# get the unique unplaced scaffolds and assign them randomly to a position
# in an assembled chromosome.  We set it up so that markers on the same scaffold
# will not extend beyond the end of the chromosome
set.seed(5)
rando_coords <- gen_coords %>%
  filter(str_detect(Chrom, "NW_") | is.na(Chrom)) %>%
  group_by(Chrom) %>%
  mutate(max_pos = max(Pos)) %>%
  select(Chrom, max_pos) %>%
  distinct() %>%
  mutate(
    # this gets the chromosome they are assigned to
    ass_chrom = sample(chrom_lengths$chrom, size = n(), prob = chrom_lengths$num_bases, replace = TRUE)
  ) %>% # now we do some stuff to place the scaffold randomly on the chromosome
  left_join(chrom_lengths, by = join_by(ass_chrom == chrom)) %>%
  mutate(
    ass_pos = floor(runif(n = n(), min = 1, max = num_bases - max_pos - 1 ))
  ) %>%
  select(Chrom, ass_chrom, ass_pos) %>%
  ungroup()

# now we have the information to randomly stick the scaffold markers onto
# chromosomes in the genome.  We do this by joining things onto gen_coords
gen_coords2 <- gen_coords %>%
  left_join(rando_coords, by = join_by(Chrom)) %>%
  mutate(
    Pos = case_when(
      is.na(ass_pos) ~ Pos,
      TRUE ~ ass_pos + Pos
    ),
    ass_chrom = case_when(
      is.na(ass_chrom) ~ Chrom,
      TRUE ~ ass_chrom
    )
  ) %>%
  mutate(Chrom = ass_chrom) %>%
  select(Locus, Chrom, Pos)

```

## Now, a series of functions to operate on each nest and do CKMRsim things

```{r}
#' @param X a tibble like that nested in a single row of ckmr_sim_nests
#' @param GC2 essentially the gen_coords2 object from above
ckmr_afreqs_ready <- function(X, GC2) {
  tmp <- X %>%
    select(-(repunit:sample_type)) %>%
    pivot_longer(
      -indiv,
      names_to = c("Locus", "gene_copy"),
      values_to = "Allele",
      names_pattern = "^(.*)_([12])$"
    ) %>%
    filter(!is.na(Allele)) %>%
    count(Locus, Allele) %>%
    group_by(Locus) %>%
    filter(n() > 1) %>% # filter out monomorphic loci
    mutate(
      Freq = n / sum(n),
    ) %>%
  ungroup() %>%
  left_join(GC2, by = join_by(Locus)) %>%
  select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq))

  no_pos_posse <- tmp %>% 
    filter(is.na(Chrom))

  if(nrow(no_pos_posse) > 0) {
    warning("Some loci without genomic positions: ", paste(unique(no_pos_posse$Locus), collapse = ", "), "\n\n")
    warning("Those loci were dropped, and the analysis should survive, but you should see what's up!")
  }

  # then, return thing without the no_pos_posse markers
  tmp %>%
    filter(!is.na(Chrom)) %>%
    mutate(
      AlleIdx = NA,
      LocIdx = NA
    ) %>%
    CKMRsim::reindex_markers()

}

# here is one to create a ckmr object.
# Y is simply the output of ckmr_afreqs_ready()
ckmr_create <- function(Y) {
  create_ckmr(
    D = Y,
    kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
    ge_mod_assumed = ge_model_TGIE,
    ge_mod_true = ge_model_TGIE,
    ge_mod_assumed_pars_list = list(epsilon = 0.005),
    ge_mod_true_pars_list = list(epsilon = 0.005)
  )
}

# This does the simulate_Qij step and then also estimates the FPRs.
ckmr_simQ <- function(C) {
  simulate_Qij(
    C, 
    calc_relats = c("PO", "FS", "HS", "U"),
    sim_relats = c("PO", "FS", "HS", "U"),
    unlinked = FALSE,
    pedigree_list = pedigrees
  )
}

```

This is just here while I develop this:
```{r}
X <- ckmr_sim_nests$data[[5]]
GC2 <- gen_coords2

Y <- ckmr_afreqs_ready(X, GC2)

C <- ckmr_create(Y)

Q <- ckmr_simQ(C)
```

### 2024-03-08 ckmr-sim-evaluation.Rmd
Update the paper a bit.

@@
 # This does the simulate_Qij step and then also estimates the FPRs.
-ckmr_simQ <- function(C) {
+ckmr_simQ <- function(C, unlinked = FALSE) {
   simulate_Qij(
     C,
+    reps = 5e4,
     calc_relats = c("PO", "FS", "HS", "U"),
     sim_relats = c("PO", "FS", "HS", "U"),
-    unlinked = FALSE,
+    unlinked = unlinked,
     pedigree_list = pedigrees
@@
 
-Q <- ckmr_simQ(C)
+Q_link <- ckmr_simQ(C)
+Q_unl <- ckmr_simQ(C, unlinked = TRUE)
+
+# get FPRs and FNRs for a variety of relationships.
+
+# first for resolving unrelated individuals from PO and FS
+mc_sample_simple(
+  Q = Q_link,
+  nu = c("PO", "FS"),
+  de = "U",
+  tr = "U",
+  method = "IS",
+  FNRs = seq(0.01, 0.3, by = 0.01)
+)
 ```

### 2024-03-08 tex/body-text-of-paper.tex
Update the paper a bit.

@@
 Genetic markers have long been used in the 
-management of Pacific salmon, and salmon conservation and management has been
-a driving force behind a number of advances in molecular ecology,
+management of Pacific salmon, and salmon conservation and management have been
+behind a number of advances in molecular ecology,
 from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
@@
 PBT is that the genetic markers used for PBT could also be useful for GSI (and
-vice versa).  This
+vice versa).
 
@@
 (fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California. NEED MORE HERE. Certainly we should cite
-\citet{yoshiyama1998historical}.
+in freshwater every month of the year in California. \comment{NEED MORE HERE. Certainly we should cite
+\citet{yoshiyama1998historical}.}
 
@@
 
-More verbiage
+\comment{More verbiage}
 
@@
 and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
-using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (CITE).
+using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays \comment{(CITE)}.
 In order to type these markers using amplicon sequencing, we followed the methods in
@@
 frequently inconsistent with genotypes produced with the SNPtype assays and any loci that 
-did not amplify well or took over PCRs when run with the other loci in the panel.  We then 
-used Freebayes \citep{garrison2012haplotype} to identify non mulltinucleotide or complex variants
-in addition the the SNPs targeted by the original TaqMan assays.
+did not amplify well or those that amplified so well that they were grossly over-represented
+among the reads on the sequencer.  We then 
+used Freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
+microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
+that were considered multinucleotide or complex variants by Freebayes.
 
@@
 microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
-NEED TO DESCRIBE.
+\comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
+candidate to then follow up on with amplicon sequencing, etc.}
 
-After we designed these loci, we compiled the reference sequences for these and the
-other loci and mapped them to the Chinook v1.0 genome assembly \citep{christensen2018chinook}
-with both the Burrows-Wheeler Aligner (CITE) and the BLAST-like Alignment tool \citep{kent2002blat}.
-Genome locations found with both aligners were assumed to be the most likely locations
-of these target amplicons in the genome.
-
-\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}
-
-We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish described in \citet{thompson2020complex}.
-
-Need more here.
+After designing and validating an amplicon-sequencing approach to type both the
+existing SNPtype assays and the product of the additional microhaplotype
+discovery process we identified their location with the genome. We mapped the reference
+sequences for these markers 
+to the Chinook v2.0 genome assembly \citep{christensen2018chinook}
+with both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
+We report the genomic coordinates these target amplicons found with both alignment methods.
 
@@
 
+\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}
+
+We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
+between spring-run and fall-run fish described in \citet{thompson2020complex}.
+\comment{Need more here. I personally am of the thought that we should not include these
+for GSI, but should use them for identifying the genotypes of fish assigned to populations
+using the rest of the markers.}
+
+\subsection*{Power for genetic stock identification and population assignment}
+
+\subsection*{Power for relationship inference}
+
 \section*{Results}
@@
 \subsection*{Genetic Markers}
+
+\subsubsection*{Conversion of existing SNPtype assays}
+
+Of \comment{XX} existing SNPtype assays, we were able to convert \comment{YY}
+into reliable amplicon-sequenced assays that were included in the panel.
+
+\subsubsection*{Additional microhaplotype discovery}
+
+We tested \comment{XX} candidates from our additional microhaplotype discovery and
+retained \comment{185} of those in the panel.
+
+Of these \comment{185} loci that originated from SNPtype assays or from our additional
+microhaplotype discovery, \comment{179} of them mapped identically to a single
+location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
+Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
+the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
+that was identical in both bwa mem and BLAT, two of them had secondary alignments that
+disagreed slightly between BLAT and bwa mem, and one of them had multiple
+non-primary alignments (\comment{SuppTable}).
 

### 2024-03-14 001-map-of-samples.Rmd
Number the Rmds, (but not their output dirs)
Update the 001-map for the new collections

--- 2024-01-07 map-of-samples.Rmd
@@
 input_list <- list(
+  oregon_rivers = "inputs/Rivers_OR/Rivers_OR.shp"
 )
@@
   st_transform(., st_crs(state_prov))
+
+or_rivers <- st_read(input_list$oregon_rivers)
+```
+
+## Crop the basemap, state lines coast lines
+
+```{r}
+# important to put them in this order and named like this
+domain <- c(
+  xmin = -127.9,
+  xmax = -117.4,
+  ymin = 36,
+  ymax = 43.25
+)
+sf_use_s2(FALSE)
+nat_crop <- crop(nat.earth, y = ext(domain))
+state_subset <- st_crop(state_prov, domain)
+coastline_cropped <- st_crop(coastline, domain)
+
 ```
@@
 
+# same with Blue Creek
+blue_area <- c(
+  xmin = -124,
+  xmax = -123.5,
+  ymin = 41.2,
+  ymax = 41.8
+)
+
+blue_candi <- all_rivers %>%
+  filter(str_detect(gnis_name, "Blue Creek")) %>%
+  st_crop(blue_area)
+
+# rogue river:
+rogue <- or_rivers %>%
+  filter(name == "Rogue") %>%
+  mutate(gnis_name = "Rogue River")
+
 our_rivers <- all_rivers %>%
@@
     "Trinity River",
-    "Klamath River"
+    "Klamath River",
+    "Eel River",
+    "Smith River"
     #"Deer Creek",
@@
   ) %>%
-  bind_rows(mill_candi)
+  bind_rows(mill_candi) %>%
+  bind_rows(blue_candi) %>%
+  bind_rows(rogue)
 
@@
   geom_sf(data = our_rivers, aes(colour = gnis_name)) +
+  geom_sf(data = coastline_cropped) +
   geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
@@
 That looks good.  So let us proceed.
-
-## Crop
-```{r}
-# important to put them in this order and named like this
-domain <- c(
-  xmin = -127.9,
-  xmax = -117.4,
-  ymin = 36,
-  ymax = 42.6
-)
-sf_use_s2(FALSE)
-nat_crop <- crop(nat.earth, y = ext(domain))
-state_subset <- st_crop(state_prov, domain)
-coastline_cropped <- st_crop(coastline, domain)
-
-```
 
@@
 mm <- base_map + 
-  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", size = 0.4) +
+  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", linewidth = 0.4) +
   geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
@@
 inset_world <- ggplot() +
-  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", size = 0.1) +
-  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
+  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", linewidth = 0.1) +
+  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", linewidth = 0.1) +
   geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3, linewidth = 0.1) +
@@
   draw_plot(mm) +
-  draw_plot(inset_world, x = 0.181, y = 0.139, width = 0.20, height = 0.17)
+  draw_plot(inset_world, x = 0.215, y = 0.139, width = 0.20, height = 0.17)
 
@@
 
-file.rename(from = output_list$cropmap, to = output_list$texmap)
+file.copy(from = output_list$cropmap, to = output_list$texmap)
 

### 2024-03-15 002-GSI_and_Fst.Rmd
Trying some GenePop Stuff, but bailing on it now

--- 2024-02-13 002-GSI_and_Fst.Rmd
@@
 input_list <- list(
-  final_baseline = "data/cali-chinook-baseline.rds",
+  #final_baseline = "data/cali-chinook-baseline.rds",
+  final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
   pop_labels = "inputs/reference-collection-names.csv",
@@
 # non-winter Sacto fish of unknown provenance
-full_base <- read_rds(input_list$final_baseline) %>%
+full_base <- read_csv(input_list$final_baseline) %>%
   filter(collection != "SacNonW")
@@
   select(-repunit) %>%
-  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
+  left_join(pop_labels %>% select(old_name, collection, repunit), by = join_by(old_name)) %>%
   select(-old_name) %>%
-  select(indiv, repunit, collection, sample_type, everything())
+  select(indiv, repunit, collection, sample_type, everything()) %>%
+  filter(!duplicated(indiv))  # necessary cuz clemento stuck mill and deer together into a single code
 
@@
 
+# Remove this once I get the full meta data
+fbmet <- fbmet %>% filter(!is.na(WATERSHED))
+
 # then extract bits of information from the data and set sparkline paths
@@
     N = map_int(.x = data, .f = function(x) nrow(x)),
+    N_date = map_int(.x = data, .f = function(x) sum(!(is.na(x$collection_date)) & yday(x$collection_date) != 1) ),  # we use YYYY-01-01 to denote year is known but day is missing in the data repository, so here we filter those out.
     `Run Timing` = map_chr(.x = data, .f = function(x) x$run_timing[1]),
@@
   ) %>%
-  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, ESU, N, `Sampling Months`, `Year Range`) %>%
+  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, ESU, N, N_date, `Sampling Months`, `Year Range`) %>%
+  mutate(map_text = str_replace(map_text, ",.*$", "")) %>% # this removes the abbreviations after the commas on the map names
+  mutate(N_date = replace_na(N_date, 0L)) %>%
   rename(
@@
     `ESU$^3$` = ESU,
-    `$N$` = N
-  )
+    `$N$` = N,
+    `$N_\\mathrm{day}$` = N_date
+  ) %>%
+  arrange(factor(Code, levels = collection_order))
 
@@
 Now that we have the `upper_tri_fst` values in the right format, we just need
-to add the WGS values to the lower triangle part of the matrix.
+to add values to the lower triangle part of the matrix.
 
-# Adding the WGS Fst values and making a figure
+# Adding genic differentiation values to lower triangle and making a figure
 
+We are going to use genepop to compute p-values for differentiation.  To do
+that we need to make a genepop file.
 ```{r}
-# for now I am just going to put the WGS Fst's into the lower diagonal, but I think it will be better, ultimately,
-# to but the bootstrapped CI's in there.
-wgs_mat <- read_rds("../anderson-et-al-chinook-ecotypic-genetic-architecture/outputs/001/full-wgs-fst-mat.rds")
+gpt <- full_base2 %>%
+  select(-repunit, -sample_type) %>%
+  pivot_longer(
+    cols = -c(indiv, collection),
+    names_to = c("locus", "gene_copy"),
+    values_to = "allele",
+    names_pattern = "(.*)[_.]([12])$"
+  ) %>%
+  group_by(locus) %>%
+  mutate(alle_int = replace_na(as.integer(factor(allele)), 0L)) %>%
+  ungroup() %>%
+  mutate(alle_str = sprintf("%02d", alle_int)) %>%
+  select(-allele, -alle_int) %>%
+  pivot_wider(
+    names_from = gene_copy,
+    values_from = alle_str
+  ) %>%
+  mutate(geno = str_c(`1`, `2`)) %>%
+  select(-`1`, -`2`) %>%
+  pivot_wider(
+    names_from = locus,
+    values_from = geno
+  ) %>%
+  select(-indiv)
+
+gpt_nest = gpt %>%
+  mutate(collection2 = str_c(collection, ","), .after = collection) %>%
+  group_by(collection) %>%
+  nest()
+
+# now we write that out to a file
+outf <- "results/GSI_and_Fst/genepop.txt"
+cat("Baseline data set in genepop format\n    ", file = outf)
+cat(names(gpt)[-1], sep = ",", file = outf, append = TRUE)
+cat("\n", file = outf, append = TRUE)
+for(x in gpt_nest$data) {
+    cat("POP\n", file = outf, append = TRUE); 
+    write_tsv(x, col_names = FALSE, file = outf, append = TRUE)
+  }
+```
+
+Now, we could use the genepop package to get results from that file, but it
+is crazy slow, so we do not do this.
+```r
+td_file <- "results/GSI_and_Fst/genic_tests.txt"
+genepop::test_diff(inputFile = outf, pairs = TRUE, outputFile = td_file)
+```
+
+OK! HERE IS WHAT WE WILL DO: MAKE THIS PARALLELIZABLE.  WE WILL MODIFY THE ABOVE
+TO MAKE A GENEPOP FILE FOR EVERY LOCUS, AND THEN WE WILL JUST RUN OVER ALL THOSE
+LOCI USING mclappy ON 8 CORES ON MY LAPTOP, TO MAKE IT FASTER.  THEN, WE WILL PRESENT
+THE RESULTS AS A TINY-MULTIPLE HISTOGRAM OF THE P-VALUES IN EACH CELL.  KINDA COOL!
+
+Making a figure.
+```{r}
 
@@
 # and then save that
-ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 7, height = 14)
+ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 10, height = 19)
 

### 2024-03-15 development/genepop-conversion.Rmd
Add genepop conversion to dev directory
Just didnt want to lose it.

--- 2024-02-13 002-GSI_and_Fst.Rmd
@@
 ---
-title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
-author: "Eric C. Anderson"
-date: "Last Updated: `r Sys.Date()`"
-output:
-  html_document:
-    df_print: paged
-    toc: true
-    toc_float: true
-
+title: "Genepop Conversion Etc"
+output: html_notebook
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE)
-knitr::opts_knit$set(root.dir = here::here())
-start_time <- Sys.time()
+# Making a genepop file
 
-USING_SNAKEMAKE <- FALSE
-NEG_USING_SMK <- TRUE
-if(exists("snakemake")) {
-  USING_SNAKEMAKE <- TRUE
-  NEG_USING_SMK <- FALSE
+I ended up not using this, but we might want to, ultimately, if we want to look at
+LD, etc.
+
+```{r}
+gpt <- full_base2 %>%
+  select(-repunit, -sample_type) %>%
+  pivot_longer(
+    cols = -c(indiv, collection),
+    names_to = c("locus", "gene_copy"),
+    values_to = "allele",
+    names_pattern = "(.*)[_.]([12])$"
+  ) %>%
+  group_by(locus) %>%
+  mutate(alle_int = replace_na(as.integer(factor(allele)), 0L)) %>%
+  ungroup() %>%
+  mutate(alle_str = sprintf("%02d", alle_int)) %>%
+  select(-allele, -alle_int) %>%
+  pivot_wider(
+    names_from = gene_copy,
+    values_from = alle_str
+  ) %>%
+  mutate(geno = str_c(`1`, `2`)) %>%
+  select(-`1`, -`2`) %>%
+  pivot_wider(
+    names_from = locus,
+    values_from = geno
+  ) %>%
+  select(-indiv)
+
+gpt_nest = gpt %>%
+  mutate(collection2 = str_c(collection, ","), .after = collection) %>%
+  group_by(collection) %>%
+  nest()
+
+# now we write that out to a file
+outf <- "results/GSI_and_Fst/genepop.txt"
+cat("Baseline data set in genepop format\n    ", file = outf)
+cat(names(gpt)[-1], sep = ",", file = outf, append = TRUE)
+cat("\n", file = outf, append = TRUE)
+for(x in gpt_nest$data) {
+    cat("POP\n", file = outf, append = TRUE); 
+    write_tsv(x, col_names = FALSE, file = outf, append = TRUE)
   }
 ```
-
-```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
-# currently, this block does not get evaluated, while developing,
-# but later when we snakemake-ize it all, we will evaluate it.
-if(USING_SNAKEMAKE) {
-  input_list <- snakemake@input
-  output_list <- snakemake@output
-}
-```
-
-# Input and Output Paths
-
-```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
-# inputs:
-input_list <- list(
-  final_baseline = "data/cali-chinook-baseline.rds",
-  pop_labels = "inputs/reference-collection-names.csv",
-  map_notations = "inputs/map-notations.tsv"
-)
-# outputs:
-output_list <- list(
-  assig_table = "results/GSI_and_Fst/assignment-table-full-baseline.pdf",
-  fst_table = "results/GSI_and_Fst/full-baseline-but-no-lfar-Fst-table.pdf",
-  assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
-  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf",
-  tex_table = "tex/inputs/samples-table.tex",
-  gsi_fst_fig = "results/GSI_and_Fst/gsi_and_fst_fig.pdf",
-  gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
-  gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf"
-)
-# you can create the necessary output directories like this:
-dump <- lapply(output_list, function(x)
-  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
-)
-```
-
-# Introduction
-
-Here we assess power of the baseline for population assignment
-
-```{r, message=FALSE, warning=FALSE}
-library(tidyverse)
-library(rubias)
-library(cowplot)
-library(hierfstat)
-library(lubridate)
-
-# get the full baseline with the new markers, etc.  We will toss out some
-# non-winter Sacto fish of unknown provenance
-full_base <- read_rds(input_list$final_baseline) %>%
-  filter(collection != "SacNonW")
-
-# read the pop labels for proper sorting
-pop_labels <- read_csv(input_list$pop_labels)
-
-# rename those collections and repunits
-full_base2 <- full_base %>%
-  rename(old_name = collection) %>%
-  select(-repunit) %>%
-  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
-  select(-old_name) %>%
-  select(indiv, repunit, collection, sample_type, everything())
-
-# finally, get the order of collections and repunits we want
-collection_order <- unique(pop_labels$collection)
-```
-
-# Make a table summary of all the samples
-
-I would like to break things down by general sampling location, run type, number of samples,
-month-range of sampling, and range of years.
-```{r}
-meta <- read_csv("data/baseline_repository_meta_complete.csv.gz") %>%
-  rename(indiv = NMFS_DNA_ID...1) %>%
-  select(
-    indiv, REPORTED_LIFE_STAGE, PHENOTYPE, COLLECTION_DATE, ESTIMATED_DATE, WATERSHED, TRIB_1,
-    TRIB_2, WATER_NAME, REACH_SITE, HATCHERY, LATITUDE_F,
-    LONGITUDE_F, LOCATION_COMMENTS_F
-  ) %>%
-  mutate(
-    collection_date = mdy(COLLECTION_DATE),
-    month = month(collection_date),
-    year = year(collection_date)
-  )
-
-# get the notations on the map, and, in particular get the distinct values of
-# abbrv and name_text
-map_notations <- read_tsv(input_list$map_notations)
-
-map_notations_dist1 <- map_notations %>%
-  distinct(abbrv, repository_name, map_text, run_timing, run_timing_for_table)
-
-# join the baseline to the meta data and standardize names and nest everything on abbrv and repository name
-fbmet <- full_base2 %>%
-  select(1:4) %>%
-  left_join(meta, by = join_by(indiv)) %>%
-  rename(abbrv = collection) %>%
-  mutate(
-    repository_name = case_when(
-      !is.na(HATCHERY) ~ HATCHERY,
-      TRUE ~ WATER_NAME
-    )
-  ) %>%
-  left_join(map_notations_dist1, by = join_by(abbrv, repository_name)) %>%
-  group_by(abbrv, repository_name, run_timing_for_table, map_text, WATERSHED) %>%
-  nest()
-
-# then extract bits of information from the data and set sparkline paths
-fbmet2 <- fbmet %>%
-  mutate(
-    N = map_int(.x = data, .f = function(x) nrow(x)),
-    `Run Timing` = map_chr(.x = data, .f = function(x) x$run_timing[1]),
-    min_year = map_int(.x = data, .f = function(x) base::min(x$year, na.rm = TRUE)),
-    max_year = map_int(.x = data, .f = function(x) base::max(x$year, na.rm = TRUE)),
-    sparkline_path = str_c("tex/images/months-", abbrv, "_", str_replace_all(repository_name, " ", "-"), ".pdf")
-  )
-
-# then make the images.  Put them right into the tex directory
-source("R/month_sample_sparkline.R")
-
-dir.create("tex/images", recursive = TRUE, showWarnings = FALSE)
-
-fbmet3 <- fbmet2 %>%
-  mutate(
-    ggplot = map2(.x = data, .y = sparkline_path, .f = function(x, y) month_sample_sparkline(tib = x, path = y))
-  )
-
-# now, add all the columns we might need.
-fbmet4 <- fbmet3 %>%
-  left_join(map_notations %>% select(abbrv, repository_name, map_text, basin, ESU, `Mainstem River` ))
-
-# and now do some formatting
-fbmet5 <- fbmet4 %>%
-  ungroup() %>%
-  mutate(
-    `Sampling Months` = sprintf("\\raisebox{-0.12 em}{\\includegraphics[height=1.02em]{../%s}}", sparkline_path),
-    `Year Range` = sprintf("%d--%d", min_year, max_year)
-  ) %>%
-  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, ESU, N, `Sampling Months`, `Year Range`) %>%
-  rename(
-    `Mainstem River$^1$` = `Mainstem River`,
-    Code = abbrv,
-    `Location Name$^2$` = map_text,
-    Run = run_timing_for_table,
-    `ESU$^3$` = ESU,
-    `$N$` = N
-  )
-
-# a hack to get an hline under the heading:
-fbmet5$`Mainstem River$^1$`[1] <- str_c("\\hline ", fbmet5$`Mainstem River$^1$`[1])
-
-# Finally write out the table
-write_delim(fbmet5, delim = "&", eol = "\\tabularnewline\n", file = output_list$tex_table)
-
-```
-
-# Do the self-assignment
-
-In this round, we go directly to assigning to repunit, and then to the max-likelihood
-collection within that repunit.
-
-```{r}
-full_sa <- self_assign(full_base2, gen_start_col = 5)
-```
-
-```{r}
-top_ass <- full_sa %>%
-  group_by(indiv, collection, repunit, inferred_repunit) %>%
-  mutate(repu_sclike = sum(scaled_likelihood)) %>%
-  group_by(indiv) %>%
-  filter(repu_sclike == max(repu_sclike)) %>%
-  arrange(indiv, desc(scaled_likelihood)) %>%
-  slice(1) %>%
-  ungroup() %>%
-  mutate(coll_f = factor(collection, levels = collection_order))
-```
-
-And, once again, we want to investigate the ones that are in the wrong repunit, and look at their
-scaled_repu_likelihoods, the others.
-```{r}
-source("R/colors.R")
-ta_corr <- top_ass %>%
-  filter(repunit == inferred_repunit)
-ta_wrong <- top_ass %>%
-  filter(repunit != inferred_repunit) %>%
-  arrange(collection, repu_sclike) %>%
-  group_by(collection) %>%
-  mutate(y = 190 + (70/9) * 1:n())
-
-ggplot() +
-  geom_histogram(
-    data = ta_corr, 
-    mapping = aes(x = repu_sclike, fill = inferred_repunit)
-  ) +
-  geom_point(
-    data = ta_wrong,
-    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
-    size = 3, 
-    shape = 21
-  ) +
-  facet_wrap(~coll_f, ncol = 3) +
-  scale_fill_manual(values = repunit_colors)
-```
-Which looks to me like there are really only 7 here that we are not getting right.
-Those others we will filter out and make a table of them:
-
-```{r}
-tafilt <- top_ass %>%
-  filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
-```
-
-Here we calculate the number of correctly and incorrectly assigned fish:
-```{r}
-tafilt %>%
-  mutate(correct = repunit == inferred_repunit) %>%
-  count(correct) %>%
-  mutate(fract = n / sum(n))
-```
-
-And now we will make a table of those:
-```{r}
-# get the function
-source("R/taf-prep.R")
-source("R/table-as-figure.R")
-
-# get the RC_groups
-RC_groups <- pop_labels %>%
-  select(-old_name) %>%
-  distinct() %>%
-  rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
-  select(external_group, internal_group, label)
-
-X2 <- tafilt %>%
-  mutate(
-    collection_f = factor(collection, levels = collection_order),
-    inferred_coll_f = factor(inferred_collection, levels = collection_order)
-  ) %>%
-  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
-  taf_prep()
-
-# get the result
-TAF2 <- table_as_figure(
-  X = X2, 
-  RC_groups = RC_groups,
-  external_colors = run_time_colors,
-  internal_colors = repunit_colors, 
-  plot_margins = c(1, 0.1, 0.7, 0.1)
-)
-
-# and plot it
-g <- plot_grid(
-  TAF2$full_plot,
-  plot_grid(
-    get_legend(TAF2$for_external_legend),
-    get_legend(TAF2$for_internal_legend),
-    nrow = 1
-  ),
-  nrow = 2,
-  rel_heights = c(7,3)
-)
-
-g
-```
-
-That is a considerable improvement. Let's write it out:
-```{r}
-ggsave(g, filename = output_list$assig_table, width = 8, height = 10)
-```
-
-# Calculating Fst between the collections from the microhaplotype markers
-
-We use the 'hierfstat' package to get the pairwise
-$F_\mathrm{ST}$ values (Weir & Cockerham method) between the collections from these microhaplotype data.
-
-In order to do this we need to convert the data to a lumped format.
-This means making integers of the alleles and then catenating them.
-One locus has 11 allele, so we should use two digits for each allele.  We drop the
-LFAR from this, because those were chosen specifically to be weird in the late fall.
-```{r, cache=TRUE}
-full_base_long <- full_base2 %>% 
-  select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
-  mutate(
-    Pop = as.integer(factor(collection, levels = collection_order)),
-    .before = collection
-  ) %>%
-  select(-repunit, -collection, -sample_type) %>%
-  pivot_longer(
-    cols = c(-indiv, -Pop),
-    names_to = c("locus", "gene_copy"),
-    names_pattern = "^(.*)_([12])$"
-  )
-
-hfs_ref_long <- full_base_long %>%
-  group_by(locus) %>%
-  mutate(
-    alle_int =  sprintf("%02d", as.integer(factor(value)))
-  ) %>%
-  ungroup() %>%
-  select(-value)
-
-# then we can lump the alleles 
-hfs_ref_wide <- hfs_ref_long %>%
-  group_by(indiv, Pop, locus) %>%
-  summarise(geno = str_c(alle_int[1], alle_int[2])) %>%
-  ungroup() %>%
-  mutate(geno = ifelse(str_detect(geno, "NA"), NA_character_, geno)) %>%
-  mutate(geno = as.integer(geno)) %>%
-  pivot_wider(
-    names_from = c(locus),
-    values_from = geno
-  ) %>%
-  arrange(Pop, indiv)
-
-hfstat_dat <- hfs_ref_wide %>% 
-  select(-indiv) %>%
-  as.data.frame()
-
-FST <- pp.fst(dat = hfstat_dat, diploid = TRUE)
-
-# let's also bootstrap that:
-FST_boot <- boot.ppfst(
-  dat = hfstat_dat,
-  nboot = 1000,
-  quant = c(0.025,0.975),
-  diploid = TRUE
-)
-
-# now, get the values out of it and put them in the upper diagonal
-# of a matrix for plotting with the WGS values.
-fst_mat <- FST$fst.pp 
-colnames(fst_mat) <- collection_order
-```
-
-```{r}
-
-all_values_tib <- fst_mat %>%
-  as_tibble() %>%
-  mutate(row_label = collection_order, .before = SRW) %>%
-  pivot_longer(
-    -row_label, 
-    names_to = "col_label",
-    values_to = "fst"
-  )
-
-# to get it to the right format we make it look like rubias
-# output and the pipe it into taf_prep()
-upper_tri_fst <- all_values_tib %>%
-  filter(!is.nan(fst)) %>%
-  mutate(
-    collection_f = factor(row_label, levels = collection_order),
-    inferred_coll_f = factor(col_label, levels = collection_order)
-  ) %>% 
-  select(-row_label, -col_label) %>%
-  mutate(n = sprintf("%0.3f", fst)) %>%
-  select(-fst) %>%
-  taf_prep()
-
-```
-
-Now that we have the `upper_tri_fst` values in the right format, we just need
-to add the WGS values to the lower triangle part of the matrix.
-
-# Adding the WGS Fst values and making a figure
-
-```{r}
-# for now I am just going to put the WGS Fst's into the lower diagonal, but I think it will be better, ultimately,
-# to but the bootstrapped CI's in there.
-wgs_mat <- read_rds("../anderson-et-al-chinook-ecotypic-genetic-architecture/outputs/001/full-wgs-fst-mat.rds")
-
-# once again, make it looks like rubias output and taf_prep() it.
-lower_tri_diag <- wgs_mat %>%
-  mutate(
-    collection_f = factor(pop1_f, levels = collection_order),
-    inferred_coll_f = factor(pop2_f, levels = collection_order),
-    n = fst
-  ) %>%
-  select(collection_f, inferred_coll_f, n) %>%
-  complete(collection_f, inferred_coll_f) %>%
-  filter(as.integer(collection_f) >= as.integer(inferred_coll_f)) %>%  # this gets lower tri + diagonal
-  taf_prep()
-
-# then we put those together and make a figure
-full_fst <- bind_rows(
-  upper_tri_fst,
-  lower_tri_diag
-) %>%
-  arrange(row_label, col_label)
-
-# get the result
-TAF_FST <- table_as_figure(
-  X = full_fst, 
-  RC_groups = RC_groups,
-  external_colors = run_time_colors,
-  internal_colors = repunit_colors,
-  Xs_on_diagonal = TRUE,
-  plot_margins = c(0.7, 0.1, 0.1, 0.1)
-)
-
-# and plot it
-TAF_FST$full_plot
-
-```
-
-And we can save that too:
-```{r}
-ggsave(TAF_FST$full_plot, file = output_list$fst_table, width = 6, height = 5.1)
-```
-
-# Put them together into a single figure for the paper
-
-```{r}
-gsi_fst_fig_for_paper <- plot_grid(
-  TAF2$full_plot,
-  plot_grid(
-    get_legend(TAF2$for_external_legend),
-    get_legend(TAF2$for_internal_legend),
-    nrow = 1
-  ),
-  TAF_FST$full_plot,
-  nrow = 3,
-  rel_heights = c(6,1.5,6)
-)
-
-# and then save that
-ggsave(gsi_fst_fig_for_paper, file = output_list$gsi_fst_fig, width = 7, height = 14)
-
-# let's crop that down for use, too
-CALL <- paste("pdfcrop ", output_list$gsi_fst_fig, collapse = " ")
-system(CALL)
-
-file.rename(from = output_list$gsi_fst_fig, to = output_list$gsi_fst_fig_tex)
-```

### 2024-03-15 002-GSI_and_Fst.Rmd
Update for new pops.  Also revamp FST fig.
(to have scatter plots in lower diagonals.)

@@
     `$N$` = N,
-    `$N_\\mathrm{day}$` = N_date
+    `$N_\\mathrm{day}^4$` = N_date
   ) %>%
@@
 
+Rubias_for_later <- X2 # keeping this for later...
+
 # get the result
@@
 LFAR from this, because those were chosen specifically to be weird in the late fall.
-```{r, cache=TRUE}
+```{r}
 full_base_long <- full_base2 %>% 
@@
 
-# Adding genic differentiation values to lower triangle and making a figure
+# Adding allele frequency scatter plots to the lower diagonal
 
-We are going to use genepop to compute p-values for differentiation.  To do
-that we need to make a genepop file.
+I thought about doing differentiation tests, or histograms of per-locus
+Fst values, but then I realized that it actually might be way more informative
+to put small-multiple scatter plots of allele frequencies in each of the lower-diagonal
+cells of the Fst figure.  I think it might look cool.
+
+We will also standardize each sample size (at each locus) to the minimum
+number of observed genotypes in a population.
+
+## Calculate allele frequencies for each population and standardize to lowest sample size
+
+Straightforward tidyverse stuff:
 ```{r}
-gpt <- full_base2 %>%
+
+# long format with no missing data
+long_genos <- full_base2 %>%
   select(-repunit, -sample_type) %>%
@@
   ) %>%
+  filter(!is.na(allele))
+
+# now get the minimum sample sizes for each
+min_sampled <- long_genos %>%
+  count(collection, locus) %>%
   group_by(locus) %>%
-  mutate(alle_int = replace_na(as.integer(factor(allele)), 0L)) %>%
-  ungroup() %>%
-  mutate(alle_str = sprintf("%02d", alle_int)) %>%
-  select(-allele, -alle_int) %>%
-  pivot_wider(
-    names_from = gene_copy,
-    values_from = alle_str
+  summarise(min_ngc = min(n))
+
+# having looked at those, if we keep anything with
+# at least 38 gene copies, we have most of the loci.
+# So, we will use this to slice sample some individuals from
+# each population. We have to do a bunch of rigamoral because
+# slice_sample requires that the n argument is a constant.  BOGUS!
+set.seed(5)
+long_genos_subsampled <- long_genos %>%
+  left_join(min_sampled, by = "locus") %>%
+  filter(min_ngc >= 38) %>%
+  group_by(collection, locus, min_ngc) %>%
+  nest() %>%
+  mutate(
+    ssd = map2(.x = data, .y = min_ngc, .f = function(x, y) slice_sample(x, n = y))
   ) %>%
-  mutate(geno = str_c(`1`, `2`)) %>%
-  select(-`1`, -`2`) %>%
-  pivot_wider(
-    names_from = locus,
-    values_from = geno
+  select(-data) %>%
+  unnest(ssd) %>%
+  select(-min_ngc)
+
+# get allele frequencies
+alle_freqs <- long_genos_subsampled %>%
+  count(collection, locus, allele) %>%
+  group_by(collection, locus) %>%
+  mutate(freq = n / sum(n)) %>%
+  select(-n) %>%
+  ungroup()
+
+# get all pairwise comparisons of population allele frequencies
+backbone <- expand_grid(
+  row_coll =  unique(alle_freqs$collection),
+  col_coll = unique(alle_freqs$collection),
+  alle_freqs %>% distinct(locus, allele)
+)
+
+# now get the freqs on there, order the collections and ensure
+# that we are on the diagonal and the lower triangle, and then also get
+# the positions for each of the points.  We will allow that the whole
+# scatterplot can be inset into the cell by a fractional amount inset
+inset <- 0.15
+pair_freqs <- backbone %>%
+  left_join(alle_freqs, by = join_by(row_coll == collection, locus, allele)) %>%
+  rename(row_freq = freq) %>%
+  left_join(alle_freqs, by = join_by(col_coll == collection, locus, allele)) %>%
+  rename(col_freq = freq) %>%
+  mutate(
+    row_coll = factor(row_coll, levels = collection_order),
+    col_coll = factor(col_coll, levels = collection_order)
   ) %>%
-  select(-indiv)
+  filter(as.integer(col_coll) <= as.integer(row_coll)) %>%
+  replace_na(list(row_freq = 0.0, col_freq = 0.0)) %>%
+  mutate( # this is just setting things as in table_as_figure
+      ymin = nlevels(row_coll) - as.integer(row_coll) + inset,
+      ymax = nlevels(row_coll) - as.integer(row_coll) + 1 - inset,
+      xmin = as.integer(col_coll) - 1 + inset,
+      xmax = as.integer(col_coll) - inset
+    ) %>%
+  mutate(  # here we get the x-y positions for the scatter points
+    x = xmin + col_freq * (xmax - xmin),
+    y = ymin + row_freq * (ymax - ymin)
+  )
 
-gpt_nest = gpt %>%
-  mutate(collection2 = str_c(collection, ","), .after = collection) %>%
-  group_by(collection) %>%
-  nest()
-
-# now we write that out to a file
-outf <- "results/GSI_and_Fst/genepop.txt"
-cat("Baseline data set in genepop format\n    ", file = outf)
-cat(names(gpt)[-1], sep = ",", file = outf, append = TRUE)
-cat("\n", file = outf, append = TRUE)
-for(x in gpt_nest$data) {
-    cat("POP\n", file = outf, append = TRUE); 
-    write_tsv(x, col_names = FALSE, file = outf, append = TRUE)
-  }
 ```
-
-Now, we could use the genepop package to get results from that file, but it
-is crazy slow, so we do not do this.
-```r
-td_file <- "results/GSI_and_Fst/genic_tests.txt"
-genepop::test_diff(inputFile = outf, pairs = TRUE, outputFile = td_file)
-```
-
-OK! HERE IS WHAT WE WILL DO: MAKE THIS PARALLELIZABLE.  WE WILL MODIFY THE ABOVE
-TO MAKE A GENEPOP FILE FOR EVERY LOCUS, AND THEN WE WILL JUST RUN OVER ALL THOSE
-LOCI USING mclappy ON 8 CORES ON MY LAPTOP, TO MAKE IT FASTER.  THEN, WE WILL PRESENT
-THE RESULTS AS A TINY-MULTIPLE HISTOGRAM OF THE P-VALUES IN EACH CELL.  KINDA COOL!
 
@@
 
-# once again, make it looks like rubias output and taf_prep() it.
-lower_tri_diag <- wgs_mat %>%
-  mutate(
-    collection_f = factor(pop1_f, levels = collection_order),
-    inferred_coll_f = factor(pop2_f, levels = collection_order),
-    n = fst
-  ) %>%
-  select(collection_f, inferred_coll_f, n) %>%
-  complete(collection_f, inferred_coll_f) %>%
-  filter(as.integer(collection_f) >= as.integer(inferred_coll_f)) %>%  # this gets lower tri + diagonal
-  taf_prep()
+# we need to get colors and things for the fill.  We just take the
+# rubias output that was run through taf_prep and we keep the lower
+# diagonal and we set all the cell_labels to NA.
+lower_tri_base <- Rubias_for_later %>%
+  filter(as.integer(col_label) <= as.integer(row_label)) %>%
+  mutate(cell_label = NA_character_)
 
@@
   upper_tri_fst,
-  lower_tri_diag
+  lower_tri_base
 ) %>%
@@
   internal_colors = repunit_colors,
-  Xs_on_diagonal = TRUE,
+  Xs_on_diagonal = FALSE,
   plot_margins = c(0.7, 0.1, 0.1, 0.1)
@@
 
-# and plot it
-TAF_FST$full_plot
+# then plot it and add the scatter points on the lower diagonal
+fst_with_scatters <- TAF_FST$full_plot +
+  geom_point(
+    data = pair_freqs,
+    mapping = aes(x = x, y = y),
+    size = 0.025,
+    colour = "black"
+  )
 
@@
 ```{r}
-ggsave(TAF_FST$full_plot, file = output_list$fst_table, width = 6, height = 5.1)
+ggsave(fst_with_scatters, file = output_list$fst_table, width = 6, height = 5.1)
 ```
@@
   ),
-  TAF_FST$full_plot,
+  fst_with_scatters,
   nrow = 3,

### 2024-03-15 tex/body-text-of-paper.tex
Update for new pops.  Also revamp FST fig.
(to have scatter plots in lower diagonals.)

@@
 included in the collections from the location.  Color codes are as follows: \Wball~=~Winter run,
-\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run. Abbreviations used in location names as folows: R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
+\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run. 
+Names of each location are followed by the codes used to refer to fish from each location, with
+{\tt XXF/S} indicating {\tt XXF} and {\tt XXS} are used to denote fall-run and spring-run fish
+from location {\tt XX}.  Abbreviations used in location names as follows: R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
 \begin{center}
@@
 {\small 
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} lllllrcr}
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} llllllrcr}
 \hline\hline\vspace*{0.4ex}
@@
 $^3$ESU = Evolutionarily Significant Unit; CVFR = Central Valley Fall Run ESU; CVSR = Central Valley Spring Run ESU; CVWR = 
-Central Valley Winter Run ESU; CC = California Coastal ESU; UKTR = Upper Klamath and Trinity Rivers ESU.
+Central Valley Winter Run ESU; CC = California Coastal ESU; UKTR = Upper Klamath and Trinity Rivers ESU. \\
+$^4N_\mathrm{day}$: the number of samples for which the specific sampling day was recorded.
 }
@@
 sequences for these markers 
-to the Chinook v2.0 genome assembly \citep{christensen2018chinook}
+to the Chinook v2.0 genome assembly (\url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/})
 with both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
@@
 \newcommand{\examplecap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
-$F_\mathrm{ST}$ values ({\bf b)} for the collections in the baseline.  The codes of the populations
+$F_\mathrm{ST}$ values and allele frequencies ({\bf b)} for the collections in the baseline.  The codes of the populations
 appear on the outer margins of the table atop cells colored according to the run-timing group
@@
 unit are colored according to the reporting unit.  For self-assignment tallies, the row corresponds
-to the true source collection and the column corresponds to the assigned collection. For $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
-data and the lower triangle was calculated with the whole genome sequence data (where available) 
-from \citet{thompson2020complex}.  NOTE: I THINK WE SHOULD MAKE THE LOWER 
-DIAGONAL THE P-VALUE FROM A TEST OF DIFFERENTIATION OR SOMETHING LIKE THAT.}
+to the true source collection and the column corresponds to the assigned collection. In ({\bf b}), for $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
+data. Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
+in the collections.  These frequencies were estimated by, for each locus, randomly
+subsampling individuals from each collection to the minimum number of observed alleles in any collection
+at the locus.  }
 \begin{center}

### 2024-03-15 tex/inputs/samples-table.tex
Update for new pops.  Also revamp FST fig.
(to have scatter plots in lower diagonals.)

@@
-Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&Sampling Months&Year Range\tabularnewline
-\hline Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
-Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
-Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
-Sacramento R.&MDF&Deer Ck.&Fall&CVFR&29&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
-Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
-Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
-Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
-Sacramento R.&MDF&Mill Ck.&Fall&CVFR&53&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
-Sacramento R.&MDS&Mill Ck.&Spring&CVSR&51&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
-Russian R.&RRF&Warm Springs H.&Fall&CC&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
-Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
-San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
-Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
-Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&$N_\mathrm{day}^4$&Sampling Months&Year Range\tabularnewline
+\hline Rogue R.&CRHS&Cole Rivers H.&Spring&SONCC&50&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CRHS_Cole-Rivers-Hatchery.pdf}}&2004--2010\tabularnewline
+Smith R.&SmRF&Smith R.&Fall&SONCC&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SmRF_Smith-River.pdf}}&2010--2011\tabularnewline
+Klamath R.&BCkF&Blue Ck.&Fall&SONCC&56&56&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCkF_Blue-Creek.pdf}}&2008--2008\tabularnewline
+Klamath R.&IGHF&Iron Gate H.&Fall&UKTR&93&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-IGHF_Iron-Gate-Hatchery.pdf}}&2017--2017\tabularnewline
+Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Eel R.&ERF&Eel R.&Fall&CCC&93&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-ERF_Van-Arsdale-Fisheries-Station.pdf}}&2014--2014\tabularnewline
+Russian R.&RRF&Warm Springs H.&Fall&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
+Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&109&19&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
+Sacramento R.&BCS&Butte Ck.&Spring&CVSR&114&114&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
+Sacramento R.&MDS&Deer Ck.&Spring&CVSR&32&32&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
+Sacramento R.&MDS&Mill Ck.&Spring&CVSR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
+Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
+Sacramento R.&BCF&Butte Ck.&Fall&CVFR&44&44&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
+Sacramento R.&MDF&Deer Ck.&Fall&CVFR&22&22&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&MDF&Mill Ck.&Fall&CVFR&48&48&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
+San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
+Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&252&252&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline

### 2024-03-15 R/colors.R

@@ -1,8 +1,9 @@
 
 # These are for the genetic-marker-resolvable genomic backgrounds
 repunit_colors <- c(
+  `SO-Ncal-Coast` = "beige",
   `Klamath-Trinity` = "#6e8b3d",
-  `Coastal` = "#79cdcd",
+  `Cent. Cal. Coast` = "#79cdcd",
   `CV-Winter` = "#ffff99",            # a "cool-ish" color for winter
   `CV-Spring` = "#99CC33",             # green like spring
   `CV-Fall` = "#F06292",

### 2024-03-15 R/month_sample_sparkline.R

@@ -14,6 +14,7 @@ month_sample_sparkline <- function(tib, path) {
   )
 
   t2 <- tib %>%
+    filter(yday(collection_date) != 1) %>%  # this is for cases where the exact day of sampling is unknown, and denoted YYYY-01-01
     filter(!is.na(month)) %>%
     mutate(month = as.integer(month)) %>%
     count(month, run_timing) %>%

### 2024-03-18 ckmr-sim-evaluation.Rmd
Got a CKMR figure done.
Still need to get it into the text.

--- 2024-02-14 ckmr-sim-evaluation.Rmd
@@
 ---
-title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
+title: "Assessing Power of the California Microhap Baseline for PBT and relationship inference"
 author: "Eric C. Anderson"
@@
   input_list <- list(
-    final_baseline = "data/cali-chinook-baseline.rds",
+    final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
     pop_labels = "inputs/reference-collection-names.csv",
-    locus_info = "inputs/ChinookFullPanelLociTable.csv",
-    chrom_lengths = "inputs/Otsh_v1.0_chrom_lengths.tsv"
+    locus_info = "inputs/ChinookFullPanelLociTable_Feb2024Updated.xlsx"
+  )
+  output_list <- list(
+    ckmr_figure = "results/ckmr-sim-eval/fpr-fnr-figure.pdf",
+    ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
+    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf"
   )
 }
+
+# you can create the necessary output directories like this:
+dump <- lapply(output_list, function(x)
+  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
+)
 ```
@@
 library(CKMRsim)
+library(readxl)
+library(parallel)
+library(grid)
+library(gridExtra)
 
@@
 # we take its position to be the start position of the amplicon
-gen_coords <- read_csv(input_list$locus_info) %>%
-  select(Locus, AmpliconGenomeV1Location) %>%
-  separate(AmpliconGenomeV1Location, into = c("Chrom", "pos_str"), sep = ":") %>%
+gen_coords <- read_excel(input_list$locus_info) %>%
+  select(Locus, AmpliconGenomveV2Location) %>%
+  separate(AmpliconGenomveV2Location, into = c("Chrom", "pos_str"), sep = ":") %>%
   separate(pos_str, into = c("Start", "Stop"), sep = "-", convert = TRUE) %>%
@@
   arrange(Chrom, Pos)
-
-# now we have to do some funny stuff to deal with things on unplaced scaffolds.
-# We will just sprinkle scaffolds into the existing chromosome lengths from the
-# Otsh_v1.0 genome
-chrom_lengths <- read_tsv(input_list$chrom_lengths)
-
-# get the unique unplaced scaffolds and assign them randomly to a position
-# in an assembled chromosome.  We set it up so that markers on the same scaffold
-# will not extend beyond the end of the chromosome
-set.seed(5)
-rando_coords <- gen_coords %>%
-  filter(str_detect(Chrom, "NW_") | is.na(Chrom)) %>%
-  group_by(Chrom) %>%
-  mutate(max_pos = max(Pos)) %>%
-  select(Chrom, max_pos) %>%
-  distinct() %>%
-  mutate(
-    # this gets the chromosome they are assigned to
-    ass_chrom = sample(chrom_lengths$chrom, size = n(), prob = chrom_lengths$num_bases, replace = TRUE)
-  ) %>% # now we do some stuff to place the scaffold randomly on the chromosome
-  left_join(chrom_lengths, by = join_by(ass_chrom == chrom)) %>%
-  mutate(
-    ass_pos = floor(runif(n = n(), min = 1, max = num_bases - max_pos - 1 ))
-  ) %>%
-  select(Chrom, ass_chrom, ass_pos) %>%
-  ungroup()
-
-# now we have the information to randomly stick the scaffold markers onto
-# chromosomes in the genome.  We do this by joining things onto gen_coords
-gen_coords2 <- gen_coords %>%
-  left_join(rando_coords, by = join_by(Chrom)) %>%
-  mutate(
-    Pos = case_when(
-      is.na(ass_pos) ~ Pos,
-      TRUE ~ ass_pos + Pos
-    ),
-    ass_chrom = case_when(
-      is.na(ass_chrom) ~ Chrom,
-      TRUE ~ ass_chrom
-    )
-  ) %>%
-  mutate(Chrom = ass_chrom) %>%
-  select(Locus, Chrom, Pos)
 
@@
 #' @param X a tibble like that nested in a single row of ckmr_sim_nests
-#' @param GC2 essentially the gen_coords2 object from above
+#' @param GC2 essentially the gen_coords object from above
 ckmr_afreqs_ready <- function(X, GC2) {
@@
     D = Y,
-    kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
+    kappa_matrix = kappas[c("PO", "FS", "HS", "AN", "U"), ],
     ge_mod_assumed = ge_model_TGIE,
@@
 # This does the simulate_Qij step and then also estimates the FPRs.
-ckmr_simQ <- function(C) {
+ckmr_simQ <- function(C, unlinked = FALSE, num_mc_samples = 1e5) {
   simulate_Qij(
     C,
-    calc_relats = c("PO", "FS", "HS", "U"),
-    sim_relats = c("PO", "FS", "HS", "U"),
-    unlinked = FALSE,
+    reps = num_mc_samples,  # 100,000 monte carlo samples by default
+    calc_relats = c("PO", "FS", "HS", "AN", "U"),
+    sim_relats = c("PO", "FS", "HS", "AN", "U"),
+    unlinked = unlinked,
     pedigree_list = pedigrees
@@
 
-This is just here while I develop this:
+And here is a function that does the above and then spits out some FPRs and FNRs:
 ```{r}
-X <- ckmr_sim_nests$data[[5]]
-GC2 <- gen_coords2
+
+get_FPRs_and_FNRs <- function(X, GC2 = gen_coords) {
 
@@
 
-Q <- ckmr_simQ(C)
+  # throttle text output to console
+  options(CKMRsim.discard_stdout = TRUE)
+  options(CKMRsim.linkage_verbosity = 0)
+
+  # Simulate some Q_ij's. Suppress output messages that the CKMRsim package
+  # produces
+  Q_link <- ckmr_simQ(C) #%>% suppressMessages()
+  Q_unl <- ckmr_simQ(C, unlinked = TRUE) #%>% suppressMessages()
+
+  # get FPRs and FNRs for a variety of relationships.  Return in a big tibble
+
+  ret <- list(
+    # first for resolving unrelated individuals from PO and FS (and HS, for which there is no hope!)
+    all_v_unrel = mc_sample_simple(
+      Q = Q_unl,
+      Q_for_fnrs = Q_link,
+      nu = c("PO", "FS", "HS"),
+      de = "U",
+      tr = "U",
+      method = "IS",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    ),
+    # then, vanilla MCMC to see how readily we can distinguish FS from PO,
+    # and then, of course FS from HS and PO from A, etc.
+    po_vs_fs = mc_sample_simple(
+      Q = Q_link,
+      nu = c("PO"),
+      de = "FS",
+      tr = "FS",
+      method = "vanilla",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    ),
+    fs_vs_po = mc_sample_simple(
+      Q = Q_link,
+      nu = c("FS"),
+      de = "PO",
+      tr = "PO",
+      method = "vanilla",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    ),
+    fs_vs_hs = mc_sample_simple(
+      Q = Q_link,
+      nu = c("FS"),
+      de = "HS",
+      tr = "HS",
+      method = "vanilla",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    ),
+    po_vs_hs = mc_sample_simple(
+      Q = Q_link,
+      nu = c("PO"),
+      de = "HS",
+      tr = "HS",
+      method = "vanilla",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    ),
+    po_vs_an = mc_sample_simple(
+      Q = Q_link,
+      nu = c("PO"),
+      de = "AN",
+      tr = "AN",
+      method = "vanilla",
+      FNRs = seq(0.01, 0.3, by = 0.01)
+    )
+  ) %>%
+    bind_rows(.id = "what_it_was")
+
+  ret
+}
+
 ```
+
+I tried ripping through everything using furrr::future_map() but that did not
+actually do it in parallel.  So we are going to just mclapply it:
+
+```{r}
+set.seed(5)
+tmp_list <- mclapply(
+  X = ckmr_sim_nests$data, 
+  FUN = function(x) get_FPRs_and_FNRs(x),
+  mc.cores = 8
+)
+
+ckmr_sim_results <- tibble(
+  ckmr_sim_nests,
+  fprs = tmp_list
+)
+```
+
+## Make a plot with all those results
+
+```{r}
+# unnest it
+ckmr_sim_tib <- ckmr_sim_results %>%
+  select(-data) %>%
+  unnest(fprs) %>%
+  ungroup()
+
+# just print out the different scenarios simulated
+ckmr_sim_tib %>%
+  distinct(mc_method, numerator, denominator, true_relat)
+```
+
+Now, make a figure of those results.
+
+```{r}
+# first, try a simple way of categorizing the errors
+error_types_tib <- tribble(
+  ~"scenario", ~"Error Category",
+  "IS-FS-U-U", "Mistaking U for FS",
+  "IS-HS-U-U", "Mistaking U for HS",
+  "IS-PO-U-U", "Mistaking U for PO",
+  "vanilla-FS-HS-HS", "Mistaking HS for FS",
+  "vanilla-FS-PO-PO", "Mistaking PO for FS",
+  "vanilla-PO-AN-AN", "Mistaking AN for PO",
+  "vanilla-PO-FS-FS", "Mistaking FS for PO",
+  "vanilla-PO-HS-HS", "Mistaking HS for PO"
+)
+
+cst2 <- ckmr_sim_tib %>%
+  mutate(scenario = str_c(mc_method, numerator, denominator, true_relat, sep = "-")) %>%
+  left_join(error_types_tib, by = join_by(scenario))
+
+source("R/shift_legend.R")
+
+# first do the importance sampling ones
+g <- ggplot(cst2 %>% filter(mc_method == "IS"), aes(x = FNR, y = FPR, colour = `Error Category`, linetype = `Error Category`)) +
+  geom_line() +
+  #geom_segment(aes(x = FNR, xend = FNR, y = FPR - 2 * se, yend = FPR + 2 * se)) +
+  facet_wrap(~ ckmr_sim_group, nrow = 4) +
+  scale_y_continuous(trans = "log10") +
+  theme_bw() + 
+  theme(legend.direction = "horizontal") +
+  xlab("False Negative Rate") +
+  ylab("False Positive Rate (Log10-scaled)") +
+  guides(
+    colour = guide_legend(
+      title.position = "top",
+      label.position = "bottom",
+      nrow = 2
+      )
+  ) +
+  guides(
+    linetype = guide_legend(
+      title.position = "top",
+      label.position = "bottom",
+      nrow = 2
+      )
+  )
+
+a <- grid.arrange(shift_legend(g))
+```
+
+Now, let's look at the vanilla ones:
+```{r}
+g2 <- ggplot(
+  cst2 %>% filter(mc_method == "vanilla" & FPR > 0 & FPR - 2 * sqrt(FPR * (1 - FPR) / 5e4) > 0.8e-05), 
+  aes(x = FNR, y = FPR)) +
+  geom_segment(
+    aes(
+      x = FNR, 
+      xend = FNR,
+      y = FPR - 2 * sqrt(FPR * (1 - FPR) / 5e4),
+      yend = FPR + 2 * sqrt(FPR * (1 - FPR) / 5e4),
+      colour = `Error Category`
+    ),
+    linewidth = 0.2,
+    show.legend = FALSE
+  ) +
+  geom_line(aes(colour = `Error Category`, linetype = `Error Category`)) +
+  facet_wrap(~ ckmr_sim_group, nrow = 4) +
+  scale_y_continuous(trans = "log10") +
+  theme_bw() + 
+  theme(legend.direction = "horizontal") +
+  xlab("False Negative Rate") +
+  ylab("False Positive Rate (Log10-scaled)") +
+  guides(
+    colour = guide_legend(
+      title.position = "top",
+      label.position = "bottom",
+      nrow = 2),
+    linetype = guide_legend(
+      title.position = "top",
+      label.position = "bottom",
+      nrow = 2)
+  ) +
+  theme(legend.direction = "horizontal")
+
+a2 <- grid.arrange(shift_legend(g2))
+
+```
+
+Then cowplot it:
+```{r}
+Both <- cowplot::plot_grid(a, a2, nrow = 1, labels = c("a)", "b)"))
+
+ggsave(Both, filename = output_list$ckmr_figure, width = 12, height = 7)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$ckmr_figure, collapse = " ")
+system(CALL)
+
+file.rename(from = output_list$ckmr_figure_crop, to = output_list$ckmr_figure_tex)
+````

### 2024-03-18 tex/body-text-of-paper.tex
Got a CKMR figure done.
Still need to get it into the text.

@@
 
+\subsection*{Population genetic analyses}
+
 \subsection*{Power for genetic stock identification and population assignment}
@@
 %%%%%%%%%%%%%%%
-\subsection*{Genetic Variation}
+\subsection*{Population genetic analyses}
 
@@
 data. Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
-in the collections.  These frequencies were estimated by, for each locus, randomly
+in the collections---the $x$ value is allele frequencies from the collection in the column and the
+$y$ value is from the collection in the row. Frequencies for all alleles are shown, with values within
+each cell ranging from 0 to 1.  These frequencies were estimated by, for each locus, randomly
 subsampling individuals from each collection to the minimum number of observed alleles in any collection

### 2024-03-18 R/shift_legend.R

@@ -1,0 +1,63 @@
+# this function unabashedly taken from
+# https://stackoverflow.com/questions/54438495/shift-legend-into-empty-facets-of-a-faceted-plot-in-ggplot2
+# Thanks to Double Beep!
+
+require(gtable)
+require(cowplot)
+
+shift_legend <- function(p){
+
+  # check if p is a valid object
+  if(!"gtable" %in% class(p)){
+    if("ggplot" %in% class(p)){
+      gp <- ggplotGrob(p) # convert to grob
+    } else {
+      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
+      return(p)
+    }
+  } else {
+    gp <- p
+  }
+
+  # check for unfilled facet panels
+  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
+  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
+  empty.facet.panels <- facet.panels[empty.facet.panels]
+  if(length(empty.facet.panels) == 0){
+    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
+    return(p)
+  }
+
+  # establish extent of unfilled facet panels (including any axis cells in between)
+  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
+  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
+                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
+  names(empty.facet.panels) <- c("t", "l", "b", "r")
+
+  # extract legend & copy over to location of unfilled facet panels
+  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
+  if(length(guide.grob) == 0){
+    message("There is no legend present. Returning original plot.")
+    return(p)
+  }
+  gp <- gtable_add_grob(x = gp,
+                        grobs = gp[["grobs"]][[guide.grob]],
+                        t = empty.facet.panels[["t"]],
+                        l = empty.facet.panels[["l"]],
+                        b = empty.facet.panels[["b"]],
+                        r = empty.facet.panels[["r"]],
+                        name = "new-guide-box")
+
+  # squash the original guide box's row / column (whichever applicable)
[…]

### 2024-03-19 tex/body-text-of-paper.tex
Add the CKMRsim results figure

@@
 
-\subsection*{Power for relationship inference}
-
+Results of power analysis for GSI and also Fst estimates are in Figure~\ref{fig:gsi}
+%%%%%%%%%%%%%%%%%%%%%% 
 \begin{figure}
-\newcommand{\examplecap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
+\newcommand{\gsicap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
 $F_\mathrm{ST}$ values and allele frequencies ({\bf b)} for the collections in the baseline.  The codes of the populations
@@
 \end{center}
-\caption[\examplecap]{\examplecap}
-\label{fig:example}
+\caption[\gsicap]{\gsicap}
+\label{fig:gsi}
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+\subsection*{Power for relationship inference}
+
+The results are summarized in Fig.~\ref{fig:fprs}.
+%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\newcommand{\fprcap}{\footnotesize False positive rates as a function of false negative
+rates for distinguishing between different relationships.  {\bf (a)} False positive rates calculated
+using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),
+full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  False positive rates calculated using regular Monte Carlo
+sampling of avuncular pairs (AN: aunt-niece, uncle-nephew, etc.) or full-siblings (FS) or half-siblings (HS) being 
+incorrectly categorized as parent-offspring (PO); as well as FPRs of parent-offspring (PO) pairs or half-sibling (HS) 
+pairs being incorrectly inferred as full siblings (FS).  In both {\em a} and {\em b}, 100,000 Monte Carlo replicates 
+were simulated.  Importance sampling allows estimation of very small false positive rates for unrelated pairs, but 
+cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
+($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2 \times s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  }
+\begin{center}
+\includegraphics[width=\textwidth]{images/fpr-fnr-figure-crop.pdf}
+\end{center}
+\caption[\fprcap]{\fprcap}
+\label{fig:fprs}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
 

### 2024-03-20 002-GSI_and_Fst.Rmd
Added pop gen summaries

@@
 
-# let's also bootstrap that:
-FST_boot <- boot.ppfst(
-  dat = hfstat_dat,
-  nboot = 1000,
-  quant = c(0.025,0.975),
-  diploid = TRUE
-)
-
 # now, get the values out of it and put them in the upper diagonal
@@
 ```
+
+# Population Genetic Summaries
+
+We can do all of our standard population genetic summaries here using `full_base2`.
+
+It is pretty standard, but I think that i am going to standardize sample sizes.
+
+```{r}
+fbl_no_miss <- full_base2 %>%
+  select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
+  pivot_longer(
+    cols = c(-(indiv:sample_type)),
+    names_to = c("locus", "gene_copy"),
+    names_pattern = "^(.*)_([12])$"
+  ) %>%
+  group_by(indiv, locus) %>%
+  mutate(isHz = value[1] != value[2]) %>%
+  ungroup() %>%
+  filter(!is.na(value) & !is.na(isHz))
+
+# now we just summarize by collection and locus
+simple_summaries <- fbl_no_miss %>%
+  group_by(repunit, collection, locus) %>%
+  summarise(
+    num_diploids = n() / 2.0,
+    obs_Hz = mean(isHz),
+    num_alleles = n_distinct(value)
+  ) %>%
+  ungroup()
+
+# here, we get the minimum number of diploids at each locus
+min_gene_copies <- simple_summaries %>%
+  group_by(locus) %>%
+  summarise(min_gc = min(num_diploids) * 2)
+
+# here is a function to take N samples of size s from a vector
+# and return the mean number of distinct elements in it:
+subsample_num_alleles <- function(v, s, N=1000) {
+  samples <- lapply(
+    1:N, 
+    function(z) {
+      sample(x = v, size = s, replace = FALSE) %>% n_distinct()
+    }) %>% unlist()
+
+    # now, return a tibble with columns mean_num_alle_subsamp an fract_polymorphic_subsamp
+    tibble(
+      mean_num_alle_subsamp = mean(samples),
+      fract_polymorphic_subsamp = mean(samples > 1)
+    )
+}
+
+# we will now use that min_diploid number to downsample
+# each collection randomly, without replacement 1000 times and 
+# calculate the mean number of alleles.
+sub_sampled_locus_num_alle <- fbl_no_miss %>%
+  left_join(min_gene_copies, by = join_by(locus)) %>%
+  group_by(repunit, collection, locus) %>%
+  summarise(
+    mean_num_alleles = list(subsample_num_alleles(value, min_gc[1], N = 100))
+  ) %>%
+  ungroup() %>%
+  unnest(mean_num_alleles)
+
+# and now we can calculate the means over loci of those
+subby <- sub_sampled_locus_num_alle %>%
+  group_by(repunit, collection) %>%
+  summarise(
+    mean_num_alle = mean(mean_num_alle_subsamp),
+    mean_fract_poly = mean(fract_polymorphic_subsamp)
+  )
+
+# here we will get the estimated allele freqs and, from those, calculate
+# expected heterozygosity as 1 - \sum_alleles freq^2
+expHz <- fbl_no_miss %>%
+  group_by(repunit, collection, locus, value) %>%
+  summarise(
+    n = n()
+  ) %>%
+  mutate(freq = n / sum(n)) %>%
+  group_by(repunit, collection, locus) %>%
+  summarise(exp_Hz = 1 - sum(freq ^2)) %>%
+  ungroup()
+
+# now, join that on there:
+locus_summaries <- simple_summaries %>%
+  left_join(expHz, by = join_by(repunit, collection, locus))
+
+# and now we can summarise there into means and SDs
+mean_Hzs_etc <- locus_summaries %>%
+  group_by(repunit, collection) %>%
+  summarise(
+    mean_num_dip = mean(num_diploids),
+    mean_expHz = mean(exp_Hz),
+    mean_obsHz = mean(obs_Hz),
+  )
+
+# OK, join the subsampled numbers onto those
+tibble_for_paper <- subby %>%
+  left_join(mean_Hzs_etc, by = join_by(repunit, collection)) %>%
+  mutate(collection = factor(collection, levels = unique(pop_labels$collection))) %>%
+  arrange(collection) %>%
+  ungroup() %>%
+  select(collection, mean_num_dip, mean_num_alle, mean_fract_poly, mean_expHz, mean_obsHz) %>%
+  mutate(
+    collection = as.character(collection),
+    mean_num_dip = sprintf("%.1f", mean_num_dip),
+    mean_num_alle = sprintf("%.2f", mean_num_alle),
+    mean_fract_poly = sprintf("%.2f", mean_fract_poly),
+    mean_expHz = sprintf("%.3f", mean_expHz),
+    mean_obsHz = sprintf("%.3f", mean_obsHz)
+  ) %>%
+  rename(
+    Code = collection,
+    `$\\bar{N}$` = mean_num_dip,
+    `$\\bar{N}_\\mathrm{A,ss}$` = mean_num_alle,
+    `$\\bar{P}_\\mathrm{poly,ss}$` = mean_fract_poly,
+    `$\\bar{\\mathrm{H}}_\\mathrm{exp}$` = mean_expHz,
+    `$\\bar{\\mathrm{H}}_\\mathrm{obs}$` = mean_obsHz,
+  )
+
+tibble_for_paper$Code[1] <- sprintf("\\hline %s", tibble_for_paper$Code[1])
+
+# then write that table out:
+write_delim(
+  tibble_for_paper,
+  delim = "&",
+  file = "tex/inputs/popgen-summary.tex", 
+  eol = "\\tabularnewline\n", 
+  quote = "none"
+)
+```
+
+# Finally, get the histogram of the total number of alleles per locus
+
+```{r}
+numa <- fbl_no_miss %>%
+  group_by(locus) %>%
+  summarise(
+    num_alle = n_distinct(value)
+  ) %>%
+  count(num_alle) %>%
+  bind_rows(tibble(num_alle = 9, n = 0))
+
+gbp <- ggplot(numa, aes(x = factor(num_alle), y = n)) +
+  geom_col(fill = "blue") +
+  theme_bw() +
+  xlab("Total number of alleles") +
+  ylab("Number of loci")
+
+ggsave(gbp, filename = "tex/images/num-alle-barplot.pdf", width = 4, height = 3)
+```

### 2024-03-20 tex/body-text-of-paper.tex
Added pop gen summaries

@@
 this column correspond to run-timing groups as described in the caption of Fig.~\protect\ref{fig:map}.}
+\label{tab:samples}
 {\small 
@@
 %%%%%%%%%%%%%%%
-\subsection*{Population genetic analyses}
+\subsection*{Population genetic summaries}
+
+Barplot of the total number of alleles per locus in the whole data set
+is Figure~\ref{fig:num-alle}.
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\newcommand{\numacap}{\footnotesize Number of loci with different
+total numbers of alleles in the data set.}
+\begin{center}
+\includegraphics[width=\columnwidth]{images/num-alle-barplot.pdf}
+\end{center}
+\caption[\numacap]{\numacap}
+\label{fig:num-alle}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+And the table of the popgen summary is Table~\ref{tab:pg-summ}.
+%%%%%%%%%%%%%%
+\begin{table}
+\caption{\footnotesize Simple population genetic summaries. All quantities are means
+over loci.  $\bar{N}$ is the number of fish with observed genotypes; $\bar{N}_\mathrm{A,ss}$ is the
+average number of alleles and $\bar{P}_\mathrm{poly,ss}$ the fraction of polymorphic loci,
+after sub-sampling to the smallest sample size per locus; $\bar{H}_\mathrm{exp}$ and
+$\bar{H}_\mathrm{obs}$ are expected and observed heterozygosity, respectively. Population
+codes are as given in Table~\ref{tab:samples}}
+\label{tab:pg-summ}
+{\small
+\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrrr}
+\hline\hline
+\input{inputs/popgen-summary.tex}
+\end{tabular*}
+}
+\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\end{table}
+%%%%%%%%%%%%%%%
 
@@
 cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
-($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2 \times s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  }
+($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  }
 \begin{center}

### 2024-03-20 tex/body-text-of-paper.tex
Change text size in tables
Added # of SNPtype loci that we were able to convert and info for SNPtype assays (assays have not been published in their SNPtype form, only in TaqMan form, so just added info for manufacturers of SNPtype assays).

@@
 \label{tab:samples}
-{\small 
+{\footnotesize 
 \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} llllllrcr}
@@
 and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
-using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays \comment{(CITE)}.
+using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (Standard BioTools Inc., South San Francisco, CA).
 In order to type these markers using amplicon sequencing, we followed the methods in
@@
 
-Of \comment{XX} existing SNPtype assays, we were able to convert \comment{YY}
+Of the 96 existing SNPtype assays, we were able to convert 79
 into reliable amplicon-sequenced assays that were included in the panel.
@@
 We tested \comment{XX} candidates from our additional microhaplotype discovery and
-retained \comment{185} of those in the panel.
+retained \comment{106} of those in the panel.
 
@@
 \label{tab:lfar-freqs}
-{\small
+{\footnotesize
 \begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrr}
@@
 \label{tab:pg-summ}
-{\small
+{\footnotesize
 \begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrrr}

### 2024-04-09 002-GSI_and_Fst.Rmd
Add GSI table with the LFAR

@@
   final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
+  lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
   pop_labels = "inputs/reference-collection-names.csv",
@@
   assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
-  assig_table_lfar_no_miss = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar-and-everyone-has-the-lfar.pdf",
   tex_table = "tex/inputs/samples-table.tex",
@@
   filter(collection != "SacNonW")
+
+lfar_genos <- read_csv(input_list$lfar_genos)
 
@@
 ggsave(g, filename = output_list$assig_table, width = 8, height = 10)
+```
+
+# Now, we repeat that while adding the LFAR markers in there
+
+We do these as an additional step to get a sense for how much
+improvement we get from adding the LFAR in there.
+
+First we need to join the LFARs on there:
+```{r}
+# get just the 3 LFAR markers we are using:
+lfar_columns <- paste(
+  rep(c(
+  "NC_037130.1:1062935-1063235",
+  "NC_037130.1:828619-828919",
+  "NC_037130.1:864908-865208"
+  ), each = 2),
+  rep(c(1,2), 3),
+  sep = "_"
+)
+
+lfar3genos <- lfar_genos[c("Sample_ID", lfar_columns)]
+
+base_with_lfar <- full_base2 %>%
+  left_join(lfar3genos, by = join_by(indiv == Sample_ID))
+```
+
+Then, all the step from above:
+```{r}
+# Do self-assignment:
+full_plus_lfar_sa <- self_assign(base_with_lfar, gen_start_col = 5)
+
+top_ass_w_lfar <- full_plus_lfar_sa %>%
+  group_by(indiv, collection, repunit, inferred_repunit) %>%
+  mutate(repu_sclike = sum(scaled_likelihood)) %>%
+  group_by(indiv) %>%
+  filter(repu_sclike == max(repu_sclike)) %>%
+  arrange(indiv, desc(scaled_likelihood)) %>%
+  slice(1) %>%
+  ungroup() %>%
+  mutate(coll_f = factor(collection, levels = collection_order))
+
+```
+
+investigate the ones that are in the wrong repunit, and look at their
+scaled_repu_likelihoods, the others.
+```{r}
+ta_corr_lfar <- top_ass_w_lfar %>%
+  filter(repunit == inferred_repunit)
+ta_wrong_lfar <- top_ass_w_lfar %>%
+  filter(repunit != inferred_repunit) %>%
+  arrange(collection, repu_sclike) %>%
+  group_by(collection) %>%
+  mutate(y = 190 + (70/9) * 1:n())
+
+ggplot() +
+  geom_histogram(
+    data = ta_corr_lfar,
+    mapping = aes(x = repu_sclike, fill = inferred_repunit)
+  ) +
+  geom_point(
+    data = ta_wrong_lfar,
+    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
+    size = 3,
+    shape = 21
+  ) +
+  facet_wrap(~coll_f, ncol = 3) +
+  scale_fill_manual(values = repunit_colors)
+```
+
+Filter out the ones that look like clear mis-labelling and then
+count up the mis-assigned fish:
+```{r}
+tafilt_w_lfar <- top_ass_w_lfar %>%
+  filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
+```
+
+Here we calculate the number of correctly and incorrectly assigned fish:
+```{r}
+tafilt_w_lfar %>%
+  mutate(correct = repunit == inferred_repunit) %>%
+  count(correct) %>%
+  mutate(fract = n / sum(n))
+```
+
+Having the LFAR in there improves things. Good.
+
+Make a table of those:
+```{r}
+
+X2_lfar <- tafilt_w_lfar %>%
+  mutate(
+    collection_f = factor(collection, levels = collection_order),
+    inferred_coll_f = factor(inferred_collection, levels = collection_order)
+  ) %>%
+  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
+  taf_prep()
+
+Rubias_for_later_lfar <- X2_lfar # keeping this for later...
+
+# get the result
+TAF2_lfar <- table_as_figure(
+  X = X2_lfar,
+  RC_groups = RC_groups,
+  external_colors = run_time_colors,
+  internal_colors = repunit_colors,
+  plot_margins = c(1, 0.1, 0.7, 0.1)
+)
+
+# and plot it
+g_lfar <- plot_grid(
+  TAF2_lfar$full_plot,
+  plot_grid(
+    get_legend(TAF2_lfar$for_external_legend),
+    get_legend(TAF2_lfar$for_internal_legend),
+    nrow = 1
+  ),
+  nrow = 2,
+  rel_heights = c(7,3)
+)
+
+g_lfar
+```
+
+Adding the LFAR definitely improves the LFAR assignment, though
+it still does not get it perfect.
+```{r}
+ggsave(g_lfar, filename = output_list$assig_table_lfar_all, width = 10, height = 12)
 ```

### 2024-04-17 tex/body-text-of-paper.tex
Added sentence to Discussion
More discussion additions.

@@
 
+We describe a novel panel of microhaplotype genetic markers for Chinook salmon 
+in California that includes multiallelic gene regions with high variability for 
+relationship inference and gene regions identified in whole-genome sequence data 
+for increased power for population identification. We show how this set of markers
+provides sufficient power for highly accurate identification of parent and offspring
+pairs and full siblings in California Chinook salmon, and also allows near-perfect 
+identification of individuals to population of origin. 
 When, and if, sequencing costs drop sufficiently, it may be possible to include

### 2024-04-18 tex/body-text-of-paper.tex
More summary paragraph

@@
 pairs and full siblings in California Chinook salmon, and also allows near-perfect 
-identification of individuals to population of origin. 
+identification of individuals to population or genetic group of origin. This set of 
+genetic markers lays the foundation for a comprehensive genetic monitoring and evaluation effort
+that facilitates multiple types of inference and is flexible and extensible.
 When, and if, sequencing costs drop sufficiently, it may be possible to include

### 2024-04-22 ckmr-sim-evaluation.Rmd
Start supplement.  Write more

@@
     final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
+    lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
     pop_labels = "inputs/reference-collection-names.csv",
@@
 file.rename(from = output_list$ckmr_figure_crop, to = output_list$ckmr_figure_tex)
-````
+```

### 2024-04-22 tex/body-text-of-paper.tex
commit body text to pull down other edits
Start supplement.  Write more

@@
 
+Basic data summaries were calculated for each collection
+using simple operations in the R programming language, Version 4.2
+\citep{rcore}. We start with a simple tally of the number of alleles
+found at each locus.  To provide a summary of the genetic variation and
+missing data we report 1) the average over loci of the number of fish
+in which each locus was successfully genotyped ($\bar{N}$);
+2) the average number of alleles per locus with all collections
+downsampled to the smallest number of genotyped individuals at each
+locus ($\bar{N}_{A,ss}$);  3) the fraction of polymorphic
+loci in each collection after subsampling to the smallest number of genotyped
+fish ($\bar{P}_{poly,ss}$); 4) the expected heterozygosity (average over loci
+of one minus the frequencies of homozygotes at each locus expected from the
+estimated allele frequencies; and 5) the observed heterozygosity (average over
+loci of the fraction of heterozygous genotypes.
+When subsampling each collection to the minimum number of genotyped
+fish across all collections at a locus, we subsampled individuals without
+replacement, taking the average of 1,000 different random subsamples.
+
+We calculated \citet{weir1984estimating}'s pairwise $F_\mathrm{ST}$ between
+all collections, using the {\footnotesize\tt pp.fst} function from the R
+package 'hierfstat' \citep{hierfstat}. Finally, we evaluated population
+structure in the data by performing unsupervised clustering of the
+the complete data set (\comment{With or without the LFAR? Which did Anthony use?}) 
+using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
+We performed 20 replicate runs at each value of the assumed number of
+subpopulations, $K$, in $\{2, 3, 4, 5, 6\}$ 
+\comment{I think we should also see what $K=7$ turns up}
+using default settings of the program.
+
 \subsection*{Power for genetic stock identification and population assignment}
@@
 the San Joaquin River (Fig.~\ref{fig:lfar-assoc}).
-%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\newcommand{\lfarcap}{\footnotesize Negative log base 10 of association $p$-values for
-individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
-with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
-to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
-upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
-panel is the comparison of late-fall to San Joaquin River fall. }
-\includegraphics[width=\textwidth]{images/lfar-assoc-faceted.jpg}
-\caption[\lfarcap]{\lfarcap}
-\label{fig:lfar-assoc}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
+
 The Manhattan plot of Fig.~\ref{fig:lfar-assoc} shows sporadic SNPs with low association
@@
 
-Barplot of the total number of alleles per locus in the whole data set
-is Figure~\ref{fig:num-alle}.
-%%%%%%%%%%%%%%%%%%%%%% 
-\begin{figure}
-\newcommand{\numacap}{\footnotesize Number of loci with different
-total numbers of alleles in the data set.}
-\begin{center}
-\includegraphics[width=\columnwidth]{images/num-alle-barplot.pdf}
-\end{center}
-\caption[\numacap]{\numacap}
-\label{fig:num-alle}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%
+The total number of alleles per locus varied between 2 and 10 (Fig.~\ref{fig:num-alle}). Biallelic 
+loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
+showed more than two alleles, providing significantly more power for relationship
+inference than a comparably sized panel of biallelic SNPs.
 
@@
 cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
-($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  }
+($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.
+This also is the reason that many of the lines terminate   }
 \begin{center}

### 2024-04-22 tex/supplemental-figures.tex
Start supplement.  Write more

%!TEX root = supplement.tex

%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\textwidth]{images/lfar-assoc-faceted.jpg}
\caption[
	Negative log base 10 of association $p$-values for
	individual SNPs for late-fall versus fall run
]{
	\footnotesize Negative log base 10 of association $p$-values for
	individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
	with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
	to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
	upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
	panel is the comparison of late-fall to San Joaquin River fall. 
}
\label{fig:lfar-assoc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{images/num-alle-barplot.pdf}
\end{center}
\caption[Number of loci with different
total numbers of alleles]{\footnotesize Number of loci with different
total numbers of alleles in the data set.}
\label{fig:num-alle}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

### 2024-04-22 tex/supplemental-text.tex
Start supplement.  Write more

--- 2024-04-17 tex/body-text-of-paper.tex
@@
-%!TEX root = main.tex
+%!TEX root = supplement.tex
 
-\section*{Introduction}
+\section{Supplemental Text}
 
@@
 vice versa).
-
-Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
-structure within the Central Valley of California. Chinook salmon of the two main river basins---the 
-Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
-within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
-(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California. \comment{NEED MORE HERE. Certainly we should cite
-\citet{yoshiyama1998historical}.}
-
- \section*{Methods}
-
-\comment{More verbiage}
-
-\subsection*{Population Sampling}
-
-We sampled fish for the reference baselines from XX locations across California.
-These locations included sites in the two
-main tributaries (the Sacramento River and the San Joaquin River) within
-California's Central Valley, the Klamath-Trinity basin in Northern California, and
-3 rivers of the California coast (Russian, Eel, etc)
-(Fig.~\ref{fig:map}).
-%%%%%%%%%%%%%
-\begin{figure}
-\newcommand{\mapcap}{\footnotesize  A map of sampling locations represented in the reference 
-baseline.  The colored circles aside
-each location name represent the different run-timing groups of Chinook salmon
-included in the collections from the location.  Color codes are as follows: \Wball~=~Winter run,
-\LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run. 
-Names of each location are followed by the codes used to refer to fish from each location, with
-{\tt XXF/S} indicating {\tt XXF} and {\tt XXS} are used to denote fall-run and spring-run fish
-from location {\tt XX}.  Abbreviations used in location names as follows: R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
-\begin{center}
-\includegraphics[width=\columnwidth]{images/map-crop.pdf}
-\end{center}
-\caption[\mapcap]{\mapcap}
-\label{fig:map}
-\end{figure}
-%%%%%%%%%%%%%%%
-Collections in each location were separated into fish from the different run-timing ecotypes according
-to a variety of criteria. The last vestiges of winter-run Chinook salmon are maintained at the Livingston
-Stone National Fish Hatchery for supplementation (released as juveniles into the river) and as a captive broodstock;
-sampling of this population was performed by hatchery staff during spawning. In Central Valley Rivers with both 
-spring and fall ecotypes (i.e. Mill, Deer and Butte Creeks), previous research has documented distinct/disparate/unique
-spawning times (Julian date) for each ecotype (REF?). As such, sampling targeted carcasses found in each of these 
-rivers at times representative of their run-timing designation, which we were able to further verify using their genotypes
-at the RoSA markers. At the time of sampling, only fall-run Chinook were found in the San Joaquin River. Run-timing 
-designation at the Feather River Hatchery is more complicated but is managed at a fine-scale by hatchery practices, 
-where early returning fish in May and April (expressing the Spring-run life history) are visibly tagged and re-released 
-to the river. Subsequently in the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
-as Spring-run broodstock. Individual-based analyses of RoSA genotypes have demonstrated
-
-(We need to talk about how these were split into spring run, fall-run, winter 
-run, etc.
-In total, the samples in the baseline represent fish from five different Evolutionarily Significant 
-Units (ESUs) of Chinook salmon \citep{waples1991pacific} A summary of the different collections 
-%%%%%%%%%%%%%%%%
-\begin{table*}
-\caption{\footnotesize Collections of fish in the reference baseline.  The Sampling Months column shows
-a pictorial representation of the proportion of each collection sampled in each month.  Colors in
-this column correspond to run-timing groups as described in the caption of Fig.~\protect\ref{fig:map}.}
-\label{tab:samples}
-{\footnotesize 
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} llllllrcr}
-\hline\hline\vspace*{0.4ex}
-\input{inputs/samples-table.tex}
-\end{tabular*}
-}
-\rule{3cm}{0.3pt}
-
-{\footnotesize
-$^1$R. = River.\\
-$^2$Ck. = Creek; NFH = National Fish Hatchery; R. = River; H. = Hatchery.\\
-$^3$ESU = Evolutionarily Significant Unit; CVFR = Central Valley Fall Run ESU; CVSR = Central Valley Spring Run ESU; CVWR = 
-Central Valley Winter Run ESU; CC = California Coastal ESU; UKTR = Upper Klamath and Trinity Rivers ESU. \\
-$^4N_\mathrm{day}$: the number of samples for which the specific sampling day was recorded.
-}
-\hrule\vspace*{0.3ex}\hrule
-\end{table*}
-%%%%%%%%%%%%%%%%%%%%%
-
-\subsection*{Genetic Markers}
-
-To assemble the markers for this reference baseline, we drew upon several different
-sources, as detailed in the separate sections below.
-
-\subsubsection*{Conversion of existing SNPtype assays}
-
-A number of SNP markers from earlier discovery efforts \citep{clemento2011discovery} had already 
-been validated as useful for GSI
-and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
-using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (Standard BioTools Inc., South San Francisco, CA).
-In order to type these markers using amplicon sequencing, we followed the methods in
-\citet{campbell2015genotyping} and added the Illumina small RNA sequencing primer and 
-read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus 
-Specific Primers respectively.  We then dropped any loci that produced genotypes that were 
-frequently inconsistent with genotypes produced with the SNPtype assays and any loci that 
-did not amplify well or those that amplified so well that they were grossly over-represented
-among the reads on the sequencer.  We then 
-used Freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
-microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by Freebayes.
-
-\subsubsection*{Additional microhaplotype discovery}
-
-We undertook a new discovery process to develop additional multiallelic
-microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
-\comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
-candidate to then follow up on with amplicon sequencing, etc.}
-
-After designing and validating an amplicon-sequencing approach to type both the
-existing SNPtype assays and the product of the additional microhaplotype
-discovery process we identified their location with the genome. We mapped the reference
-sequences for these markers 
-to the Chinook v2.0 genome assembly (\url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/})
-with both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
-We report the genomic coordinates these target amplicons found with both alignment methods.
-
-\subsubsection*{Identifying and Designing markers from the Late-Fall Associated Region}
-
-To identify loci with large allele frequency differences between late-fall and fall-run Chinook salmon, 
-we first conducted a simple case-control GWAS using the low-coverage whole genome sequencing 
-data of \citet{thompson2020complex}.  We conducted two separate association-study comparisons.  
-The first involved the 16 late-fall run fish as cases with 16 fall-run fish from the Feather River Hatchery as controls. 
-The second involved the same 16 late-fall fish as cases, but used 16 fall-run fish from the San 
-Joaquin River as controls.  The association studies were performed for each chromosome with 
-ANGSD \citep{pmid21663684,korneliussen_angsd_2014}, using the following command: {\footnotesize\tt angsd -yBin \{ybin\}  -r \{chrom\} 
--minMapQ 30 -minQ 20 -minInd 12 -doAsso 1 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP\_pval 1e-6  
--out \{prefix\}  -bam \{bamlist\}}, where {\tt \{ybin\} } is the path to a file indicating the cases and 
-controls, 
-{\tt \{chrom\} } indicates the chromosome name, {\tt \{prefix\} } is the prefix to use for output files, and 
-{\tt \{bamlist\} } is the path to the text file holding paths to the BAM files, one for each individual, ordered as in {\tt \{ybin\} }.
-
-There was only one peak in association scores that was shared between both comparisons (see 
-Results). From this peak we identified candidates for followup genotyping by including the 10 
-SNPs with the lowest association p-values, and supplementing those with additional SNPs that 
-showed:
-\begin{itemize}
-\item An estimated absolute allele frequency difference (between late-fall and all 32 fall-run fish)  $>0.75$,  with an approximate lower confidence interval of that figure of 0.5,  or
-\item An estimated absolute frequency difference of >0.5 with a lower confidence interval >0.25 and 
-with either fall-run or late-fall run being nearly fixed for one of the alleles, or
-\item An annotation from snpEff \citep{cingolani2012program} of “HIGH” or “MODERATE”, or
-\item An absolute frequency difference $>0.48$ and a lower confidence interval of $>0.25$ and a 
-genomic  coordinate $> 2.2$~Mb.
-\end{itemize}
-The final condition was implemented to gather several SNPs within the flanking region to the right of 
-the main peak of association.
-
-For the candidate markers found by the above process, we designed amplification primers using 
-Primer3 \citep{koressaar2007enhancements,untergasser2012primer3} and tested amplification in a 
-subset of Chinook salmon 
-samples. The primer pairs that amplified sequence that reliably mapped back to the association-peak region were then typed
-in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring 
-run, and 188 fall/late-fall run. We divided the fish identified as fall/late-fall via the 125-locus genetic 
-stock identification data set into fall run and late fall run according to their sampling date at Keswick: 
-99 fish arriving before Dec 26 were classified as fall run and 89 fish arriving after Dec 26 were 
-classified as late-fall run.   The allele frequencies of the target SNPs at the remaining amplicons 
-were 
-estimated for each of these four groups: winter, spring, fall, and late-fall, and we retained
-amplicons showing large allele frequency  differences between late-fall fish and all other runs.
-
-\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}
-
-We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish described in \citet{thompson2020complex}.
-\comment{Need more here. I personally am of the thought that we should not include these
-for GSI, but should use them for identifying the genotypes of fish assigned to populations
-using the rest of the markers.}
-
-\subsection*{Population genetic analyses}
-
-\subsection*{Power for genetic stock identification and population assignment}
-
-\subsection*{Power for relationship inference}
-
-\section*{Results}
-
-\subsection*{Genetic Markers}
-
-\subsubsection*{Conversion of existing SNPtype assays}
-
-Of the 96 existing SNPtype assays, we were able to convert 79
-into reliable amplicon-sequenced assays that were included in the panel.
-
-\subsubsection*{Additional microhaplotype discovery}
-
-We tested \comment{XX} candidates from our additional microhaplotype discovery and
-retained \comment{106} of those in the panel.
-
-Of these \comment{185} loci that originated from SNPtype assays or from our additional
-microhaplotype discovery, \comment{179} of them mapped identically to a single
-location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
-Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
-the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
-that was identical in both bwa mem and BLAT, two of them had secondary alignments that
-disagreed slightly between BLAT and bwa mem, and one of them had multiple
-non-primary alignments (\comment{SuppTable}).
-
-\subsubsection*{Late-fall Associated Region Markers}
-
-The case-control association test revealed a single peak on chromosome 34 with large allele frequency differences
-between Central Valley late-fall and fall-run fish.  This peak was evident both in the comparison of
-late-fall fish to fall-run fish in the Feather River Hatchery and of late-fall fish to the fall-run fish in
-the San Joaquin River (Fig.~\ref{fig:lfar-assoc}).
-%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\newcommand{\lfarcap}{\footnotesize Negative log base 10 of association $p$-values for
-individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
-with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
-to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
-upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
-panel is the comparison of late-fall to San Joaquin River fall. }
-\includegraphics[width=\textwidth]{images/lfar-assoc-faceted.jpg}
-\caption[\lfarcap]{\lfarcap}
-\label{fig:lfar-assoc}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
-The Manhattan plot of Fig.~\ref{fig:lfar-assoc} shows sporadic SNPs with low association
-$p$-values.  These are not shared between the two different late-fall to fall comparisons, and
-sites that we investigated were concentrated in areas of poor mapping suggesting that their
-$p$-values were artifactual.  By contrast, a large number of sites near the prominent peak on
-chromosome 34 had large allele frequency differences between late-fall and fall run.  Taking the
-top 10 most associated SNPs, and including additional ones from the filtering criteria given in the
-Methods, yielded 49 candidate markers for which we designed primers for amplication.  Only 8
-of these 49 primer pairs yielded reliable amplification and mapping.   Of these 8, two of the SNPs 
-showed inconsistent genotyping and were removed from consideration.  Of the remaining 6, only 3 
-showed marked allele frequency differences ($> 0.7$) between late-fall and fall run.  These three 
-SNPs (Chr34:828,768,  Chr34:865,057, and Chr34:1,063,084) also had pronounced differences in 
-allele frequency between 
-late-fall and spring or winter run.
-
-Analysis of the mapping of amplicons for Chr34:828,768 
-showed that many ($\approx$89\%) of the reads were off-target (aligning to other chromosomes, 
-etc.), making it costly (in terms of sequencing effort) to include the marker in the baseline panel. 
-Consequently it was dropped.  The frequency of the different alleles at the two remaining loci
-within different reporting units in the baseline are shown in Table~\ref{tab:lfar-freqs}.
-%%%%%%%%%%%%%%
-\begin{table}
-\caption{\footnotesize Allele frequencies across reporting units of the two late-fall associated
-markers.  Frequencies are given for the alleles most common amongst late-fall Chinook: nucleotide 
-{\tt G} at locus Chr34:865,057 and {\tt A} at Chr34:1,063,084.  $N$ is the total number of gene 
-copies available for estimating the allele frequency.}
-\label{tab:lfar-freqs}
-{\footnotesize
-\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrr}
-\hline\hline
-& \multicolumn{2}{c}{\underline{Chr34:865,057}} & \multicolumn{2}{c}{\underline{Chr34:1,063,084}} \\
-Reporting Unit & freq. & $N$ & freq. & $N$ \\ \hline
-\input{inputs/lfar-freqs.tex}
-\end{tabular*}
-}
-\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
-\end{table}
-%%%%%%%%%%%%%%%
-\subsection*{Population genetic summaries}
-
-Barplot of the total number of alleles per locus in the whole data set
-is Figure~\ref{fig:num-alle}.
-%%%%%%%%%%%%%%%%%%%%%% 
-\begin{figure}
-\newcommand{\numacap}{\footnotesize Number of loci with different
-total numbers of alleles in the data set.}
-\begin{center}
-\includegraphics[width=\columnwidth]{images/num-alle-barplot.pdf}
-\end{center}
-\caption[\numacap]{\numacap}
-\label{fig:num-alle}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%
-
-And the table of the popgen summary is Table~\ref{tab:pg-summ}.
-%%%%%%%%%%%%%%
-\begin{table}
-\caption{\footnotesize Simple population genetic summaries. All quantities are means
-over loci.  $\bar{N}$ is the number of fish with observed genotypes; $\bar{N}_\mathrm{A,ss}$ is the
-average number of alleles and $\bar{P}_\mathrm{poly,ss}$ the fraction of polymorphic loci,
-after sub-sampling to the smallest sample size per locus; $\bar{H}_\mathrm{exp}$ and
-$\bar{H}_\mathrm{obs}$ are expected and observed heterozygosity, respectively. Population
-codes are as given in Table~\ref{tab:samples}}
-\label{tab:pg-summ}
-{\footnotesize
-\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrrr}
-\hline\hline
-\input{inputs/popgen-summary.tex}
-\end{tabular*}
-}
-\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
-\end{table}
-%%%%%%%%%%%%%%%
-
-\subsection*{Power for genetic stock identification and population assignment}
-
-Results of power analysis for GSI and also Fst estimates are in Figure~\ref{fig:gsi}
-%%%%%%%%%%%%%%%%%%%%%% 
-\begin{figure}
-\newcommand{\gsicap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
-$F_\mathrm{ST}$ values and allele frequencies ({\bf b)} for the collections in the baseline.  The codes of the populations
-appear on the outer margins of the table atop cells colored according to the run-timing group
-of the collection.  Interior cells in which the column and row correspond to the same reporting
-unit are colored according to the reporting unit.  For self-assignment tallies, the row corresponds
-to the true source collection and the column corresponds to the assigned collection. In ({\bf b}), for $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
-data. Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
-in the collections---the $x$ value is allele frequencies from the collection in the column and the
-$y$ value is from the collection in the row. Frequencies for all alleles are shown, with values within
-each cell ranging from 0 to 1.  These frequencies were estimated by, for each locus, randomly
-subsampling individuals from each collection to the minimum number of observed alleles in any collection
-at the locus.  }
-\begin{center}
-\includegraphics[width=\columnwidth]{images/gsi_and_fst_fig-crop.pdf}
-\end{center}
-\caption[\gsicap]{\gsicap}
-\label{fig:gsi}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%
-
-\subsection*{Power for relationship inference}
-
-The results are summarized in Fig.~\ref{fig:fprs}.
-%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\newcommand{\fprcap}{\footnotesize False positive rates as a function of false negative
-rates for distinguishing between different relationships.  {\bf (a)} False positive rates calculated
-using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),
-full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  False positive rates calculated using regular Monte Carlo
-sampling of avuncular pairs (AN: aunt-niece, uncle-nephew, etc.) or full-siblings (FS) or half-siblings (HS) being 
-incorrectly categorized as parent-offspring (PO); as well as FPRs of parent-offspring (PO) pairs or half-sibling (HS) 
-pairs being incorrectly inferred as full siblings (FS).  In both {\em a} and {\em b}, 100,000 Monte Carlo replicates 
-were simulated.  Importance sampling allows estimation of very small false positive rates for unrelated pairs, but 
-cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
-($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  }
-\begin{center}
-\includegraphics[width=\textwidth]{images/fpr-fnr-figure-crop.pdf}
-\end{center}
-\caption[\fprcap]{\fprcap}
-\label{fig:fprs}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
-
-\section*{Discussion}
-
-We describe a novel panel of microhaplotype genetic markers for Chinook salmon 
-in California that includes multiallelic gene regions with high variability for 
-relationship inference and gene regions identified in whole-genome sequence data 
-for increased power for population identification. We show how this set of markers
-provides sufficient power for highly accurate identification of parent and offspring
-pairs and full siblings in California Chinook salmon, and also allows near-perfect 
-identification of individuals to population of origin. 
-When, and if, sequencing costs drop sufficiently, it may be possible to include
-all different populations of a single species within a single standardized reference baseline
-that performs equally well at broad and regional scales. A statistical approach for population
-assignment from low coverage whole genome sequences appeared recently
-\citep{desaixINPRESSpopulation}, and its authors noted that such a data type could
-be ideal for producing reference baselines simultaneously applicable to broad scale
-coverage and high resolution within sub-regions.   For the present, however,
-baselines tailored to specific regions are essential for regional management questions.
-It should be noted that recent

### 2024-04-24 tex/body-text-of-paper.tex
Adding to the methods

@@
 using simple operations in the R programming language, Version 4.2
-\citep{rcore}. We start with a simple tally of the number of alleles
+\citep{rcore}. We started with a simple tally of the number of alleles
 found at each locus.  To provide a summary of the genetic variation and
-missing data we report 1) the average over loci of the number of fish
-in which each locus was successfully genotyped ($\bar{N}$);
-2) the average number of alleles per locus with all collections
+missing data we report
+\begin {itemize}
+\item the average over loci of the number of fish
+in which each locus was successfully genotyped, $\bar{N}$;
+\item the average number of alleles per locus with all collections
 downsampled to the smallest number of genotyped individuals at each
-locus ($\bar{N}_{A,ss}$);  3) the fraction of polymorphic
+locus, $\bar{N}_\mathrm{A,ss}$,
+\item the fraction of polymorphic
 loci in each collection after subsampling to the smallest number of genotyped
-fish ($\bar{P}_{poly,ss}$); 4) the expected heterozygosity (average over loci
+fish, $\bar{P}_\mathrm{poly,ss}$,
+\item the expected heterozygosity as the average over loci
 of one minus the frequencies of homozygotes at each locus expected from the
-estimated allele frequencies; and 5) the observed heterozygosity (average over
-loci of the fraction of heterozygous genotypes.
+estimated allele frequencies, $\bar{H}_\mathrm{exp}$, and 
+\item the observed heterozygosity as the average over
+loci of the fraction of heterozygous genotypes, $\bar{H}_\mathrm{obs}$.
+\end{itemize}
 When subsampling each collection to the minimum number of genotyped
@@
 package 'hierfstat' \citep{hierfstat}. Finally, we evaluated population
-structure in the data by performing unsupervised clustering of the
+structure in the data by performing unsupervised clustering of
 the complete data set (\comment{With or without the LFAR? Which did Anthony use?}) 
@@
 \subsection*{Power for genetic stock identification and population assignment}
+
+To assess the power of the marker panel for GSI, we used the R package
+'rubias' \citep{moran2019bayesian}.  We used the function {\tt self\_assign()}
+to perform assignment of each individual in the reference baseline to one of the
+collections within the baseline.  Each fish was removed from the reference baseline
+when calculating the likelihood that it originated from the collection it actually belongs
+to, using a classical leave-one-out approach to eliminate the inflated estimates of
+power that might occur without such a procedure \citep{anderson2008improved}.
+The scaled likelihoods of collection membership returned by
+{\tt self\_assign()} were summed within reporting units, and fish were
+assigned to the reporting unit with the largest value of that sum.  Subsequently,
+fish were assigned to the collection within that reporting unit with the highest
+scaled likelihood.
 

### 2024-05-07 002-GSI_and_Fst.Rmd
Start supplement.  Write more
Committing 002 before doing a lot of pruning of it
I want to use the new baseline with the hapstrs and I also want to just do it all with the LFAR included (I don't really care about the Fst's being inflated from the LFAR---we kind of want that, ultimately, I think.)

@@
 
+repu_likes <- full_plus_lfar_sa %>%
+  group_by(indiv, collection, repunit, inferred_repunit) %>%
+  summarise(repu_sclike = sum(scaled_likelihood)) %>%
+  ungroup()
+
 top_ass_w_lfar <- full_plus_lfar_sa %>%
@@
 
-investigate the ones that are in the wrong repunit, and look at their
-scaled_repu_likelihoods, the others.
+## Identifying fish erroneously included in some collections
+
+The occurrence of winter-run-assigned fish amongst the spring run collections is
+a clear indicator that there are some fish in the collections that do not belong
+with one another.  The same occurs amongst the fall and the spring run, I expect.
+
+Our goal here is to examine the likelihoods with which fish are assigned to
+different reporting units and come up with a common-sense and defensible approach
+for discarding individuals that don't belong in the collections they were sampled
+from.
+
+We start by tallying up the misassigned fish---those that have a repunit and an
+inferred repunit that differ:
+```{r}
+misassigned_ones <- top_ass_w_lfar %>%
+  filter(repunit != inferred_repunit) %>%
+  count(repunit, inferred_repunit)
+
+misassigned_ones
+```
+
+For each of these 6 pairs, I want to make a scatterplot that shows the repunit
+likelihoods of all the fish _from either the repunit or the inferred_repunit_.
+The x-axis will be the likelihood for the repunit and the y-axis will be the
+likelihood for the inferred_repunit.  So, there is some weird prep to get
+this all together.  
+```{r}
+MO2 <- misassigned_ones %>%
+  rename(
+    Foc_repunit = repunit,
+    Foc_inferred_repunit = inferred_repunit
+  )
+
+tmp <- lapply(1:nrow(MO2), function(i) {
+  repu_likes %>%
+    filter(
+      (repunit == MO2$Foc_repunit[i] | repunit == MO2$Foc_inferred_repunit[i]) &
+        (inferred_repunit == MO2$Foc_repunit[i] | inferred_repunit == MO2$Foc_inferred_repunit[i])
+    ) %>%
+    mutate(
+     Foc_repunit =  MO2$Foc_repunit[i],
+     Foc_inferred_repunit = MO2$Foc_inferred_repunit[i],
+     .before = repunit
+    )
+}) %>%
+  bind_rows() %>%
+  arrange(indiv, Foc_repunit, Foc_inferred_repunit)
+
+for_plots <- tmp %>%
+  mutate(
+    rep = ifelse(Foc_repunit == repunit, "rep", "inf"),
+    inf = ifelse(Foc_inferred_repunit == inferred_repunit, "inf", "rep")
+  ) %>%
+  #count(Foc_repunit, Foc_inferred_repunit, repunit, inferred_repunit, rep, inf)
+  select(-repunit, -inferred_repunit) %>%
+  pivot_wider(names_from = c(rep, inf), names_sep = "_to_", values_from = repu_sclike) %>%
+  mutate(condition = str_c(Foc_repunit, " misassigned to ", Foc_inferred_repunit), .after = Foc_inferred_repunit)
+```
+
+Now we should be able to facet wrap those:
+```{r}
+ggplot(for_plots) +
+  geom_point(aes(x = inf_to_rep, y = inf_to_inf, colour = Foc_inferred_repunit)) +
+  geom_point(aes(x = rep_to_rep, y = rep_to_inf, colour = Foc_repunit)) +
+  facet_wrap(~condition) +
+  scale_colour_manual(values = repunit_colors)
+
+```
+
+So, for each of the six plots, I will need four different repu_sclike's:
+For example, for the `repunit = CV-Fall` and `inferred_repunit = CV-Spring` case
+we will need to have the repu_sclikes for CV-Fall fish being calculated for CV-Fall
+and CV-Spring, and for the CV-Spring fish being calculated for CV-Spring and CV-Fall.
+
+Now, we want to use this to semi-join to get all the likelihoods for these scenarios,
+whether or not they were misassigned. 
+```{r}
+focal_comparisons <- repu_likes %>%
+  semi_join(misassigned_ones, by = join_by(repunit, inferred_repunit))
+```
+
+And we are going to want to compare these to all the likelihoods of being
+from the actual repunit that the fish was col
+
+Now, I want to look at these different comparisons _pairwise_, so we will
+just pivot it wider and then look at different columns in there.  This isn't
+super "tidy" but I think it should work OK:
+```{r}
+focal_comparisons %>%
+  pivot_wider(names_from = inferred_repunit, values_from = repu_sclike)
+```
+
+And I want to look at the observed distribution of these scores in for each
+comparison, colored by repunit.
+```{r}
+ggplot(focal_comparisons, aes(x = )
+```
+
+First thing I am going to do is plot histograms of the,
+max likelihoods for all of the ones from each repunit.
+```{r}
+repunit_order <- unique(pop_labels$repunit)
+repu_likes %>%
+  mutate(
+    repunit_f = factor(repunit, levels = repunit_order),
+    inferred_repunit_f = factor(inferred_repunit, levels = repunit_order),
+  ) %>%
+  ggplot() +
+  geom_histogram(aes(x = repu_sclike, fill = repunit_f), colour = "black", linewidth = 0.16) + 
+  facet_grid(repunit_f ~ inferred_repunit_f) + 
+  scale_fill_manual(values = repunit_colors) +
+  theme_bw()
+```
+
 ```{r}
@@
 ggsave(gbp, filename = "tex/images/num-alle-barplot.pdf", width = 4, height = 3)
+
+numa
 ```

### 2024-05-14 002-GSI_and_Fst.Rmd
commit before deleting a bunch

@@
 
-USING_SNAKEMAKE <- FALSE
-NEG_USING_SMK <- TRUE
+```
+
+# Input and Output Paths
+
+```{r}
 if(exists("snakemake")) {
-  USING_SNAKEMAKE <- TRUE
-  NEG_USING_SMK <- FALSE
-}
-```
-
-```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
-# currently, this block does not get evaluated, while developing,
-# but later when we snakemake-ize it all, we will evaluate it.
-if(USING_SNAKEMAKE) {
   input_list <- snakemake@input
   output_list <- snakemake@output
-}
-```
-
-# Input and Output Paths
-
-```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
-# inputs:
+} else {
   input_list <- list(
@@
     final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
+    final_baseline_apr = "data/subRoSA_baseline_rubias_APR2024_w_mixture_revised_rosa_added.csv",
     lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
@@
   )
+}
+
 # you can create the necessary output directories like this:
@@
 # non-winter Sacto fish of unknown provenance
-full_base <- read_csv(input_list$final_baseline) %>%
+###full_base <- read_csv(input_list$final_baseline) %>%
+###  filter(collection != "SacNonW")
+
+full_base_apr <- read_csv(input_list$final_baseline_apr) %>%
   filter(collection != "SacNonW")
+
+# remove the rosa hapstr from it
+full_base <- full_base_apr %>%
+  select(-hapstr)
+
+# capture that rosa hapstr with some meta data from future use
+rosa_hapstrs <- full_base_apr %>%
+  select(indiv:hapstr)
 
@@
 collection_order <- unique(pop_labels$collection)
+
 ```
@@
 
+# Include the LFAR in there in a version
+
+We do these as an additional step to get a sense for how much
+improvement we get from adding the LFAR in there.
+
+First we need to join the LFARs on there:
+```{r}
+# get just the 3 LFAR markers we are using:
+lfar_columns <- paste(
+  rep(c(
+    "NC_037130.1:1062935-1063235",
+    "NC_037130.1:828619-828919",
+    "NC_037130.1:864908-865208"
+  ), each = 2),
+  rep(c(1,2), 3),
+  sep = "_"
+)
+
+lfar3genos <- lfar_genos[c("Sample_ID", lfar_columns)]
+
+base_with_lfar <- full_base2 %>%
+  left_join(lfar3genos, by = join_by(indiv == Sample_ID))
+```
+
+# Process the RoSA hapstrs to include at some point
+
+We are not going to delve on the different between E1 and E2 here.  That is for the
+subRoSA paper.  But we want to get the genotypes simply as EE, EL, LL as done in the
+science paper.
+
+Let us have a quick look at the hapstrs:
+```{r}
+rosa_hapstrs %>%
+  count(hapstr) %>%
+  arrange(desc(n))
+```
+
+We will explode those into a vector of characters at positions
+`c(1, 3, 4, 6:10)`, thus dropping the E2-specific variants as
+well as the Tasha2 SNP that effectively tags the duplication, leaving
+us with the 8 core RoSA SNPs.
+```{r}
+H8 <- rosa_hapstrs %>%
+  select(indiv, hapstr) %>%
+  mutate(
+    hvec = map(hapstr, function(x) {y <- strsplit(x, "")[[1]]; y[c(1, 4, 5, 7:11)]}),
+    hapstr8 = map_chr(hvec, function(x) paste(x, collapse = "", sep = ""))
+  )
+
+# Let's count up those hapstr8's
+H8 %>%
+  count(hapstr8) %>%
+  arrange(desc(n))
+```
+
+From that it is clear that there have been some recombinations in here, but we will
+just go ahead and call RoSA genotype by the majority of the non-missing ones. 
+```{r}
+H8_strip <- H8 %>%
+  mutate(
+    stripped = map(hvec, function(x) x[!(x %in% c("?", "N"))]),
+    sglom = map_chr(stripped, function(x) paste(x, collapse = "", sep = ""))
+    )
+
+H8_strip %>%
+  count(sglom) %>%
+  arrange(desc(n))
+
+```
+
+OK, we will call something and EL if it has two or more Hs (the idea being that a single
+H could be a genotyping error). Note that there are no haplotypes mixing L's and E's together.
+Here is a function that will determine the RoSA genotype
+from a vector of H, L, and E parts. 
+```{r}
+call_rosa <- function(x) {
+  # things that were all ?'s get NA
+  if(length(x) == 0) {
+    return(NA_character_)
+  }
+
+  nE <- sum(x == "E")
+  nL <- sum(x == "L")
+  nH <- sum(x == "H")
+
+  if(nH >= 2) { 
+    return("EL")
+  } else if(nL > 0 && nE == 0) {
+    return("LL")
+  } else if(nE > 0 && nL == 0) {
+    return("EE")
+  } else {
+    return("Weird")
+  }
+
+}
+
+H8_genos <- H8_strip %>%
+  mutate(rosa_geno = map_chr(stripped, call_rosa))
+```
+
+Count these up so we can see what we called LL, EL, and EE:
+```{r}
+H8_genos %>%
+  count(sglom, rosa_geno) %>%
+  arrange(desc(n))
+```
+
+Now we will slim it down for future use:
+```{r}
+rosa_genos <- H8_genos %>%
+  select(indiv, rosa_geno)
+```
+
 # Do the self-assignment
 
-In this round, we go directly to assigning to repunit, and then to the max-likelihood
-collection within that repunit.
+We go directly to assigning to repunit, and then to the max-likelihood
+collection within that repunit.  We include the lfar in here.
 
@@
 ```{r}
-top_ass <- full_sa %>%
-  group_by(indiv, collection, repunit, inferred_repunit) %>%
-  mutate(repu_sclike = sum(scaled_likelihood)) %>%
-  group_by(indiv) %>%
-  filter(repu_sclike == max(repu_sclike)) %>%
-  arrange(indiv, desc(scaled_likelihood)) %>%
-  slice(1) %>%
-  ungroup() %>%
-  mutate(coll_f = factor(collection, levels = collection_order))
+source("R/get_top_assignments.R")
+
+top_ass <- get_top_assignments(full_sa, collection_order)
+```
+
+Plot that initial table:
+```{r}
+source("R/plot_self_assigment_table.R")
+
+no_lfar_ass_table <- plot_self_assignment_table(top_ass, pop_labels, collection_order)
+
+no_lfar_ass_table_80 <- top_ass %>% 
+  filter(scaled_likelihood > 0.8) %>%
+  plot_self_assignment_table(pop_labels, collection_order)
+
+no_lfar_ass_table_Rosa <- top_ass %>%
+  left_join(rosa_genos, by = join_by(indiv)) %>%
+  plot_self_assignment_table(pop_labels, collection_order)
+```
+
+And then do the same for the data with the LFAR:
+```{r}
+lfar_sa <- self_assign(base_with_lfar, gen_start_col = 5)
+lfar_top_ass <- get_top_assignments(lfar_sa, collection_order)
+lfar_ass_table <- plot_self_assignment_table(lfar_top_ass, pop_labels, collection_order)
+
+lfar_ass_table_80 <- lfar_top_ass %>% 
+  filter(scaled_likelihood > 0.8) %>%
+  plot_self_assignment_table(pop_labels, collection_order)
+
+lfar_ass_table_Rosa <- lfar_top_ass %>%
+  left_join(rosa_genos, by = join_by(indiv)) %>%
+  plot_self_assignment_table(pop_labels, collection_order)
 ```
@@
 ```{r}
-ggplot(focal_comparisons, aes(x = )
+ggplot(focal_comparisons, aes(x = ))
 ```

### 2024-05-14 002-GSI_and_Fst.Rmd
Big overhaul just before I gave lab meeting about this

@@
     assig_table_lfar_all = "results/GSI_and_Fst/assignment-table-full-baseline-plus-lfar.pdf",
+    ass_table80 = "results/GSI_and_Fst/ass-table-80.pdf",
+    rosa_gsi_table = "results/GSI_and_Fst/rosa-gsi-table.pdf",
     tex_table = "tex/inputs/samples-table.tex",
@@
     gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
-    gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf"
+    gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf",
+    ass_table80_crop = "results/GSI_and_Fst/ass-table-80-crop.pdf",
+    ass_table80_tex = "tex/images/ass-table-80-crop.pdf",
+    rosa_gsi_table_crop = "results/GSI_and_Fst/rosa-gsi-table-crop.pdf",
+    rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf"
   )
@@
 
-no_lfar_ass_table <- plot_self_assignment_table(top_ass, pop_labels, collection_order)
-
-no_lfar_ass_table_80 <- top_ass %>% 
+no_lfar_ass_table <- plot_self_assignment_table(top_ass$top, pop_labels, collection_order)
+
+no_lfar_ass_table_80 <- top_ass$top %>% 
   filter(scaled_likelihood > 0.8) %>%
@@
 
-no_lfar_ass_table_Rosa <- top_ass %>%
+no_lfar_ass_table_Rosa <- top_ass$top %>%
   left_join(rosa_genos, by = join_by(indiv)) %>%
@@
 lfar_top_ass <- get_top_assignments(lfar_sa, collection_order)
-lfar_ass_table <- plot_self_assignment_table(lfar_top_ass, pop_labels, collection_order)
-
-lfar_ass_table_80 <- lfar_top_ass %>% 
+lfar_ass_table <- plot_self_assignment_table(lfar_top_ass$top, pop_labels, collection_order)
+
+# make a version with no legend on it
+lfar_ass_table_no_legend <- plot_self_assignment_table(lfar_top_ass$top, pop_labels, collection_order, no_legend = TRUE)
+
+lfar_ass_table_80 <- lfar_top_ass$top %>% 
   filter(scaled_likelihood > 0.8) %>%
@@
 
-lfar_ass_table_Rosa <- lfar_top_ass %>%
+lfar_ass_table_Rosa <- lfar_top_ass$top %>%
   left_join(rosa_genos, by = join_by(indiv)) %>%
@@
 
-And, once again, we want to investigate the ones that are in the wrong repunit, and look at their
-scaled_repu_likelihoods, the others.
-```{r}
-source("R/colors.R")
-ta_corr <- top_ass %>%
-  filter(repunit == inferred_repunit)
-ta_wrong <- top_ass %>%
-  filter(repunit != inferred_repunit) %>%
-  arrange(collection, repu_sclike) %>%
-  group_by(collection) %>%
-  mutate(y = 190 + (70/9) * 1:n())
-
-ggplot() +
-  geom_histogram(
-    data = ta_corr, 
-    mapping = aes(x = repu_sclike, fill = inferred_repunit)
-  ) +
-  geom_point(
-    data = ta_wrong,
-    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
-    size = 3, 
-    shape = 21
-  ) +
-  facet_wrap(~coll_f, ncol = 3) +
-  scale_fill_manual(values = repunit_colors)
-```
-Which looks to me like there are really only 7 here that we are not getting right.
-Those others we will filter out and make a table of them:
-
-```{r}
-tafilt <- top_ass %>%
-  filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
-```
-
-Here we calculate the number of correctly and incorrectly assigned fish:
-```{r}
-tafilt %>%
-  mutate(correct = repunit == inferred_repunit) %>%
-  count(correct) %>%
-  mutate(fract = n / sum(n))
-```
-
-And now we will make a table of those:
-```{r}
-# get the function
-source("R/taf-prep.R")
-source("R/table-as-figure.R")
-
-# get the RC_groups
-RC_groups <- pop_labels %>%
-  select(-old_name) %>%
-  distinct() %>%
-  rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
-  select(external_group, internal_group, label)
-
-X2 <- tafilt %>%
-  mutate(
-    collection_f = factor(collection, levels = collection_order),
-    inferred_coll_f = factor(inferred_collection, levels = collection_order)
-  ) %>%
-  count(collection_f, inferred_coll_f, .drop = FALSE) %>%
-  taf_prep()
-
-Rubias_for_later <- X2 # keeping this for later...
-
-# get the result
-TAF2 <- table_as_figure(
-  X = X2, 
-  RC_groups = RC_groups,
-  external_colors = run_time_colors,
-  internal_colors = repunit_colors, 
-  plot_margins = c(1, 0.1, 0.7, 0.1)
-)
-
-# and plot it
-g <- plot_grid(
-  TAF2$full_plot,
-  plot_grid(
-    get_legend(TAF2$for_external_legend),
-    get_legend(TAF2$for_internal_legend),
-    nrow = 1
-  ),
-  nrow = 2,
-  rel_heights = c(7,3)
-)
-
-g
-```
-
-That is a considerable improvement. Let's write it out:
-```{r}
-ggsave(g, filename = output_list$assig_table, width = 8, height = 10)
-```
-
-# Now, we repeat that while adding the LFAR markers in there
-
-We do these as an additional step to get a sense for how much
-improvement we get from adding the LFAR in there.
-
-First we need to join the LFARs on there:
-```{r}
-# get just the 3 LFAR markers we are using:
-lfar_columns <- paste(
-  rep(c(
-    "NC_037130.1:1062935-1063235",
-    "NC_037130.1:828619-828919",
-    "NC_037130.1:864908-865208"
-  ), each = 2),
-  rep(c(1,2), 3),
-  sep = "_"
-)
-
-lfar3genos <- lfar_genos[c("Sample_ID", lfar_columns)]
-
-base_with_lfar <- full_base2 %>%
-  left_join(lfar3genos, by = join_by(indiv == Sample_ID))
-```
-
-Then, all the step from above:
-```{r}
-# Do self-assignment:
-full_plus_lfar_sa <- self_assign(base_with_lfar, gen_start_col = 5)
-
-repu_likes <- full_plus_lfar_sa %>%
-  group_by(indiv, collection, repunit, inferred_repunit) %>%
-  summarise(repu_sclike = sum(scaled_likelihood)) %>%
-  ungroup()
-
-top_ass_w_lfar <- full_plus_lfar_sa %>%
-  group_by(indiv, collection, repunit, inferred_repunit) %>%
-  mutate(repu_sclike = sum(scaled_likelihood)) %>%
-  group_by(indiv) %>%
-  filter(repu_sclike == max(repu_sclike)) %>%
-  arrange(indiv, desc(scaled_likelihood)) %>%
-  slice(1) %>%
-  ungroup() %>%
-  mutate(coll_f = factor(collection, levels = collection_order))
-
+### Get some raw numbers
+
+```{r}
+# total number of fish
+tot_fish <- nrow(lfar_top_ass$top)
+tot_fish
+
+# Get the number and fraction correct
+lfar_top_ass$top %>%
+  count(repunit == inferred_repunit) %>%
+  mutate(fract = n/tot_fish)
+```
+
+```{r}
+over80 <- nrow(lfar_top_ass$top %>% filter(repu_sclike > 0.8))
+
+lfar_top_ass$top %>% filter(repu_sclike > 0.8) %>% count(repunit == inferred_repunit) %>% mutate(fract = n / over80)
+```
+
+Make a plot for > 80%
+```{r}
+
+# and then save that
+ggsave(lfar_ass_table_80$plot, file = output_list$ass_table80, width = 10, height = 10.5)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$ass_table80, collapse = " ")
+system(CALL)
+
+file.rename(from = output_list$ass_table80_crop, to = output_list$ass_table80_tex)
+```
+
+Make a plot for the byRosa:
+```{r}
+
+# and then save that
+ggsave(lfar_ass_table_Rosa$plot, file = output_list$rosa_gsi_table, width = 10, height = 12.5)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$rosa_gsi_table, collapse = " ")
+system(CALL)
+
+file.rename(from = output_list$rosa_gsi_table_crop, to = output_list$rosa_gsi_table_tex)
 ```
@@
 ```{r}
-misassigned_ones <- top_ass_w_lfar %>%
+misassigned_ones <- lfar_top_ass$top %>%
   filter(repunit != inferred_repunit) %>%
@@
 
+pick_em <- lapply(1:nrow(MO2), function(i) {
+
+  FR <- MO2$Foc_repunit[i]
+  FIR <- MO2$Foc_inferred_repunit[i]
+
+  lfar_top_ass$repu_likes %>%
+    ungroup() %>%
+    filter(
+      (repunit == FR & inferred_repunit == FIR) |
+        (inferred_repunit == FR & repunit == FIR)
+    ) %>%
+    mutate(
+      Foc_repunit = FR,
+      Foc_inferred_repunit = FIR,
+      .before = indiv
+    ) 
+}) %>%
+    bind_rows()
+```
+
+OLD STUFF BELOW HERE
+```{r}
 tmp <- lapply(1:nrow(MO2), function(i) {
-  repu_likes %>%
+  lfar_top_ass$repu_likes %>%
     filter(
@@
   ) %>%
+  ungroup() %>%
   #count(Foc_repunit, Foc_inferred_repunit, repunit, inferred_repunit, rep, inf)
@@
 
+I include the LFAR in their because I want to have it.  We note this is not for
+demographic inference or anything.  It is just a summary of the differentiation
+of the markers...
+
 In order to do this we need to convert the data to a lumped format.
@@
 ```{r}
-full_base_long <- full_base2 %>% 
+full_base_long <-  base_with_lfar %>% 
   select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
@@
 # long format with no missing data
-long_genos <- full_base2 %>%
+long_genos <- base_with_lfar %>%
 select(-repunit, -sample_type) %>%
@@
 # diagonal and we set all the cell_labels to NA.
-lower_tri_base <- Rubias_for_later %>%
+lower_tri_base <- lfar_ass_table$RFL %>%
 filter(as.integer(col_label) <= as.integer(row_label)) %>%
@@
 arrange(row_label, col_label)
+
+# Get the RC_groups
+RC_groups <- pop_labels %>%
+    select(-old_name) %>%
+    distinct() %>%
+    rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
+    select(external_group, internal_group, label)
 
@@
 gsi_fst_fig_for_paper <- plot_grid(
-TAF2$full_plot,
+  fst_with_scatters,
   plot_grid(
-get_legend(TAF2$for_external_legend),
-get_legend(TAF2$for_internal_legend),
+    get_legend(lfar_ass_table$for_external_legend),
+    get_legend(lfar_ass_table$for_internal_legend),
     nrow = 1
   ),
-fst_with_scatters,
+  lfar_ass_table_no_legend$plot,
   nrow = 3,

### 2024-05-14 tex/body-text-of-paper.tex
Big overhaul just before I gave lab meeting about this

@@
 
+Thus, any panel of markers to be used for GSI (or PBT) should also be evaluated on its
+utility for PBT (or GSI).  PBT has been remarkably successful in fisheries
+management, having largely replaced coded wire tags in large parts of the Snake River and
+Columbia River basin (CITATION), and having been employed to dramatically further our
+understanding of the genetic inheritance of key traits in salmonids
+\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of whole
+family of statistical genetic methods employing relationship inference to learn about
+populations.  For example inference of the full siblings amongst a sample of fish
+provides information about the effective number of adults producing offspring
+\citep{waples2011inbreeding,wang2023estimating}, which can
+be a valuable source of information when sampling of the adults is not possible.
+
 Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
@@
 
+In the following we present a new reference baseline for both GSI and PBT in California.
+We describe how we identified and compiled the markers, including several new markers
+that are particularly effective at distinguishing some of the closely related groups,
+such as late-fall and fall-run fish.  We also provide a comprehensive analysis of this
+set of markers for inference of parents and siblings within different groups of populations.
+
  \section*{Methods}
-
-\comment{More verbiage}
 
@@
 
-We sampled fish for the reference baselines from XX locations across California.
+We sampled fish for the reference baselines from 13 locations within California, and one 
+within Oregon.
 These locations included sites in the two
 main tributaries (the Sacramento River and the San Joaquin River) within
-California's Central Valley, the Klamath-Trinity basin in Northern California, and
-3 rivers of the California coast (Russian, Eel, etc)
-(Fig.~\ref{fig:map}).
+California's Central Valley, the Klamath and Trinity River basins in Northern California, and
+3 rivers of the California coast (Russian, Eel, and Smith)
+(Fig.~\ref{fig:map}, Table~\ref{tab:samples}).
 %%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%
-Collections in each location were separated into fish from the different run-timing ecotypes according
-to a variety of criteria. The last vestiges of winter-run Chinook salmon are maintained at the Livingston
+
+Collections in each location were separated into fish from the different adult migration timing ecotypes according
+to a variety of criteria.  Winter-run Chinook salmon are propagated at the Livingston
 Stone National Fish Hatchery for supplementation (released as juveniles into the river) and as a captive broodstock;
 sampling of this population was performed by hatchery staff during spawning. In Central Valley Rivers with both 
-spring and fall ecotypes (i.e. Mill, Deer and Butte Creeks), previous research has documented distinct/disparate/unique
-spawning times (Julian date) for each ecotype (REF?). As such, sampling targeted carcasses found in each of these 
-rivers at times representative of their run-timing designation, which we were able to further verify using their genotypes
-at the RoSA markers. At the time of sampling, only fall-run Chinook were found in the San Joaquin River. Run-timing 
-designation at the Feather River Hatchery is more complicated but is managed at a fine-scale by hatchery practices, 
-where early returning fish in May and April (expressing the Spring-run life history) are visibly tagged and re-released 
+spring-run and fall-run ecotypes (i.e.
+Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype (REF?).
+\comment{As such, sampling targeted carcasses found in each of these
+rivers at times representative of their run-timing designation}, which we were able to further confirm using their genotypes
+within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}.
+At the time of sampling, only fall-run Chinook were found in the San Joaquin River so the samples there were
+easily categorized. Run-timing 
+designation at the Feather River Hatchery is more complicated but is managed by the hatchery practices in place: 
+early returning fish in May and April (expressing the spring-run life history) are visibly tagged and re-released 
 to the river. Subsequently in the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
-as Spring-run broodstock. Individual-based analyses of RoSA genotypes have demonstrated
+as spring-run broodstock, the rest are included in the fall-run broodstock.
+We sampled putative spring-run and fall-run fish from these two broodstocks.
+\comment{Need info about sampling in the Klamath, and on the coast}.
 
-(We need to talk about how these were split into spring run, fall-run, winter 
-run, etc.
 In total, the samples in the baseline represent fish from five different Evolutionarily Significant 
-Units (ESUs) of Chinook salmon \citep{waples1991pacific} A summary of the different collections 
+Units (ESUs) of Chinook salmon \citep{waples1991pacific}.
 %%%%%%%%%%%%%%%%
@@
 did not amplify well or those that amplified so well that they were grossly over-represented
-among the reads on the sequencer.  We then 
+among the reads on the sequencer.  We 
 used Freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
@@
 
-We undertook a new discovery process to develop additional multiallelic
-microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook.
+We undertook a further discovery process to develop additional multiallelic
+microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon.
 \comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
@@
 
+\subsubsection*{Localizing markers within the genome}
+
+The existing SNPtype assays and the additional microhaplotype markers were
+developed in the absence of a well-assembled genome for Chinook salmon
+\comment{(Is this true?)}. As a consequence, many of these markers had
+been used previously without certainty about their location in the genome.
 After designing and validating an amplicon-sequencing approach to type both the
 existing SNPtype assays and the product of the additional microhaplotype
-discovery process we identified their location with the genome. We mapped the reference
-sequences for these markers 
-to the Chinook v2.0 genome assembly (\url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/})
-with both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
-We report the genomic coordinates these target amplicons found with both alignment methods.
+discovery process we identified the locations of these markers within the genome by mapping the
+consensus sequences (roughly 300 bp) around the markers
+to both the Chinook salmon version 1 (Otsh\_v1.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_002872995.1/}) and version 2  (Otsh\_v2.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/}) genome assemblies,
+using both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
+A similar mapping exercise was performed for the remaining markers so that we can
+report their locations in both Otsh\_v1.0 and Otsh\_v2.0.
 
@@
 we first conducted a simple case-control GWAS using the low-coverage whole genome sequencing 
-data of \citet{thompson2020complex}.  We conducted two separate association-study comparisons.  
-The first involved the 16 late-fall run fish as cases with 16 fall-run fish from the Feather River Hatchery as controls. 
+data of \citet{thompson2020complex}, mapped to Otsh\_v1.0.  We conducted
+two separate association-study comparisons.  
+The first involved the 16 late-fall-run fish as cases with 16 fall-run fish from the Feather River Hatchery as controls. 
 The second involved the same 16 late-fall fish as cases, but used 16 fall-run fish from the San 
@@
 with either fall-run or late-fall run being nearly fixed for one of the alleles, or
-\item An annotation from snpEff \citep{cingolani2012program} of “HIGH” or “MODERATE”, or
+\item An annotation from snpEff \citep{cingolani2012program} of “High” or “Moderate”, or
 \item An absolute frequency difference $>0.48$ and a lower confidence interval of $>0.25$ and a 
@@
 
-We include amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish described in \citet{thompson2020complex}.
-\comment{Need more here. I personally am of the thought that we should not include these
-for GSI, but should use them for identifying the genotypes of fish assigned to populations
-using the rest of the markers.}
+We include in the reference baseline amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
+between spring-run and fall-run fish described in \citet{thompson2020complex}.  These are useful markers to
+have in the panel, as typing them routinely will create a data base for understanding the spatial and temporal
+distribution of the alleles at this region; however we have omitted these markers for calculating likelihoods for
+GSI, since their distribution across populations in the Feather River has been clearly and evidently
+influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.
 
@@
 Basic data summaries were calculated for each collection
-using simple operations in the R programming language, Version 4.2
+using simple operations in the R programming language, Version 4.3
 \citep{rcore}. We started with a simple tally of the number of alleles
 found at each locus.  To provide a summary of the genetic variation and
-missing data we report
+missing data we report:
 \begin {itemize}
@@
 in which each locus was successfully genotyped, $\bar{N}$;
-\item the average number of alleles per locus with all collections
+\item the average number of alleles per locus in each collection with all collections
 downsampled to the smallest number of genotyped individuals at each
-locus, $\bar{N}_\mathrm{A,ss}$,
+locus, $\bar{N}_\mathrm{A,ss}$;
 \item the fraction of polymorphic
 loci in each collection after subsampling to the smallest number of genotyped
-fish, $\bar{P}_\mathrm{poly,ss}$,
+fish, $\bar{P}_\mathrm{poly,ss}$;
 \item the expected heterozygosity as the average over loci
 of one minus the frequencies of homozygotes at each locus expected from the
-estimated allele frequencies, $\bar{H}_\mathrm{exp}$, and 
+estimated allele frequencies, $\bar{H}_\mathrm{exp}$; and 
 \item the observed heterozygosity as the average over
@@
 structure in the data by performing unsupervised clustering of
-the complete data set (\comment{With or without the LFAR? Which did Anthony use?}) 
+the complete data set (\comment{With or without the LFAR? Which did Anthony use? I would
+like to do it with the LFAR.}) 
 using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
@@
 \comment{I think we should also see what $K=7$ turns up}
-using default settings of the program.
+using default settings of the program.  STRUCTURE output was
+a summarized and visualized using CLUMPAK version~\comment{??}
+\citep{kopelman2015clumpak}.
 
@@
 fish were assigned to the collection within that reporting unit with the highest
-scaled likelihood.
+scaled likelihood.  We considered two criteria for assignment.  In the first,
+all fish are assigned, in the second only fish with a summed (i.e., summed
+over all collections in a reporting unit)
+scaled likelihood greater than 0.8 were assigned.
+
+Finally, we investigated the matrix of assignments and misassignments
+broken down according to the genotype of the fish at the markers within
+the RoSA.  For this, \comment{Need to describe how I categorized
+genotypes}.
 
 \subsection*{Power for relationship inference}
+
+To explore the utility of the marker panel for relationship inference, we used
+the R package `CKMRsim'
+(\url{https://github.com/eriqande/CKMRsim})
+to estimate the false-positive rates (FPRs) expected at
+different values of the false negative rate (FNR) for a variety of different pairwise
+relationships in different populations.  For these purposes, we divided the
+collections into 10 different groups.  The members of each group are
+relatively similar For all cases, we assumed that the genotyping error rate
+was 1\% per locus and the genotyping error model was the default
+"True-genotype-independent" model.
+
+For errors involving misidentifying
+unrelated (U) pairs as parent-offspring (PO), full-sibling (FS), or half-sibling (HS)
+pairs we used importance sampling to estimate the very small FPRs.  The
+FNRs were estimated in each case using simple Monte Carlo.  For the FS and
+HS cases the FNRs were calculated from Monte Carlo sampling while taking account
+of the physical linkage of the markers by inputting their genomic positions into
+'CKMRsim' and using the MENDEL \citep{lange2013mendel} software to simulate genotypes of related pairs
+in the presence of physical linkage.  For PO pairs (and U pairs),
+FNRs were calculated without reference to the genomic locations of the markers
+because physical linkage does not
+affect the distribution of log-likelihood ratios in those two relationship types,
+because there is no variation, anywhere
+in the genome, in the number of gene copies shared identical by descent between
+a parent and its child or between two unrelated individuals.
+
+For errors involving misidentifications between the relationship
+types of avuncular (AN: aunt-niece, uncle-nephew, etc.), PO, FS, and HS,
+we used regular Monte Carlo to estimate the FPRs and FNRs, because
+importance sampling cannot be used in those cases, and, since there are
+expected to be far fewer such relationships than the number of unrelated
+pairs expected, the relevant FPRs are not so tiny and can be estimated
+(at least down to rates of $10^{-3}$) reliably.  For AN, FS, and HS, both the
+FPRs and FNRs were estimated while taking account of physical linkage
+of the markers, as described above.
 
@@
 
+\subsubsection*{Localizing markers within the genome}
+
 Of these \comment{185} loci that originated from SNPtype assays or from our additional
@@
 disagreed slightly between BLAT and bwa mem, and one of them had multiple
-non-primary alignments (\comment{SuppTable}).
+non-primary alignments (Table~S1 \comment{This is the Excel File.  It needs to
+be edited so as to: 1) have a column
+giving the category or provenance of the markers (i.e. LFAR or MHAP or SNPTYPE, etc.) 2) retain
+only the 2 LFAR markers we actually use. 3) consider renaming the E2 markers since
+winter-run-greb1L-diagnostic is confusing...actually, I would consider removing those
+entirely---we can introduce them in the subRoSA paper. We don't use them in this paper and I don't
+really want to do the explanation of them again.}).
+
+Mapping to Otsh\_v1.0 was similar, except that a greater number of markers mapped to unplaced
+scaffolds in Otsh\_v1.0 than in Otsh\_v2.0, likely reflecting the more complete nature of the
+second assembly.
 
@@
 chromosome 34 had large allele frequency differences between late-fall and fall run.  Taking the
-top 10 most associated SNPs, and including additional ones from the filtering criteria given in the
+10 SNPs with the lowest association $p$-values, and including additional ones from the filtering criteria given in the
 Methods, yielded 49 candidate markers for which we designed primers for amplication.  Only 8
@@
 showed marked allele frequency differences ($> 0.7$) between late-fall and fall run.  These three 
-SNPs (Chr34:828,768,  Chr34:865,057, and Chr34:1,063,084) also had pronounced differences in 
-allele frequency between 
+SNPs (located at: Chr34:828,768,  Chr34:865,057, and Chr34:1,063,084 in the Otsh\_v1.0 assembly)
+also had pronounced differences in allele frequency between 
 late-fall and spring or winter run.
@@
 etc.), making it costly (in terms of sequencing effort) to include the marker in the baseline panel. 
-Consequently it was dropped.  The frequency of the different alleles at the two remaining loci
+\comment{Consequently it was dropped.}  The frequency of the different alleles at the two remaining loci
 within different reporting units in the baseline are shown in Table~\ref{tab:lfar-freqs}.
@@
 loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
-showed more than two alleles, providing significantly more power for relationship
-inference than a comparably sized panel of biallelic SNPs.
+showed more than two alleles, which has been shown to provide significantly more
+power for relationship inference than a comparably sized panel of biallelic SNPs
+\citep{baetscher2018microhaplotypes}.
 
-And the table of the popgen summary is Table~\ref{tab:pg-summ}.
+The population-genetic summary statistics are presented in Table~\ref{tab:pg-summ}.
 %%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%
+The average number of fish with non-missing data in each population varied from
+a low of 39.2 (BCF) to a high of 275.2 (CHLF).  The average number of alleles
+per locus, standardized to the smallest sample size, ranged from 2.35 (SRW) to
+2.68 (RRF), with those two populations also showing the most extreme values for the
+standardized fraction of polymorphic markers at 0.9 and 0.98, respectively.
+SRW also had the lowest values for expected and observed heterozygosity:
+0.349 and 0.340, respectively. While the highest for expected heterozygosity
+was 0.412 (SmRF) and for observed heterozygosity, 0.412 (SmRF).
+
+The estimated pairwise $F_\mathrm{ST}$ values between the collections
+varied from a low of 0.0 to a high of 0.273 (Figure~\ref{fig:gsi}a).
+Notably, most of the largest values
+of $F_\mathrm{ST}$ occurred between SRW and another collection.
+
+The program STRUCTURE clustered fish at all values of $K$ from 2 to \comment{7}.
+(Fig.~\ref{fig:struct}).  The clusters generally corresponded to groupings of related populations, and confirmed
+{\em a priori} knowledge about which collections in the baseline could and should be
+grouped together into reporting units.  At every value of $K$ CLUMPAK discerned at least
+11 or 20 replicates in the major mode (Fig.~\ref{fig:struct}). Clustering solutions in the minor
+modes appear in Fig.~\ref{fig:minor-modes}.
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\newcommand{\structcap}{\footnotesize Major modes found by STRUCTURE and summarized
+by CLUMPAK out of 20 replicates for each value of $K$, the assumed number of
+subpopulations. \comment{We kind of need to make our own figure here that doesn't take up
+as much whitespace and also gives the number of occurrences of each mode.} }
+\begin{center}
+\includegraphics[width=\columnwidth]{images/major-modes-crop.pdf}
+\end{center}
+\caption[\structcap]{\structcap}
+\label{fig:struct}
+\end{figure}
 
@@
 
-Results of power analysis for GSI and also Fst estimates are in Figure~\ref{fig:gsi}
+Leave-one-out cross-validation (self-assignment) analysis demonstrates that this
+reference baseline provides a high degree of accuracy for distinguishing the major
+groups of fish in California.  The results of the analysis are summarized
+in a self-assignment matrix (Figure~\ref{fig:gsi}). 
 %%%%%%%%%%%%%%%%%%%%%% 
 \begin{figure}
-\newcommand{\gsicap}{\footnotesize Self-assignment rates  ({\bf a)} and pairwise 
-$F_\mathrm{ST}$ values and allele frequencies ({\bf b)} for the collections in the baseline.  The codes of the populations
+\newcommand{\gsicap}{\footnotesize Pairwise $F_\mathrm{ST}$ values and allele frequencies({\bf a)} 
+and self-assignment rates
+ ({\bf b)}  for the collections in the baseline.  The codes of the populations
 appear on the outer margins of the table atop cells colored according to the run-timing group
 of the collection.  Interior cells in which the column and row correspond to the same reporting
-unit are colored according to the reporting unit.  For self-assignment tallies, the row corresponds
-to the true source collection and the column corresponds to the assigned collection. In ({\bf b}), for $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
+unit are colored according to the reporting unit.  In ({\bf a}), for $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
 data. Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
@@
 subsampling individuals from each collection to the minimum number of observed alleles in any collection
-at the locus.  }
+at the locus.  In ({\bf b}), for self-assignment tallies, the row corresponds
+to the true source collection and the column corresponds to the assigned collection.
+\comment{Hey! Didn't we drop one of the LFARs? I see three, still.}}
 \begin{center}
@@
 %%%%%%%%%%%%%%%%%%%%
+This figure shows the collections divided into seven different reporting groups:
+SO-Ncal-Coast,  Klamath-Trinity, Cent. Cal. Coast,
+CV-Winter, CV-Spring, CV-Fall, and  CV-Late-Fall.  \comment{Really, where and how
+do we want to introduce these reporting units?} These groupings correspond
+largely to the divisions in the data found by STRUCTURE, and they also correspond
+with divisions amongst the stocks as defined for management.
+
+Overall, of the 1613 fish in the reference baseline, 1587 (98.4\%) were correctly
+assigned to their reporting unit of origin, while 26 (1.6\%) were incorrectly assigned.
+Four of those 26 misassignments which involved fish being incorrectly assigned to
+SRW, were clearly fish from SRW that had been incorrectly sampled into a different
+collection, likely due to straying. (This is evident because SRW are so distinct that it is not possible
+to incorrectly identify them).  Of the remaining 22 misassignments,  9 involved fish from the
+CHLF (late fall run) collection being misassigned to the CV-Fall reporting unit and 7 were
+fish from the MDS (Mill-Deer spring run) collection being assigned to the CV-Fall reporting unit.
+It is possible that some of these misassignments represent CV-Fall fish that were incorrectly
+sampled as CHLF or MDS; however, unlike with SRW, that we could not conclude that
+confidently, as there is some degree of overlap in likelihood values between the two.
+
+While fish from CHLF were found to misassign at a rate of 3.1\% (9/291) and MDS at a rate 
+of 8.6\% (7/81) to CV-Fall, are seen to misassign at fairly high rates to CV-Fall, the
+converse is not evident---that is, fish from the CV-Fall are misassigned to CHLF at a rate
+of only 0.6\% (3/501) and to the CV-Spring reporting unit (to which MDS belongs) at a rate
+of only 0.4\% (2/501).
+
+When assignments are only accepted with a scaled likelihood greater than 0.8,
+18 assignments are discarded, but not all of those were incorrect.  At the 0.8 threshold,
+the total fraction of correct assignments is 98.8\% (Fig.~\ref{fig:eighty}).
 

### 2024-05-14 tex/supplemental-figures.tex
Big overhaul just before I gave lab meeting about this

@@
 %%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{images/minor-modes-crop.pdf}
+\end{center}
+\caption[STRUCTURE minor modes found by CLUMPAK]{\footnotesize At each value of $K$
+for which a minor mode was found, these plots show them}
+\label{fig:minor-modes}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{images/ass-table-80-crop.pdf}
+\end{center}
+\caption[Assignment table for fish with scaled likelihood $ > 0.8$]{\footnotesize Assignment table
+like that in Fig.~\ref{fig:gsi}b in the paper, but constrained so that only fish assigning
+to a reporting unit with scaled likelihood greater than 0.8 are included.}
+\label{fig:eighty}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{images/rosa-gsi-table-crop.pdf}
+\end{center}
+\caption[Assignment table by RoSA genotype]{\footnotesize Assignment table
+like that in Fig.~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
+at the RoSA}
+\label{fig:rosa-gsi}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%

### 2024-05-14 R/colors.R

@@ -21,3 +21,8 @@ run_time_colors <- c(
 )
 
 
+rosa_colors <- c(
+  EE = "gold",
+  EL = "tan2",
+  LL = "blue"
+)

### 2024-05-14 R/get_top_assignments.R

@@ -1,0 +1,28 @@
+
+
+#' Get the top assignments from the self_assign() output
+#'
+#' We assign to repunit, then assign to the most likely collection
+#' within the most likely repunit
+#' @param full_sa The full output tibble from the self_assign() function
+#' @param collection_order A vector giving the order you want the collections
+#' sorted.  It creates a factor coll_f from this.
+get_top_assignments <- function(full_sa, collection_order) {
+
+  repu_likes <- full_sa %>%
+    group_by(indiv, collection, repunit, inferred_repunit) %>%
+    mutate(repu_sclike = sum(scaled_likelihood))
+
+  top <- repu_likes %>%
+    group_by(indiv) %>%
+    filter(repu_sclike == max(repu_sclike)) %>%
+    arrange(indiv, desc(scaled_likelihood)) %>%
+    slice(1) %>%
+    ungroup() %>%
+    mutate(coll_f = factor(collection, levels = collection_order))
+
+  list(
+    top = top,
+    repu_likes = repu_likes
+  )
+}

### 2024-05-14 R/plot_self_assigment_table.R

@@ -1,0 +1,79 @@
+
+#' from the top-assignments make a colorful assignment table
+#'
+#' The top assignments are like those that come out of get_top_assignments()
+#' or a filtered version thereof
+#' @param pop_labels  The pop_labels tibble
+#' @param top_ass the top assignments. If it has a column "rosa_geno" then this
+#' will trigger counting things broken down by rosa_geno, too.
+#' @param collection_order A vector giving the order you want the collections
+#' sorted.
+#' @return Returns a list.  Component "plot" of which is the ggplot of the assignment
+#' table, and component RFL is the Rubias_for_later output---basically
+#' the top assignments with some factors in there for getting the order right.
+plot_self_assignment_table <- function(top_ass, pop_labels, collection_order, no_legend = FALSE) {
+
+  # get the function
+  source("R/taf-prep.R")
+  source("R/table-as-figure.R")
+
+  # get the RC_groups
+  RC_groups <- pop_labels %>%
+    select(-old_name) %>%
+    distinct() %>%
+    rename(external_group = run_timing_group, internal_group = repunit, label = collection) %>%
+    select(external_group, internal_group, label)
+
+  if(!("rosa_geno" %in% names(top_ass))) {
+    X2 <- top_ass %>%
+      mutate(
+        collection_f = factor(collection, levels = collection_order),
+        inferred_coll_f = factor(inferred_collection, levels = collection_order)
+      ) %>%
+      count(collection_f, inferred_coll_f, .drop = FALSE) %>%
+      taf_prep()
+  } else {
+    X2 <- top_ass %>%
+      mutate(
+        collection_f = factor(collection, levels = collection_order),
+        inferred_coll_f = factor(inferred_collection, levels = collection_order),
+        rosa_geno_f = factor(rosa_geno, levels = c("EE", "EL", "LL"))
+      ) %>%
+      count(collection_f, inferred_coll_f, rosa_geno_f, .drop = FALSE) %>%
+      taf_prep()
+  }
+
+  Rubias_for_later <- X2 # keeping this for later...
+
+  # get the result
+  TAF2 <- table_as_figure(
+    X = X2,
+    RC_groups = RC_groups,
+    external_colors = run_time_colors,
+    internal_colors = repunit_colors,
+    plot_margins = c(1, 0.1, 0.7, 0.1)
+  )
+
+  # and plot it
+  if(!no_legend) {
[…]

### 2024-05-14 R/table-as-figure.R

@@ -41,6 +41,15 @@ table_as_figure <- function(
     plot_margins = c(0, 0, 1.5, 0)
 ) {
 
+
+  # catch whether rosa_genos are in here or not
+  HasRosa <- FALSE
+  if("rosa_geno_f" %in% names(X)) HasRosa <- TRUE
+  fill_alpha <- 1.0
+  if(HasRosa) {
+    fill_alpha <- 1.0  # ultilmately I don't find the transparency works. I just use black for the text
+  }
+
   # first, expand X to include the xmin/max and ymin/ymax for each internal cell
   N <- length(levels(X$row_label))
   X2 <- X %>%
@@ -95,7 +104,7 @@ table_as_figure <- function(
     filter(val != max(val))
 
   # then put these together into a big ggplot
-  full_plot <- ggplot() +
+  all_but_cell_labels <- ggplot() +
     geom_rect(  # outer edge filled rectangles
       data = Z,
       mapping = aes(
@@ -124,17 +133,8 @@ table_as_figure <- function(
         ymin = ymin,
         ymax = ymax,
         fill = cell_fill
-      )
-    ) +
-    geom_text( # interior cell labels
-      data = X2,
-      mapping = aes(
-        x = (xmin + xmax) / 2,
-        y = (ymin + ymax) / 2,
-        label = cell_label
       ),
-      vjust = 0.5,
-      hjust = 0.5
+      alpha = fill_alpha
     ) +
     geom_segment(  # get thick lines around the interior perimeter
       data = perim_thick,
@@ -153,6 +153,36 @@ table_as_figure <- function(
       plot.margin = unit(plot_margins, "lines")
     )
 
+
+  if(!HasRosa) {  # Standard plot without breaking it out by rosa_geno
+    full_plot <- all_but_cell_labels +
+      geom_text( # interior cell labels
+        data = X2,
+        mapping = aes(
+          x = (xmin + xmax) / 2,
+          y = (ymin + ymax) / 2,
+          label = cell_label
+        ),
+        vjust = 0.5,
+        hjust = 0.5
+      )
+  } else {  # otherwise we are breaking it out by rosa_geno
+    source("R/colors.R")
+    full_plot <- all_but_cell_labels +
+      geom_text( # interior cell labels
+        data = X2 %>% mutate(rosa_int = as.integer(rosa_geno_f) - 1L, rosa_hjust = rosa_int / 2),
+        mapping = aes(
[…]

### 2024-05-15 002-GSI_and_Fst.Rmd
Make zeros blank in assign tables.
Follow up on ERF and RRF hets.

@@
   input_list <- list(
-    #final_baseline = "data/cali-chinook-baseline.rds",
     final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
@@
 ```{r}
-# get just the 3 LFAR markers we are using:
+# get just the 2 LFAR markers we are using.  We are dropping 828619 because
+# it really doesn't do much, and so many reads are off-target.
 lfar_columns <- paste(
@@
     "NC_037130.1:1062935-1063235",
-    "NC_037130.1:828619-828919",
     "NC_037130.1:864908-865208"
   ), each = 2),
-  rep(c(1,2), 3),
+  rep(c(1,2), 2),
   sep = "_"
@@
 
-### Get some raw numbers
+## Get some raw numbers
 
@@
 
+## Make a Plot for scaled likelihood > 0.8
+
 Make a plot for > 80%
@@
 ```
+
+## Make a plot with numbers broken out by the RoSA genotype
 
@@
 
-## Identifying fish erroneously included in some collections
+## Follow up on those Early alleles found in RRF and ERF
 
-The occurrence of winter-run-assigned fish amongst the spring run collections is
-a clear indicator that there are some fish in the collections that do not belong
-with one another.  The same occurs amongst the fall and the spring run, I expect.
+The analysis shows 4 fish from RRF and ERF with heterozygous genoypes (EL).
+This has not been previously reported, so we really want to check up on those.
+It might be that they had lots of missing data.
 
-Our goal here is to examine the likelihoods with which fish are assigned to
-different reporting units and come up with a common-sense and defensible approach
-for discarding individuals that don't belong in the collections they were sampled
-from.
-
-We start by tallying up the misassigned fish---those that have a repunit and an
-inferred repunit that differ:
+So, I want to grab their RoSA genotype strings so that we can verify.
 ```{r}
-misassigned_ones <- lfar_top_ass$top %>%
-  filter(repunit != inferred_repunit) %>%
-  count(repunit, inferred_repunit)
-
-misassigned_ones
+# here are the IDs fo the hets from those coastal pops
+coastal_hets <- lfar_top_ass$top %>%
+  left_join(rosa_genos, by = join_by(indiv)) %>%
+  filter(collection %in% c("ERF", "RRF")) %>%
+  select(indiv, rosa_geno) %>%
+  filter(rosa_geno %in% c("EE", "EL"))
+coastal_hets
 ```
 
-For each of these 6 pairs, I want to make a scatterplot that shows the repunit
-likelihoods of all the fish _from either the repunit or the inferred_repunit_.
-The x-axis will be the likelihood for the repunit and the y-axis will be the
-likelihood for the inferred_repunit.  So, there is some weird prep to get
-this all together.  
+Now, let's get the raw genotype strings (the "hapstrs") off those guys:
 ```{r}
-MO2 <- misassigned_ones %>%
-  rename(
-    Foc_repunit = repunit,
-    Foc_inferred_repunit = inferred_repunit
-  )
-
-pick_em <- lapply(1:nrow(MO2), function(i) {
-
-  FR <- MO2$Foc_repunit[i]
-  FIR <- MO2$Foc_inferred_repunit[i]
-
-  lfar_top_ass$repu_likes %>%
-    ungroup() %>%
-    filter(
-      (repunit == FR & inferred_repunit == FIR) |
-        (inferred_repunit == FR & repunit == FIR)
-    ) %>%
-    mutate(
-      Foc_repunit = FR,
-      Foc_inferred_repunit = FIR,
-      .before = indiv
-    ) 
-}) %>%
-    bind_rows()
+rosa_hapstrs %>%
+  semi_join(coastal_hets, by = join_by(indiv))
 ```
 
-OLD STUFF BELOW HERE
-```{r}
-tmp <- lapply(1:nrow(MO2), function(i) {
-  lfar_top_ass$repu_likes %>%
-    filter(
-      (repunit == MO2$Foc_repunit[i] | repunit == MO2$Foc_inferred_repunit[i]) &
-        (inferred_repunit == MO2$Foc_repunit[i] | inferred_repunit == MO2$Foc_inferred_repunit[i])
-    ) %>%
-    mutate(
-      Foc_repunit =  MO2$Foc_repunit[i],
-      Foc_inferred_repunit = MO2$Foc_inferred_repunit[i],
-      .before = repunit
-    )
-}) %>%
-  bind_rows() %>%
-  arrange(indiv, Foc_repunit, Foc_inferred_repunit)
-
-for_plots <- tmp %>%
-  mutate(
-    rep = ifelse(Foc_repunit == repunit, "rep", "inf"),
-    inf = ifelse(Foc_inferred_repunit == inferred_repunit, "inf", "rep")
-  ) %>%
-  ungroup() %>%
-  #count(Foc_repunit, Foc_inferred_repunit, repunit, inferred_repunit, rep, inf)
-  select(-repunit, -inferred_repunit) %>%
-  pivot_wider(names_from = c(rep, inf), names_sep = "_to_", values_from = repu_sclike) %>%
-  mutate(condition = str_c(Foc_repunit, " misassigned to ", Foc_inferred_repunit), .after = Foc_inferred_repunit)
-```
-
-Now we should be able to facet wrap those:
-```{r}
-ggplot(for_plots) +
-  geom_point(aes(x = inf_to_rep, y = inf_to_inf, colour = Foc_inferred_repunit)) +
-  geom_point(aes(x = rep_to_rep, y = rep_to_inf, colour = Foc_repunit)) +
-  facet_wrap(~condition) +
-  scale_colour_manual(values = repunit_colors)
-
-```
-
-So, for each of the six plots, I will need four different repu_sclike's:
-For example, for the `repunit = CV-Fall` and `inferred_repunit = CV-Spring` case
-we will need to have the repu_sclikes for CV-Fall fish being calculated for CV-Fall
-and CV-Spring, and for the CV-Spring fish being calculated for CV-Spring and CV-Fall.
-
-Now, we want to use this to semi-join to get all the likelihoods for these scenarios,
-whether or not they were misassigned. 
-```{r}
-focal_comparisons <- repu_likes %>%
-  semi_join(misassigned_ones, by = join_by(repunit, inferred_repunit))
-```
-
-And we are going to want to compare these to all the likelihoods of being
-from the actual repunit that the fish was col
-
-Now, I want to look at these different comparisons _pairwise_, so we will
-just pivot it wider and then look at different columns in there.  This isn't
-super "tidy" but I think it should work OK:
-```{r}
-focal_comparisons %>%
-  pivot_wider(names_from = inferred_repunit, values_from = repu_sclike)
-```
-
-And I want to look at the observed distribution of these scores in for each
-comparison, colored by repunit.
-```{r}
-ggplot(focal_comparisons, aes(x = ))
-```
-
-First thing I am going to do is plot histograms of the,
-max likelihoods for all of the ones from each repunit.
-```{r}
-repunit_order <- unique(pop_labels$repunit)
-repu_likes %>%
-  mutate(
-    repunit_f = factor(repunit, levels = repunit_order),
-    inferred_repunit_f = factor(inferred_repunit, levels = repunit_order),
-  ) %>%
-  ggplot() +
-  geom_histogram(aes(x = repu_sclike, fill = repunit_f), colour = "black", linewidth = 0.16) + 
-  facet_grid(repunit_f ~ inferred_repunit_f) + 
-  scale_fill_manual(values = repunit_colors) +
-  theme_bw()
-```
-
-```{r}
-ta_corr_lfar <- top_ass_w_lfar %>%
-  filter(repunit == inferred_repunit)
-ta_wrong_lfar <- top_ass_w_lfar %>%
-  filter(repunit != inferred_repunit) %>%
-  arrange(collection, repu_sclike) %>%
-  group_by(collection) %>%
-  mutate(y = 190 + (70/9) * 1:n())
-
-ggplot() +
-  geom_histogram(
-    data = ta_corr_lfar,
-    mapping = aes(x = repu_sclike, fill = inferred_repunit)
-  ) +
-  geom_point(
-    data = ta_wrong_lfar,
-    mapping = aes(x = repu_sclike, y = y, fill = inferred_repunit),
-    size = 3,
-    shape = 21
-  ) +
-  facet_wrap(~coll_f, ncol = 3) +
-  scale_fill_manual(values = repunit_colors)
-```
-
-Filter out the ones that look like clear mis-labelling and then
-count up the mis-assigned fish:
-```{r}
-tafilt_w_lfar <- top_ass_w_lfar %>%
-filter(!(repunit != inferred_repunit & repu_sclike > 0.99))
-```
-
-Here we calculate the number of correctly and incorrectly assigned fish:
-```{r}
-tafilt_w_lfar %>%
-  mutate(correct = repunit == inferred_repunit) %>%
-  count(correct) %>%
-  mutate(fract = n / sum(n))
-```
-
-Having the LFAR in there improves things. Good.
-
-Make a table of those:
-```{r}
-
-X2_lfar <- tafilt_w_lfar %>%
-mutate(
-collection_f = factor(collection, levels = collection_order),
-inferred_coll_f = factor(inferred_collection, levels = collection_order)
-) %>%
-count(collection_f, inferred_coll_f, .drop = FALSE) %>%
-taf_prep()
-
-Rubias_for_later_lfar <- X2_lfar # keeping this for later...
-
-# get the result
-TAF2_lfar <- table_as_figure(
-X = X2_lfar,
-RC_groups = RC_groups,
-external_colors = run_time_colors,
-internal_colors = repunit_colors,
-plot_margins = c(1, 0.1, 0.7, 0.1)
-)
-
-# and plot it
-g_lfar <- plot_grid(
-TAF2_lfar$full_plot,
-plot_grid(
-get_legend(TAF2_lfar$for_external_legend),
-get_legend(TAF2_lfar$for_internal_legend),
-nrow = 1
-),
-nrow = 2,
-rel_heights = c(7,3)
-)
-
-g_lfar
-```
-
-Adding the LFAR definitely improves the LFAR assignment, though
-it still does not get it perfect.
-```{r}
-ggsave(g_lfar, filename = output_list$assig_table_lfar_all, width = 10, height = 12)
-```
+Those are definitely hets in the left half of the markers.  One of them is also
+hets at the right half.  The three remaining ones are L homozygotes on the second
+half.  So, there was been some recombination within these.
 
@@
 ```{r}
-fbl_no_miss <- full_base2 %>%
+fbl_no_miss <- base_with_lfar %>%
   select(-starts_with("NC_037130.1")) %>%  # drop the LFARs

### 2024-05-15 tex/body-text-of-paper.tex
Make zeros blank in assign tables.
Follow up on ERF and RRF hets.

@@
 The existing SNPtype assays and the additional microhaplotype markers were
-developed in the absence of a well-assembled genome for Chinook salmon
-\comment{(Is this true?)}. As a consequence, many of these markers had
+developed prior to the advent of a well-assembled genome for Chinook salmon.
+As a consequence, many of these markers had
 been used previously without certainty about their location in the genome.
@@
 per locus, standardized to the smallest sample size, ranged from 2.35 (SRW) to
-2.68 (RRF), with those two populations also showing the most extreme values for the
-standardized fraction of polymorphic markers at 0.9 and 0.98, respectively.
+2.68 (RRF), with  SRW also having the smallest value of the
+standardized fraction of polymorphic markers at 0.9 and RRF sharing the highest value
+(0.97) with ERF and BCkF.
 SRW also had the lowest values for expected and observed heterozygosity:
 0.349 and 0.340, respectively. While the highest for expected heterozygosity
-was 0.412 (SmRF) and for observed heterozygosity, 0.412 (SmRF).
+was 0.412 (SmRF) and for observed heterozygosity, 0.411 (CRHS).
 

### 2024-05-15 R/plot_self_assigment_table.R

@@ -11,7 +11,7 @@
 #' @return Returns a list.  Component "plot" of which is the ggplot of the assignment
 #' table, and component RFL is the Rubias_for_later output---basically
 #' the top assignments with some factors in there for getting the order right.
-plot_self_assignment_table <- function(top_ass, pop_labels, collection_order, no_legend = FALSE) {
+plot_self_assignment_table <- function(top_ass, pop_labels, collection_order, no_legend = FALSE, zeros_are_empty = TRUE) {
 
   # get the function
   source("R/taf-prep.R")
@@ -32,6 +32,13 @@ plot_self_assignment_table <- function(top_ass, pop_labels, collection_order, no
       ) %>%
       count(collection_f, inferred_coll_f, .drop = FALSE) %>%
       taf_prep()
+
+    # remove 0's if desired
+    if(zeros_are_empty == TRUE) {
+      X2 <- X2 %>%
+        mutate(cell_label = ifelse(cell_label == "0", "", cell_label))
+    }
+
   } else {
     X2 <- top_ass %>%
       mutate(
@@ -41,6 +48,13 @@ plot_self_assignment_table <- function(top_ass, pop_labels, collection_order, no
       ) %>%
       count(collection_f, inferred_coll_f, rosa_geno_f, .drop = FALSE) %>%
       taf_prep()
+
+    # here, we remove zeros only from the cells that are ALL zero
+    if(zeros_are_empty == TRUE) {
+      X2 <- X2 %>%
+        group_by(row_label, col_label) %>%
+        mutate(cell_label = ifelse(rep(all(cell_label == "0"), n()), "", cell_label))
+    }
   }
 
   Rubias_for_later <- X2 # keeping this for later...

### 2024-05-20 tex/body-text-of-paper.tex
Changes from overleaf

@@
 management of Pacific salmon, and salmon conservation and management have been
-behind a number of advances in molecular ecology,
+at the forefront of a number of advances in molecular ecology,
 from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
@@
 \citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
-Two broadly applicable innovations that have been
+Two broadly applicable techniques that have been
 actively fostered by the Pacific salmon research community are genetic
@@
 
- In the 1980's, electrophoretically
+ In the 1980s, electrophoretically
 detectable genetic variation, in the form of allozymes
@@
 main tributaries (the Sacramento River and the San Joaquin River) within
-California's Central Valley, the Klamath and Trinity River basins in Northern California, and
+California's Central Valley, and the two main tributaries (Klamath and Trinity Rivers) in the Klamath basin in Northern California, and
 3 rivers of the California coast (Russian, Eel, and Smith)
@@
 microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by Freebayes.
+that were considered multinucleotide or complex variants by Freebayes. The resulting genetic markers are he
 
@@
 grouped together into reporting units.  At every value of $K$ CLUMPAK discerned at least
-11 or 20 replicates in the major mode (Fig.~\ref{fig:struct}). Clustering solutions in the minor
+11 of 20 replicates in the major mode (Fig.~\ref{fig:struct}). Clustering solutions in the minor
 modes appear in Fig.~\ref{fig:minor-modes}.
@@
 reference baseline provides a high degree of accuracy for distinguishing the major
-groups of fish in California.  The results of the analysis are summarized
+groups of Chinook salmon in California.  The results of the analysis are summarized
 in a self-assignment matrix (Figure~\ref{fig:gsi}). 

### 2024-05-20 tex/body-text-of-paper.tex
small edits

--- 2024-05-15 tex/body-text-of-paper.tex
@@
 understanding of the genetic inheritance of key traits in salmonids
-\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of whole
+\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of a whole
 family of statistical genetic methods employing relationship inference to learn about
@@
 \begin{figure*}
-\newcommand{\fprcap}{\footnotesize False positive rates as a function of false negative
-rates for distinguishing between different relationships.  {\bf (a)} False positive rates calculated
+\newcommand{\fprcap}{\footnotesize False positive rates (FPRs) as a function of false negative
+rates (FNRs) for distinguishing between different relationships.  {\bf (a)} FPRs calculated
 using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),
-full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  False positive rates calculated using regular Monte Carlo
+full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  FPRs calculated using regular Monte Carlo
 sampling of avuncular pairs (AN: aunt-niece, uncle-nephew, etc.) or full-siblings (FS) or half-siblings (HS) being 
@@
 ($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.
-This also is the reason that many of the lines terminate   }
+This also is the reason that many of the lines terminate.
+}
 \begin{center}

### 2024-05-21 tex/body-text-of-paper.tex
add paragraph (overleaf)

--- 2024-05-20 tex/body-text-of-paper.tex
@@
 microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by Freebayes. The resulting genetic markers are he
+that were considered multinucleotide or complex variants by Freebayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
 
@@
 We undertook a further discovery process to develop additional multiallelic
-microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon.
+microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon. A total of 125 microhaplotype markers were developed for Chinook salmon 
 \comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
@@
 
-The existing SNPtype assays and the additional microhaplotype markers were
+The SNPlicons targets and the additional microhaplotype markers were
 developed prior to the advent of a well-assembled genome for Chinook salmon.
@@
 baselines tailored to specific regions are essential for regional management questions.
-It should be noted that recent
+
+Accurate identification of the distinct ecotypes of Chinook salmon in the California Central Valley is of utmost importance for designing and implementing effective conservation and management actions.
+This has been challenging, given the recent common ancestors of these ecotypes and ongoing migration between subbasins.
+We describe the first set of genetic markers that identifies the late-fall-run salmon ecotype, which occurs only in the CCV and shares a genomic background with the more common CCV fall-run salmon ecotype.
+Moreover, we show how the FRH spring-run lineage is easily identifiable through a combination of traditional GSI and the characterization of functional genetic markers in the RoSA.
+Finally, although previous work has described GSI capabilities that distinguish the natural-spawning CCV spring-run lineages from each other and their fall-run counterparts with moderately high accuracy, we demonstrate near complete accuracy in distinguishing these 'stocks'.
+Moreover, the few fish that are apparently misidentified (Fig.
+3) are likely to be primarily migrants and not true misidentifications.
+For example, the three fish that were field-characterized as spring-run from Butte Creek, but are genetically identified as winter-run salmon are clearly carry the winter-run genomic background and could not realistically be misidentified on the basis of the genetic data.
+This marker set and baseline reference dataset also provides unprecedented power for identifying fish from not only the other ESUs in California, but even the individual populations within those ESUs (Fig.
+3).
+Additionally, previous genetic marker sets for California Chinook salmon had sufficient power for inferring parent-offspring relationships, but only when both parents were sampled and genotyped.
+This marker set considerably increases the capacity for relationship inference in California salmon, by providing sufficient power for parent-offspring relationship inference, when only one parent was sampled, as well as identifying pairs of full siblings when no parents are sampled.
+Groups of full siblings larger than two are identified even more accurately than expressed in Figure 4.

### 2024-05-28 tex/body-text-of-paper.tex
More overleaf changes

--- 2024-05-21 tex/body-text-of-paper.tex
@@
 revealed that these allozyme markers
-exhibited different allele frequencies among major stocks of Chinook salmon on the West Coast.
+exhibited different allele frequencies among major lineages, or stocks, of Chinook salmon on the West Coast.
 These allele frequency differences make GSI possible, and
@@
 
-The ascendance over the last decade of parentage based tagging (PBT) as a management
+The emergence over the last decade of parentage based tagging (PBT) as a management
 tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
@@
 utility for PBT (or GSI).  PBT has been remarkably successful in fisheries
-management, having largely replaced coded wire tags in large parts of the Snake River and
-Columbia River basin (CITATION), and having been employed to dramatically further our
+management, having been used for over a decade in the management of Chinook salmon
+and steelhead trout in major basins of the Columbia River \cite{steele2019parentage,horn2023multigeneration}, and having been employed to dramatically further our
 understanding of the genetic inheritance of key traits in salmonids
-\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of whole
+\citep{abadia2013large,beulke2023distinct}. However, PBT is just one subset of a whole
 family of statistical genetic methods employing relationship inference to learn about
@@
 
+Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important fisheries.
+Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities during many different seasons and geographic locations.
+However, recent declines in population sizes, from the Yukon River in Arctic north, to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect less productive stocks.
+The co-occurence of fish from relatively productive and relict populations reaches its paragon in California, with the largest remaining ocean fisheries for Chinook salmon targeting the Central Valley fall-run stock that spawns in different tributaries of the same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks.
+In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are very closely related \cite{clemento2014evaluation} and share recent common ancestry that is independent of ongoing migration between these lineages.
+As such, GSI has been particularly challenging in the CCV, with at least one of the distinct ecotypes, late-fall-run, unresolvable with previous GSI marker sets \cite{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the harvested fall-run stock.
+
 In the following we present a new reference baseline for both GSI and PBT in California.
@@
 within Oregon.
+This set of populations and stocks includes all of the previously described lineages of Chinook salmon in the southern portion of the range and most of the populations that are known to have significant genetic differentiation in this region.
 These locations included sites in the two
@@
 
+Although previous work using genome-scale data has shown that many of these populations or stocks can be resolved \cite{thompson2020complex,meek2020identifying},
+such data is not practical for most large-scale applications of GSI and PBT, particularly because
+of the much higher costs, longer turnaround time and challenges of producing highly
+replicable datasets with reduced representation and low-coverage whole genome sequencing methods \cite{ali2016rad}. 
 Accurate identification of the distinct ecotypes of Chinook salmon in the California Central Valley is of utmost importance for designing and implementing effective conservation and management actions.

### 2024-05-31 002-GSI_and_Fst.Rmd
Out locus summaries by collection

@@
     rosa_gsi_table_crop = "results/GSI_and_Fst/rosa-gsi-table-crop.pdf",
-    rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf"
+    rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf",
+    pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv"
   )
@@
 
+While we are at it, let's look at the distribution of the E2 alleles across
+collections.
+```{r}
+# count up genos at each variant
+marks <- paste("s", 1:12, sep = "")
+exploded <- rosa_hapstrs %>%
+  filter(!(collection %in% c("ColeRHsp", "SmithFall"))) %>%
+  separate(hapstr, into = paste("s", 0:12, sep = ""), sep = "") %>%
+  select(-s0)
+rh_counts <-  exploded %>%
+  pivot_longer(cols = starts_with("s"), names_to = "variant", values_to = "geno") %>%
+  mutate(variant_f = factor(variant, levels = marks)) %>%
+  count(repunit, collection, variant_f, geno)
+
+# look just at s2, s3, and s6
+rh_counts %>%
+  filter(variant_f %in% c("s2", "s3", "s6")) %>%
+  filter(geno %in% c("H", "W"))
+```
+
+So, there is only one more E2-het outside of the Sacto, and that is from TRHsp.
+
+Let's go back and find that:
+```{r}
+exploded %>% 
+  filter(repunit == "klamathspring" & s6 == "H") %>%
+  left_join(rosa_hapstrs, by = join_by(indiv))
+
+```
+
 # Calculating Fst between the collections from the microhaplotype markers
@@
 # calculate the mean number of alleles.
+set.seed(1)
 sub_sampled_locus_num_alle <- fbl_no_miss %>%
@@
 
+## Locus specific pop-gen summaries
+
+Now, we also are going to make summaries for each locus in each collection.
+Carlos wanted a supplemental data table that has Hexp and Hobs for each
+collection.  I figure we might as well add the number of diploids and the
+number of alleles in there as well.   All of that information is in
+`locus_summaries` in a tidy format, but we will make it wide for this.
+Carlos wanted to add it to the Supp-Data-1-Amplicon-info.xlsx, but I think
+that is actually a little crazy because we are going to have so many columns in this
+output. So, I will put it into a CSV file.
+```{r}
+by_locus_table <- locus_summaries %>%
+  select(locus, collection, num_diploids, num_alleles, exp_Hz, obs_Hz) %>%
+  pivot_wider(
+    names_from = collection,
+    values_from = num_diploids:obs_Hz,
+    names_sep = "."
+  )
+
+write_csv(by_locus_table, file = output_list$pop_gen_by_loc_coll)
+```
+
 # Finally, get the histogram of the total number of alleles per locus

### 2024-05-31 tex/body-text-of-paper.tex
changes made in overleaf

@@
 
-To assemble the markers for this reference baseline, we drew upon several different
-sources, as detailed in the separate sections below.
+We assembled genetic markers for this reference baseline dataset from several different
+efforts, as detailed in the separate sections below.
+
+\subsubsection*{Novel microhaplotype discovery}
+
+We undertook a process to discover and validate  multiallelic
+microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon, with a focus on Central Valley populations.
+
+We identified candidate loci with variation among multiple populations and with multiple single-nucleotide polymorphisms using reduced representation genome sequencing of 2--4 individuals each from a number of
+populations ranging from California to British 
+Columbia and the interior Columbia River (\comment{Table SX}).
+Genomic DNA was extracted using Qiagen DNeasy 96 Blood and Tissue Kits on a BioRobot 3000 
+(Qiagen, Inc). We used double-digest restriction-site associated DNA sequencing (ddRAD) and with different size selections (400 bp and 500 bp 
+fragments) to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
+\citet{peterson2012double} with the modifications described in \citep{baetscher2018microhaplotypes}.
+Sequencing was performed on a MiSeq (Illumina, Inc) using $2 \times 300$-cycle paired-end sequencing.
+
+Stacks v1.45 \cite{catchen2013stacks} was used to analyze ddRAD sequence data.
+As there was not a well assembled Chinook salmon genome at the time, a {\em de novo} assembly was performed.
+Initial filtering retained RAD loci with two or more SNPs that
+amplified in at least 10 samples. After this initial filtering, there were 3,931 RAD loci retained in the ddRAD data with 400 bp size selection and
+4,794 retained in the data with 500 bp size selection. 
+Gene regions were then selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation in the winter-run 
+population and at least one other.
+A 100 bp window was used because the success of multiplex PCR and MiSeq sequencing is diminished in amplicons larger than 100 bp. Primer3 
+software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.  These gene regions were amplified and variants
+
+FreeBayes output was filtered to remove indels and retain variants with a Quality score of 30 or greater and a read depth of 10 or higher.
+Further filtering and analysis of the loci was conducted in MICROHAPLOT (Ng et al., DOI: 10.5281/zenodo.820110).
+Loci that did not conform to Mendelian segregation, that is, loci that produced more than two haplotypes within an individual were removed.
+Hardy-Weinberg equilibrium was assessed using MICROHAPLOT.
 
@@
 that were considered multinucleotide or complex variants by Freebayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
-
-\subsubsection*{Additional microhaplotype discovery}
-
-We undertook a further discovery process to develop additional multiallelic
-microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon. A total of 125 microhaplotype markers were developed for Chinook salmon 
-\comment{NEED TO DESCRIBE. How did we look for these, etc. What constituted a
-candidate to then follow up on with amplicon sequencing, etc.}
 
@@
 We include in the reference baseline amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish described in \citet{thompson2020complex}.  These are useful markers to
+between spring-run and fall-run fish on Chromosome Ots28 described in \citet{thompson2020complex}.  These are useful markers to
 have in the panel, as typing them routinely will create a data base for understanding the spatial and temporal
@@
 influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.
+
+\subsection*{Amplicon sequencing and microhaplotype calling}
+
+Amplicon sequencing was performed with the Genotyping-in-Thousands by sequencing (GT-seq) method from Campbell et al. (2015). All loci were 
+multiplexed with primer concentrations ranging from 0.083 uM to 0.5 uM to increase uniformity of read depths across all loci. Sample normalization was 
+performed using the SequalPrep Normalization Plate Kit (Applied Biosystems) according to the manufacturer's instructions. Samples were pooled by plate 
+and then purified using Agencourt AMPure XP magnetic beads. Libraries were quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New 
+England BioLabs) before dilution and pooling for sequencing. All samples were sequenced on a MiSeq using $2\times 75$-cycle paired-end sequencing protocols. 
+Primer sequences and concentrations for making the primer pool are in Supplemental Table 1.
+
+For evaluation of candidate loci, dual-barcoded sequences were used to identify individual tissue samples using the MiSeq analysis software (Illumina).
+The paired-end sequence reads were joined together using Fast Length Adjustment of SHort reads (FLASH: Magoc and Salzberg 2011) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
+Mapped reads were converted to BAM files using SAMtools (Li et al.
+2009) and FreeBayes (Garrison and Marth 2012) was used to call variants.
+
+For genotyping of project samples, final filtering at each locus required a read depth of 20 and a read depth ratio of 0.30 (i.e., for heterozygotes the lower read depth haplotype must have 30\% of the total reads at that locus).
+MICROHAPLOT was used to call alleles and genotype processing was performed with R scripts.
+
+Before parentage analysis began each locus was assessed for the presence of null alleles using R scripts (eric citation).
+Twelve loci appeared to have null alleles in the winter-run population and were removed in all subsequent analyses.
+Additionally, two loci were monomorphic in the winter-run, but contained multiple alleles in other populations.
+The two monomorphic loci were removed in all winter-run analyses resulting in a final dataset of genotypes for 111 microhaplotype loci for this project.
+For more diverse salmon populations, it's likely that nearly all of the 125 microhaplotype loci could be used without Hardy-Weinberg Equilibirum or null allele issues.
 

### 2024-05-31 ckmr-sim-evaluation.Rmd
Add comparison of CKMR logls with all SNPs vs...
just the top SNPs from each amplicon.

@@
     pop_labels = "inputs/reference-collection-names.csv",
-    locus_info = "inputs/ChinookFullPanelLociTable_Feb2024Updated.xlsx"
+    locus_info = "inputs/ChinookFullPanelLociTable_May2024Updated.xlsx"
   )
@@
     ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
-    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf"
+    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
+    ckmr_comp = "results/ckmr-sim-eval/cmkr-comp-figure.pdf",
+    ckmr_comp_crop = "results/ckmr-sim-eval/cmkr-comp-figure-crop.pdf",
+    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf"
+
   )
@@
 ```
+
+# Logl Ratio Plot Comparisons
+
+Neil suggested, and Carlos and I agreed, that one way to see they improvement in
+relationship inference obtained by using multiallelic microhaps is to look at the
+separation of the different relationship log-likelihood ratios when using just the
+best (MAF closest to 0.5) SNP from each locus, or when using all the SNPs/alleles.
+We deciced to do this for the hatchery populations in California, as well as the
+winter run.  I am going to implement this in a few functions, each of which operate
+on the one of the ckmr_sim_nests.
+
+```{r}
+#' function to get just the best SNP out of each microhap
+#' 
+#' This works on a single collection at a time.
+#' @param X a tibble like that nested in a single row of ckmr_sim_nests
+#' @return Returns a tibble like X, but with just the best SNP from
+#' each amplicon (note that it might also drop some monomorphic amplicons).
+retain_top_snp <- function(X) {
+  # first, get the microhaps in long format
+  tmp <- X %>%
+    select(-(repunit:sample_type)) %>%
+    pivot_longer(
+      -indiv,
+      names_to = c("Locus", "gene_copy"),
+      values_to = "Allele",
+      names_pattern = "^(.*)_([12])$"
+    ) %>%
+    filter(!is.na(Allele))
+
+  # then explode it out to individual SNPs
+  exploded <- tmp %>%
+    mutate(
+      snp_num = map(Allele, function(x) 1:nchar(x)),
+      snp_alle = map(Allele, function(x) strsplit(x, "")[[1]])
+    ) %>%
+    unnest(cols = c(snp_num, snp_alle))
+
+  # now, get the highest maf snp at each locus in a tibble
+  highest_mafs <- exploded %>%
+    count(Locus, snp_num, snp_alle) %>%
+    group_by(Locus, snp_num) %>%
+    mutate(af = n / sum(n)) %>%
+    filter(af <= 0.5) %>%
+    ungroup() %>%
+    arrange(Locus, desc(af)) %>%
+    group_by(Locus) %>%
+    slice(1) %>%
+    ungroup()
+
+  # then pull those out of the "exploded" variable with a semi_join
+  # and put them back into wide format, left join it onto the meta
+  # data
+  ret <- exploded %>%
+    semi_join(highest_mafs, by = join_by(Locus, snp_num)) %>%
+    select(indiv, Locus, gene_copy, snp_alle) %>%
+    pivot_wider(
+      names_from = c(Locus, gene_copy),
+      values_from = snp_alle,
+      names_sep = "_"
+    ) %>%
+    left_join(
+      X %>% select(indiv, repunit, collection, sample_type),
+      .,
+      by = join_by(indiv)
+    )
+
+  ret
+
+}
+
+```
+
+So, that is how we pick out the top SNP.  But now we also need a function that is
+going to give us the set of all the simulated logl-ratios that we need (whether it
+is all the microhap alleles or just the top SNPs).  We can build off of what
+we did in a function above.
+```{r}
+get_logl_rats <- function(X, GC2 = gen_coords) {
+
+  Y <- ckmr_afreqs_ready(X, GC2)
+
+  C <- ckmr_create(Y)
+
+  # throttle text output to console
+  options(CKMRsim.discard_stdout = TRUE)
+  options(CKMRsim.linkage_verbosity = 0)
+
+  # Simulate some Q_ij's. Suppress output messages that the CKMRsim package
+  # produces
+  Q_link <- ckmr_simQ(C) #%>% suppressMessages()
+
+  # now extract the logl_ratios for PO vs U and return those
+  extract_logls(Q_link, numer = c(PO = 1), denom = c(U = 1))
+}
+```
+
+And, with those two functions, it would appear that we have eveything we need to get the
+simulation results we will want.  Once again we will mclapply over the populations.
+
+```{r}
+# collections / groups we want to do
+pops <- c("ck-FRH", "ck-IGH", "ck-SRW", "ck-TRH")
+
+# let's put the microhaps and the best SNPs together in a long format.
+input_tibbles <- bind_rows(
+  ckmr_sim_nests %>%
+    filter(ckmr_sim_group %in% pops) %>%
+    mutate(condition = "microhaps", .before = ckmr_sim_group),
+  ckmr_sim_nests %>%
+    filter(ckmr_sim_group %in% pops) %>%
+    mutate(tops = map(data, retain_top_snp)) %>%
+    mutate(data = tops) %>%
+    select(-tops) %>%
+    mutate(condition = "top_snps", .before = ckmr_sim_group)
+)
+
+Qsims <- mclapply(input_tibbles$data, get_logl_rats, mc.cores = 8)
+
+AllSims <- tibble(
+  condition = input_tibbles$condition,
+  ckmr_sim_group = input_tibbles$ckmr_sim_group,
+  Qsims = Qsims
+) %>% 
+  unnest(Qsims) %>%
+  select(condition, ckmr_sim_group, true_relat, logl_ratio)
+```
+
+Now, we can plot all of those.  Let's see if we can get all four onto one page easily enough.
+```{r}
+asg <- ggplot(AllSims, aes(x = logl_ratio, fill = true_relat)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(~ ckmr_sim_group + condition, ncol = 1) +
+  theme_bw() +
+  labs(fill = "True\nRelationship") +
+  xlab("PO vs U Log-likelihood Ratio")
+
+# this fails for some reason (ggsave never returns)
+# ggsave(asg, path = "boing.png", height = 15, width = 9)
+
+# So, instead I just print the thing in the Rstudio window and export it
+# at size 13" high by 10" wide. Then I moved it to where it belongs in the tex directory.
+asg
+```

### 2024-05-31 tex/supplemental-figures.tex
Add comparison of CKMR logls with all SNPs vs...
just the top SNPs from each amplicon.

@@
 %%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{images/cmkr-comp-figure-crop.pdf}
+\end{center}
+\caption[Comparison of microhaplotypes vs.~best SNP in each amplicon]{\footnotesize 
+A comparison of the distribution of log-likelihood ratios when using all the alleles at each
+amplicon typed as microhaplotypes versus using just a single (the most informative) SNP from each
+amplicon.  Results shown for four hatchery collections in California.  The density plots show
+the distribution of log-likelihood ratios for Parent-Offspring vs Unrelated for four different relationships:
+PO = parent offspring; FS = full-sibling; HS = Half sibling; AN = avuncular (i.e., aunt-niece, etc.)
+The distributions for AN and HS overlap completely.  Note that the overlap between FS and PO occurs
+for the PO vs U likelihood ratio, but is nearly eliminated with the PO vs FS likelihood ratio, allowing these
+two relationships to be resolved accurately.}
+\label{fig:ckmr-comp}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%

### 2024-05-31 tex/supplemental-figures.tex
Update caption in supplement

@@
 A comparison of the distribution of log-likelihood ratios when using all the alleles at each
-amplicon typed as microhaplotypes versus using just a single (the most informative) SNP from each
-amplicon.  Results shown for four hatchery collections in California.  The density plots show
+amplicon typed as microhaplotypes ("microhaps" in the panel headers) versus using just a single (the most informative) SNP from each
+amplicon ("top\_snps" in the panel headers).  Results shown for four hatchery collections in California 
+(ck-FRH: Feather River Hatchery [spring and fall]; ck-IGH: Iron Gate Hatchery; ck-SRW: Sacramento Winter Run;
+ck-TRH: Trinity River Hatchery [spring and fall]).  The density plots show
 the distribution of log-likelihood ratios for Parent-Offspring vs Unrelated for four different relationships:

### 2024-06-06 tex/body-text-of-paper.tex
Add comparison of CKMR logls with all SNPs vs...
just the top SNPs from each amplicon.
more from overleaf

@@
 spring-run and fall-run ecotypes (i.e.
-Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype (REF?).
-\comment{As such, sampling targeted carcasses found in each of these
-rivers at times representative of their run-timing designation}, which we were able to further confirm using their genotypes
-within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}.
-At the time of sampling, only fall-run Chinook were found in the San Joaquin River so the samples there were
+Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype (\comment{REF?}).
+As such, sampling targeted carcasses found in each of these
+rivers at times representative of their run-timing designation, which we were able to further confirm using their genotypes
+within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}. At the time of sampling (2016), only
+fall-run Chinook salmon adults were found in the San Joaquin River, so the samples there were
 easily categorized. Run-timing 
@@
 %%%%%%%%%%%%%%%%%%%%
+And we can also point readers to the supplement figure \ref{fig:ckmr-comp}.
 

### 2024-06-06 002-GSI_and_Fst.Rmd
Revamp rosa geno calling
Wanted to deal with recombinants differently and went to a more stringent definition of EE, EL, and LL

@@
 
-We are not going to delve on the different between E1 and E2 here.  That is for the
+We are not going to delve on the difference between E1 and E2 here.  That is for the
 subRoSA paper.  But we want to get the genotypes simply as EE, EL, LL as done in the
@@
 
-From that it is clear that there have been some recombinations in here, but we will
-just go ahead and call RoSA genotype by the majority of the non-missing ones. 
+From that it is clear that there have been some recombinations in here. But there is a bit of
+junk in here.  It looks to me like there are some sites that get N's and some that get ?'s.  Both
+of those are missing data.  I am going to be stringent here and toss anything with missing data,
+which is only about 150 of them, so that should be fine. It will make downstream processing
+much easier, since the number of categories will be a bit smaller. (And we can be quite confident
+in these assignments)
 ```{r}
-H8_strip <- H8 %>%
-  mutate(
-    stripped = map(hvec, function(x) x[!(x %in% c("?", "N"))]),
-    sglom = map_chr(stripped, function(x) paste(x, collapse = "", sep = ""))
-    )
-
-H8_strip %>%
-  count(sglom) %>%
-  arrange(desc(n))
-
+H8_nomiss <- H8 %>%
+  filter(str_count(hapstr8, pattern = "\\?|N") == 0)
 ```
 
-OK, we will call something and EL if it has two or more Hs (the idea being that a single
-H could be a genotyping error). Note that there are no haplotypes mixing L's and E's together.
-Here is a function that will determine the RoSA genotype
-from a vector of H, L, and E parts. 
+How many were filtered out there?
 ```{r}
-call_rosa <- function(x) {
-  # things that were all ?'s get NA
-  if(length(x) == 0) {
-    return(NA_character_)
-  }
-
-  nE <- sum(x == "E")
-  nL <- sum(x == "L")
-  nH <- sum(x == "H")
-
-  if(nH >= 2) { 
-    return("EL")
-  } else if(nL > 0 && nE == 0) {
-    return("LL")
-  } else if(nE > 0 && nL == 0) {
-    return("EE")
-  } else {
-    return("Weird")
-  }
-
-}
-
-H8_genos <- H8_strip %>%
-  mutate(rosa_geno = map_chr(stripped, call_rosa))
+nrow(H8) - nrow(H8_nomiss)
 ```
 
-Count these up so we can see what we called LL, EL, and EE:
+Let's look at them:
+```{r, df.print = 20}
+H8_nomiss %>%
+  count(hapstr8) %>%
+  arrange(desc(n))
+```
+
+This is interesting.  I am struck by the 12 instances of EEHEEEEE---that suggests to me that
+it is not simply the result of genotyping error.  I suspect that one of the haplotypes
+has an L at that position.  The cleanest thing here might be to simply go for the
+things that are perfectly pure across all the sites, cuz there is a much lower chance
+of the region being broken up by recombination or gene conversion there.  So, that is what
+I am going to do.  And we can look at the ones that are not monolithic (possible recombinants, etc.
+as well.)
+
 ```{r}
-H8_genos %>%
-  count(sglom, rosa_geno) %>%
-  arrange(desc(n))
+H8_monolith <- H8_nomiss %>%
+  filter(hapstr8 %in% c("LLLLLLLL", "EEEEEEEE", "HHHHHHHH"))
+
+H8_recombs <- H8_nomiss %>%
+  filter(!(hapstr8 %in% c("LLLLLLLL", "EEEEEEEE", "HHHHHHHH")))
+
+# call genos from H8_monolith
+H8_genos <- H8_monolith %>%
+  mutate(
+    rosa_geno = recode(
+      hapstr8,
+      LLLLLLLL = "LL",
+      EEEEEEEE = "EE",
+      HHHHHHHH = "EL"
+    )
+  )
 ```

### 2024-06-06 002-GSI_and_Fst.Rmd
Add summary of "recombinant" haplotypes
Remove spurious fixed allele diffs in scatters
Just an issue with calling NAs zero.  Fixed it.

@@
 
+## Let's do a quick summary of the apparent recombinants
+
+```{r, rows.print=30}
+H8_recombs %>%
+  left_join(
+    full_base2 %>% select(indiv, repunit, collection),
+    by = join_by(indiv)
+  ) %>% 
+  arrange(hapstr8)
+```
+
 # Do the self-assignment
@@
 
-## Follow up on those Early alleles found in RRF and ERF
-
-The analysis shows 4 fish from RRF and ERF with heterozygous genoypes (EL).
-This has not been previously reported, so we really want to check up on those.
-It might be that they had lots of missing data.
-
-So, I want to grab their RoSA genotype strings so that we can verify.
-```{r}
-# here are the IDs fo the hets from those coastal pops
-coastal_hets <- lfar_top_ass$top %>%
-  left_join(rosa_genos, by = join_by(indiv)) %>%
-  filter(collection %in% c("ERF", "RRF")) %>%
-  select(indiv, rosa_geno) %>%
-  filter(rosa_geno %in% c("EE", "EL"))
-coastal_hets
-```
-
-Now, let's get the raw genotype strings (the "hapstrs") off those guys:
-```{r}
-rosa_hapstrs %>%
-  semi_join(coastal_hets, by = join_by(indiv))
-```
-
-Those are definitely hets in the left half of the markers.  One of them is also
-hets at the right half.  The three remaining ones are L homozygotes on the second
-half.  So, there was been some recombination within these.
+## Look at the distribution of the E2 alleles across collections
 
@@
 inset <- 0.15
-pair_freqs <- backbone %>%
+pf_init <- backbone %>%
   left_join(alle_freqs, by = join_by(row_coll == collection, locus, allele)) %>%
@@
   ) %>%
-filter(as.integer(col_coll) <= as.integer(row_coll)) %>%
-replace_na(list(row_freq = 0.0, col_freq = 0.0)) %>%
+  filter(as.integer(col_coll) <= as.integer(row_coll))
+
+# now, here is a somewhat more complex step.  The problem is that some collections have
+# no observations of any loci, it seems.  In those cases, we want the NA values to stand.
+# Those cases can be recognized because every allele at a locus in them is NA. In cases,
+# where not all of the alleles have NA for frequency, we do want to set the NAs to 0.
+pf0s <- pf_init %>%
+  group_by(row_coll, col_coll, locus) %>%
+  mutate(
+    allNA_rowfreq = all(is.na(row_freq)),
+    allNA_colfreq = all(is.na(col_freq)),
+  ) %>%
+  ungroup() %>%
+  mutate(
+    row_freq = ifelse(is.na(row_freq) & !allNA_rowfreq, 0.0, row_freq),
+    col_freq = ifelse(is.na(col_freq) & !allNA_colfreq, 0.0, col_freq)
+  )
+
+# finally, do some more setting up of things:
+pair_freqs <- pf0s %>% 
   mutate( # this is just setting things as in table_as_figure

### 2024-06-06 tex/body-text-of-paper.tex

@@
 
+We note that the RoSA markers provide the differential signal necessary to distinguish the
+Feather River spring-run stock from the other CCV stocks with a predominantly fall-run
+genomic background, as they do not share genome-wide ancestry with the other spring-run
+stocks/lineages in the basin. In this context, it is clear that the 'fall-run' hatchery
+program in the Feather River both spawns and produces many spring-run fish and RoSA heterozygotes.
+Similarly, in other regions, the RoSA markers are the only way to distinguish fall-
+and spring-run fish, as they otherwise share genomic background. 
+In California, only the Sacramento and  Klamath basins have documented spring-run
+salmon and their historical occurrence in other basins has been unclear. We find no
+instance of spring-run fish in these coastal basins, but did identify one copy of the early-lineage
+RoSA haplotype in the Eel River. Moreover, we identified three additional copies of a
+recombinant haplotype that carries half of the RoSA SNPs from both the E and L lineages in the Eel and Russian rivers.
+No other clear recombinants were identified in the study populations, indicated that it arose in the California Coastal Chinook Salmon lineage, and provides further evidence of past presence of RoSA haplotypes associated with early migration.
+
 \subsection*{Power for relationship inference}

### 2024-06-07 tex/abstract.tex
Updates from Overleaf

@@
 
-We need to write an abstract.
+Genetic methods have become an essential component of ecological investigation
+and conservation planning for fish and wildlife. Among these methods is the use
+of genetic marker data to identify individuals to
+populations, or stocks, of origin. More recently, methods that involve genetic
+pedigree reconstruction have also become common. We present, here, a novel set of
+multi-allelic microhaplotype genetic markers for Chinook salmon that have sufficient information
+content to accurately undertake methods that require identifications
+to populations and family groups. We show how this set of microhaplotypes
+provides definitive power to identify all known lineages of Chinook salmon in
+California. The inclusion of genetic loci that have known associations with phenotype
+and that were identified as outliers in examination of whole genome sequence data,
+allows resolution of stocks that are not genetically differentiated but
+are phenotypically distinct and managed as such. 
+This same set of multiallelic genetic markers have ample variation to
+accurately identify parent-offspring and full-sibling
+pairs in all California populations, including the genetically depauperate winter-run
+lineage.
+Validation of this marker panel in salmon populations not previously studied with
+modern genetic methods, also revealed some novel biological insights, including
+the presence of a single copy of a haplotype for a phenotype that has not
+been documented in that part of the species range, and a clear signal of mixed
+ancestry for a salmon population that is on the geographic margins of the 
+primary evolutionary lineages present in California.

### 2024-06-18 tex/body-text-of-paper.tex

@@
 
-\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}
+\subsubsection*{Markers potentially associated with phenotypes \citet{thompson2020complex}}
+
+Additionally, we include here genetic markers that characterize variation in two genes, VGLL3 and SIX6,
+that have been found to be associated with phenotypes related to reproductive maturity in both
+Atlantic \citep{barson2015sex} and Pacific \citep{waters2021heterogeneous} salmonids. We include
+them both to provide additional variation for population genetic analyses and to
+monitor populations for potential phenotypic patterns.
 
@@
 influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.
+\comment{Describe additional E1/E2 markers (additional markers within the RoSA region) and
+Tasha Marker (because it is close to the duplication noted
+by Thompson et al.). We don't go into detail here---keeping it vague because it is
+going to be described in the subRoSA paper, but don't mention that.}
+
+We also include a marker that targets the sex-determining region on the Y chromosome in Chinook salmon
+and that characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}
 
@@
 structure in the data by performing unsupervised clustering of
-the complete data set (\comment{With or without the LFAR? Which did Anthony use? I would
-like to do it with the LFAR.}) 
+the complete data set
 using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
 We performed 20 replicate runs at each value of the assumed number of
-subpopulations, $K$, in $\{2, 3, 4, 5, 6\}$ 
-\comment{I think we should also see what $K=7$ turns up}
+subpopulations, $K$, in $\{2, 3, 4, 5, 6, 7\}$ 
 using default settings of the program.  STRUCTURE output was
@@
 broken down according to the genotype of the fish at the markers within
-the RoSA.  For this, \comment{Need to describe how I categorized
-genotypes}.
+the RoSA.  For this, \comment{Eric. Describe this from what you have in the notebook}.
 
@@
 
-We tested \comment{XX} candidates from our additional microhaplotype discovery and
-retained \comment{106} of those in the panel.
+We tested 229 candidates from our additional microhaplotype discovery and
+initially retained the 125 markers described in \citep{thompson2020complex}
+for use in stock identification within
+the Klamath Basin.  However, further evaluation of those markers in other
+populations led us to drop some, and retain 106 of those markers in the current panel.
 
@@
 giving the category or provenance of the markers (i.e. LFAR or MHAP or SNPTYPE, etc.) 2) retain
-only the 2 LFAR markers we actually use. 3) consider renaming the E2 markers since
-winter-run-greb1L-diagnostic is confusing...actually, I would consider removing those
-entirely---we can introduce them in the subRoSA paper. We don't use them in this paper and I don't
-really want to do the explanation of them again.}).
+only the 2 LFAR markers we actually use. 3) Definitely rename the E2 markers to be something other
+than winter-run-greb1L-diagnostic}).
 
@@
 cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
-($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.
-This also is the reason that many of the lines terminate.
-}
+($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$
+is the estimated standard error of the mean of the Monte Carlo sample.  This also is the reason that many of the lines terminate.   }
 \begin{center}
@@
 And we can also point readers to the supplement figure \ref{fig:ckmr-comp}.
+The power analyses for this set of markers demonstrate that they have ample variation
+for accurate identification of parent-offspring (PO) and full-sibling (FS) pairs in almost all
+realistic situations. The predicted false positive rates for both PO and FS pairs are
+exceedingly low for even stringent false negative thresholds (e.g. 0.05). 
+For example, the false positive rate for FS identification is less than one error in 10 million
+comparisons of potential siblings in the California Chinook salmon population with the lowest
+heterozygosity (SRW) at a false negative rate of 0.05. However, HS pairs can not be accurately
+distinguished from unrelated pairs in any populations without very large false negative rates
+and in very modest-sized studies.
+We note, however, that the presence in a dataset of genotypes from individuals that are related, but not in
+ the relationship category proposed, will result in some additional misidentifications.
+ These 'errors' will all involve related fish being identified as such, but in the incorrect relationship category.
+The cross-assignment frequencies predicted for some common relationship categories, and for realistic FPR and FNR, are found in (Fig 4b).
+
+\textbf{Although some additional errors might be
+expected when related but non-FS individuals are sampled,  this number of comparisons exceeds
+those necessary in almost all realistic scenarios.
+We note that, as in previous work, it is nearly impossible to identify HS pairs accurately
+with data from this, or any standard marker dataset. However, the use of properly LD-filtered
+genome-scale data should allow such discrimination in future work.
+Accurate relationship
+inference in even the most genetically depauperate Chinook salmon population in California
+and rangewide (Seeb et al. 2007, Clemento et al. 2014) is therefore possible with this marker set.}
+
+While there is not a currently available method (like that of \citeauthor{anderson2006power}) for estimating error rates in
+inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which require much more statistical power, it is clear that PO-trio inference with these markers will be essentially error free.
 

### 2024-06-18 extras/Renaming-amplicons/Amplicon-renaming.Rmd
Make a subdir within extras

---
title: "Renaming The Amplicons and Tidying up the Amplicon Information Table"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2024-06-18 extras/Renaming-amplicons/Amplicon-renaming.Rmd
Done with revamping the locus names

@@
 
-This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.
+Before releasing this, we want to have a decent, consistent naming convention for
+all of the amplicons.  We want a short, unique, descriptive name that will tell us
+what kind of marker it is (snplicon vs microhap, etc), and then we want to have
+position information in there too.
 
-Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.
+Here is the convention that we came up with:
+
+General pattern = `Ots_type00X_ChrNum_V2StartPos`, where:
+
+- `Ots` is just `Ots`
+- `type` is a brief descriptor (shoot for 4 characters) for example:
+    + SNPtype = `scon`
+    + Microhap = `mhap`
+    + RoSA = `rosa`
+    + SexID = `sexy`
+    + LFAR = `lfar`
+    + WRAP = `wrap`
+    + VGLL = `vgll`
+    + Six6 = `sixx`
+- 00X is a three digit left zero-padded, base-1 number within type ordered by genome V2 location.
+- ChrNum is a two digit left zero padded Chromosome number, except for sexy which is
+  going to be just the scaffold name, with the underscore removed.
+- V2StartPos is just where it maps in Otsh_V2.
+
+So, let's get crackin' on slurping stuff out of the Excel file and making these modifications.
+
+## Doing it
 
 ```{r}
-plot(cars)
+library(tidyverse)
+library(readxl)
+
+# read it in and change some of the names
+orig <- read_excel("ChinookFullPanelLociTable_May2024Updated.xlsx") %>%
+  rename(
+    OtherName = Locus,
+    Otsh_v1.0_coordinates = AmpliconGenomeV1Location,
+    Otsh_v2.0_coordinates = AmpliconGenomveV2Location,
+    Alternate_Otsh_v1.0_coordinates = `AlternateGenomeV1Location(s)`,
+    Alternate_Otsh_v2.0_coordinates = `AlternateGenomeV2Location(s)`,
+    LocusOrigin = LocusType,
+    OriginalReference = `Original Reference`,
+    MappingNotes = Notes
+  )
+
+# now, reorder the columns so that more information is front loaded
+orig2 <- orig %>%
+  select(
+    LocusOrigin,
+    OriginalReference,
+    OtherName,
+    ForwardPrimerSequence,
+    ReversePrimerSequence,
+    ReferenceSequence,
+    SNPLocationsInReferenceSequence,
+    TrimmedAmpliconSequence,
+    Otsh_v2.0_coordinates,
+    Alternate_Otsh_v2.0_coordinates,
+    Otsh_v1.0_coordinates,
+    Alternate_Otsh_v1.0_coordinates,
+    everything()
+  )
 ```
 
-Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
+Now we are going to get columns of chromosome number (or name if unmapped) and
+start positions, so that we can order things by genome coordinate.
 
-When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
+```{r}
+# get the chromosome numbers (pulled from NCBI page at https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/)
+chroms <- read_tsv("Otsh_v2.0_chroms.tsv") %>%
+  mutate(Otsh_v2.0_chromosome = str_replace(Chromosome, "LG", ""), .before = Chromosome) %>%
+  select(RefSeq, Otsh_v2.0_chromosome)
 
-The preview shows you a rendered HTML copy of the contents of the editor.
-Consequently, unlike *Knit*, *Preview* does not run any R code chunks.
-Instead, the output of the chunk when it was last run in the editor is displayed.
+orig3 <- orig2 %>%
+  extract(Otsh_v2.0_coordinates, into = c("RefSeq", "AmpliconStartPos"), regex = "^(.*):([0-9]+)-[0-9]+$", remove = FALSE, convert = TRUE) %>%
+  left_join(chroms, by = join_by(RefSeq)) %>%
+  mutate(Otsh_v2.0_chromosome = ifelse(is.na(Otsh_v2.0_chromosome), str_replace(RefSeq, "_", ""), Otsh_v2.0_chromosome)) %>%
+  select(Otsh_v2.0_chromosome, AmpliconStartPos, everything()) %>%
+  select(-RefSeq) %>%
+  arrange(Otsh_v2.0_chromosome, AmpliconStartPos)
+
+```
+
+Now we
+will be able to rename everything all at once, easily.
+
+```{r}
+final <- orig3 %>%
+  mutate(
+    tmpname = case_match(
+      LocusOrigin,
+      "LFAR" ~ "lfar",
+      "Microhap" ~ "mhap",
+      "RoSA" ~ "rosa",
+      "SNPtype" ~ "scon",
+      "Sex ID" ~ "sexy",
+      "Six6" ~ "sixx",
+      "VGLL3" ~ "vgll",
+      "WRAP" ~ "wrap"
+    )
+  ) %>% 
+  group_by(tmpname) %>%
+  mutate(
+    AmpliconName = str_c(
+      "Ots_", tmpname,
+      sprintf("%03d", 1:n()), "_", 
+      Otsh_v2.0_chromosome, "_", 
+      sprintf("%08d",AmpliconStartPos)),
+    .before = Otsh_v2.0_chromosome
+  ) %>%
+  ungroup() %>%
+  select(-tmpname)
+```
+
+All right!  That pretty much does it.  Let's just print it out.
+
+```{r}
+write_csv(final, "../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv", na = "")
+```

### 2024-06-18 extras/Renaming-amplicons/Amplicon-renaming.Rmd
Cleaning up some stuff

@@
 
-So, let's get crackin' on slurping stuff out of the Excel file and making these modifications.
+### As an aside:
+
+To process these things in the snakemake mega-simple workflow, we will add a new
+field that goes within `genome` or `target_fasta` in the config file which is called
+`rename:` and that give tsv file with columns named `orig` and `new` which are the
+original names and the desired new names, and we will make one extra rule in there
+somehow to deal with it.  We will keep the VCFs with the original names, but we need
+a rule that will make a renamed VCF, and we also need a renamed Microhaplot output.
+
+So we won't change the target fastas.  This way we have a data set that is compatible
+with old stuff, and we also have one that we can use with the new names.
 
 ## Doing it
+
+So, let's get crackin' on slurping stuff out of the Excel file and making these modifications.
 

### 2024-06-21 extras/Applying-New-Names-To-Baseline/applying-new-names-to-baseline.Rmd
Beginnging to apply new locus names to the baseline
But I am going to have to get a more complete baseline from Anthony.

---
title: "Applying the new names to the baseline"
output: html_notebook
---

We are just going to take the most recent version of the 
baseline and munge it to change the names of the loci.
```{r}
library(tidyverse)

amplicons <- read_csv("../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")

new_names <- amplicons %>%
  select(AmpliconName, OtherName)

base <- read_csv("subRoSA_baseline_rubias_JUN2024_w_mixture_complete-rosa_lfar_vgll3six6_added.csv") %>%
  pivot_longer(
    cols = -(indiv:sample_type),
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "^(.+)_([12])$"
  ) %>%
  left_join(new_names, by = join_by(locus == OtherName))

```

OK, now, let's check and see which ones don't have new names:

```{r}
base %>%
  filter(is.na(AmpliconName)) %>%
  count(AmpliconName, locus)
```
OK, cool.  Those are the LFAR's that we dropped.  We expect that they won't have new names
because we don't really want them in there.

Let's see if there are any AmpliconName's that don't appear in base.

```{r}
amplicons %>%
  anti_join(base, by = join_by(AmpliconName))
```

### 2024-06-28 tex/abstract.tex
Minor edits

--- 2023-12-15 tex/abstract.tex
@@
 
-We need to write an abstract.
+Genetic methods have become an essential component of ecological investigation
+and conservation planning for fish and wildlife. Among these methods is the use
+of genetic marker data to identify individuals to
+populations, or stocks, of origin. More recently, methods that involve genetic
+pedigree reconstruction to identify relationships between individuals within populations
+have become common. We present, here, a novel set of
+multi-allelic microhaplotype genetic markers for Chinook salmon which provide unprecedented
+resolution for population discrimination and relationship identification for a rapidly and
+economically assayed panel of markers. We show how this set of microhaplotypes
+provides definitive power to identify all known lineages of Chinook salmon in
+California. The inclusion of genetic loci that have known associations with phenotype
+and that were identified as outliers in examination of whole genome sequence data,
+allows resolution of stocks that are not highly genetically differentiated but
+are phenotypically distinct and managed as such. 
+This same set of multiallelic genetic markers have ample variation to
+accurately identify parent-offspring and full-sibling
+pairs in all California populations, including the genetically depauperate winter-run
+lineage.
+Validation of this marker panel in salmon populations not previously studied with
+modern genetic methods, also reveals novel biological insights, including
+the presence of a single copy of a haplotype for a phenotype that has not
+been documented in that part of the species range, and a clear signal of mixed
+ancestry for a salmon population that is on the geographic margins of the 
+primary evolutionary lineages present in California.

### 2024-06-28 tex/author-title-etc.tex
More overleaf changes
Minor edits

@@
 %%%%% INFORMATION AND  A FEW OTHER THINGS
-\newcommand{\myTitle}{A multipurpose microhaplotype panel for analysis of California Chinook salmon}
+\newcommand{\myTitle}{A multipurpose microhaplotype panel for genetic analysis of California Chinook salmon}
 \title{\myTitle}
@@
 Clemento$^{*}$, Matthew A.
-Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
-\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
+Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F.
+Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}
+\author{Eric C.
+Anderson$^{*}$, Anthony J.
+Clemento$^{*}$, Matthew A.
+Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F.
+Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}
 
@@
 \newcommand{\myAffiliations}{
-$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA. $^\dagger$Current address: Mac's home address in AK.
+$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA.
+$^\dagger$Current address: Mac's home address in AK.
+$^\ddag$Current address: Get Neil's address.
 }

### 2024-06-28 tex/body-text-of-paper.tex
Minor edits

@@
 
-The emergence over the last decade of parentage based tagging (PBT) as a management
-tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
+Over the last decade an additional genetic method, parentage based tagging, or PBT,
+\citep{anderson2005description,steele2019parentage}
+has become established as an extremely valuable management
+tool for Pacific salmon.  The availability of such family-based methods adds another factor to
+consider when developing a GSI baseline.
 Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
@@
 vice versa).  
+Thus, any panel of markers to be used for GSI (or PBT) should also be evaluated on its
+utility for PBT (or GSI).
 
-Thus, any panel of markers to be used for GSI (or PBT) should also be evaluated on its
-utility for PBT (or GSI).  PBT has been remarkably successful in fisheries
+PBT has been remarkably successful in fisheries
 management, having been used for over a decade in the management of Chinook salmon

### 2024-07-02 extras/Applying-New-Names-To-Baseline/applying-new-names-to-baseline.Rmd
Committing code to rename loci

@@
 
-base <- read_csv("subRoSA_baseline_rubias_JUN2024_w_mixture_complete-rosa_lfar_vgll3six6_added.csv") %>%
+csv <- read_csv("subRoSA_baseline_rubias_with_rosa_wrap_lfar_vgll3six6_and_mixture_28JUN2024.csv")
+
+# for some reason, there are 22 fish with identical sample_IDs.  
+dupies <- csv %>%
+  count(Sample_ID) %>%
+  filter(n > 1) %>%
+  pull(Sample_ID)
+
+# there doesn't seem to be too much any difference in how much missing data each copy of
+# each duplicated fish has:
+mat <- csv %>% filter(Sample_ID %in% dupies) %>% as.matrix()
+rowSums(is.na(mat))
+
+# So, we will just drop the second instance of each:
+csv2 <- csv %>%
+  filter(!duplicated(Sample_ID))
+
+base <- csv2 %>%
   pivot_longer(
-    cols = -(indiv:sample_type),
+    cols = -(Sample_ID:sample_type),
     names_to = c("locus", "gene_copy"),
@@
 
+Let us count how many things match up here.
+```{r}
+base %>%
+  filter(!is.na(AmpliconName) & !is.na(locus)) %>%
+  count(locus, AmpliconName)
+
+```
+
+193 sounds about right.  Those are probably just the RoSAs and the SexID marker.
+
 OK, now, let's check and see which ones don't have new names:
@@
 ```
-OK, cool.  Those are the LFAR's that we dropped.  We expect that they won't have new names
+OK, cool.  Those look like the LFAR's and the WRAPs that we do not include in the
+final baseline.  We expect that they won't have new names
 because we don't really want them in there.
@@
 ```
+Only the RoSA's and the SexID marker.  OK.  That is looking good.
+
+So, let's put it all back together again with the other names.
+```{r}
+wide <- base %>%
+  filter(!is.na(locus) & !is.na(AmpliconName)) %>% # drops the RoSA stuff
+  select(-locus) %>%
+  pivot_wider(names_from = c(AmpliconName, gene_copy), values_from = allele, names_sep = "_")
+
+# but we want to put things in coordinate sorted order
+loci <- amplicons$AmpliconName # get locus names
+loci <- loci[!loci == "Ots_sexy001_NW024608692.1_00004816"]  # remove the sexID locus, cuz we don't have it
+loci <- loci[!str_detect(loci, "rosa")]  # drop the RoSAs cuz they are in there as the genotype string
+
+loccols <- paste(rep(loci, each = 2), c(1, 2), sep = "_")
+
+wide_ordered_no_rosa <- wide[, c("Sample_ID", "repunit", "collection", "sample_type", loccols)]
+
+# and then we also want to add the rosa genotypes on there, AND remove the SacNonW fish
+final <- wide_ordered_no_rosa %>%
+  left_join(csv2 %>% select(Sample_ID, hapstr), by = join_by(Sample_ID)) %>%
+  mutate(rosa_genotypes_str = hapstr, .after = sample_type) %>%
+  select(-hapstr) %>%
+  filter(!(collection == "SacNonW"))
+```
+
+Now, we can write that out.
+
+```{r}
+dir.create("results", showWarnings = FALSE)
+write_csv(final, "results/subRoSA_baseline_rubias_loci_renamed_July_2024.csv")
+```

### 2024-07-02 001-map-of-samples.Rmd
Clean up Snakemake stuff in 001

@@
 start_time <- Sys.time()
-
-USING_SNAKEMAKE <- FALSE
-NEG_USING_SMK <- TRUE
-if(exists("snakemake")) {
-  USING_SNAKEMAKE <- TRUE
-  NEG_USING_SMK <- FALSE
-}
-```
-
-```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
-# currently, this block does not get evaluated, while developing,
-# but later when we snakemake-ize it all, we will evaluate it.
-if(USING_SNAKEMAKE) {
-  input_list <- snakemake@input
-  output_list <- snakemake@output
-}
 ```
@@
 ```{r}
+if(exists("snakemake")) {
+  input_list <- snakemake@input
+  output_list <- snakemake@output
+} else {
   # inputs:
@@
   )
-# you can create the necessary output directories like this:
+}
+
+# we create the necessary output directories like this:
 dump <- lapply(output_list, function(x)
@@
 
-Now look add stuff to that:
+Now add stuff to that:
 

### 2024-07-02 extras/Applying-New-Names-To-Baseline/applying-new-names-to-baseline.Rmd
writing the loci-renamed baseline file.
(Also, I copied that to data)
Change sample column name to indiv

@@
   select(-hapstr) %>%
-  filter(!(collection == "SacNonW"))
+  filter(!(collection == "SacNonW")) %>%
+  rename(indiv = Sample_ID)
 ```
@@
 dir.create("results", showWarnings = FALSE)
-write_csv(final, "results/subRoSA_baseline_rubias_loci_renamed_July_2024.csv")
+write_csv(final, "results/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz")
 ```

### 2024-07-02 002-GSI_and_Fst.Rmd
Update 002 to use the baseline with new locus names

@@
   input_list <- list(
-    final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
+    final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
     final_baseline_apr = "data/subRoSA_baseline_rubias_APR2024_w_mixture_revised_rosa_added.csv",
@@
 
-# get the full baseline with the new markers, etc.  We will toss out some
-# non-winter Sacto fish of unknown provenance
-###full_base <- read_csv(input_list$final_baseline) %>%
-###  filter(collection != "SacNonW")
-
-full_base_apr <- read_csv(input_list$final_baseline_apr) %>%
-  filter(collection != "SacNonW")
-
-# remove the rosa hapstr from it
-full_base <- full_base_apr %>%
-  select(-hapstr)
+# read in the full baseline
+full_base <- read_csv(input_list$final_baseline)
 
 # capture that rosa hapstr with some meta data from future use
-rosa_hapstrs <- full_base_apr %>%
-  select(indiv:hapstr)
+rosa_genos0 <- full_base %>%
+  select(indiv:rosa_genotypes_str)
 
@@
 
-# Include the LFAR in there in a version
+# Get a version with no LFAR
 
-We do these as an additional step to get a sense for how much
-improvement we get from adding the LFAR in there.
+We might want to do a run with no LFAR in there just to see how much
+they add to the power for identifying the late fall.
 
-First we need to join the LFARs on there:
 ```{r}
-# get just the 2 LFAR markers we are using.  We are dropping 828619 because
-# it really doesn't do much, and so many reads are off-target.
-lfar_columns <- paste(
-  rep(c(
-    "NC_037130.1:1062935-1063235",
-    "NC_037130.1:864908-865208"
-  ), each = 2),
-  rep(c(1,2), 2),
-  sep = "_"
-)
 
-lfar3genos <- lfar_genos[c("Sample_ID", lfar_columns)]
+# get the positions of the LFARs:
+lfar_idxs <- which(str_detect(names(full_base2), "lfar"))
 
-base_with_lfar <- full_base2 %>%
-  left_join(lfar3genos, by = join_by(indiv == Sample_ID))
+# then make a version of the baseline without them
+base_with_no_lfar <- full_base2[, -lfar_idxs]
 ```
@@
 ```{r}
-rosa_hapstrs %>%
-  count(hapstr) %>%
+rosa_genos0 %>%
+  count(rosa_genotypes_str) %>%
   arrange(desc(n))
@@
 ```{r}
-H8 <- rosa_hapstrs %>%
-  select(indiv, hapstr) %>%
+H8 <- rosa_genos0 %>%
+  select(indiv, rosa_genotypes_str) %>%
   mutate(
-    hvec = map(hapstr, function(x) {y <- strsplit(x, "")[[1]]; y[c(1, 4, 5, 7:11)]}),
+    hvec = map(rosa_genotypes_str, function(x) {y <- strsplit(x, "")[[1]]; y[c(1, 4, 5, 7:11)]}),
     hapstr8 = map_chr(hvec, function(x) paste(x, collapse = "", sep = ""))
@@
 of those are missing data.  I am going to be stringent here and toss anything with missing data,
-which is only about 150 of them, so that should be fine. It will make downstream processing
+which is only about 71 of them, so that should be fine. It will make downstream processing
 much easier, since the number of categories will be a bit smaller. (And we can be quite confident
@@
 ```
-
-This is interesting.  I am struck by the 12 instances of EEHEEEEE---that suggests to me that
-it is not simply the result of genotyping error.  I suspect that one of the haplotypes
-has an L at that position.  The cleanest thing here might be to simply go for the
-things that are perfectly pure across all the sites, cuz there is a much lower chance
-of the region being broken up by recombination or gene conversion there.  So, that is what
-I am going to do.  And we can look at the ones that are not monolithic (possible recombinants, etc.
-as well.)
 
@@
 We go directly to assigning to repunit, and then to the max-likelihood
-collection within that repunit.  We include the lfar in here.
+collection within that repunit.  We don't include the lfar in here, at first.
 
 ```{r}
-full_sa <- self_assign(full_base2, gen_start_col = 5)
+# gen_start_col is 6 because we have the rosa_genotypes_str column in there
+full_sa <- self_assign(base_with_no_lfar, gen_start_col = 6)
 ```
@@
 ```{r}
-lfar_sa <- self_assign(base_with_lfar, gen_start_col = 5)
+lfar_sa <- self_assign(full_base2, gen_start_col = 6)
 lfar_top_ass <- get_top_assignments(lfar_sa, collection_order)
@@
 marks <- paste("s", 1:12, sep = "")
-exploded <- rosa_hapstrs %>%
-  filter(!(collection %in% c("ColeRHsp", "SmithFall"))) %>%
-  separate(hapstr, into = paste("s", 0:12, sep = ""), sep = "") %>%
+exploded <- rosa_genos0 %>%
+  separate(rosa_genotypes_str, into = paste("s", 0:12, sep = ""), sep = "") %>%
   select(-s0)
@@
 
-So, there is only one more E2-het outside of the Sacto, and that is from TRHsp.
-
+So, in California, there is only one more E2-het outside of the Sacto, and that is from TRHsp. 
 Let's go back and find that:
@@
   filter(repunit == "klamathspring" & s6 == "H") %>%
-  left_join(rosa_hapstrs, by = join_by(indiv))
+  left_join(rosa_genos0, by = join_by(indiv))
 
+```
+
+However, at ColeRivers we have 12 instances of heterozygotes at the s6 E2 position. Let's count things up from
+ColeRivers:
+```{r}
+rosa_genos0 %>% 
+  filter(collection == "ColeRHsp") %>%
+  count(rosa_genotypes_str)
 ```
@@
 This means making integers of the alleles and then catenating them.
-One locus has 11 allele, so we should use two digits for each allele.  We drop the
-LFAR from this, because those were chosen specifically to be weird in the late fall.
+One locus has 11 allele, so we should use two digits for each allele.
 ```{r}
-full_base_long <-  base_with_lfar %>% 
+full_base_long <-  full_base2 %>% 
   select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
@@
 # long format with no missing data
-long_genos <- base_with_lfar %>%
+long_genos <- full_base2 %>%
 select(-repunit, -sample_type) %>%
@@
 ```{r}
-fbl_no_miss <- base_with_lfar %>%
+fbl_no_miss <- full_base2 %>%
   select(-starts_with("NC_037130.1")) %>%  # drop the LFARs

### 2024-07-02 tex/inputs/samples-table.tex
Update 002 to use the baseline with new locus names

@@
 Russian R.&RRF&Warm Springs H.&Fall&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
-Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&109&19&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
-Sacramento R.&BCS&Butte Ck.&Spring&CVSR&114&114&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
-Sacramento R.&MDS&Deer Ck.&Spring&CVSR&32&32&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
-Sacramento R.&MDS&Mill Ck.&Spring&CVSR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&20&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
+Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
+Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
+Sacramento R.&MDS&Mill Ck.&Spring&CVSR&50&50&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
 Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
 Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
-Sacramento R.&BCF&Butte Ck.&Fall&CVFR&44&44&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
-Sacramento R.&MDF&Deer Ck.&Fall&CVFR&22&22&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
-Sacramento R.&MDF&Mill Ck.&Fall&CVFR&48&48&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
+Sacramento R.&MDF&Deer Ck.&Fall&CVFR&24&24&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&MDF&Mill Ck.&Fall&CVFR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
 San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
-Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&252&252&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
+Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline

### 2024-07-02 extras/Renaming-amplicons/Amplicon-renaming.Rmd
Revamped Amplicon Info to have SNP positions...
...relative to the actual amplicon that maps, so that we can calculate the Otsh_v2.0 position easily.

@@
 
+## Deal with the different between ReferenceSequence and TrimmedAmpliconSequence
+
+So, the `RefereneSequence` is sometimes the same as the TrimmedAmpliconSequence, and
+sometimes it is not. Mostly it is not the same for the WRAP and LFAR and RoSA, etc.
+So, this is not a huge problem.  In cases where things are not the same, the ReferenceSequence
+was the sequence I sent to Anthony to design the primers, while the TrimmedAmpliconSequence
+is what is actually amplified and what actually gets mapped to the Otsh_V1.0 genome.
+So, when we want to figure out the actual position in the genome of particular variants,
+we need to know their position in the TrimmedAmpliconSequence, but we actually
+only have it in the ReferenceSequence.
+
+So, we have to do some modifications here to make sure we have the position
+in the variants in the actualy amplicon sequence.  We can do that pretty
+easily.
+
+I am going to rename things a little bit so ReferenceSequence becomes DesignSequence, and
+TrimmedAmpliconSequence becomes "AmpliconSequence".
+
+```{r}
+# we first need a function to modify the snp locations
+# pos is the current list column of positions
+# as is the amplicon start position in the reference sequence
+new_locations <- function(pos, as, ae) {
+  amp_length = ae - as
+  pvec <- lapply(str_split(pos, ", *"), as.integer)
+  altered_pvec <- map2(.x = pvec, .y = as, .f = function(x, y) x - y + 1)
+  thinned_pvec <- map2(.x = altered_pvec, .y = amp_length, .f = function(x, y) x[x > 0 & x <= y])
+  ret <- map_chr(.x = thinned_pvec, .f = function(x) paste(x, collapse = ", "))
+  unlist(ret)
+}
+
+tmp <- final %>%
+  mutate(
+    ampli_start = str_locate(ReferenceSequence, TrimmedAmpliconSequence)[,1],
+    ampli_end = str_locate(ReferenceSequence, TrimmedAmpliconSequence)[,2],
+    .before = AmpliconName
+  ) %>%
+  mutate(
+    SNPLocationsInAmpliconSequence = pmap_chr(
+      .l = list(pos = SNPLocationsInReferenceSequence, as = ampli_start, ae = ampli_end),
+      .f = new_locations
+    ),
+    .after = SNPLocationsInReferenceSequence
+  )
+
+# then, pick out columns in the right order
+real_final <- tmp %>% 
+  select(
+    AmpliconName, 
+    Otsh_v2.0_chromosome, 
+    AmpliconStartPos, 
+    LocusOrigin, 
+    OriginalReference, 
+    OtherName, 
+    ForwardPrimerSequence, 
+    ReversePrimerSequence, 
+    TrimmedAmpliconSequence,
+    SNPLocationsInAmpliconSequence, 
+    Otsh_v2.0_coordinates, 
+    Alternate_Otsh_v2.0_coordinates, 
+    Otsh_v1.0_coordinates, 
+    Alternate_Otsh_v1.0_coordinates, 
+    MappingNotes,
+    ReferenceSequence, 
+    SNPLocationsInReferenceSequence, 
+  ) %>%
+  rename(
+    AmpliconSequence = TrimmedAmpliconSequence,
+    DesignSequence = ReferenceSequence,
+    SNPLocationsInDesignSequence = SNPLocationsInReferenceSequence
+  ) %>%
+  mutate(
+    SNPLocationsInAmpliconSequence = ifelse(SNPLocationsInAmpliconSequence == "NA", NA, SNPLocationsInAmpliconSequence)
+  )
+```
+
+## Print that dude out
+
 All right!  That pretty much does it.  Let's just print it out.
@@
 ```{r}
-write_csv(final, "../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv", na = "")
+write_csv(real_final, "../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv", na = "")
+write_csv(real_final, "../../tex/supp_data/Supp-Data-1-Amplicon-info.csv", na = "")
 ```

### 2024-07-02 tex/body-text-of-paper.tex
Re-calc LFAR freqs with locus-renamed baseline

@@
 markers.  Frequencies are given for the alleles most common amongst late-fall Chinook: nucleotide 
-{\tt G} at locus Chr34:865,057 and {\tt A} at Chr34:1,063,084.  $N$ is the total number of gene 
-copies available for estimating the allele frequency.}
+{\tt G} at locus Chr34:954,054 and {\tt A} at Chr34:1,151,868.  $N$ is the total number of gene 
+copies available for estimating the allele frequency. Genome coordinates of SNPs are from Otsh\_v2.0.}
 \label{tab:lfar-freqs}
@@
 \hline\hline
-& \multicolumn{2}{c}{\underline{Chr34:865,057}} & \multicolumn{2}{c}{\underline{Chr34:1,063,084}} \\
+& \multicolumn{2}{c}{\underline{Chr34:954,054}} & \multicolumn{2}{c}{\underline{Chr34:1,151,868}} \\
 Reporting Unit & freq. & $N$ & freq. & $N$ \\ \hline

### 2024-07-05 tex/body-text-of-paper.tex
Edits in the Introduction

@@
 management, having been used for over a decade in the management of Chinook salmon
-and steelhead trout in major basins of the Columbia River \cite{steele2019parentage,horn2023multigeneration}, and having been employed to dramatically further our
+and steelhead trout in major basins of the Columbia River \citep{steele2019parentage,horn2023multigeneration}, and having been employed to dramatically further our
 understanding of the genetic inheritance of key traits in salmonids
@@
 family of statistical genetic methods employing relationship inference to learn about
-populations.  For example inference of the full siblings amongst a sample of fish
+populations.  For example, inference of the full siblings amongst a sample of fish
 provides information about the effective number of adults producing offspring
@@
 be a valuable source of information when sampling of the adults is not possible.
+Accordingly, marker panels should be evaluated not only on their capacity to
+identify the parent-offspring relationships employed in PBT, but also on their
+ability to resolve full-sibling relationships.
 
-Here, we present a reference baseline for Chinook salmon, targeted to the complex population 
-structure within the Central Valley of California. Chinook salmon of the two main river basins---the 
+Here, we present a reference baseline for Chinook salmon, focusing on the regional scale of
+rivers within the state of California, and particularly targeted to the complex population 
+structure of Chinook salmon within California's Central Valley. Chinook salmon of the two main river basins---the 
 Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
@@
 (fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California. \comment{NEED MORE HERE. Certainly we should cite
-\citet{yoshiyama1998historical}.}
+in freshwater every month of the year in California \citep{fisher1994past}.
 
-Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important fisheries.
-Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities during many different seasons and geographic locations.
-However, recent declines in population sizes, from the Yukon River in Arctic north, to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect less productive stocks.
+Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important fisheries \citep{myers1998status}.
+Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities during many different seasons and geographic locations \citep{healey1991life}.
+However, recent declines in population sizes, from the Yukon River in Arctic north, to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect less productive stocks \comment{(CITE)}.
 The co-occurence of fish from relatively productive and relict populations reaches its paragon in California, with the largest remaining ocean fisheries for Chinook salmon targeting the Central Valley fall-run stock that spawns in different tributaries of the same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks.
-In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are very closely related \cite{clemento2014evaluation} and share recent common ancestry that is independent of ongoing migration between these lineages.
-As such, GSI has been particularly challenging in the CCV, with at least one of the distinct ecotypes, late-fall-run, unresolvable with previous GSI marker sets \cite{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the harvested fall-run stock.
 
-In the following we present a new reference baseline for both GSI and PBT in California.
+In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are closely related \citep{clemento2014evaluation} and share recent common ancestry that is independent of ongoing migration between these lineages.
+As such, GSI has been particularly challenging in the CCV, with at least one of the distinct ecotypes---late-fall-run---unresolvable with previous GSI marker sets \citep{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the harvested fall-run stock \comment{(CITE)}.
+
+In the following we present a new reference baseline that includes 1636 Chinook salmon individuals
+from 17 collections, genotyped at 204 amplicons from throughout the genome,
+scored as microhaplotypes \citep{baetscher2018microhaplotypes}. This baseline is
+highly effective for GSI within California and the markers within it provide ample power
+for PBT.
 We describe how we identified and compiled the markers, including several new markers
-that are particularly effective at distinguishing some of the closely related groups,
-such as late-fall and fall-run fish.  We also provide a comprehensive analysis of this
+that are particularly effective at distinguishing between closely related groups of fish,
+such as late-fall and fall-run Chinook salmon in the CCV.  We also provide a comprehensive analysis of this
 set of markers for inference of parents and siblings within different groups of populations.

### 2024-07-05 extras/Renaming-amplicons/Amplicon-renaming.Rmd
Add genomic locations plot to the supplement

@@
 # get the chromosome numbers (pulled from NCBI page at https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/)
-chroms <- read_tsv("Otsh_v2.0_chroms.tsv") %>%
+chroms0 <- read_tsv("Otsh_v2.0_chroms.tsv") %>%
   mutate(Otsh_v2.0_chromosome = str_replace(Chromosome, "LG", ""), .before = Chromosome) %>%
+  select(RefSeq, Otsh_v2.0_chromosome, `Size (bp)`)
+
+chroms <- chroms0 %>%
   select(RefSeq, Otsh_v2.0_chromosome)
@@
   arrange(Otsh_v2.0_chromosome, AmpliconStartPos)
+
+```
+
+Before I go on, I am going to write out the Otsh_v2.0 chromosome lenghts for later use.
+```{r}
+chroms0 %>%
+  rename(
+    num_bases = `Size (bp)`,
+    chrom = Otsh_v2.0_chromosome
+  ) %>%
+  write_tsv("Otsh_v2.0_chrom_lengths.tsv")
 

### 2024-07-05 tex/supplemental-figures.tex
Add genomic locations plot to the supplement

@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%% 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{images/genomic-locations-plot.pdf}
+\end{center}
+\caption[Genomic locations of amplicons]{\footnotesize Genomic locations of amplicons in the
+Otsh\_v2.0 assembly of the Chinook salmon genome.  Color shows type of marker (see main text).
+The sex-ID marker is not included here because it aligns to a scaffold that is not part of a
+named chromosome/linkage group in the assembly.}
+\label{fig:num-alle}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
 

### 2024-07-07 ckmr-sim-evaluation.Rmd
Get ckmr running with renamed amplicons

@@
   input_list <- list(
-    final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
+    final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
     lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
     pop_labels = "inputs/reference-collection-names.csv",
-    locus_info = "inputs/ChinookFullPanelLociTable_May2024Updated.xlsx"
+    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv"
   )
@@
 # non-winter Sacto fish of unknown provenance
-full_base <- read_csv(input_list$final_baseline) %>%
-  filter(collection != "SacNonW")
-
-lfar_genos <- read_csv(input_list$lfar_genos)
+full_base <- read_csv(input_list$final_baseline)
 
@@
 # we take its position to be the start position of the amplicon
-gen_coords <- read_excel(input_list$locus_info) %>%
-  select(Locus, AmpliconGenomveV2Location) %>%
-  separate(AmpliconGenomveV2Location, into = c("Chrom", "pos_str"), sep = ":") %>%
-  separate(pos_str, into = c("Start", "Stop"), sep = "-", convert = TRUE) %>%
-  rename(Pos = Start) %>%
-  select(-Stop) %>%
-  arrange(Chrom, Pos)
+gen_coords <- read_csv(input_list$locus_info) %>%
+  rename(Locus = AmpliconName, Chrom = Otsh_v2.0_chromosome, Pos = AmpliconStartPos) %>%
+  select(Locus, Chrom, Pos)
 

### 2024-07-08 extras/Renaming-amplicons/add-species-id.Rmd
Rename markers to include the coho marker in there

---
title: "Renaming all Amplicons to add the Coho-species-id marker"
output: html_notebook
---

After running through the Amplicon-renaming.Rmd notebook, we realized that we
really should add the species ID marker in there.   Doing so wil mean
renumbering all the new names for the scon markers.  No worries.  We can do that.

We will pull it from the renamed input file. Should not be a big deal.

```{r}
library(tidyverse)

amplicons <- read_csv("../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")

amp2 <- amplicons %>%
  mutate(LocusOrigin = ifelse(OtherName == "OkiOts_120255-113", "Coho", LocusOrigin))

# now we just rename all these guys:
renamed <- amp2 %>%
  mutate(
    tmpname = case_match(
      LocusOrigin,
      "Coho" ~ "coho",
      "LFAR" ~ "lfar",
      "Microhap" ~ "mhap",
      "RoSA" ~ "rosa",
      "SNPtype" ~ "scon",
      "Sex ID" ~ "sexy",
      "Six6" ~ "sixx",
      "VGLL3" ~ "vgll",
      "WRAP" ~ "wrap"
    )
  ) %>% 
  group_by(tmpname) %>%
  mutate(
    AmpliconName = str_c(
      "Ots_", tmpname,
      sprintf("%03d", 1:n()), "_", 
      Otsh_v2.0_chromosome, "_", 
      sprintf("%08d",AmpliconStartPos)),
    .before = Otsh_v2.0_chromosome
  ) %>%
  ungroup() %>%
  select(-tmpname)

# I checked and that is all good with just the change in the coho one (and renumbering).

# So, we can write it out now.
write_csv(renamed, "../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")
```

Now, let's just go ahead and record the number of amplicons of each type.
```{r}
renamed %>%
  count(LocusOrigin)
```

And make sure that comports with how we have named them all:
```{r}
renamed %>%
  extract(AmpliconName, into = "categ", regex = "Ots_([a-z]+).*$") %>%
  count(categ)
```

Yep.  That is all good.

### 2024-07-08 003-ckmr-sim-evaluation.Rmd
Rename some notebooks

--- 2024-07-07 ckmr-sim-evaluation.Rmd
@@
     final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
-    lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
     pop_labels = "inputs/reference-collection-names.csv",

### 2024-07-08 004-lfar-allele-freqs.Rmd
Rename some notebooks

---
title: "LFAR Allele Freqs"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
    pop_labels = "inputs/reference-collection-names.csv",
    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv"
  )
  # outputs:
  output_list <- list(
    lfar_tex_table = "tex/inputs/lfar-freqs.tex"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Introduction

The goal here is simply to get the allele frequenccies in different reporting units
at the LFAR markers into a table format.

First, get the baseline and attach the collection and repunit names that we want.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_csv(input_list$final_baseline)

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything())

# finally, get the order of collections and repunits we want
repunit_order <- rev(unique(pop_labels$repunit))
```

Now, whittle that down to just the LFAR and pivot it longer:
```{r}
lfar_base <- full_base2 %>%
  select(indiv:collection, starts_with("Ots_lfar")) %>%
  pivot_longer(
    cols = starts_with("Ots_lfar"), 
    names_to = c("locus", "gene_copy"),
    names_pattern = "^(.*)_([12])",
    values_to = "allele"
  )
```

And now we count up the alleles within repunits
```{r}
# first, get the multiallelic counts
lfar_cnts_MA <- lfar_base %>%
  filter(!is.na(allele)) %>%
  count(repunit, locus, allele)

# now we want to pick out the actual bases that are relevant and count the up
lfar_cnts <- lfar_cnts_MA %>%
  mutate(
    base = case_when(
      locus == "Ots_lfar001_34_00954022" ~ str_sub(allele, 1, 1),
      locus == "Ots_lfar002_34_01151770" ~ str_sub(allele, 2, 2)
    )
  ) %>%
  group_by(repunit, locus, base) %>%
  summarise(n = sum(n)) %>%
  ungroup() %>%
  complete(repunit, locus, base, fill = list(n = 0L)) %>%
  group_by(repunit, locus) %>%
  mutate(
    tot_n = sum(n),
    fract = n / tot_n
  ) %>%
  ungroup()

lfar_cnts
```

Now, we just want to focus on the alleles that are most frequent in the late-fall.
```{r}
focal <- lfar_cnts %>%
  filter(repunit == "CV-Late-Fall") %>%
  group_by(locus) %>%
  filter(n == max(n))

alle_tib <- lfar_cnts %>%
  semi_join(focal, by = join_by(locus, base)) %>%
  mutate(
    repunit_f = factor(repunit, levels = repunit_order),
    f = sprintf("%.3f", fract)
  ) %>%
  select(-repunit, -base, -n, -fract) %>%
  pivot_wider(names_from = locus, values_from = c(f, tot_n)) %>%
  arrange(repunit_f) %>%
  select(repunit_f, `f_Ots_lfar001_34_00954022`, `tot_n_Ots_lfar001_34_00954022`, `f_Ots_lfar002_34_01151770`, `tot_n_Ots_lfar002_34_01151770`)

alle_tib
```

Now, we will print this out without any column names so that we can use it in a TeX table.
```{r}
write_delim(alle_tib, delim = "&", eol = "\\tabularnewline\n", file = output_list$lfar_tex_table, col_names = FALSE)
```

And, as a final hurrah, we need to get the actual positions of those bases in Otsh_v2.0
in order to make the TeX column headers.

```{r}
locus_info <- read_csv(input_list$locus_info)

lfar_info <- locus_info %>%
  filter(str_detect(AmpliconName, "lfar"))

# for Ots_lfar001_34_00954022 the focal SNP is the first one
# for Ots_lfar002_34_01151770 the focal SNPs is the second one

# the positions in the amplicons are like this:
lfar_info %>%
  select(AmpliconName, SNPLocationsInAmpliconSequence)
```

So, 33 for the first and 66 for the second. And we can get the SNP positions
in the genome like this:
```{r}
lfar_info %>%
  mutate(
    Otsh_v2.0_snp_pos = case_when(
      AmpliconName == "Ots_lfar001_34_00954022" ~ AmpliconStartPos + 33 - 1,
      AmpliconName == "Ots_lfar002_34_01151770" ~ AmpliconStartPos + 99 - 1
    ),
    .after = Otsh_v2.0_chromosome
  )
```

So, we put those in the header of the TeX table for these.

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2024-07-08 005-plotting-genome-locations-of-markers.Rmd
Rename some notebooks

---
title: "Plotting Marker Locations in Genome"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    locus_info = "inputs/Calif-Chinook-Amplicon-Panel-Information.csv",
    chr_len = "inputs/Otsh_v2.0_chrom_lengths.tsv"
  )
  # outputs:
  output_list <- list(
    pdf = "results/genomic-locations-plot/gen-loc.pdf",
    pdfcrop = "results/genomic-locations-plot/gen-loc-crop.pdf",
    pdftex = "tex/images/genomic-locations-plot.pdf"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Introduction

We desire a simple plot for the supplement showing us the location of different markers,
colored according to their category.  Should be pretty easy.  We will use the Otsh_v2.0
coordinates.

```{r}
library(tidyverse)
library(ggchicklet)  # for geom_rrect().  Get it with remotes::install_github("hrbrmstr/ggchicklet")

amplicons <- read_csv(input_list$locus_info) %>%
  mutate(Type = LocusOrigin, .before = AmpliconName)

chroms <- read_tsv(input_list$chr_len) %>%
  filter(chrom != "MT") %>%
  mutate(chrom_dbl = as.numeric(chrom))
```

First, we want to plot the chromosomes as rounded rectangles.

```{r}
chroms2 <- chroms %>%
  mutate(
    chrom_y = 35 - chrom_dbl,
    mb = num_bases / 1e6
  )

hh <- 0.35  # the half-height of the chromosome

base <- ggplot() +
  geom_rrect(
    data = chroms2, 
    mapping = aes(ymin = chrom_y - hh, ymax = chrom_y + hh, xmin = 0, xmax = mb),
    fill = NA,
    colour = "gray"
    ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  ) +
  scale_x_continuous(breaks = seq(10, 100, by = 10)) +
  scale_y_continuous(breaks = 1:34, labels = 34:1)
```

Now, we get the markers on there
```{r}
amplicons2 <- amplicons %>%
  filter(!str_detect(Otsh_v2.0_chromosome, "NW")) %>%
  mutate(
    chrom_y = 35 - as.numeric(Otsh_v2.0_chromosome),
    mb = AmpliconStartPos / 1e6,
    Type = factor(Type, levels = c("SNPtype", "Microhap", "RoSA", "LFAR", "WRAP", "VGLL3", "Six6"))
  )

final <- base +
  geom_segment(
    data = amplicons2,
    mapping = aes(x = mb, xend = mb, y = chrom_y - hh, yend = chrom_y + hh, colour = Type),
    lineend = "butt"
  ) +
  scale_color_brewer(type = "qual", palette = "Set1") +
  xlab("Genomic coordinate (in megabases)") +
  ylab("Chromsome (in Otsh_v2.0)")

final
```

Write the thing out to a PDF file.
```{r}
ggsave(final, file = output_list$pdf, width = 7, height = 8.5)
```

Then pdfcrop it and put it where we can include it in the Supplement.
```{r}
# let's crop that down for use, too
CALL <- paste("pdfcrop ", output_list$pdf, collapse = " ")
system(CALL)

file.copy(from = output_list$pdfcrop, to = output_list$pdftex, overwrite = TRUE)
```

### 2024-07-08 002-GSI_and_Fst.Rmd
Re-ran all notebooks.  Some editing.

@@
   input_list <- list(
-    final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
-    final_baseline_apr = "data/subRoSA_baseline_rubias_APR2024_w_mixture_revised_rosa_added.csv",
-    lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
+    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
     pop_labels = "inputs/reference-collection-names.csv",
@@
   select(indiv:rosa_genotypes_str)
-
-lfar_genos <- read_csv(input_list$lfar_genos)
 
@@
 
-# long format with no missing data
+# long format with no missing data.  We filter out NA locus to get rid of the rosa_geno_str which
+# we are not using
 long_genos <- full_base2 %>%
@@
     values_to = "allele",
-names_pattern = "(.*)[_.]([12])$"
+    names_pattern = "(.*)[_]([12])$"
   ) %>%
-filter(!is.na(allele))
+  filter(!is.na(allele) & !is.na(locus))
 

### 2024-07-08 003-ckmr-sim-evaluation.Rmd
Re-ran all notebooks.  Some editing.

@@
   input_list <- list(
-    final_baseline = "data/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz",
+    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
     pop_labels = "inputs/reference-collection-names.csv",

### 2024-07-08 005-plotting-genome-locations-of-markers.Rmd
Re-ran all notebooks.  Some editing.

@@
     mb = AmpliconStartPos / 1e6,
-    Type = factor(Type, levels = c("SNPtype", "Microhap", "RoSA", "LFAR", "WRAP", "VGLL3", "Six6"))
+    Type = factor(Type, levels = c("SNPtype", "Microhap", "RoSA", "LFAR", "WRAP", "VGLL3", "Six6", "Coho"))
   )
@@
 ```
+
+# Citations
+
+::: {#refs}
+:::
+
+# Session Info
+
+```{r}
+sessioninfo::session_info()
+```
+
+# Running Time
+
+Running the code and rendering this notebook required approximately this much time:
+
+```{r}
+Sys.time() - start_time
+```

### 2024-07-08 tex/body-text-of-paper.tex
Re-ran all notebooks.  Some editing.

@@
 
-\subsubsection*{Identifying and Designing markers from the Late-Fall Associated Region}
+\subsubsection*{Identifying the Late-Fall Associated Region and designing markers therein}
 
@@
 at the locus.  In ({\bf b}), for self-assignment tallies, the row corresponds
-to the true source collection and the column corresponds to the assigned collection.
-\comment{Hey! Didn't we drop one of the LFARs? I see three, still.}}
+to the true source collection and the column corresponds to the assigned collection.}
 \begin{center}

### 2024-07-08 tex/supplemental-figures.tex
Re-ran all notebooks.  Some editing.

@@
 total numbers of alleles]{\footnotesize Number of loci with different
-total numbers of alleles in the data set.}
+total numbers of alleles in the data set. The one amplicon with only
+one allele is {\tt Ots\_coho001\_05\_32691399}, which is fixed for alternate
+in Chinook vs.\ coho salmon. It is helpful in identifying coho samples that are
+misidentified during sampling as Chinook salmon.}
 \label{fig:num-alle}

### 2024-07-08 tex/abstract.tex
Editing the paper

@@
 lineage.
-Validation of this marker panel in salmon populations not previously studied with
+Validation of this marker panel in coastal salmon populations not previously studied with
 modern genetic methods, also reveals novel biological insights, including

### 2024-07-08 tex/body-text-of-paper.tex
Editing the paper

@@
 at the forefront of a number of advances in molecular ecology,
-from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
+from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing,baetscher2018microhaplotypes}
 to statistical methodology
-\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
+\citep{smouse1990genetic,anderson2002model,pella2006gibbs,anderson2010assessing}.
 Two broadly applicable techniques that have been
@@
 \citet{milner1985genetic} soon showed that proportions of stocks in
-the Washington state coastal troll fishery could be estimated by GSI.
+the Washington state coastal troll fishery could be estimated by GSI\@.
 Since that time, with the development of novel molecular markers, and, now
@@
 spatial scale may include only a few populations from each sub-region.  Furthermore,
-baselines tailored to a specific region often assemble markers that specifically
+baselines tailored to a specific region often assemble markers that
 show allele frequency differences between the closely related---and hence difficult
@@
 be a valuable source of information when sampling of the adults is not possible.
-Accordingly, marker panels should be evaluated not only on their capacity to
-identify the parent-offspring relationships employed in PBT, but also on their
-ability to resolve full-sibling relationships.
+Accordingly, marker panels should also be evaluated on their
+capacity to resolve full-sibling relationships.
 
@@
 
-Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important fisheries \citep{myers1998status}.
-Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities during many different seasons and geographic locations \citep{healey1991life}.
-However, recent declines in population sizes, from the Yukon River in Arctic north, to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect less productive stocks \comment{(CITE)}.
-The co-occurence of fish from relatively productive and relict populations reaches its paragon in California, with the largest remaining ocean fisheries for Chinook salmon targeting the Central Valley fall-run stock that spawns in different tributaries of the same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks.
+Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important 
+fisheries \citep{myers1998status}. Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities in many 
+different seasons and geographic locations \citep{healey1991life}. However, recent declines in population sizes, from the Yukon River in Arctic north, 
+to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect 
+less productive stocks \comment{(CITE)}. The co-occurence of fish from relatively productive and relict populations reaches its paragon in California, 
+with the largest remaining ocean fisheries for Chinook salmon targeting the Central Valley fall-run stock that spawns in different tributaries of the 
+same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks \comment{(CITE)}.
 
-In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are closely related \citep{clemento2014evaluation} and share recent common ancestry that is independent of ongoing migration between these lineages.
-As such, GSI has been particularly challenging in the CCV, with at least one of the distinct ecotypes---late-fall-run---unresolvable with previous GSI marker sets \citep{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the harvested fall-run stock \comment{(CITE)}.
+In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are closely related 
+\citep{clemento2014evaluation} and share recent common ancestry that is independent of ongoing migration between these lineages. As such, GSI 
+has been particularly challenging in the CCV, with at least one of the distinct ecotypes, the late-fall-run, unresolvable with previous GSI marker sets 
+\citep{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the 
+harvested fall-run stock \comment{(CITE)}.
 
@@
 
-We sampled fish for the reference baselines from 13 locations within California, and one 
+We compiled samples of fish for the reference baselines from 13 locations within California, and one 
 within Oregon.
@@
 California's Central Valley, and the two main tributaries (Klamath and Trinity Rivers) in the Klamath basin in Northern California, and
-3 rivers of the California coast (Russian, Eel, and Smith)
+three rivers of the California coast (Russian, Eel, and Smith)
 (Fig.~\ref{fig:map}, Table~\ref{tab:samples}).
@@
 spring-run and fall-run ecotypes (i.e.
-Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype (\comment{REF?}).
+Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype \comment{(CITE)}.
 As such, sampling targeted carcasses found in each of these
 rivers at times representative of their run-timing designation, which we were able to further confirm using their genotypes
-within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}. At the time of sampling (2016), only
-fall-run Chinook salmon adults were found in the San Joaquin River, so the samples there were
+within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}.
+In the San Joaquin River, in 2016, at the time of sampling, only
+fall-run Chinook salmon adults were found, so the samples there were
 easily categorized. Run-timing 
-designation at the Feather River Hatchery is more complicated but is managed by the hatchery practices in place: 
+designation at the Feather River Hatchery is more complicated but it is managed by the hatchery practices in place: 
 early returning fish in May and April (expressing the spring-run life history) are visibly tagged and re-released 
@@
 
+The genetic markers in the baseline include newly discovered microhaplotypes, SNP assays
+that were translated into amplicons, markers associated with phenotypes, as well
+as species- and sex-specific markers.
 We assembled genetic markers for this reference baseline dataset from several different
-efforts, as detailed in the separate sections below.
+efforts, detailed in the separate sections below.
 

### 2024-07-09 tex/body-text-of-paper.tex
A boatload of editing on the paper

@@
 %%%%%%%%%%%%%%%%
-\begin{table*}
+\begin{sidewaystable*}
 \caption{\footnotesize Collections of fish in the reference baseline.  The Sampling Months column shows
@@
 \label{tab:samples}
-{\footnotesize 
+{\small 
 \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} llllllrcr}
@@
 
-{\footnotesize
+{\scriptsize
 $^1$R. = River.\\
@@
 \hrule\vspace*{0.3ex}\hrule
-\end{table*}
+\end{sidewaystable*}
 %%%%%%%%%%%%%%%%%%%%%
@@
 as species- and sex-specific markers.
-We assembled genetic markers for this reference baseline dataset from several different
-efforts, detailed in the separate sections below.
+These genetic markers were compiled from several different
+efforts, as detailed in the sections below.
 
@@
 
-We undertook a process to discover and validate  multiallelic
-microhaplotype \citep{baetscher2018microhaplotypes} markers in California Chinook salmon, with a focus on Central Valley populations.
-
-We identified candidate loci with variation among multiple populations and with multiple single-nucleotide polymorphisms using reduced representation genome sequencing of 2--4 individuals each from a number of
+We identified candidate loci with multiple single-nucleotide polymorphisms (SNPs) among multiple
+populations using reduced-representation genome sequencing of 2--4 Chinook salmon, each, from a number of
 populations ranging from California to British 
@@
 Genomic DNA was extracted using Qiagen DNeasy 96 Blood and Tissue Kits on a BioRobot 3000 
-(Qiagen, Inc). We used double-digest restriction-site associated DNA sequencing (ddRAD) and with different size selections (400 bp and 500 bp 
+(Qiagen, Inc). We used double-digest restriction-site associated DNA sequencing (ddRAD) with different size selections (400 bp and 500 bp 
 fragments) to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
@@
 
-Stacks v1.45 \cite{catchen2013stacks} was used to analyze ddRAD sequence data.
-As there was not a well assembled Chinook salmon genome at the time, a {\em de novo} assembly was performed.
-Initial filtering retained RAD loci with two or more SNPs that
-amplified in at least 10 samples. After this initial filtering, there were 3,931 RAD loci retained in the ddRAD data with 400 bp size selection and
+Stacks v1.45 \cite{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome 
+at the time, a {\em de novo} assembly was performed.  Initial filtering retained RAD loci with two or more SNPs present
+in at least 10 samples. After this initial filtering, there were 3,931 RAD loci retained in the ddRAD data with 400 bp size selection and
 4,794 retained in the data with 500 bp size selection. 
-Gene regions were then selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation in the winter-run 
-population and at least one other.
+Gene regions were then selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation amongst the winter run and at least one other population.
 A 100 bp window was used because the success of multiplex PCR and MiSeq sequencing is diminished in amplicons larger than 100 bp. Primer3 
-software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.  These gene regions were amplified and variants
+software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.
 
-FreeBayes output was filtered to remove indels and retain variants with a Quality score of 30 or greater and a read depth of 10 or higher.
-Further filtering and analysis of the loci was conducted in MICROHAPLOT (Ng et al., DOI: 10.5281/zenodo.820110).
-Loci that did not conform to Mendelian segregation, that is, loci that produced more than two haplotypes within an individual were removed.
-Hardy-Weinberg equilibrium was assessed using MICROHAPLOT.
+These gene regions were PCR-amplified and sequenced on a MiSeq sequencer using a $2\times 75$-cycle kit.
+The sequences were mapped using
+bwa-mem \citep{bwa-mem2009} to the reference sequences
+from the de novo assembly for each locus, sorted with samtools \citep{li2009sequence} and provided to
+freeyayes \citep{garrison2012haplotype} for variant calling. 
+freebayes output was filtered to remove indels and retain variants with a Phred-scaled quality score of 30
+or greater and a read depth of 10 or higher. Further filtering and analysis of the loci, including
+assessment of Hardy-Weinberg equilibrium \citep{hardy1908mendelian} was conducted in MICROHAPLOT (Ng et al., DOI: 10.5281/zenodo.820110).
+Loci that produced more than two haplotypes within an individual, as well as those clearly violating Hardy-Weinberg equilibrium within populations were removed.
 
@@
 
-A number of SNP markers from earlier discovery efforts \citep{clemento2011discovery} had already 
+A number of SNP markers from an earlier discovery effort \citep{clemento2011discovery} had already 
 been validated as useful for GSI
@@
 among the reads on the sequencer.  We 
-used Freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
+used freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
 microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by Freebayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
+that were considered multinucleotide or complex variants by freebayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
 
-\subsubsection*{Localizing markers within the genome}
-
-The SNPlicons targets and the additional microhaplotype markers were
-developed prior to the advent of a well-assembled genome for Chinook salmon.
-As a consequence, many of these markers had
-been used previously without certainty about their location in the genome.
-After designing and validating an amplicon-sequencing approach to type both the
-existing SNPtype assays and the product of the additional microhaplotype
-discovery process we identified the locations of these markers within the genome by mapping the
-consensus sequences (roughly 300 bp) around the markers
-to both the Chinook salmon version 1 (Otsh\_v1.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_002872995.1/}) and version 2  (Otsh\_v2.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/}) genome assemblies,
-using both the Burrows-Wheeler Aligner \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
-A similar mapping exercise was performed for the remaining markers so that we can
-report their locations in both Otsh\_v1.0 and Otsh\_v2.0.
-
-\subsubsection*{Identifying the Late-Fall Associated Region and designing markers therein}
+\subsubsection*{Identifying the Late-Fall Associated Region (LFAR) and designing markers therein}
 
 To identify loci with large allele frequency differences between late-fall and fall-run Chinook salmon, 
-we first conducted a simple case-control GWAS using the low-coverage whole genome sequencing 
+we first conducted a simple case-control GWAS using the low-coverage whole-genome sequencing 
 data of \citet{thompson2020complex}, mapped to Otsh\_v1.0.  We conducted
@@
 {\tt \{chrom\} } indicates the chromosome name, {\tt \{prefix\} } is the prefix to use for output files, and 
-{\tt \{bamlist\} } is the path to the text file holding paths to the BAM files, one for each individual, ordered as in {\tt \{ybin\} }.
+{\tt \{bamlist\} } is the path to the text file holding paths to the binary alignment map (BAM) files
+of aligned sequence data, one for each individual, ordered as in {\tt \{ybin\}}.
 
 There was only one peak in association scores that was shared between both comparisons (see 
-Results). From this peak we identified candidates for followup genotyping by including the 10 
-SNPs with the lowest association p-values, and supplementing those with additional SNPs that 
+Results). From this peak we identified candidates for follow-up genotyping by including the 10 
+SNPs with the lowest association $p$-values, and supplementing those with additional SNPs that 
 showed:
 \begin{itemize}
-\item An estimated absolute allele frequency difference (between late-fall and all 32 fall-run fish)  $>0.75$,  with an approximate lower confidence interval of that figure of 0.5,  or
-\item An estimated absolute frequency difference of >0.5 with a lower confidence interval >0.25 and 
+\item An estimated absolute allele frequency difference between late-fall and all 32 fall-run fish, $|d| >0.75$,  with an approximate lower confidence interval of $|d|$ greater than 0.5,  or
+\item $|d| >0.5$ with a lower confidence interval $>$0.25 and 
 with either fall-run or late-fall run being nearly fixed for one of the alleles, or
 \item An annotation from snpEff \citep{cingolani2012program} of “High” or “Moderate”, or
-\item An absolute frequency difference $>0.48$ and a lower confidence interval of $>0.25$ and a 
+\item $|d|>0.48$ and a lower confidence interval of $>$0.25 and a 
 genomic  coordinate $> 2.2$~Mb.
@@
 in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring 
-run, and 188 fall/late-fall run. We divided the fish identified as fall/late-fall via the 125-locus genetic 
-stock identification data set into fall run and late fall run according to their sampling date at Keswick: 
-99 fish arriving before Dec 26 were classified as fall run and 89 fish arriving after Dec 26 were 
+run, and 188 fall/late-fall run. We divided the fish identified as fall/late-fall (via the 125-locus genetic 
+stock identification data set of \citealt{thompson2020complex} into fall run and late-fall run according
+to their sampling date at Keswick:  99 fish arriving before Dec 26 were classified as fall run and 89 fish arriving after Dec 26 were 
 classified as late-fall run.   The allele frequencies of the target SNPs at the remaining amplicons 
@@
 
-\subsubsection*{Markers potentially associated with phenotypes \citet{thompson2020complex}}
+\subsubsection*{Winter-run associated polymorphisms (WRAPs)}
 
-Additionally, we include here genetic markers that characterize variation in two genes, VGLL3 and SIX6,
+Using the whole genome sequencing data of \citet{thompson2020complex}, we also sought
+genomic regions that might provide diagnostic markers for the the Sacramento winter-run
+Chinook salmon.  This ecotype is highly differentiated from all other Chinook salmon populations
+throughout its genome, which complicates the process of finding diverged genome regions using
+a simple association study, with only 16 samples, as done to identify the LFAR.  Consequently we pursued a bespoke
+analysis to find regions we could expect to have large allele frequency differences between
+the winter-run and all other ecotypes of Chinook salmon in California's Central Valley.   This analysis
+screened for regions with large allele frequency differences across large blocks of the genome, and is
+described in the supplemental text section \ref{sec:wrap-methods}.
+
+\subsubsection*{Run-timing associated markers}
+
+We include six amplicons that capture the eight  SNPs within the ``Region of Strongest Assocation'' (RoSA)
+between spring-run and fall-run fish on Chromosome 28 listed in Table~S3 of \citet{thompson2020complex}.  
+We add three more amplicons in the same RoSA region to the panel, as well as one more amplicon that
+allows genotyping of snp670329, which was discovered and described by \citet{thompson2019anthropogenic}.
+Including all these markers in the panel allows them to be routinely typed, creating
+ a data base for understanding the spatial and temporal
+distribution of the alleles in this genomic region; however we have omitted these markers for calculating likelihoods for
+GSI, since their distribution across populations in the Feather River has been clearly and evidently
+influenced by hatchery practices (see Results), it is not necessary to include them for accurate assignment,
+and because the 12 SNPs within these three amplicons are typically in profound linkage disequilibrium (LD)
+making them inappropriate for use as separate markers in most software for population assignment or GSI, that
+typically assumes markers are not in LD within the component populations of the baseline.
+
+\subsubsection*{Additional phenotypically associated markers}
+
+The baseline also includes three amplicons with genetic variation
+in two genes, VGLL3 and SIX6,
 that have been found to be associated with phenotypes related to reproductive maturity in both
@@
 
-We include in the reference baseline amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish on Chromosome Ots28 described in \citet{thompson2020complex}.  These are useful markers to
-have in the panel, as typing them routinely will create a data base for understanding the spatial and temporal
-distribution of the alleles at this region; however we have omitted these markers for calculating likelihoods for
-GSI, since their distribution across populations in the Feather River has been clearly and evidently
-influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.
-\comment{Describe additional E1/E2 markers (additional markers within the RoSA region) and
-Tasha Marker (because it is close to the duplication noted
-by Thompson et al.). We don't go into detail here---keeping it vague because it is
-going to be described in the subRoSA paper, but don't mention that.}
+Finally, we include a marker that targets the sex-determining region ({\em sdY} on the Y chromosome in Chinook salmon.
+This marker characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}.
 
-We also include a marker that targets the sex-determining region on the Y chromosome in Chinook salmon
-and that characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}
+\subsection*{Localizing markers within the genome}
 
-\subsection*{Amplicon sequencing and microhaplotype calling}
+The novel microhaplotype markers and the SNPlicon targets were
+developed prior to the advent of a well-assembled genome for Chinook salmon.
+As a consequence, many of these markers had
+been used previously without certainty about their location in the genome.
+After designing and validating an amplicon-sequencing approach to type both of these
+groups of markers we identified their locations within the genome by mapping the
+consensus sequences (roughly 100--300 bp) around the markers
+to both the Chinook salmon version 1 (Otsh\_v1.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_002872995.1/}) and version 2  (Otsh\_v2.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/}) genome assemblies,
+using both bwa-mem \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
+A similar mapping exercise was performed for the remaining markers so that we can
+report their locations in both Otsh\_v1.0 and Otsh\_v2.0.
 
-Amplicon sequencing was performed with the Genotyping-in-Thousands by sequencing (GT-seq) method from Campbell et al. (2015). All loci were 
-multiplexed with primer concentrations ranging from 0.083 uM to 0.5 uM to increase uniformity of read depths across all loci. Sample normalization was 
-performed using the SequalPrep Normalization Plate Kit (Applied Biosystems) according to the manufacturer's instructions. Samples were pooled by plate 
-and then purified using Agencourt AMPure XP magnetic beads. Libraries were quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New 
-England BioLabs) before dilution and pooling for sequencing. All samples were sequenced on a MiSeq using $2\times 75$-cycle paired-end sequencing protocols. 
+\subsection*{Amplicon sequencing, microhaplotype calling, and further quality control}
+
+Amplicon sequencing was performed with the Genotyping-in-Thousands by sequencing (GT-seq)
+method \citep{campbell2015genotyping}. All loci were 
+multiplexed with primer concentrations ranging from 0.083 uM to 0.5 uM to increase
+uniformity of read depths across all loci. Sample normalization was 
+performed using the SequalPrep Normalization Plate Kit (Applied Biosystems) according
+to the manufacturer's instructions. Samples were pooled by plate 
+and then purified using Agencourt AMPure XP magnetic beads.
+Libraries were quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New 
+England BioLabs) before dilution and pooling for sequencing.
+All samples were sequenced on a MiSeq using $2\times 75$-cycle paired-end sequencing protocols. 
 Primer sequences and concentrations for making the primer pool are in Supplemental Table 1.
 
-For evaluation of candidate loci, dual-barcoded sequences were used to identify individual tissue samples using the MiSeq analysis software (Illumina).
-The paired-end sequence reads were joined together using Fast Length Adjustment of SHort reads (FLASH: Magoc and Salzberg 2011) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
+For evaluation of candidate loci, dual-barcoded sequences were used
+to identify individual tissue samples using the MiSeq analysis software (Illumina).
+The paired-end sequence reads were joined together using Fast Length Adjustment
+of SHort reads (FLASH: \citealt{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
 Mapped reads were converted to BAM files using SAMtools (Li et al.
-2009) and FreeBayes (Garrison and Marth 2012) was used to call variants.
+2009) and freebayes (Garrison and Marth 2012) was used to call variants.
 
-For genotyping of project samples, final filtering at each locus required a read depth of 20 and a read depth ratio of 0.30 (i.e., for heterozygotes the lower read depth haplotype must have 30\% of the total reads at that locus).
+For genotyping of baseline samples, final filtering at each locus required a read depth of 20 and a read depth ratio of 0.30 (i.e., for heterozygotes the lower read depth haplotype must have 30\% of the total reads at that locus).
 MICROHAPLOT was used to call alleles and genotype processing was performed with R scripts.
 
-Before parentage analysis began each locus was assessed for the presence of null alleles using R scripts (eric citation).
-Twelve loci appeared to have null alleles in the winter-run population and were removed in all subsequent analyses.
-Additionally, two loci were monomorphic in the winter-run, but contained multiple alleles in other populations.
-The two monomorphic loci were removed in all winter-run analyses resulting in a final dataset of genotypes for 111 microhaplotype loci for this project.
-For more diverse salmon populations, it's likely that nearly all of the 125 microhaplotype loci could be used without Hardy-Weinberg Equilibirum or null allele issues.
+Each locus was assessed for the presence of null alleles using the R package `whoa' 
+(\url{https://github.com/eriqande/whoa}; {\em see} \citealt{hendricks2018recent}).
 
@@
 
-Basic data summaries were calculated for each collection
+Basic data summaries were calculated for each population/collection
+in the reference baseline
 using simple operations in the R programming language, Version 4.3
@@
 found at each locus.  To provide a summary of the genetic variation and
-missing data we report:
+missing data we further summarized:
 \begin {itemize}
@@
 all collections, using the {\footnotesize\tt pp.fst} function from the R
-package 'hierfstat' \citep{hierfstat}. Finally, we evaluated population
+package `hierfstat' \citep{hierfstat}. Finally, we evaluated population
 structure in the data by performing unsupervised clustering of
@@
 using default settings of the program.  STRUCTURE output was
-a summarized and visualized using CLUMPAK version~\comment{??}
+a summarized and visualized using CLUMPAK
 \citep{kopelman2015clumpak}.
@@
 To assess the power of the marker panel for GSI, we used the R package
-'rubias' \citep{moran2019bayesian}.  We used the function {\tt self\_assign()}
+`rubias' \citep{moran2019bayesian}.  We used the function {\tt self\_assign()}
 to perform assignment of each individual in the reference baseline to one of the
@@
 fish were assigned to the collection within that reporting unit with the highest
-scaled likelihood.  We considered two criteria for assignment.  In the first,
+scaled likelihood.  We considered two thresholds for assignment.  In the first,
 all fish are assigned, in the second only fish with a summed (i.e., summed
@@
 broken down according to the genotype of the fish at the markers within
-the RoSA.  For this, \comment{Eric. Describe this from what you have in the notebook}.
+the RoSA.  For this we identified the RoSA genotype (EE, EL, or LL, see
+\citealt{thompson2020complex}) from the concordance of the genotypes
+at all 8 markers described in \citep{thompson2020complex}, requiring no missing
+data at any of those 8 markers, and discarding data from 12 individuals
+in which the E or L haplotypes had clearly recombined.
 
@@
 collections into 10 different groups.  The members of each group are
-relatively similar For all cases, we assumed that the genotyping error rate
+relatively, genetically similar.  For all cases, we assumed that the genotyping error rate
 was 1\% per locus and the genotyping error model was the default
@@
 of the physical linkage of the markers by inputting their genomic positions into
-'CKMRsim' and using the MENDEL \citep{lange2013mendel} software to simulate genotypes of related pairs
+`CKMRsim' and using the package's interface to the MENDEL \citep{lange2013mendel}
+software to simulate genotypes of related pairs
 in the presence of physical linkage.  For PO pairs (and U pairs),
@@
 we used regular Monte Carlo to estimate the FPRs and FNRs, because
-importance sampling cannot be used in those cases, and, since there are
+importance sampling cannot be used in those cases. Since there are
 expected to be far fewer such relationships than the number of unrelated
-pairs expected, the relevant FPRs are not so tiny and can be estimated
-(at least down to rates of $10^{-3}$) reliably.  For AN, FS, and HS, both the
+pairs, the relevant FPRs are large enough that they can be reliably estimated
+(at least down to rates of $10^{-3}$) without importance sampling.
+For AN, FS, and HS, both the
 FPRs and FNRs were estimated while taking account of physical linkage
@@
 
-\subsubsection*{Conversion of existing SNPtype assays}
-
-Of the 96 existing SNPtype assays, we were able to convert 79
-into reliable amplicon-sequenced assays that were included in the panel.
+A summary, which includes genomic location, consensus sequence, and primer sequences,
+of all the amplicons used in the baseline is available
+in Data Supplement~1.
 
@@
 
-We tested 229 candidates from our additional microhaplotype discovery and
+We tested the 229 candidates from our novel microhaplotype discovery process and
 initially retained the 125 markers described in \citep{thompson2020complex}
@@
 the Klamath Basin.  However, further evaluation of those markers in other
-populations led us to drop some, and retain 106 of those markers in the current panel.
+populations led us to remove some of the markers.
+Twelve loci appeared to have null alleles in the winter-run population and were removed.
+Additionally, two loci were monomorphic in the winter-run, even though they contained
+multiple alleles in other populations, and an additional 5 amplicons were deemed
+difficult to score in routine application of the baseline, and were removed, leaving 106
+novel microhaplotype amplicons in the current panel.
 
-\subsubsection*{Localizing markers within the genome}
+\subsubsection*{Conversion of existing SNPtype assays}
 
-Of these \comment{185} loci that originated from SNPtype assays or from our additional
-microhaplotype discovery, \comment{179} of them mapped identically to a single
-location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
-Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
-the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
-that was identical in both bwa mem and BLAT, two of them had secondary alignments that
-disagreed slightly between BLAT and bwa mem, and one of them had multiple
-non-primary alignments (Table~S1 \comment{This is the Excel File.  It needs to
-be edited so as to: 1) have a column
-giving the category or provenance of the markers (i.e. LFAR or MHAP or SNPTYPE, etc.) 2) retain
-only the 2 LFAR markers we actually use. 3) Definitely rename the E2 markers to be something other
-than winter-run-greb1L-diagnostic}).
-
-Mapping to Otsh\_v1.0 was similar, except that a greater number of markers mapped to unplaced
-scaffolds in Otsh\_v1.0 than in Otsh\_v2.0, likely reflecting the more complete nature of the
-second assembly.
+Of the 96 existing SNPtype assays, we were able to convert 78
+into reliable amplicon-sequenced assays that were included in the panel.
 
@@
 $p$-values.  These are not shared between the two different late-fall to fall comparisons, and
-sites that we investigated were concentrated in areas of poor mapping suggesting that their
+sites that we investigated were concentrated in areas of poor mapping, suggesting that their
 $p$-values were artifactual.  By contrast, a large number of sites near the prominent peak on
@@
 etc.), making it costly (in terms of sequencing effort) to include the marker in the baseline panel. 
-\comment{Consequently it was dropped.}  The frequency of the different alleles at the two remaining loci
+Consequently it was dropped from the panel.  The two remaining loci, which are at coordinates
+Chr34:865,057, and Chr34:1,063,084 in the Otsh\_v1.0 assembly are at coordinates
+Chr34:954,054 and  Chr34:1,151,868 in the Otsh\_v2.0 assembly, and
+the frequency of the different alleles at the two remaining loci
 within different reporting units in the baseline are shown in Table~\ref{tab:lfar-freqs}.
@@
 %%%%%%%%%%%%%%%
+
+\subsection*{Localizing markers within the genome}
+
+Of the 184 loci that originated from our novel
+microhaplotype discovery or from SNPtype assays, 179 of them mapped identically to a single
+location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
+Of the remaining 5 loci, one of them was mapped to the same location by BLAT and bwa mem, but
+the alignments differed in length, two of them had a single secondary alignment on a different chromosome
+that was identical in both bwa mem and BLAT, one of them had only a fragment mapping to
+the secondary alignment, and one of them had multiple
+non-primary alignments (Supp.~Data~1).
+
+Mapping to Otsh\_v1.0 was similar, except that a greater number of markers mapped to unplaced
+scaffolds in Otsh\_v1.0 than in Otsh\_v2.0, likely reflecting the more complete nature of the
+second assembly.
+
 \subsection*{Population genetic summaries}

### 2024-07-09 002-GSI_and_Fst.Rmd
Add reporting units to Table 1

@@
 fbmet4 <- fbmet3 %>%
-  left_join(map_notations %>% select(abbrv, repository_name, map_text, basin, ESU, `Mainstem River` ))
+  left_join(map_notations %>% select(abbrv, repository_name, map_text, basin, ESU, `Mainstem River` )) %>%
+  left_join(pop_labels %>% select(collection, repunit), by = join_by(abbrv == collection))
 
@@
   ) %>%
-  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, ESU, N, N_date, `Sampling Months`, `Year Range`) %>%
+  select(`Mainstem River`, abbrv, map_text, run_timing_for_table, repunit, ESU, N, N_date, `Sampling Months`, `Year Range`) %>%
   mutate(map_text = str_replace(map_text, ",.*$", "")) %>% # this removes the abbreviations after the commas on the map names
@@
     Run = run_timing_for_table,
-    `ESU$^3$` = ESU,
+    `Reporting Unit$^4$` = repunit,
+    `ESU$^4$` = ESU,
     `$N$` = N,
-    `$N_\\mathrm{day}^4$` = N_date
+    `$N_\\mathrm{day}^5$` = N_date
   ) %>%

### 2024-07-09 tex/body-text-of-paper.tex
Add reporting units to Table 1

@@
 {\small 
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} llllllrcr}
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} lllllllrcr}
 \hline\hline\vspace*{0.4ex}
@@
 $^2$Ck. = Creek; NFH = National Fish Hatchery; R. = River; H. = Hatchery.\\
-$^3$ESU = Evolutionarily Significant Unit; CVFR = Central Valley Fall Run ESU; CVSR = Central Valley Spring Run ESU; CVWR = 
+$^3$Reporting Unit = an aggregation of the collections into groups relevant to management and
+which are accurately resoluble with the genetic data in the baseline.
+$^4$ESU = Evolutionarily Significant Unit; CVFR = Central Valley Fall Run ESU; CVSR = Central Valley Spring Run ESU; CVWR = 
 Central Valley Winter Run ESU; CC = California Coastal ESU; UKTR = Upper Klamath and Trinity Rivers ESU. \\
-$^4N_\mathrm{day}$: the number of samples for which the specific sampling day was recorded.
+$^5N_\mathrm{day}$: the number of samples for which the specific sampling day was recorded.
 }

### 2024-07-09 tex/inputs/samples-table.tex
Capitalize C in NCal in repunit names

--- 2024-01-07 tex/inputs/samples-table.tex
@@
-Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&Sampling Months&Year Range\tabularnewline
-\hline Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
-Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
-Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
-Sacramento R.&MDF&Deer Ck.&Fall&CVFR&29&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
-Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
-Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
-Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
-Sacramento R.&MDF&Mill Ck.&Fall&CVFR&53&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
-Sacramento R.&MDS&Mill Ck.&Spring&CVSR&51&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
-Russian R.&RRF&Warm Springs H.&Fall&CC&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
-Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
-San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
-Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
-Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Mainstem River$^1$&Code&Location Name$^2$&Run&Reporting Unit$^4$&ESU$^4$&$N$&$N_\mathrm{day}^5$&Sampling Months&Year Range\tabularnewline
+\hline Rogue R.&CRHS&Cole Rivers H.&Spring&SO-NCal-Coast&SONCC&50&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CRHS_Cole-Rivers-Hatchery.pdf}}&2004--2010\tabularnewline
+Smith R.&SmRF&Smith R.&Fall&SO-NCal-Coast&SONCC&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SmRF_Smith-River.pdf}}&2010--2011\tabularnewline
+Klamath R.&BCkF&Blue Ck.&Fall&SO-NCal-Coast&SONCC&56&56&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCkF_Blue-Creek.pdf}}&2008--2008\tabularnewline
+Klamath R.&IGHF&Iron Gate H.&Fall&Klamath-Trinity&UKTR&93&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-IGHF_Iron-Gate-Hatchery.pdf}}&2017--2017\tabularnewline
+Trinity R.&TRS&Trinity R. H.&Spring&Klamath-Trinity&UKTR&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Trinity R.&TRF&Trinity R. H.&Fall&Klamath-Trinity&UKTR&41&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
+Eel R.&ERF&Eel R.&Fall&Cent. Cal. Coast&CCC&93&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-ERF_Van-Arsdale-Fisheries-Station.pdf}}&2014--2014\tabularnewline
+Russian R.&RRF&Warm Springs H.&Fall&Cent. Cal. Coast&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
+Sacramento R.&SRW&Livingston Stone NFH&Winter&CV-Winter&CVWR&111&20&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
+Sacramento R.&BCS&Butte Ck.&Spring&CV-Spring&CVSR&116&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
+Sacramento R.&MDS&Deer Ck.&Spring&CV-Spring&CVSR&34&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
+Sacramento R.&MDS&Mill Ck.&Spring&CV-Spring&CVSR&50&50&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&FRHS&Feather R. H.&Spring&CV-Fall&CVSR&202&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
+Sacramento R.&FRHF&Feather R. H.&Fall&CV-Fall&CVFR&106&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
+Sacramento R.&BCF&Butte Ck.&Fall&CV-Fall&CVFR&45&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
+Sacramento R.&MDF&Deer Ck.&Fall&CV-Fall&CVFR&24&24&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&MDF&Mill Ck.&Fall&CV-Fall&CVFR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
+San Joaquin R.&SJRF&San Joaquin R.&Fall&CV-Fall&CVFR&79&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
+Sacramento R.&CHLF&Coleman NFH&Late-fall&CV-Late-Fall&CVFR&261&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline

### 2024-07-09 R/colors.R

@@ -1,7 +1,7 @@
 
 # These are for the genetic-marker-resolvable genomic backgrounds
 repunit_colors <- c(
-  `SO-Ncal-Coast` = "beige",
+  `SO-NCal-Coast` = "beige",
   `Klamath-Trinity` = "#6e8b3d",
   `Cent. Cal. Coast` = "#79cdcd",
   `CV-Winter` = "#ffff99",            # a "cool-ish" color for winter

### 2024-07-10 tex/inputs/samples-table.tex
Change name of Cent. Cal. Coast to have hyphens

@@
 Trinity R.&TRF&Trinity R. H.&Fall&Klamath-Trinity&UKTR&41&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
-Eel R.&ERF&Eel R.&Fall&Cent. Cal. Coast&CCC&93&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-ERF_Van-Arsdale-Fisheries-Station.pdf}}&2014--2014\tabularnewline
-Russian R.&RRF&Warm Springs H.&Fall&Cent. Cal. Coast&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
+Eel R.&ERF&Eel R.&Fall&Cent-Cal-Coast&CCC&93&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-ERF_Van-Arsdale-Fisheries-Station.pdf}}&2014--2014\tabularnewline
+Russian R.&RRF&Warm Springs H.&Fall&Cent-Cal-Coast&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
 Sacramento R.&SRW&Livingston Stone NFH&Winter&CV-Winter&CVWR&111&20&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline

### 2024-07-10 R/colors.R

@@ -3,7 +3,7 @@
 repunit_colors <- c(
   `SO-NCal-Coast` = "beige",
   `Klamath-Trinity` = "#6e8b3d",
-  `Cent. Cal. Coast` = "#79cdcd",
+  `Cent-Cal-Coast` = "#79cdcd",
   `CV-Winter` = "#ffff99",            # a "cool-ish" color for winter
   `CV-Spring` = "#99CC33",             # green like spring
   `CV-Fall` = "#F06292",

### 2024-07-10 tex/body-text-of-paper.tex
Capitalize C in NCal in repunit names
continue editing paper.

@@
 
-Finally, we include a marker that targets the sex-determining region ({\em sdY} on the Y chromosome in Chinook salmon.
+We include a marker that targets the sex-determining region ({\em sdY} on the Y chromosome in Chinook salmon.
 This marker characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}.
+And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
+coho salmon and Chinook salmon.  This is helpful when samplers have unwitting
+collected coho salmon from ocean catches.
 
@@
 
+\subsubsection*{WRAP}
+
+\comment{Gotta add the WRAP discovery and filtering stuff.}
+
 \subsection*{Localizing markers within the genome}
@@
 
-The total number of alleles per locus varied between 2 and 10 (Fig.~\ref{fig:num-alle}). Biallelic 
+The total number of alleles per locus varied between 1 and 10 (Fig.~\ref{fig:num-alle}), with the
+only locus with a single allele being the species-specific marker Ots\_coho001\_05\_32691399
+used to distinguish coho from Chinook salmon. Biallelic
 loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
@@
 The average number of fish with non-missing data in each population varied from
-a low of 39.2 (BCF) to a high of 275.2 (CHLF).  The average number of alleles
-per locus, standardized to the smallest sample size, ranged from 2.35 (SRW) to
-2.68 (RRF), with  SRW also having the smallest value of the
-standardized fraction of polymorphic markers at 0.9 and RRF sharing the highest value
-(0.97) with ERF and BCkF.
+a low of 40.2 (TRF) to a high of 279.6 (CHLF).  The average number of alleles
+per locus, standardized to the smallest sample size, ranged from 2.32 (SRW) to
+2.64 (RRF), with  SRW also having the smallest value of the
+standardized fraction of polymorphic markers at 0.88 and RRF sharing the highest value
+(0.96) with FHRF and BCkF.
 SRW also had the lowest values for expected and observed heterozygosity:
-0.349 and 0.340, respectively. While the highest for expected heterozygosity
-was 0.412 (SmRF) and for observed heterozygosity, 0.411 (CRHS).
+0.338 and 0.329, respectively. While the highest for expected heterozygosity
+was 0.401 (SmRF) and for observed heterozygosity, 0.399 (CRHS).
 
 The estimated pairwise $F_\mathrm{ST}$ values between the collections
-varied from a low of 0.0 to a high of 0.273 (Figure~\ref{fig:gsi}a).
+varied from a low of 0.0 to a high of 0.288 (Figure~\ref{fig:gsi}a).
 Notably, most of the largest values
 of $F_\mathrm{ST}$ occurred between SRW and another collection.
-
-The program STRUCTURE clustered fish at all values of $K$ from 2 to \comment{7}.
-(Fig.~\ref{fig:struct}).  The clusters generally corresponded to groupings of related populations, and confirmed
-{\em a priori} knowledge about which collections in the baseline could and should be
-grouped together into reporting units.  At every value of $K$ CLUMPAK discerned at least
-11 of 20 replicates in the major mode (Fig.~\ref{fig:struct}). Clustering solutions in the minor
-modes appear in Fig.~\ref{fig:minor-modes}.
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-\newcommand{\structcap}{\footnotesize Major modes found by STRUCTURE and summarized
-by CLUMPAK out of 20 replicates for each value of $K$, the assumed number of
-subpopulations. \comment{We kind of need to make our own figure here that doesn't take up
-as much whitespace and also gives the number of occurrences of each mode.} }
-\begin{center}
-\includegraphics[width=\columnwidth]{images/major-modes-crop.pdf}
-\end{center}
-\caption[\structcap]{\structcap}
-\label{fig:struct}
-\end{figure}
-
-\subsection*{Power for genetic stock identification and population assignment}
-
-Leave-one-out cross-validation (self-assignment) analysis demonstrates that this
-reference baseline provides a high degree of accuracy for distinguishing the major
-groups of Chinook salmon in California.  The results of the analysis are summarized
-in a self-assignment matrix (Figure~\ref{fig:gsi}). 
 %%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%
-This figure shows the collections divided into seven different reporting groups:
-SO-Ncal-Coast,  Klamath-Trinity, Cent. Cal. Coast,
-CV-Winter, CV-Spring, CV-Fall, and  CV-Late-Fall.  \comment{Really, where and how
-do we want to introduce these reporting units?} These groupings correspond
-largely to the divisions in the data found by STRUCTURE, and they also correspond
-with divisions amongst the stocks as defined for management.
+
+The program STRUCTURE clustered fish at all values of $K$ from 2 to 7.
+(Fig.~\ref{fig:struct}).  The clusters generally corresponded to groupings of related populations, and confirmed
+{\em a priori} knowledge about which collections in the baseline could and should be
+grouped together into reporting units.  At every value of $K$ CLUMPAK discerned \comment{at least
+11 of 20 replicates} in the major mode (Fig.~\ref{fig:struct}).   
+\comment{Need to add more description of the patterns, especially at $K=7$.}
+\comment{Clustering solutions in the minor
+modes appear in Fig.~\ref{fig:minor-modes}, which likely needs to be updated.  Get all this done
+when AC is back and can vet it.}
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\newcommand{\structcap}{\footnotesize Major modes found by STRUCTURE and summarized
+by CLUMPAK out of 20 replicates for each value of $K$, the assumed number of
+subpopulations. }
+\begin{center}
+\includegraphics[width=\columnwidth]{images/clumpak-crop.pdf}
+\end{center}
+\caption[\structcap]{\structcap}
+\label{fig:struct}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection*{Power for genetic stock identification and population assignment}
+
+Leave-one-out cross-validation (self-assignment) analysis demonstrates that this
+reference baseline provides a high degree of accuracy for distinguishing the major
+groups of Chinook salmon in California.  The results of the analysis are summarized
+in a self-assignment matrix (Figure~\ref{fig:gsi}b).
+This figure shows the collections divided into the seven different reporting groups:
+SO-NCal-Coast,  Klamath-Trinity, Cent.\ Cal.\ Coast,
+CV-Winter, CV-Spring, CV-Fall, and  CV-Late-Fall.  These groupings correspond
+largely to the divisions in the data found by STRUCTURE, they also correspond
+with divisions amongst the stocks as defined for management, and, as seen in the
+figure, they correspond with groups of populations that can be reliably distinguished
+from one another for population assignment.
 

### 2024-07-10 002-GSI_and_Fst.Rmd
Done with major editing push

@@
 no_lfar_ass_table_80 <- top_ass$top %>% 
-  filter(scaled_likelihood > 0.8) %>%
+  filter(repu_sclike > 0.8) %>%
   plot_self_assignment_table(pop_labels, collection_order)
@@
 lfar_ass_table_80 <- lfar_top_ass$top %>% 
-  filter(scaled_likelihood > 0.8) %>%
+  filter(repu_sclike > 0.8) %>%
   plot_self_assignment_table(pop_labels, collection_order)
@@
 
+Look at the distribution there:
 ```{r}
-over80 <- nrow(lfar_top_ass$top %>% filter(repu_sclike > 0.8))
+lfar_top_ass$top %>%
+  filter(repunit != inferred_repunit) %>%
+  count(collection, inferred_repunit)
+```
 
-lfar_top_ass$top %>% filter(repu_sclike > 0.8) %>% count(repunit == inferred_repunit) %>% mutate(fract = n / over80)
+Get some more values for the paper:
+
+From CHLF to Fall
+```{r}
+lfar_top_ass$top %>%
+  filter(collection == "CHLF") %>%
+  count(repunit == inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+
+```
+
+From Fall to CHLF
+```{r}
+lfar_top_ass$top %>%
+  filter(repunit == "CV-Fall") %>%
+  count(repunit,inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+
+```
+
+```{r}
+lfar_top_ass$top %>%
+  filter(collection == "MDS") %>%
+  count(repunit,inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+
+```
+
+#### Now, look at the >80% criterion.
+```{r}
+# first, the number and fraction of fish that are still assigned
+lfar_top_ass$top %>% 
+  mutate(gt80 = repu_sclike > 0.8) %>%
+  count(gt80, repunit == inferred_repunit)
+```
+
+```{r}
+lfar_top_80 <- lfar_top_ass$top %>% 
+  filter(repu_sclike > 0.8)
+
+lfar_top_80 %>%
+  count(repunit == inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+```
+
+Drop the clear winter run strays
+```{r}
+
+lfar_top_80 %>%
+  filter(!((inferred_collection == "SRW") & (inferred_repunit != repunit))) %>%
+  count(repunit == inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+```
+
+And particularly look at misassignments to the CHLF and CV-Spring
+
+From Fall to other repunits
+```{r}
+lfar_top_80 %>%
+  filter(repunit == "CV-Fall") %>%
+  count(repunit,inferred_repunit) %>%
+  mutate(fract = n / sum(n))
+
 ```

### 2024-07-10 tex/body-text-of-paper.tex
Done with major editing push

@@
 \begin{figure}
-\newcommand{\gsicap}{\footnotesize Pairwise $F_\mathrm{ST}$ values and allele frequencies({\bf a)}
+\newcommand{\gsicap}{\footnotesize Pairwise $F_\mathrm{ST}$ values and allele frequencies ({\bf a)}
 and self-assignment rates
@@
 of the collection.  Interior cells in which the column and row correspond to the same reporting
-unit are colored according to the reporting unit.  In ({\bf a}), for $F_\mathrm{ST}$ values, the upper triangle was computed using the baseline microhaplotype
-data. Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
+unit are colored according to the reporting unit.  In ({\bf a}), for $F_\mathrm{ST}$ values,
+the upper triangle was computed using all the genetic markers except for those in the
+RoSA on chromosome 28. 
+Each cell in the lower triangle shows a small scatter plot of estimated allele frequencies
 in the collections---the $x$ value is allele frequencies from the collection in the column and the
@@
 at the locus.  In ({\bf b}), for self-assignment tallies, the row corresponds
-to the true source collection and the column corresponds to the assigned collection.}
+to the true source collection and the column corresponds to the assigned collection.
+All individuals were assigned here, using a maximum-scaled-likelihood rule to reporting unit and
+then to collection within reporting unit as described in the text. For the assignment matrix resulting
+when only fish with a scaled likelihood $>0.80$ to reporting unit are assigned, see Figure~\ref{fig:eighty}}
 \begin{center}
@@
 This figure shows the collections divided into the seven different reporting groups:
-SO-NCal-Coast,  Klamath-Trinity, Cent.\ Cal.\ Coast,
+SO-NCal-Coast,  Klamath-Trinity, Cent-Cal-Coast,
 CV-Winter, CV-Spring, CV-Fall, and  CV-Late-Fall.  These groupings correspond
@@
 
-Overall, of the 1613 fish in the reference baseline, 1587 (98.4\%) were correctly
-assigned to their reporting unit of origin, while 26 (1.6\%) were incorrectly assigned.
-Four of those 26 misassignments which involved fish being incorrectly assigned to
+Overall, of the 1636 fish in the reference baseline, 1612 (98.5\%) were correctly
+assigned to their reporting unit of origin, while 24 (1.5\%) were incorrectly assigned.
+Four of those 24 misassignments which involved fish being incorrectly assigned to
 SRW, were clearly fish from SRW that had been incorrectly sampled into a different
 collection, likely due to straying. (This is evident because SRW are so distinct that it is not possible
-to incorrectly identify them).  Of the remaining 22 misassignments,  9 involved fish from the
-CHLF (late fall run) collection being misassigned to the CV-Fall reporting unit and 7 were
+to incorrectly identify them).  Of the remaining 20 misassignments,  8 involved fish from the
+CHLF (late fall run) collection being misassigned to the CV-Fall reporting unit and 6 were
 fish from the MDS (Mill-Deer spring run) collection being assigned to the CV-Fall reporting unit.
 It is possible that some of these misassignments represent CV-Fall fish that were incorrectly
-sampled as CHLF or MDS; however, unlike with SRW, that we could not conclude that
+sampled as CHLF or MDS; however, unlike with SRW,  we  cannot conclude that
 confidently, as there is some degree of overlap in likelihood values between the two.
 
-While fish from CHLF were found to misassign at a rate of 3.1\% (9/291) and MDS at a rate
-of 8.6\% (7/81) to CV-Fall, are seen to misassign at fairly high rates to CV-Fall, the
+While fish from CHLF were found to misassign at a rate of 2.7\% (8/300) and MDS at a rate
+of 7.1\% (6/84) to CV-Fall, the
 converse is not evident---that is, fish from the CV-Fall are misassigned to CHLF at a rate
-of only 0.6\% (3/501) and to the CV-Spring reporting unit (to which MDS belongs) at a rate
-of only 0.4\% (2/501).
+of only 0.6\% (3/508) and to the CV-Spring reporting unit (to which MDS belongs) at a rate
+of only 0.4\% (2/508).
 
 When assignments are only accepted with a scaled likelihood greater than 0.8,
-18 assignments are discarded, but not all of those were incorrect.  At the 0.8 threshold,
-the total fraction of correct assignments is 98.8\% (Fig.~\ref{fig:eighty}).
+20 assignments are discarded, and half of those (10) were the incorrect assignments.  Thus,
+at the 0.8 threshold, the total fraction of correct assignments is 99.1\% (Fig.~\ref{fig:eighty}).
+Notably with the $>0.8$ criterion, the misassignment rates from the abundant CV-Fall
+reporting unit to CV-Late-Fall or to CV-Spring both drop below 0.2\% (1/502).
 
-We note that the RoSA markers provide the differential signal necessary to distinguish the
-Feather River spring-run stock from the other CCV stocks with a predominantly fall-run
-genomic background, as they do not share genome-wide ancestry with the other spring-run
-stocks/lineages in the basin. In this context, it is clear that the 'fall-run' hatchery
+As seen in previous studies, FRHS, the Feather River Hatchery
+spring-run stock, is considerably differntiated from the remaining spring-run
+stocks in the Central Valley (MDS and BCS). This is reflected in the low
+misassignment rates of FRHS fish to the CV-Spring, and is also evident in the
+STRUCTURE results.   We note, however, that the RoSA markers provide
+the differential signal necessary to distinguish
+Feather River spring-migrating fish from the other CCV stocks with a predominantly fall-run
+genomic background. This is evident when the assignment matrix is enumerated in
+terms of RoSA genotypes (Figure~\ref{fig:rosa-gsi}). In this context, it is clear that the ``fall-run'' hatchery
 program in the Feather River both spawns and produces many spring-run fish and RoSA heterozygotes.
-Similarly, in other regions, the RoSA markers are the only way to distinguish fall-
+Similarly, in other regions, such as the Trinity River, the RoSA markers are the only reliable means to distinguish fall-
 and spring-run fish, as they otherwise share genomic background.
+
 In California, only the Sacramento and  Klamath basins have documented spring-run
@@
 recombinant haplotype that carries half of the RoSA SNPs from both the E and L lineages in the Eel and Russian rivers.
-No other clear recombinants were identified in the study populations, indicated that it arose in the California Coastal Chinook Salmon lineage, and provides further evidence of past presence of RoSA haplotypes associated with early migration.
+No other clear recombinants were identified in the study populations, indicating that it arose in the California Coastal Chinook Salmon lineage, and providing further evidence of past presence of RoSA haplotypes associated with early migration.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-And we can also point readers to the supplement figure \ref{fig:ckmr-comp}.
-The power analyses for this set of markers demonstrate that they have ample variation
+
+The `CKMRsim' power analyses for this set of markers demonstrates that they have ample variation
 for accurate identification of parent-offspring (PO) and full-sibling (FS) pairs in almost all
-realistic situations. The predicted false positive rates for both PO and FS pairs are
-exceedingly low for even stringent false negative thresholds (e.g. 0.05).
+realistic situations. The predicted false positive rates from unrelated pairs for both the
+parent-offspring (PO) and full-sibling (FS) relationships are
+exceedingly low for all population groups, even when using stringent false negative thresholds as
+low as 0.05 (Fig.~\ref{fig:fprs}a).
 For example, the false positive rate for FS identification is less than one error in 10 million
 comparisons of potential siblings in the California Chinook salmon population with the lowest
-heterozygosity (SRW) at a false negative rate of 0.05. However, HS pairs can not be accurately
+heterozygosity (SRW) at a false negative rate of 0.05.
+
+Half-sibling (HS) pairs cannot be accurately
 distinguished from unrelated pairs in any populations without very large false negative rates
-and in very modest-sized studies.
-We note, however, that the presence in a dataset of genotypes from individuals that are related, but not in
- the relationship category proposed, will result in some additional misidentifications.
- These 'errors' will all involve related fish being identified as such, but in the incorrect relationship category.
-The cross-assignment frequencies predicted for some common relationship categories, and for realistic FPR and FNR, are found in (Fig 4b).
+and in very modest-sized studies with very few pairwise comparisons.
 
-\textbf{Although some additional errors might be
-expected when related but non-FS individuals are sampled,  this number of comparisons exceeds
-those necessary in almost all realistic scenarios.
-We note that, as in previous work, it is nearly impossible to identify HS pairs accurately
-with data from this, or any standard marker dataset. However, the use of properly LD-filtered
-genome-scale data should allow such discrimination in future work.
-Accurate relationship
-inference in even the most genetically depauperate Chinook salmon population in California
-and rangewide (Seeb et al. 2007, Clemento et al. 2014) is therefore possible with this marker set.}
+In a reasonably large population, the vast majority of pairs of individuals are expected to
+be unrelated, or at least, effectively unrelated, sharing common ancestors only many generations
+in the past.  However, a small fraction of pairs will be more recently related, and distinguishing these
+kin pairs from target relationships (like PO and FS) must be accounted for.  `CKMRsim' provides
+facilities for assessing error rates between different kin groups, and results for these are
+summarised in (Fig.~\ref{fig:fprs}b). Here, again, focusing on the results for the genetically
+depauperate SRW, and using an FNR of 0.1 we see that the chance of incorrectly
+categorizing an HS as an FS,  or an FS as a PO is less than 1 in 100. Likewise, 
+we can expect PO to be misidentified as FS at a rate of around 1 in 500, and 
+HS or Aunts/Uncles to be misidentified as parents at a rate less than 1 in 1000.
+Though these rates are much higher than the FPRs for unrelated pairs, they should be
+compared to the number of actual non-target kin pairs expected.  For example, in a salmon
+population, the number of full aunts or uncles of a fish is unlikely to exceed 50, 
+at later life stages,  so a FPR of  $<$0.001 is comfortably low in that case.
+
+The presence of multiple alleles at many of these loci improves power for
+kin inference compared to a marker panel using only a single SNP at each
+amplicon.  To provide a visual display of the difference afforded by using
+microhaplotypes compared to single SNPs in this data set, we show the
+distribution of log-likelihood ratios for different relationships using our microhaplotype-scored
+amplicons, versus using the most heterozygous single SNP from each
+amplicon (\ref{fig:ckmr-comp}).  The figures show a considerable increase
+in separation between the relationships categories from calling genotypes
+as microhaplotypes in these amplicons.
 
 While there is not a currently available method (like that of \citeauthor{anderson2006power}) for estimating error rates in
-inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which require much more statistical power, it is clear that PO-trio inference with these markers will be essentially error free.
+inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which 
+require much more statistical power, it is clear that PO-trio inference with these markers will be essentially error free.
 
@@
 that facilitates multiple types of inference and is flexible and extensible.
+
 When, and if, sequencing costs drop sufficiently, it may be possible to include
@@
 
-Although previous work using genome-scale data has shown that many of these populations or stocks can be resolved \cite{thompson2020complex,meek2020identifying},
-such data is not practical for most large-scale applications of GSI and PBT, particularly because
+Although previous work using genome-scale data has shown that many of
+these populations or stocks can be resolved \cite{thompson2020complex,meek2020identifying},
+such data are not practical for most large-scale applications of GSI and PBT, particularly because
 of the much higher costs, longer turnaround time and challenges of producing highly
 replicable datasets with reduced representation and low-coverage whole genome sequencing methods \cite{ali2016rad}.
-Accurate identification of the distinct ecotypes of Chinook salmon in the California Central Valley is of utmost importance for designing and implementing effective conservation and management actions.
-This has been challenging, given the recent common ancestors of these ecotypes and ongoing migration between subbasins.
-We describe the first set of genetic markers that identifies the late-fall-run salmon ecotype, which occurs only in the CCV and shares a genomic background with the more common CCV fall-run salmon ecotype.
-Moreover, we show how the FRH spring-run lineage is easily identifiable through a combination of traditional GSI and the characterization of functional genetic markers in the RoSA.
-Finally, although previous work has described GSI capabilities that distinguish the natural-spawning CCV spring-run lineages from each other and their fall-run counterparts with moderately high accuracy, we demonstrate near complete accuracy in distinguishing these 'stocks'.
-Moreover, the few fish that are apparently misidentified (Fig.
-3) are likely to be primarily migrants and not true misidentifications.
-For example, the three fish that were field-characterized as spring-run from Butte Creek, but are genetically identified as winter-run salmon are clearly carry the winter-run genomic background and could not realistically be misidentified on the basis of the genetic data.
-This marker set and baseline reference dataset also provides unprecedented power for identifying fish from not only the other ESUs in California, but even the individual populations within those ESUs (Fig.
-3).
-Additionally, previous genetic marker sets for California Chinook salmon had sufficient power for inferring parent-offspring relationships, but only when both parents were sampled and genotyped.
-This marker set considerably increases the capacity for relationship inference in California salmon, by providing sufficient power for parent-offspring relationship inference, when only one parent was sampled, as well as identifying pairs of full siblings when no parents are sampled.
-Groups of full siblings larger than two are identified even more accurately than expressed in Figure 4.
+
+Accurate identification of the distinct ecotypes of Chinook salmon in the
+California Central Valley is of utmost importance for designing and implementing
+effective conservation and management actions. This has been challenging,
+given the recent common ancestors of these ecotypes and ongoing migration between subbasins.
+We describe the first replicable, reproducible panel of genetic markers that identifies the late-fall-run salmon ecotype,
+which occurs only in the CCV and shares a genomic background with the more
+common CCV fall-run salmon ecotype. Moreover, we show how the FRH spring-run lineage is
+easily identifiable through a combination of traditional GSI and the characterization
+of functional genetic markers in the RoSA. Finally, although previous work has described
+GSI capabilities that distinguish the natural-spawning CCV spring-run lineages from
+each other and their fall-run counterparts with moderately high accuracy, we demonstrate
+near complete accuracy in distinguishing these 'stocks'. Moreover, the few fish that are
+apparently misidentified (Fig. 3) are likely to be primarily migrants and not true misidentifications.
+For example, the three fish that were field-characterized as spring-run from Butte Creek, but are
+genetically identified as winter-run salmon clearly carry the winter-run genomic background and
+are strays from the winter-run stock, as these 
+could not realistically be misidentified on the basis of the genetic data.
+
+This marker set and baseline reference dataset also provides
+unprecedented power for identifying fish from not only the other ESUs in California,
+but even the individual populations within those ESUs (Fig. 3).
+
+Previous genetic marker sets for California Chinook salmon had sufficient power for inferring
+parent-offspring relationships, but only when both parents were sampled and genotyped.
+This marker set considerably increases the capacity for relationship inference in California
+salmon, by providing sufficient power for parent-offspring relationship inference, when only
+one parent is sampled, as well as identifying pairs of full siblings when no parents are sampled.
+Groups of full siblings larger than two are identified even more accurately than expressed in Fig.~\ref{fig:fprs}.
+Although some additional errors might be
+expected when related but non-target kin pairs are sampled,  the false positive rates
+for these non-targe kin pairs are sufficiently small
+for almost all realistic scenarios.
+Accurate relationship
+inference in even the most genetically depauperate Chinook salmon population in California
+and rangewide \citep{seeb2007development,clemento2014evaluation} is therefore possible with this marker set.
+We note that, as in previous work, it is nearly impossible to identify HS pairs accurately
+with data from this, or any standard marker dataset. Accurate identification of HS pairs
+typically relies on thousands of SNP markers \citep{hillary2018genetic}.

### 2024-07-10 tex/supplemental-figures.tex
Done with major editing push

@@
 like that in Fig.~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
-at the RoSA}
+at the RoSA. In each cell, the top left entry gives the number of EE (early run allele homozygotes)
+genotypes, the middle entry is the number of EL genotypes (heterozygotes), and the bottom
+right entry gives the number of LL (late-run homozygote) genotypes. }
 \label{fig:rosa-gsi}

### 2024-07-18 006-winter-run-vs-all-cv-alle-freqs.Rmd
Getting WRAP stuff added in to repo and paper/supp

---
title: "Absolute Value of Alle Freq Diff between Winter Run and all other Central Valley fish"
biblio-style: apalike
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    wrap_chr_len = "inputs/wrap/chrom_lengths.txt",
    win_non_win_abs_diffs = "stored_results/win-non-win-abs-diffs-gt0.5.rds"
  )
  # outputs:
  output_list <- list(
    mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    mh4tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    pdf = "results/genomic-locations-plot/gen-loc.pdf",
    hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
    hundy_kb4tex = "tex/images/wrap-slide-window.pdf",
    fuv2 = "results/wrap-discover/follow_up_variants2.rds",
    wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
    wrap61_4tex = "tex/images/wrap-candi.pdf"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Stuff Run on the Cluster

## Computing the allele frequencies across the whole genome

### Packages and paths 
```r
library(tidyverse)
library(viridis)
dir.create("outputs/501", recursive = TRUE, showWarnings = FALSE)
dir.create("intermediates/501", recursive = TRUE, showWarnings = FALSE)
```

## Text files with names of fall-run and spring-run fish

We do this in R.
```r
pmeta <- read_csv("data/wgs-chinook-samples.csv")

# get file with fall-run names in it
pmeta %>%
  filter(run_type == "Winter") %>%
  .$vcf_name %>%
  cat(., file = "intermediates/501/winter-run.txt", sep = "\n")

# get file with spring-run names in it
pmeta %>%
  filter(
    !str_detect(Population, "^Salmon|^Trinity"),
    run_type != "Winter") %>%
  .$vcf_name %>%
  cat(., file = "intermediates/501/non-winter-run-cv.txt", sep = "\n")
```

## Clean up the damn VCF files

Why is this still a problem?  Three of the VCF files have non-ascii gremlins in them.
We clean them up with `tr` as shown below. Doing so gave us the correct md5 hashes for the
files when we got them from Russ.  But apparently, the fixed versions were never put back
up onto Google Drive.  WTF?!  I will fix and do that now...
```sh
conda activate bioinf # to get bgzip
cd chinook_WGS_processed
for i in NC_037097.1.vcf.gz NC_037109.1.vcf.gz NC_037116.1.vcf.gz; do 
  echo $i;
  zcat $i | tr -cd '\11\12\15\40-\176' | bgzip > cleaned_$i; 
done

# when done:
for i in cleaned_NC_037*.vcf.gz; do j=${i/cleaned_/}; echo $i $j; mv $i $j; done

# after I had indexed all the VCFs I then rcloned everything back
# to the Google Drive
rclone copy --drive-shared-with-me chinook_WGS_processed gdrive-rclone:chinook_WGS_processed

```

## Do it all in a job array

**NOTE!!** I installed angsd 0.921 with conda because the latest version
fails when parsing the VCFs.

First, some prep:
```sh
# get a list of the vcf files to process
ls -l chinook_WGS_processed/*.vcf.gz | awk 'BEGIN{printf("index\tvcf\n");} {printf("%d\t%s\n", ++n, $NF);}' > intermediates/501/vcflist.txt

# get some directories where we need them:
mkdir -p intermediates/501/slurm_{out,err}
```

Then we make a quick job array script called `script/501-winter-non-winter-alle-freqs.sh`
and run it.

First test it on the first 3:
```sh
sbatch --array=1-3 script/501-winter-non-winter-alle-freqs.sh
```
That seems successful, so we crank it out on the remining ones
```sh
sbatch --array=4-41 script/501-winter-non-winter-alle-freqs.sh
```

## Filtering, alle freq diffs, and plotting

I am still doing this on the cluster within R. Our strategy
is to read the winter run and the non-winter-run-cv separately, and filter each
on the read depth requirement, separately.  Then we will do an inner join of all those.

I am requiring 62.5% or more having at least one read. So that is 10/16 for the
winter run and 50/80 for the nonwinter run.

Get the winter run:
```r
winter_files <- dir(
  path = "intermediates/501/winter-run",
  pattern = "*mafs.gz",
  full.names = TRUE
)

winter_freqs <- lapply(winter_files, function(x) {
  read_tsv(x) %>%
    filter(nInd >= 10)
}) %>%
  bind_rows()
```

Then get the non-winter-run:
```r
nonwinter_files <- dir(
  path = "intermediates/501/non-winter-run-cv",
  pattern = "*mafs.gz",
  full.names = TRUE
)

nonwinter_freqs <- lapply(nonwinter_files, function(x) {
  read_tsv(x) %>%
    filter(nInd >= 50)
}) %>%
  bind_rows()
```
Now we do the inner join:
```r
all_them <- inner_join(
  winter_freqs,
  nonwinter_freqs,
  by = c("chromo", "position"),
  suffix = c("_winter", "_nonwinter")
) %>%
  mutate(
    ave_freq = (unknownEM_winter + unknownEM_nonwinter) / 2,
    abs_diff = abs(unknownEM_winter - unknownEM_nonwinter)
  )
```
And it is quite straightforward from here. Note that we have 7,295,001 SNPs here.

First, just look at the distribution of absolute differences:
```r
g <- ggplot(all_them, aes(x = abs_diff)) +
  geom_histogram(binwidth = 0.001)
g
```
That looks good.  What if we focus on everything with an abs diff > 0.25?
```r
g +
  xlim(0.25, 1.01)
```
There is a gradual decline toward 1.00, but a tiny bump up at 1.00 itself.

This is going to be hard...
```r
g +
  xlim(0.8, 1.01)
```

Have a look at everything > 0.95 with rView:
```r
library(rView)
all_them %>%
  filter(abs_diff > 0.95) %>%
  arrange(desc(abs_diff)) %>%
  select(chromo, position, abs_diff, everything()) %>%
  count(chromo) %>%
  arrange(desc(n)) %>%
  rView()
```

So, I think what I want to do now is put everything > 0.5 into stored_results.
That still gives us 175 K SNPs to plot.
```r
dir.create("stored_results/501", showWarnings = FALSE, recursive = TRUE)
all_them %>%
  filter(abs_diff >=  0.5) %>%
  write_rds("stored_results/501/win-non-win-abs-diffs-gt0.5.rds", compress = "xz")
```

# Back to the laptop

## "Manhattan" plot

I have some code for this, but need the chromosome lengths:
We need to get the chromosome lengths for all of these.  We can pull those out of the
VCF file.
```sh
mkdir -p intermediates/501
mkdir -p outputs/501
(echo "Chromosome chrom_length"; gunzip -c data/greb1l-ish-region.vcf.gz 2> /dev/null | awk -F"=" '/^##contig/ {print $3, $4} !/^#/ {exit}' | sed 's/,length//g; s/>//g;') > inputs/wrap/chrom_lengths.txt
```

Now, filter down to abs_diff > 0.25, and get the chrom_lengths
```{r}
library(tidyverse)

wnw_lite <- read_rds(input_list$win_non_win_abs_diffs) %>%
  rename(Chromosome = chromo)

chrom_lengths <- read_table(input_list$wrap_chr_len)
```

Now, my function to make the plot. This website [https://www.r-graph-gallery.com/wp-content/uploads/2018/02/Manhattan_plot_in_R.html](https://www.r-graph-gallery.com/wp-content/uploads/2018/02/Manhattan_plot_in_R.html) had a nice discussion of it. We've store
two functions in `R/`: `my_mh_prep` and `plot_mh` and source them here: 
```{r}
source("R/manhattan_plot_funcs.R")
```

Now, use those funcs:
```{r}
wnw_prepped <- my_mh_prep(wnw_lite, chrom_lengths)

mh_plot <- plot_mh(wnw_prepped$snps, wnw_prepped$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot
```

So, as expected, there are a lot of things with large allele frequency differences.  And there are
those gaps, from before, too.  (Places where the VCFs got screwed up, somehow).

Let us also be clear that some of these SNPs might have only 10 individuals, each with just one read.  So, even though they are diploids, that is still just a sample of 10 gene copies---not 20.  Let's try filtering to 
16 individuals.

First, just look at the distribution of number of winter indivs with reads:
```{r}
wnw_lite %>%
  count(nInd_winter)
```

Plot it with sites with at least 14.
```{r}
wnw_prepped_filt <- my_mh_prep(wnw_lite %>% filter(nInd_winter >= 14), chrom_lengths)

mh_plot_filt <- plot_mh(wnw_prepped_filt$snps, wnw_prepped_filt$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot_filt
```

So, the standouts there are still interesting.

Hell, do it with 16, too:
```{r}
wnw_prepped_filt <- my_mh_prep(wnw_lite %>% filter(nInd_winter >= 16), chrom_lengths)

mh_plot_filt <- plot_mh(wnw_prepped_filt$snps, wnw_prepped_filt$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot_filt
```

Which confirms that the NC_037112.1 chromosome is the most interesting.

Let's plot a big long version of the full plot so we can 
see the points and point density better.
```{r}
ggsave(mh_plot + ylim(0.5, 1.0), filename = output_list$mh_plot,
       height = , width = 12)

file.copy(from = output_list$mh_plot, to = output_list$mh4tex, overwrite = TRUE)
```

## Zoom in on NC077112.1

```{r}
wnw_lite %>%
  filter(Chromosome == "NC_037112.1", position > 2.37e7, position < 2.65e7) %>%
  ggplot(aes(x = position, y = abs_diff)) + 
  geom_point()
```

Let's also just do a smooth for average number of sites with abs_diff > .9 per megabase.
```{r}
hundy_kb_chunks <- wnw_lite %>%
  select(Chromosome, position, abs_diff) %>%
  mutate(
    geq90 = abs_diff >= 0.9
  ) %>%
  group_by(Chromosome) %>%
  mutate(length = max(position) - min(position)) %>%
  dplyr::do(
    as_tibble(
      ksmooth(
        .$position,
        .$geq90,
        bandwidth = 1e6,
        n.points = 10 * (1 + .$length[1] / 1e6)
      )
    )
  ) %>%
  ungroup() %>%
  rename(
    `100kb_midpoint` = x,
    fract_snps_with_abs_diff_geq_0.9 = y  
  )

hundy_kb_chunks %>%
  arrange(desc(fract_snps_with_abs_diff_geq_0.9 )) %>%
  slice(1:20)

```

That is pretty compelling.  There a 1 Mb chunk on 37112 that is more interesting than anythign else.
Let's make a smoothed Manhattan plot of this:
```{r}
hundy_prepped <- hundy_kb_chunks %>%
  mutate(
    position = floor(`100kb_midpoint`),
    abs_diff = fract_snps_with_abs_diff_geq_0.9
  ) %>% 
  my_mh_prep(chrom_lengths)

hundy_mh_plot <- plot_mh(hundy_prepped$snps, hundy_prepped$axisdf) +
  xlab("Midpoint of 100 Kb interval") +
  ylab("Fraction of sites with abs_diff > 0.5 that have abs_diff >= 0.9") +
  ylim(0,NA) +
  geom_line(aes(color = color_band, group = Chromosome), size = 0.4) +
  geom_point(aes(color = color_band), size = 0.9)

hundy_mh_plot
```

## Investigate SNPs in the "six peaks"

We are going to look at SNPs in the 6 highest peaks from our smoothed 100 Kb windows,
with the possibility of designing some primers for them.

We will define those peaks as places where the smoothed fraction of sites > 0.9 is
greater than 0.124
```{r}
# filter it down and break them into groups of the different "adjacent" chunks
follow_up_chunks <- hundy_kb_chunks %>%
  filter(fract_snps_with_abs_diff_geq_0.9 > 0.124) %>%
  mutate(
    dist = `100kb_midpoint` - lag(`100kb_midpoint`, default = 200000),
    group = cumsum(Chromosome != lag(Chromosome, default = "bonkers") | dist > 100000), # this breaks the second peak on Chr 16 into two, so we will merge them
    group = case_when(
      group == 6 ~ 5L,
      group > 6 ~ group - 1L,
      TRUE ~ group
    )
  )

follow_up_chunks
```

That gives us 6 groups of adjacent (more or less) 100 Kb windows. Let us plot
those out:
```{r}
fups <- hundy_prepped$snps %>% semi_join(follow_up_chunks %>% filter(Chromosome != "NC_037103.1"), by = join_by(Chromosome, `100kb_midpoint`))

red_balloons <- hundy_mh_plot +
  geom_point(data = fups, shape = 21, fill = NA, colour = "red", size = 2)

red_balloons
```

While we are at it, let's put a copy of that into outputs
```{r}
ggsave(red_balloons, filename = output_list$hundy_kb_plot, width = 10, height = 6)

file.copy(from = output_list$hundy_kb_plot, to = output_list$hundy_kb4tex, overwrite = TRUE)
```

I want to plot the
estimated allele freq diffs from those windows (plus a buffer of 70 Kb on either side)
in a big facet_wrap.  
```{r}
follow_up_variants <- follow_up_chunks %>%
  group_by(group) %>%
  summarise(
    Chromosome = Chromosome[1],
    min_pos = min(`100kb_midpoint`) - 5e4,
    max_pos = max(`100kb_midpoint`) + 5e4
    ) %>%
  left_join(wnw_lite, by = "Chromosome") %>%
  filter(position >= min_pos - 7e4, position <= max_pos + 7e4) 
```

With that, we can then plot things. I will color by the number of winter run that
we obtained reads from:
```{r, fig.height=10, fig.width=8}

group_facet_inds <- ggplot(follow_up_variants, aes(x = position, y = abs_diff)) +
  facet_wrap(~ group, scales = "free_x") +
  geom_point(aes(colour = nInd_winter)) +
  scale_color_viridis_c()

# make another showing the SNPs to design on.  That will be: anything above 0.975
# absolute difference, or the highest abs diff, up to 8 variants.
fuv2 <- follow_up_variants %>%
  ungroup() %>%
  arrange(group, desc(abs_diff)) %>%
  group_by(group) %>%
  mutate(
    abs_diff_rank = 1:n(),
    try_design = abs_diff_rank <= 8 | abs_diff > 0.975
  )

group_facet_selecto <- ggplot(fuv2, aes(x = position / 1e06, y = abs_diff)) +
  facet_wrap(~ Chromosome, scales = "free_x") +
  geom_point(aes(fill = try_design), shape = 21, stroke = 0.05) +
  scale_fill_manual(values = c("blue", "red")) +
  xlab("Position on Chromosome (in megabases)") +
  ylab("Absolute difference in allele frequency between\nwinter-run and non-winter-run")

group_facet_selecto
```

How many is that?
```{r}
sum(fuv2$try_design)
```

Let's make a plot of that for the supplement.
```{r}
ggsave(group_facet_selecto, filename = output_list$wrap61_plot, width = 9, height = 6)

file.copy(from = output_list$wrap61_plot, to = output_list$wrap61_4tex, overwrite = TRUE)
```

## Finally, save fuv2 for later use

```{r}
write_rds(fuv2, path = output_list$fuv2, compress = "xz")
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2024-07-18 tex/supplemental-figures.tex
Boing
Getting WRAP stuff added in to repo and paper/supp

@@
 %%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=\textwidth]{images/winter-v-non-winter-mh-plot_nind_ge10.pdf}
+\caption[
+	Absolute difference in allele frequency between winter run and non-winter run fish
+]{
+	\footnotesize Absolute difference in allele frequency between winter run and non-winter run fish of the CCV.
+	The plot shows only those SNPs with at least a frequency difference of 0.5 between the two groups. Each point
+	is a SNP. The  $x$ axis shows position in the Otsh\_v1.0 genome
+	with color alternating by chromosome as indicated by RefSeq names on plot.
+}
+\label{fig:wrap-absdiff}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=\textwidth]{images/wrap-slide-window.pdf}
+\caption[
+	Sliding windows of absolute allele frequency difference between winter run and others
+]{
+	\footnotesize Values within 100 kb sliding windows of the fraction of sites with $|d|$ (absolute difference in
+	allele frequency between winter run and non-winter run fish of the CCV) greater than 0.5 that also have
+	$|d|>0.9$. The  $x$ axis shows position in the Otsh\_v1.0 genome
+	with color alternating by chromosome as indicated by RefSeq names on plot. Red circles denote sliding
+	windows chosen for further investigation for candidate markers.
+}
+\label{fig:wrap-slide-window}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=\textwidth]{images/wrap-candi.pdf}
+\caption[
+	61 candidates for primer design for winter-run-associated polymorphisms
+]{
+	\footnotesize Genomic positions and values of $|d|$ (absolute difference in
+	allele frequency between winter run and non-winter run fish of the CCV)
+	for the 61 candidate SNPs (in red) to design for winter-run-associated polymorphisms
+	(WRAPs).  Other sites are shown in black.  The $x$-axis shows position on
+	each chromsome in the Otsh\_v1.0 assembly.  The chromsome name is indicated in the facet headers by
+	RefSeq name
+}
+\label{fig:wrap-slide-window}
+\end{figure}

### 2024-07-18 tex/supplemental-text.tex
Big overhaul just before I gave lab meeting about this
A boatload of editing on the paper
Getting WRAP stuff added in to repo and paper/supp

@@
 
-\section{Supplemental Text}
+\section{Winter-run-associated polymorphism (WRAP) methods and results\label{sec:wrap-methods}}
+
+We used the whole genome sequencing data from \citet{thompson2020complex} to seek variants with
+large allele frequency differences between winter-run Chinook salmon and all the other
+Chinook salmon ecotypes in the Central Valley of California (CCV).  Because winter-run Chinook are
+already highly differentiated from all others, we did not pursue an association study as used
+for identifying the late-fall-run associated variants.  Rather, we first identified regions with a high
+density of variants with large allele frequency differences between winter-run and non-winter-run
+fish in the CCV\@. Subsequently we identified SNPs within those regions with particularly
+large allele frequency differences.  This approach was taken to avoid targeting single SNPs
+with large allele frequency differences resulting from sampling variation, which was a
+concern because we had whole genome sequencing data from only 16 winter run fish.
+
+More specifically, we calculated allele frequencies for 16 winter-run fish and 84 non-winter-run
+fish from the  \citet{thompson2020complex} variant data VCF files using ANGSD version~0.921.
+Sites were retained if at least 62.5\% of samples in each group had read data (10 of 16 for the
+winter run and 50 of 80 for the non-winter run), resulting in 7,295,001 SNPs for downstream
+analysis.  We first investigated the distribution throughout the genome of $|d|$, the absolute difference
+of the alternate allele frequency between the two groups  (Fig.~\ref{fig:wrap-absdiff}). This revealed
+that many loci had large values of $|d|$, but there were several
+regions in the genome, in particular, with prominent peaks in allele frequency difference and one
+or more SNPs apparently fixed for alternate alleles between the two sample groups.
+
+To leverage information from multiple SNPs to identify regions in the genome with large
+allele frequency differences, we calculated the fraction of SNPs with $|d| > 0.5$ that also
+had $|d| > 0.9$ within non-overlapping 100 kb sliding windows throughout the genome. This metric indicated
+several prominent peaks.  We focused on all of those
+100 kb windows in which more than 12.4\% of sites with $|d|>0.5$ also had $|d|>0.9$ which were
+also adjacent to at least one window in which $>10\%$ of sites with $|d|>0.5$ also had $|d|>0.9$ 
+(Fig.~\ref{fig:wrap-slide-window}).  Within the windows found on four different chromosomes using
+the above criteria, we then attempted to design amplicons to type the SNPs at a subset of the sites
+within each window.  We chose all sites with $|d|>0.975$, as well as the 8 SNPs on each chromosome
+with the highest values of $|d|$.
+
+\comment{ECA has to add the WRAP text here}
+
+THIS TEXT IS JUST IN HERE AS A PLACEHOLDER IN CASE WE WANT TO
+HAVE SOME SUPPLEMENTAL TEXT.
 

### 2024-07-18 R/manhattan_plot_funcs.R

@@ -1,0 +1,51 @@
+#' @param d data frame like sf_lite
+#' @param len dataframe of chromosome lengths with columns Chromome and chrom_length
+my_mh_prep <- function(d, len) {
+  # first get cumulative position.  Note that chromosomes are ordered as they should be
+  sandl <- d %>%
+    count(Chromosome) %>%
+    left_join(., len, by = "Chromosome") %>%
+    mutate(chrom_start = cumsum(chrom_length) - chrom_length) %>%
+    left_join(d, ., by = "Chromosome") %>%
+    mutate(xpos = position + chrom_start) %>%
+    mutate(cluster = ifelse(str_detect(Chromosome, "^NC_03"), Chromosome, "NWs")) %>%
+    mutate(color_band = as.factor(as.integer(factor(cluster)) %% 2)) %>% # lump the NWs together to color and make names
+    ungroup()
+
+  axisdf <- sandl %>%
+    group_by(Chromosome) %>%
+    slice(1) %>%
+    group_by(cluster) %>%
+    summarize(
+      clust_start = min(chrom_start),
+      clust_length = sum(chrom_length),
+      clust_center = (2 * clust_start + clust_length) / 2
+    )
+
+  list(snps = sandl, axisdf = axisdf)
+}
+
+
+#' a function to plot the prepped thing
+plot_mh <- function(sandl, axisdf) {
+  g <- ggplot(sandl, aes(x = xpos, y = abs_diff)) +
+    geom_point(aes(color = color_band), alpha = 0.7, size = 0.9) +
+    scale_color_manual(values = c("skyblue", "grey")) +
+
+    # custom X axis:
+    scale_x_continuous(label = axisdf$cluster, breaks = axisdf$clust_center) +
+    scale_y_continuous(expand = c(0, 0.05)) + # remove space between plot area and x axis
+
+    # Customize the theme:
+    theme_bw() +
+    theme(
+      legend.position = "none",
+      panel.border = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.x = element_blank(),
+      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
+    ) +
+    ylim(0.25, 1.0)
+
+  g
+}

### 2024-07-18 tex/supplemental-text.tex
Toss placeholder text from supp

@@
 
-\comment{ECA has to add the WRAP text here}
-
-THIS TEXT IS JUST IN HERE AS A PLACEHOLDER IN CASE WE WANT TO
-HAVE SOME SUPPLEMENTAL TEXT.
-
-Genetic markers have long been used in the 
-management of Pacific salmon, and salmon conservation and management have been
-behind a number of advances in molecular ecology,
-from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
-to statistical methodology
-\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
-Two broadly applicable innovations that have been
-actively fostered by the Pacific salmon research community are genetic
-stock identification
-(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
-and parentage based tagging
-(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
-
- In the 1980's, electrophoretically
-detectable genetic variation, in the form of allozymes
-\citep{ayala1972allozymes,allendorf1981use}, was used to
-establish a program of GSI for Chinook salmon,
-{\em Oncorhynchus tshawytscha} \citep{milner1982genetic}.  Extensive sampling
-revealed that these allozyme markers
-exhibited different allele frequencies among major stocks of Chinook salmon on the West Coast.
-These allele frequency differences make GSI possible, and
-\citet{milner1985genetic} soon showed that proportions of stocks in
-the Washington state coastal troll fishery could be estimated by GSI.
-Since that time, with the development of novel molecular markers, and, now
-with increasing capacity to sequence genomic material, the scope and scale of GSI
-has expanded considerably.
-
-By using greater numbers of more variable markers than the allozymes available
-in the 1980s it is possible to accurately
-identify the population of origin of individual fish, rather than simply estimating
-aggregated stock proportions.  It is also possible to resolve populations of fish that
-are much more closely related than before.  Furthermore, reference data sets with genotypes
-from hundreds of populations throughout the range of multiple species of salmon and
-other anadromous species
-\citep{seeb2007development,gilbey2018microsatellite,barclay2019genetic} now exist, and are routinely used to assign fish caught thousands of
-kilometers from their natal streams to their stock of origin. Applications include estimating fishery
-composition \citep{satterthwaite2015stock}, providing real-time information for genetics-informed fishery closures \citep{beacham2004dna}, assessing the spatial distribution of different stocks in the
-ocean \citep{urawa2009stock} and their temporal distribution in upstream migrations
-\citep{hess2014monitoring},  and monitoring  bycatch \citep{hasselman2016genetic} or illegal captures \citep{wilmot1999origins} in marine fisheries.
-
-Although sequencing costs continue to decline, they are high enough
-that there remains a tradeoff
-between reference baselines that include information from a large number of
-populations across a broad scale, and those that have been tailored
-to distinguish between closely related populations on a smaller, regional scale.
-Because of cost considerations, reference baselines that include populations across a broad 
-spatial scale may include only a few populations from each sub-region.  Furthermore,
-baselines tailored to a specific region often assemble markers that specifically
-show allele frequency differences between the closely related---and hence difficult
-to resolve---stocks within the region.  Consequently, regionally targeted baselines typically
-outperform broad-scale baselines in resolving populations within the region.
-
-The ascendance over the last decade of parentage based tagging (PBT) as a management
-tool for Pacific salmon adds another factor to consider when developing a GSI baseline.
-Since the first proposal \citep{anderson2005description} to replace or augment the coded-wire tag
-programme \citep{nandor2010overview}
-with PBT, it has been noted that one of the major advantages of a genetic program for
-PBT is that the genetic markers used for PBT could also be useful for GSI (and
-vice versa).
+\comment{ECA still has to add the WRAP marker-exploration steps here}

### 2024-08-16 004-lfar-and-wrap-allele-freqs.Rmd
Include full WRAP description

--- 2024-07-08 004-lfar-allele-freqs.Rmd
@@
 ---
-title: "LFAR Allele Freqs"
+title: "LFAR and WRAP Allele Freqs"
 author: "Eric C. Anderson"
@@
   output_list <- list(
-    lfar_tex_table = "tex/inputs/lfar-freqs.tex"
+    lfar_tex_table = "tex/inputs/lfar-freqs.tex",
+    wrap_tex_table = "tex/inputs/wrap-freqs.tex"
   )
@@
 
+# WRAP allele frequencies
+
+Here we do something similar to what we did above, but we use the WRAP.
+```{r}
+wrap_base <- full_base2 %>%
+  select(indiv:collection, starts_with("Ots_wrap")) %>%
+  pivot_longer(
+    cols = starts_with("Ots_wrap"), 
+    names_to = c("locus", "gene_copy"),
+    names_pattern = "^(.*)_([12])",
+    values_to = "allele"
+  )
+```
+
+Then count up the alleles in each reporting group:
+```{r}
+wrap_freqs <- wrap_base %>%
+  filter(!is.na(allele)) %>%
+  count(locus, repunit, allele) %>%
+  group_by(locus, repunit) %>%
+  mutate(freq = n / sum(n)) %>%
+  ungroup()
+```
+
+Now, for properly sorting the alleles, let's get the freqs in Winter-Run for each.
+```{r}
+alle_ord <- wrap_freqs %>%
+  filter(repunit == "CV-Winter") %>%
+  arrange(locus, desc(freq)) %>%
+  group_by(locus) %>%
+  mutate(order = 1:n()) %>%
+  ungroup() %>%
+  select(locus, allele, order)
+```
+
+This turns out to be all the alleles, too.  Good.
+
+Now we can format a table:
+```{r}
+wrap_for_table <- wrap_freqs %>%
+  left_join(alle_ord, by = join_by("locus", "allele"), relationship = "many-to-one") %>%
+  select(-n) %>%
+  pivot_wider(names_from = repunit, values_from = freq, values_fill = 0.0) %>%
+  arrange(locus, order) %>%
+  select(locus:order, `CV-Winter`, `CV-Spring`, everything())
+
+wrap_for_table
+```
+
+Write it to a table:
+```{r}
+wft2 <- wrap_for_table %>%
+  select(-order) %>%
+  extract(
+    locus, 
+    into = c("chrom", "pos"), 
+    regex = "Ots_wrap[0-9]+_([0-9]+)_([0-9]+)", 
+    convert = FALSE, 
+    remove = FALSE
+  ) %>%
+  mutate(locus = str_c("Chr", chrom, ":", pos)) %>%
+  select(-chrom, -pos) %>%
+  mutate(across(.cols = -c(locus, allele), .fns = function(x) sprintf("%.3f", x)))
+
+write_delim(wft2, delim = "&", eol = "\\tabularnewline\n", file = output_list$wrap_tex_table, col_names = FALSE)
+```
+
 # Citations

### 2024-08-16 tex/body-text-of-paper.tex
Haggling with latexdiff
small tweaks
Include full WRAP description

@@
 \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}} lllllllrcr}
-\hline\hline\vspace*{0.4ex}
+\hline\hline
+\vspace*{0.4ex}
 \input{inputs/samples-table.tex}
@@
 
-\subsubsection*{WRAP}
+\subsubsection*{Winter-run-associated polymorphisms}
 
-\comment{Gotta add the WRAP discovery and filtering stuff.}
+Though the winter run fish are easily identified, we undertook an additional effort to
+try to find markers that could be diagnostic for the winter run, distinguishing the
+winter run from all other groups of Chinook salmon in the CCV.  We were unable to
+find a single diagnostic locus; however three markers from that effort are included in
+the reference baseline.  The process of discovering them is described in the
+supplemental text section \ref{sec:wrap-methods}.
 

### 2024-08-16 tex/supplemental-tables.tex
Include full WRAP description

%!TEX root = supplement.tex

%%%%%%%%%%%%%%
\begin{table}
\caption{\footnotesize Allele frequencies across reporting units of the three winter-run associated
markers. Genome coordinates of SNPs are from Otsh\_v2.0. Alleles are listed according to the SNP nucleotide
at the variable SNPs within the amplicon.}
\label{tab:wrap-freqs}
{\footnotesize
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrrrrrr}
\hline\hline
Locus&Allele&CV-Winter&CV-Spring&CV-Fall&CV-Late-Fall&Cent-Cal-Coast&Klamath-Trinity&SO-NCal-Coast \\ \hline
\input{inputs/wrap-freqs.tex}
\end{tabular*}
}
\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
\end{table}
%%%%%%%%%%%%%%%

### 2024-08-16 tex/supplemental-text.tex
Include full WRAP description

@@
 fish in the CCV\@. Subsequently we identified SNPs within those regions with particularly
-large allele frequency differences.  This approach was taken to avoid targeting single SNPs
-with large allele frequency differences resulting from sampling variation, which was a
+large allele frequency differences.  This approach was taken to avoid targeting single, isolated SNPs
+with large allele frequency differences that may have resulted merely from sampling variation, which was a
 concern because we had whole genome sequencing data from only 16 winter run fish.
@@
 within each window.  We chose all sites with $|d|>0.975$, as well as the 8 SNPs on each chromosome
-with the highest values of $|d|$.
+with the highest values of $|d|$, yielding 61 candidate SNPs (Fig.~\ref{fig:wrap-candi}).
 
-\comment{ECA still has to add the WRAP marker-exploration steps here}
+Some of those 61 candidate SNPs were close enough that it was possible to consider amplifying them
+with PCR on 58 different short sequences.  We used Primer3 \citep{untergasser2012primer3} to return
+three possible primer-pair designs for each of the 58 sequences and then chose the primer pair
+with the fewest penalties, optimal target size, and most consistent
+melting temperatures.  One primer pair was dropped from consideration because it
+amplified a large indel that was apparent in the sequence data which would have rendered
+the sequence too large to efficiently amplify and several others were dropped because the primers
+overlapped other amplicons.  Finally, several amplicons with the lowest $|d|$ on chromosome 16
+(RefSeq NC\_037112.1) were removed from consideration, leaving us with 48 amplicons to test
+for amplification and for evaluation of allele frequencies.
+
+%% Note to self.  The next section is documented in the project at:
+%% /Users/eriq/Documents/work/assist/anthony_clemento/WRAP-crunching
+These 48 amplicons were amplified and sequenced in 192 fish---96 Feather River spring run
+and 96 winter run---on a MiSeq sequencer, and the variants were called using
+GATK.  The resulting VCF file was used to estimate allele frequency differences between
+winter run and feather river spring.  We also processed the sequence data using
+microhaplot (\url{https://github.com/ngthomas/microhaplot}) and visually inspected
+loci for consisten allele depth ratio and numbers of haplotypes.  We then chose 24 amplicons for further
+testing on the basis of allele frequency differences between Feather River spring run and
+winter run, number of haplotypes, and ease of scoring.  These 24 markers were typed on
+a variety of fish over the course of a year, and we finally chose three to include in
+our California Chinook reference baseline: one amplicon on each of chromosomes
+8, 12, and 16.  The estimated frequencies of all the alleles present in the reference baseline
+in those three amplicons appears shows that there are not fixed differences between
+winter run and all other reporting units (Table~\ref{tab:wrap-freqs}).

### 2024-08-16 tex/body-text-of-paper.tex
Update major and minor Structure modes

@@
 {\em a priori} knowledge about which collections in the baseline could and should be
-grouped together into reporting units.  At every value of $K$ CLUMPAK discerned \comment{at least
-11 of 20 replicates} in the major mode (Fig.~\ref{fig:struct}).   
+grouped together into reporting units.  At every value of $K$ CLUMPAK discerned at least
+12 of 20 replicates in the major mode (Fig.~\ref{fig:struct}).   
 \comment{Need to add more description of the patterns, especially at $K=7$.}
-\comment{Clustering solutions in the minor
+Clustering solutions in the minor
 modes appear in Fig.~\ref{fig:minor-modes}, which likely needs to be updated.  Get all this done
-when AC is back and can vet it.}
+when AC is back and can vet it.
 %%%%%%%%%%%%%%%%%%%%
@@
 by CLUMPAK out of 20 replicates for each value of $K$, the assumed number of
-subpopulations. }
+subpopulations. $K$ value is listed at lower right of each barplot, followed by the number
+of replicate runs of structure that converged to this major mode.}
 \begin{center}

### 2024-08-16 tex/supplemental-figures.tex
Include full WRAP description
Update major and minor Structure modes

@@
 total numbers of alleles in the data set. The one amplicon with only
-one allele is {\tt Ots\_coho001\_05\_32691399}, which is fixed for alternate
+one allele is {\tt Ots\_coho001\_05\_32691399}, which is fixed for alternate alleles
 in Chinook vs.\ coho salmon. It is helpful in identifying coho samples that are
@@
 \end{center}
-\caption[STRUCTURE minor modes found by CLUMPAK]{\footnotesize At each value of $K$
-for which a minor mode was found, these plots show them}
+\caption[STRUCTURE minor modes found by CLUMPAK]{\footnotesize STRUCTURE minor modes found by CLUMPAK.
+At each value of $K$ for which a minor mode was found, the plot is shown.  $K$ values and number of times each minor
+mode appeared out of 20 replicate runs of STRUCTURE appear to the lower right of each.}
 \label{fig:minor-modes}
@@
 }
-\label{fig:wrap-slide-window}
+\label{fig:wrap-candi}
 \end{figure}

### 2024-08-16 tex/supplemental-text.tex
Update major and minor Structure modes

@@
 
-\section{Winter-run-associated polymorphism (WRAP) methods and results\label{sec:wrap-methods}}
+\section{Winter-run-associated polymorphism methods and results\label{sec:wrap-methods}}
 
@@
 8, 12, and 16.  The estimated frequencies of all the alleles present in the reference baseline
-in those three amplicons appears shows that there are not fixed differences between
+in those three amplicons shows that there are not fixed differences between
 winter run and all other reporting units (Table~\ref{tab:wrap-freqs}).

### 2024-08-19 tex/body-text-of-paper.tex
Add ascertainment panel sample size table to supp

@@
 populations ranging from California to British
-Columbia and the interior Columbia River (\comment{Table SX}).
+Columbia and the interior Columbia River (Table~\ref{tab:ascert-panel}).
 Genomic DNA was extracted using Qiagen DNeasy 96 Blood and Tissue Kits on a BioRobot 3000

### 2024-08-19 tex/inputs/ascertainment-samples-supp-table.tex
Add ascertainment panel sample size table to supp

Sacramento River winter run&4 \\
Feather River spring run&2 \\
Feather River fall run&2 \\
Butte Creek spring run&2 \\
Deer Creek fall run&2\\
Eel River fall run&2 \\ 
Klamath Basin, Trinity River Hatchery spring run&2\\
Southern Oregon, Chetco River&2\\
Northern Oregon, Siletz River&2\\
Columbia River, Kalama Hatchery spring run&2\\
Upper Columbia River, Wenatchee River&2\\
Snake River, McCall Hatchery summer-spring run&2\\
Puget Sound,Kendall Hatchery spring run&2\\
Fraser River, Thompson River, Spius Creek&2\\
Fraser River, Birkenhead River&2\\

### 2024-08-19 tex/supplemental-tables.tex
Update major and minor Structure modes
Add ascertainment panel sample size table to supp

@@
 \begin{table}
-\caption{\footnotesize Allele frequencies across reporting units of the three winter-run associated
+\caption[Microhaplotype discovery panel sample sizes]{Numbers of fish from different collections/populations used in the microhaplotype discovery effort.
+$N$ is the number of diploid individuals included in the ascertainment panel.}
+\label{tab:ascert-panel}
+\begin{center}
+{\footnotesize
+\begin{tabular}{lr}
+\hline\hline
+Population&$N$ \\ \hline
+\input{inputs/ascertainment-samples-supp-table.tex}
+\\ \hline\hline
+\end{tabular}
+}
+\end{center}
+\end{table}
+%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%
+\begin{table}
+\caption[Allele frequencies of winter-run-associated polymorphisms]{Allele frequencies across reporting units of the three winter-run associated
 markers. Genome coordinates of SNPs are from Otsh\_v2.0. Alleles are listed according to the SNP nucleotide

### 2024-08-19 tex/body-text-of-paper.tex
Work in Mac's edits

@@
 
-In total, the samples in the baseline represent fish from five different Evolutionarily Significant
+In total, the samples in the baseline represent fish from five (Table~\ref{tab:samples}) different Evolutionarily Significant
 Units (ESUs) of Chinook salmon \citep{waples1991pacific}.
@@
 These gene regions were PCR-amplified and sequenced on a MiSeq sequencer using a $2\times 75$-cycle kit.
-The sequences were mapped using
-bwa-mem \citep{bwa-mem2009} to the reference sequences
-from the de novo assembly for each locus, sorted with samtools \citep{li2009sequence} and provided to
-freeyayes \citep{garrison2012haplotype} for variant calling.
-freebayes output was filtered to remove indels and retain variants with a Phred-scaled quality score of 30
+The sequences from each locus were mapped to the de novo reference sequences (produced by Stacks) with
+the Burrows-Wheeler aligner using the mem algorithm \citep{bwa-mem2009}. 
+Subsequently, alignmets were sorted with SAMtools \citep{li2009sequence} and provided to
+FreeBayes \citep{garrison2012haplotype} for variant calling.
+FreeBayes output was filtered to remove indels and retain variants with a Phred-scaled quality score of 30
 or greater and a read depth of 10 or higher. Further filtering and analysis of the loci, including

### 2024-08-26 007-whoa-on-chinook.Rmd
Commit the Whoa stuff

---
title: "Running the California Microhap Baseline through `whoa` and Deciding We Should Drop Two of the Amplicons"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
    pop_labels = "inputs/reference-collection-names.csv"
  )
  # outputs:
  output_list <- list(
    zs = "results/whoa/heterozygote-z-scores.pdf"
  )
}

# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(whoa)

# read in the full baseline
full_base <- read_csv(input_list$final_baseline)

# capture that rosa hapstr with some meta data from future use
rosa_genos0 <- full_base %>%
  select(indiv:rosa_genotypes_str)

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(old_name, collection, repunit), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything()) %>%
  filter(!duplicated(indiv))  # necessary cuz clemento stuck mill and deer together into a single code

# finally, get the order of collections and repunits we want
collection_order <- unique(pop_labels$collection)

```

Get things into long format:
```{r}
fb_long <- full_base2 %>%
  select(-sample_type, -rosa_genotypes_str) %>%
  pivot_longer(
    cols = -(indiv:collection), 
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "^(.*)_([12])$"
  )
```

Now, for every allele in the data set we define a new locus---one in which the
allele is the 1 allele and everything else is the 0 allele.

For the purposes of passing things off to whoa, we
are going to use as "loki" the combination of `locus` and `recoding`, so we will
just paste those with a dot.

Whoa says it wants loci named by Chrom-Pos but I don't think that is actually
necessary, so I will just paste recoding on with a dot to the locus name.

```{r}
RecodedAlleles <- fb_long %>%
  filter(!is.na(allele)) %>%   # it is very important to remove missing genotypes
  filter(locus != "Ots_coho001_05_32691399") %>%  # no reason to have the coho allele
  group_by(locus) %>%
  mutate(
    alle_int = as.integer(factor(allele, levels = sort(unique(allele)))),
    alle_cnt = length(levels(factor(allele)))
  ) %>%
  ungroup() %>%
  mutate(
    boing = map2(.x = alle_int, .y = alle_cnt, .f = function(x, y) {
      list(
        recoding = 1:y,
        new_alle = ifelse(x == 1:y, 1L, 0L)
      )
    })
  ) %>%
  mutate(
    recoding = map(boing, "recoding"),
    new_alle = map(boing, "new_alle")
  ) %>%
  select(-boing) %>%
  unnest(cols = c(recoding, new_alle)) %>%
  #extract(locus, into = c("chrom", "pos"), regex = "Ots_[a-z]+[0-9]+_([0-9]+)_([0-9]+)", remove = FALSE) %>%
  #mutate(loki = str_c("Ots", chrom, "_", recoding, "-", pos ), .after = locus) %>%
  mutate(loki = str_c(locus, recoding, sep = "."))
  #select(-chrom, -pos)
```

That seemed to go well.  In order to be able to get back from the new_alle and
recoding to the actual allele, in case we want to, and to verify that things worked,
we can collapse this into the different cases:
```{r}
haplo_integers <- RecodedAlleles %>%
  distinct(locus, loki, allele, alle_int, alle_cnt, recoding, new_alle) %>%
  arrange(locus, recoding, allele)

haplo_integers
```

That all looks just right.

Recall, for whoa we want an 012 matrix with missing data denoted by -1.

So, we need to turn these thing back into genotypes for the different
individuals.  Once I am done with that, I am going to nest the 012 matrices
up by repunit and collection

```{r}
Recoded_012_full_tib <- RecodedAlleles %>%
  select(repunit, collection, indiv, loki, new_alle) %>%
  group_by(repunit, collection, indiv, loki) %>%
  summarise(geno = sum(new_alle)) %>%
  ungroup() %>%
  pivot_wider(names_from = loki, values_from = geno, values_fill = -1L) %>%
  nest(tibble = -c(repunit, collection)) %>%
  mutate(
    d012 = map(tibble, function(x) x %>% select(-indiv) %>% as.matrix() %>% t())
  )
```

Cool, now we have a list column with a d012 for each collection.

Now we can make a new list column with the whoa output.
```{r}
tib_with_plots <- Recoded_012_full_tib %>%
  mutate(
    exp_and_obs = map(d012, function(x) whoa::exp_and_obs_geno_freqs(d012 = x)),
    scatter = pmap(
      list(eo = exp_and_obs, col = collection, ti = tibble),
      function(eo, col, ti) {
        whoa::geno_freqs_scatter(eo, max_plot_loci = 1000) +
          ggtitle(str_c(col, "  (", nrow(ti), ")"))
      }
    ) 
  )
```

We can see those all in the notebook like this
```{r}
tib_with_plots$scatter
```

There are clearly a few alleles that do not show up in heterozygous
form as freuquently as expected.  I suspect that these may belong
to just one or two loci.

To find which loci are involved, we will simply sort on the z-scores
for the heterozygotes.
```{r}
sorted_losers <- tib_with_plots %>%
  select(repunit, collection, exp_and_obs) %>%
  unnest(exp_and_obs) %>%
  filter(geno == "1") %>%
  arrange(z_score)

```

Looking at that, it is clear that two amplicons are heavily implicated: `Ots_mhap073_19_03686095`
and `Ots_mhap093_27_16923862`.  In the Eel river we also see `Ots_mhap053_11_42403691` pop up
once or twice amongst the most extremely low z-scores for hets, and also
we see `Ots_mhap057_13_38673256.2` at Cole Rivers.  Eventually with z-scores above
-3 we start to see more loci that are suspect are just due to random variation.

But, let's investigate those few loci that we saw in the context of all of the loci.
This makes a massive faceted thing that is really only viewable as a PDF, so we just
save it.
```{r}
zs <- sorted_losers %>%
  extract(snp, into = "locus", regex = "^(.+)\\.[0-9]+", remove = FALSE) %>%
  ggplot(aes(x = z_score)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_histogram() +
  theme_bw() +
  facet_wrap(~ locus, ncol = 10)

  ggsave(zs, filename = output_list$zs, width = 20, height = 30)
```

That provides a lovely perspective on things.  mhap073 and mhap093 are clearly pathological.
mhap053 and mhap057 show a slight downward trend in the z-scores, but it must be somewhat
variable by population, and I don't think it is enough to toss them out.

mhap073 and mhap093 have 4 and 2 alleles respectively, so tossing them won't be
a big deal.  mhap093 only has two alleles but the are named things like
`CCGCCGTTGCT` and `CTGCCGTTGCT`, so that is pretty messed up anyway.

So, I think we should toss those two loci out.

If we do toss them out, this is what the geno freq plots look like.
```{r}
tib_with_dropped_plots <- tib_with_plots %>%
  mutate(dropped_eo = map(exp_and_obs, function(x) x %>% filter(!str_detect(snp, "Ots_mhap073_19_03686095|Ots_mhap093_27_16923862")))) %>%
  mutate(
    dropped_scatters = pmap(
      list(eo = dropped_eo, col = collection, ti = tibble),
      function(eo, col, ti) {
        whoa::geno_freqs_scatter(eo, max_plot_loci = 1000) +
          ggtitle(str_c(col, "  (", nrow(ti), ")"))
      }
    ) 
  )

tib_with_dropped_plots$dropped_scatters
```

So, what we see is that most of the really egregious ones are gone, but in some
populations there are a few that look a little wonky, but those are only
in a few pops.  So, I don't think we really need to toss any others.

### 2024-09-27 tex/author-title-etc.tex
Added Mac's current address.
Hawaiian balcony edits

@@
 \newcommand{\myAuthors}{Eric C.
-Anderson$^{*,\S}$, Anthony J.
-Clemento$^{*}$, Matthew A.
-Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F.
-Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}
+Anderson$^{1,\S}$, Anthony J.
+Clemento$^{1,2}$, Matthew A.
+Campell$^{1,2\dagger}$, Devon Pearse$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
+Thompson$^{1,2\ddag}$, John Carlos Garza$^{1,2}$}
 \author{Eric C.
-Anderson$^{*}$, Anthony J.
-Clemento$^{*}$, Matthew A.
-Campell$^{*,\dagger}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, Neil F.
-Thompson$^{*,\ddag}$, John Carlos Garza$^{*}$}
+Anderson$^{1,2}$, Anthony J.
+Clemento$^{1,2}$, Matthew A.
+Campell$^{1,2,\dagger}$, Devon Pearse$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
+Thompson$^{1,2,\ddag}$, John Carlos Garza$^{1}$}
 
@@
 \newcommand{\myAffiliations}{
-$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA.
-$^\dagger$Current address: Mac's home address in AK.
-$^\ddag$Current address: Get Neil's address.
+$^1$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA, Santa Cruz, California, USA.
+$^2$Institute for Marine Sciences, University of California, Santa Cruz, USA.
+$^\dagger$Current address: Centre for Carbon, Water and Food, The University of Sydney, 380 Werombi Road, NSW 2570, Australia.
+$^\ddag$Current address: Pacific Shellfish Breeding Center, Agricultural Research Service, US Department of Agriculture, Newport, Oregon, USA.
 }
@@
 
-\renewcommand{\CorrespondingAuthor}{eric.c.anderson@noaa.gov}
+\renewcommand{\CorrespondingAuthor}{eric.anderson@noaa.gov,~carlos.garza@noaa.gov}
 
 % the email address for the corresponding author
-\newcommand{\myEmailAddress}{eric.c.anderson@noaa.gov}
+\newcommand{\myEmailAddress}{eric.anderson@noaa.gov,~carlos.garza@noaa.gov}
 \newcommand{\myEmailFootnote}{$^\S$}

### 2024-09-27 tex/body-text-of-paper.tex
Hawaiian balcony edits

@@
 rivers within the state of California, and particularly targeted to the complex population
-structure of Chinook salmon within California's Central Valley. Chinook salmon of the two main river basins---the
-Sacramento and the San Joaquin---within the Central Valley exhibit the greatest run-timing diversity
+structure of Chinook salmon within California's Central Valley (CCV). Chinook salmon of the two main CCV river basins---the
+Sacramento and the San Joaquin---exhibit the greatest run-timing diversity
 within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
@@
 fisheries \citep{myers1998status}. Moreover, because of their high degree of ecotypic variation, they have provided fishery opportunities in many
-different seasons and geographic locations \citep{healey1991life}. However, recent declines in population sizes, from the Yukon River in Arctic north,
-to the southern extent of their range in California has led to almost complete fishery closures  throughout most of the northeastern Pacific to protect
-less productive stocks \comment{(CITE)}. The co-occurence of fish from relatively productive and relict populations reaches its paragon in California,
+different seasons and geographic locations \citep{healey1991life}. However, recent declines in population sizes, from the Yukon River in the Arctic north,
+to the southern extent of their range in California, has led to multiple fishery closures to protect
+less productive stocks \citep{lindley2009caused}. The co-occurence of fish from relatively productive and relict populations reaches its paragon in California,
 with the largest remaining ocean fisheries for Chinook salmon targeting the Central Valley fall-run stock that spawns in different tributaries of the
-same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks \comment{(CITE)}.
+same river basin as the highly endangered and phenotypically distinct winter-run stock and the threatened spring-run stocks \citep{satterthwaite2015stock}.
 
@@
 has been particularly challenging in the CCV, with at least one of the distinct ecotypes, the late-fall-run, unresolvable with previous GSI marker sets
-\citep{seeb2007development,clemento2014evaluation} and with unsatisfactory power for discriminating the protected spring-run stocks from the
-harvested fall-run stock \comment{(CITE)}.
+and with unsatisfactory power for discriminating the protected spring-run stocks from the
+harvested fall-run stock \citep{seeb2007development,clemento2014evaluation}.
 
-In the following we present a new reference baseline that includes 1636 Chinook salmon individuals
-from 17 collections, genotyped at 204 amplicons from throughout the genome,
-scored as microhaplotypes \citep{baetscher2018microhaplotypes}. This baseline is
+In the following, we present a new reference baseline that includes 1,636 Chinook salmon individuals
+from 17 collections, genotyped at 204 microhaplotypes \citep{baetscher2018microhaplotypes} from throughout the genome.
+This baseline is
 highly effective for GSI within California and the markers within it provide ample power
@@
 spring-run and fall-run ecotypes (i.e.
-Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype \comment{(CITE)}.
+Mill, Deer and Butte Creeks), previous research has documented different  (but overlapping) ranges of spawning time (Julian date) for each ecotype \citep{fry1961king,yoshiyama1998historical}.
 As such, sampling targeted carcasses found in each of these
@@
 to the river. Subsequently in the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
-as spring-run broodstock, the rest are included in the fall-run broodstock.
-We sampled putative spring-run and fall-run fish from these two broodstocks.
-\comment{Need info about sampling in the Klamath, and on the coast}.
+as spring-run broodstock, and the rest are included in the fall-run broodstock. We sampled putative spring-run and fall-run fish 
+from these two broodstocks.
+Similar to broodstock collections in the CCV, samples from spring- and fall-run fish in the Trinity River were selected on the basis of spawn timing at the Trinity River Hatchery with migration-timing ecotype then confirmed by RoSA genotype.
+Elsewhere, samples were taken from live broodstock at Cole Rivers (Rogue), Rowdy Creek (Smith), Iron Gate (Klamath) and Warm Springs (Russian River) Hatcheries, where only fall-run fish have been documented.
+Samples from the Eel River were taken from live fish ascending a ladder in the upper basin and were confirmed as fall-run ecotype by RoSA genotype.
 

### 2024-09-27 tex/body-text-of-paper.tex
more tweaks

@@
 package `hierfstat' \citep{hierfstat}. Finally, we evaluated population
-structure in the data by performing unsupervised clustering of
-the complete data set
+structure in the data by performing unsupervised clustering 
 using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
+For this analysis, we omitted the run-timing associated markers.
 We performed 20 replicate runs at each value of the assumed number of
@@
 
-The program STRUCTURE clustered fish at all values of $K$ from 2 to 7.
-(Fig.~\ref{fig:struct}).  The clusters generally corresponded to groupings of related populations, and confirmed
+At value of $K$ from 2 to 7, clusters identified by the program STRUCTURE
+generally corresponded to groupings of related populations, and confirmed
 {\em a priori} knowledge about which collections in the baseline could and should be
-grouped together into reporting units.  At every value of $K$ CLUMPAK discerned at least
+grouped together into reporting units (Fig.~\ref{fig:struct}).  At every value of $K$, CLUMPAK discerned at least
 12 of 20 replicates in the major mode (Fig.~\ref{fig:struct}).   
-\comment{Need to add more description of the patterns, especially at $K=7$.}
 Clustering solutions in the minor
-modes appear in Fig.~\ref{fig:minor-modes}, which likely needs to be updated.  Get all this done
-when AC is back and can vet it.
+modes generally converged on a single alternative solution and appear in Fig.~\ref{fig:minor-modes}.
 %%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
+The late-fall run collection (CHLF) appears as a separate cluster at $K$ values
+as low as 4, with the inclusion of the LFAR markers.  This is notable, since accurately discriminating between the late-fall and
+fall-run Chinook salmon of the Central Valley has been impossible with previously used genetic marker sets.
+Another noteworthy feature appears at $K=7$, with STRUCTURE separating the fall-run reporting unit
+into two clusters (orange and red) that do not correspond exactly with the individual
+collections within the reference baseline. For example, fish from the
+orange cluster that predominates in the FRHS collection are also found in the FRHF and MDS collections.
 
@@
 As seen in previous studies, FRHS, the Feather River Hatchery
-spring-run stock, is considerably differntiated from the remaining spring-run
+spring-run stock, is considerably differentiated from the remaining spring-run
 stocks in the Central Valley (MDS and BCS). This is reflected in the low
@@
 
+\comment{Maybe discuss: A large fraction of the Feather River
+Hatchery Spring-run (FRHS) collection into a cluster (colored orange in Fig.~\ref{fig:struct}).
+A large fraction of the samples from the fall-run hatchery
+program at the Feather River (FRHF) are also members of that same orange cluster, while others
+belong to a cluster composed of other fall-run Chinook salmon. These STRUCTURE results at $K=7$,
+validate the finding that all populations outside of the }
+
 Although previous work using genome-scale data has shown that many of
@@
 typically relies on thousands of SNP markers \citep{hillary2018genetic}.
+
+Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns of population relationships that are coincident with previous work (Clemento et al.
+2014; Kinziger et al.
+2013), emphasizing the distinction between Chinook salmon
+populations in Coastal California and the Central Valley. These analyses also uncover some additional patterns, including the clear
+presence of mixed ancestry in the Coastal California population in the Russian River, which is consistent with its location, at the southernmost edge of the Coastal Chinook salmon between the mouths of the Sacramento River and the other Coastal Chinook salmon basins to the north.

### 2024-10-02 tex/body-text-of-paper.tex
Clean up discussion

@@
 highly effective for GSI within California and the markers within it provide ample power
-for PBT.
+for PBT, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program 
+\citep{garza2007large,beacham2021parentage}.
 We describe how we identified and compiled the markers, including several new markers
@@
 We describe a novel panel of microhaplotype genetic markers for Chinook salmon
-in California that includes multiallelic gene regions with high variability for
-relationship inference and gene regions identified in whole-genome sequence data
-for increased power for population identification. We show how this set of markers
-provides sufficient power for highly accurate identification of parent and offspring
-pairs and full siblings in California Chinook salmon, and also allows near-perfect
-identification of individuals to population or genetic group of origin. This set of
-genetic markers lays the foundation for a comprehensive genetic monitoring and evaluation effort
-that facilitates multiple types of inference and is flexible and extensible.
+in California that provides sufficient power for highly accurate identification 
+of parent and offspring pairs and full siblings, and also allows near-perfect
+identification of individuals to population or genetic group of origin in California. 
+This set of genetic markers includes multiallelic gene regions with high variability for
+relationship inference, gene regions identified in whole-genome sequence data
+for increased power for identification of specific populations and incorporates markers that have
+long been in use for both GSI and PBT in California (Clemento et al. 2014), by converting them
+into so-called 'SNPlicons'. This set of markers lays the foundation for a comprehensive 
+genetic monitoring and evaluation effort that facilitates multiple types of inference 
+and is flexible and extensible.
 
-When, and if, sequencing costs drop sufficiently, it may be possible to include
-all different populations of a single species within a single standardized reference baseline
-that performs equally well at broad and regional scales. A statistical approach for population
-assignment from low coverage whole genome sequences appeared recently
-\citep{desaixINPRESSpopulation}, and its authors noted that such a data type could
-be ideal for producing reference baselines simultaneously applicable to broad scale
-coverage and high resolution within sub-regions.   For the present, however,
-baselines tailored to specific regions are essential for regional management questions.
-
-\comment{Maybe discuss: A large fraction of the Feather River
-Hatchery Spring-run (FRHS) collection into a cluster (colored orange in Fig.~\ref{fig:struct}).
-A large fraction of the samples from the fall-run hatchery
-program at the Feather River (FRHF) are also members of that same orange cluster, while others
-belong to a cluster composed of other fall-run Chinook salmon. These STRUCTURE results at $K=7$,
-validate the finding that all populations outside of the }
-
-Although previous work using genome-scale data has shown that many of
-these populations or stocks can be resolved \cite{thompson2020complex,meek2020identifying},
-such data are not practical for most large-scale applications of GSI and PBT, particularly because
-of the much higher costs, longer turnaround time and challenges of producing highly
-replicable datasets with reduced representation and low-coverage whole genome sequencing methods \cite{ali2016rad}.
-
+This marker set and baseline reference dataset provide
+unprecedented power for identifying fish from all of the Chinook salmon ESUs in California,
+as well as individual populations within those ESUs (Fig.~\ref{fig:gsi}).
 Accurate identification of the distinct ecotypes of Chinook salmon in the
 California Central Valley is of utmost importance for designing and implementing
-effective conservation and management actions. This has been challenging,
-given the recent common ancestors of these ecotypes and ongoing migration between subbasins.
-We describe the first replicable, reproducible panel of genetic markers that identifies the late-fall-run salmon ecotype,
-which occurs only in the CCV and shares a genomic background with the more
-common CCV fall-run salmon ecotype. Moreover, we show how the FRH spring-run lineage is
+effective conservation and management actions. The CCV is one of the largest river
+basins on the west coast of North America and has the highest diversity of recognized Chinook salmon
+ecotypes in the species range, with four named ecotypes, two of which are protected
+under the US Endangered Species Act. This basin drains the Sierra Nevada and southern Cascade
+mountain ranges and is the source of water for several major urban areas as well as the
+largest and most valuable agriculture area in North America. As such, monitoring and
+evaluation of individual ecotypes and populations is of critical importance. This has been challenging,
+given the recent common ancestors of these ecotypes \citep{clemento2014evaluation}, and
+ongoing migration between subbasins where different ecotypes predominate.
+
+We describe the first set of genetic markers that produces easily replicable data and
+that identifies all of the ecotypes in the CCV, including the late-fall-run Chinook salmon ecotype.
+The late-fall-run occurs only in the CCV and shares a genomic background with the more
+common CCV fall-run salmon ecotype, so has been refractory to previous GSI efforts with other marker types \citep{seeb2007development,clemento2014evaluation,thompson2024genomics}.
+Moreover, we show how the FRH spring-run lineage is
 easily identifiable through a combination of traditional GSI and the characterization
@@
 
-This marker set and baseline reference dataset also provides
-unprecedented power for identifying fish from not only the other ESUs in California,
-but even the individual populations within those ESUs (Fig. 3).
+As sequencing costs drop, whole genome sequence (WGS) data may become the preferred
+data type for salmon genetics. If so, it may become possible to include
+all >200 genetically distinct Chinook salmon populations within a single standardized reference baseline
+constructed with WGS data
+that performs equally well at broad and regional scales \citep{desaixINPRESSpopulation}.
+Such an approach would be highly flexible and extensible, as it would allow for the use
+just about any marker type, including reduced representation DNA sequencing (e.g. RADseq,
+\citealt{meek2020identifying,thompson2024genomics}),
+as regional monitoring datasets would not need to be replicable.  For the present, however,
+baselines tailored to specific regions are essential for regional management questions and
+targeted sequencing approaches have proven to be the most practical for large-scale
+applications of either GSI, PBT or an integrated monitoring program \citep{beacham2021parentage}.
+
+%\comment{Maybe discuss: A large fraction of the Feather River
+%Hatchery Spring-run (FRHS) collection into a cluster (colored orange in Fig.~\ref{fig:struct}).
+%A large fraction of the samples from the fall-run hatchery
+%program at the Feather River (FRHF) are also members of that same orange cluster, while others
+%belong to a cluster composed of other fall-run Chinook salmon. These STRUCTURE results at $K=7$,
+%validate the finding that all populations outside of the }
+
+%Although previous work using genome-scale data has shown that many of
+%these populations or stocks can be resolved \cite{thompson2020complex,meek2020identifying},
+%such data are not practical for most large-scale applications of GSI and PBT, particularly because
+%of the much higher costs, longer turnaround time and challenges of producing highly
+%replicable datasets with reduced representation and low-coverage whole genome sequencing methods \cite{ali2016rad}.
 
@@
 with data from this, or any standard marker dataset. Accurate identification of HS pairs
-typically relies on thousands of SNP markers \citep{hillary2018genetic}.
+typically relies on many hundreds of microhaplotype markers \citep{baetscher2018microhaplotypes} or thousands of SNP markers \citep{hillary2018genetic}.
 
-Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns of population relationships that are coincident with previous work (Clemento et al.
-2014; Kinziger et al.
-2013), emphasizing the distinction between Chinook salmon
+Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns of population relationships that are
+coincident with previous work (Clemento et al. 2014; Kinziger et al. 2013), emphasizing the distinction between Chinook salmon
 populations in Coastal California and the Central Valley. These analyses also uncover some additional patterns, including the clear
-presence of mixed ancestry in the Coastal California population in the Russian River, which is consistent with its location, at the southernmost edge of the Coastal Chinook salmon between the mouths of the Sacramento River and the other Coastal Chinook salmon basins to the north.
+presence of mixed ancestry in the Coastal California population in the Russian River, which is consistent with its location, at the 
+southernmost edge of the Coastal Chinook salmon distribution and proximate to the mouth of the Sacramento River (the Golden Gate). In the
+Central Valley, the clustering results emphasize the genetic distinction of the SWR population, which is likely at least partially due to 
+the extreme bottleneck that it passed through 15-20 generations prior. The inclusion of the LFAR markers also resulted in the clear 
+distinction of the CHLF group, which is coincident with the long-known phenotypic distinction of this population, but represents a novel 
+genetic result. Moreover, at K = 7, the fall-run reporting group breaks into two clusters, which are distributed across almost all of the 
+spring and fall-run populations in the Central Valley, albeit not equivalently, emphasizing both current and historical  migration and 
+gene flow between them.

### 2024-10-28 tex/body-text-of-paper.tex
Edits with carlos

@@
 \LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run.
-Names of each location are followed by the codes used to refer to fish from each location, with
-{\tt XXF/S} indicating {\tt XXF} and {\tt XXS} are used to denote fall-run and spring-run fish
-from location {\tt XX}.  Abbreviations used in location names as follows: R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
+Names of each location are followed by the location codes (Tab.~\ref{tab:samples}).
+Abbreviations used in location names as follows:
+R. = River; Ck. = Creek; H. = Hatchery; NFH = National Fish Hatchery.}
 \begin{center}
@@
 \begin{figure}
-\newcommand{\gsicap}{\footnotesize Pairwise $F_\mathrm{ST}$ values and allele frequencies ({\bf a)}
-and self-assignment rates
- ({\bf b)}  for the collections in the baseline.  The codes of the populations
+\newcommand{\gsicap}{\footnotesize ({\bf a)} Pairwise $F_\mathrm{ST}$ values and allele frequencies,  
+and ({\bf b)} self-assignment rates.  The codes of the populations
 appear on the outer margins of the table atop cells colored according to the run-timing group
-of the collection.  Interior cells in which the column and row correspond to the same reporting
-unit are colored according to the reporting unit.  In ({\bf a}), for $F_\mathrm{ST}$ values,
+of the collection.  Interior cells are colored according to reporting unit.  In ({\bf a}), for $F_\mathrm{ST}$ values,
 the upper triangle was computed using all the genetic markers except for those in the
@@
 \newcommand{\fprcap}{\footnotesize False positive rates (FPRs) as a function of false negative
-rates (FNRs) for distinguishing between different relationships.  {\bf (a)} FPRs calculated
+rates (FNRs) for distinguishing between different relationships.
+Groupings for panels with more than one population are as follows: SONC = \{CRFS, SmRF, BCkF\}; TRH = TRS, TRF; CCC = \{ERF, RRF\}; CVF = BCF, MDF, SJRF); FRH = \{FRHS, FRHF\} relationship inference analysis are as follows  {\bf (a)} FPRs calculated
 using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),

### 2024-10-28 002-GSI_and_Fst.Rmd
Add code for a/b labels on fig 2

@@
   nrow = 3,
-  rel_heights = c(6,1.5,6)
+  rel_heights = c(6,1.5,6), 
+  labels = c("a)", "", "b)"),
+  label_size = 32, 
+  label_x = -0.02,
+  label_y = 1.013
 )

### 2024-10-28 003-ckmr-sim-evaluation.Rmd
Tweak the names of groups in the CKMRsim plot

@@
 
+Now name the groups a little differently and sort them so that they are sorted
+like they are in the other figures.
+```{r}
+CST0 <- ckmr_sim_tib %>%
+  mutate(
+    group = str_replace(ckmr_sim_group, "^ck-", ""),
+    group = case_match(
+      group,
+      "LF" ~ "CHLF",
+      "IGH" ~ "IGHF",
+      "SONC" ~ "SO-NCal-Coast",
+      "CCC" ~ "Cent-Cal-Coast",
+      .default = group
+    ),
+    .before = ckmr_sim_group
+  ) %>%
+  mutate(group = factor(group, levels = c("SO-NCal-Coast", "IGHF", "TRH", "Cent-Cal-Coast", "SRW", "BCS", "MDS", "FRH", "CVF", "CHLF")))
+
+```
+
 Now, make a figure of those results.
@@
 
-cst2 <- ckmr_sim_tib %>%
+cst2 <- CST0 %>%
   mutate(scenario = str_c(mc_method, numerator, denominator, true_relat, sep = "-")) %>%
@@
   #geom_segment(aes(x = FNR, xend = FNR, y = FPR - 2 * se, yend = FPR + 2 * se)) +
-  facet_wrap(~ ckmr_sim_group, nrow = 4) +
+  facet_wrap(~ group, nrow = 4) +
   scale_y_continuous(trans = "log10") +
@@
   geom_line(aes(colour = `Error Category`, linetype = `Error Category`)) +
-  facet_wrap(~ ckmr_sim_group, nrow = 4) +
+  facet_wrap(~ group, nrow = 4) +
   scale_y_continuous(trans = "log10") +

### 2024-10-28 tex/body-text-of-paper.tex
Tweak the names of groups in the CKMRsim plot

@@
 collections into 10 different groups.  The members of each group are
-relatively, genetically similar.  For all cases, we assumed that the genotyping error rate
+relatively, genetically similar, but the groups do not necessarily correspond to reporting
+units, because, in some cases, we desired FPRs and FNRs for a single
+collection.   For all cases, we assumed that the genotyping error rate
 was 1\% per locus and the genotyping error model was the default
@@
 rates (FNRs) for distinguishing between different relationships.
-Groupings for panels with more than one population are as follows: SONC = \{CRFS, SmRF, BCkF\}; TRH = TRS, TRF; CCC = \{ERF, RRF\}; CVF = BCF, MDF, SJRF); FRH = \{FRHS, FRHF\} relationship inference analysis are as follows  {\bf (a)} FPRs calculated
+Groupings for panels that are neither a previously defined reporting unit nor a single collection as follows: TRH = \{TRS, TRF\}; FRH = \{FRHS, FRHF\}; CVF = \{BCF, MDF, SJRF\}.
+{\bf (a)} FPRs calculated
 using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),

### 2024-10-29 tex/body-text-of-paper.tex

@@
 rates (FNRs) for distinguishing between different relationships.
-Groupings for panels that are neither a previously defined reporting unit nor a single collection as follows: TRH = \{TRS, TRF\}; FRH = \{FRHS, FRHF\}; CVF = \{BCF, MDF, SJRF\}.
+Groupings for panels that are neither a previously defined reporting unit nor a single collection (Table \protect\ref{tab:samples}) are as follows: TRH = \{TRS, TRF\}; FRH = \{FRHS, FRHF\}; CVF = \{BCF, MDF, SJRF\}.
 {\bf (a)} FPRs calculated

### 2024-10-30 tex/body-text-of-paper.tex
changes from overleaf

@@
 as well as individual populations within those ESUs (Fig.~\ref{fig:gsi}).
-Accurate identification of the distinct ecotypes of Chinook salmon in the
-California Central Valley is of utmost importance for designing and implementing
-effective conservation and management actions. The CCV is one of the largest river
-basins on the west coast of North America and has the highest diversity of recognized Chinook salmon
+The California Central Valley (CCV) is one of the largest river
+basins on the west coast of North America and drains the Sierra Nevada and southern Cascade
+mountain ranges. It has the highest diversity of recognized Chinook salmon
 ecotypes in the species range, with four named ecotypes, two of which are protected
-under the US Endangered Species Act. This basin drains the Sierra Nevada and southern Cascade
-mountain ranges and is the source of water for several major urban areas as well as the
-largest and most valuable agriculture area in North America. As such, monitoring and
-evaluation of individual ecotypes and populations is of critical importance. This has been challenging,
+under the US Endangered Species Act. It is also the source of water for several major urban areas as well as the
+largest and most valuable agriculture area in North America. As such, accurate identification of the distinct ecotypes of CCV Chinook salmon is of utmost importance for monitoring and
+evaluation of individual ecotypes, as well as designing and implementing
+effective conservation and management actions. This has been challenging,
 given the recent common ancestors of these ecotypes \citep{clemento2014evaluation}, and
@@
 genetically identified as winter-run salmon clearly carry the winter-run genomic background and
-are strays from the winter-run stock, as these 
-could not realistically be misidentified on the basis of the genetic data.
+are strays from the winter-run stock, as these could not realistically be misidentified on the 
+basis of the genetic data.
+
+Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns 
+of population relationships that are coincident with previous work (Clemento et al. 2014; 
+Kinziger et al. 2013), emphasizing the distinction between Chinook salmon populations 
+in Coastal California and the Central Valley. These analyses also uncover some additional patterns, 
+including the clear presence of mixed ancestry in the Coastal California population in the 
+Russian River, which is consistent with its location, at the southernmost edge of the 
+Coastal Chinook salmon distribution and proximate to the mouth of the Sacramento River 
+(the Golden Gate). In the Central Valley, the clustering results emphasize the genetic 
+distinction of the SWR population, which is likely at least partially due to 
+the extreme bottleneck that it passed through 15-20 generations prior. The inclusion 
+of the LFAR markers also resulted in the clear distinction of the CHLF group, which 
+is coincident with the long-known phenotypic distinction of this population, but represents a novel 
+genetic result. Moreover, at K = 7, the fall-run reporting group breaks into two clusters, 
+which are distributed across almost all of the spring and fall-run populations in the 
+Central Valley, albeit not equivalently, emphasizing both current and historical migration and 
+gene flow between them.
 
@@
 typically relies on many hundreds of microhaplotype markers \citep{baetscher2018microhaplotypes} or thousands of SNP markers \citep{hillary2018genetic}.
-
-Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns of population relationships that are
-coincident with previous work (Clemento et al. 2014; Kinziger et al. 2013), emphasizing the distinction between Chinook salmon
-populations in Coastal California and the Central Valley. These analyses also uncover some additional patterns, including the clear
-presence of mixed ancestry in the Coastal California population in the Russian River, which is consistent with its location, at the 
-southernmost edge of the Coastal Chinook salmon distribution and proximate to the mouth of the Sacramento River (the Golden Gate). In the
-Central Valley, the clustering results emphasize the genetic distinction of the SWR population, which is likely at least partially due to 
-the extreme bottleneck that it passed through 15-20 generations prior. The inclusion of the LFAR markers also resulted in the clear 
-distinction of the CHLF group, which is coincident with the long-known phenotypic distinction of this population, but represents a novel 
-genetic result. Moreover, at K = 7, the fall-run reporting group breaks into two clusters, which are distributed across almost all of the 
-spring and fall-run populations in the Central Valley, albeit not equivalently, emphasizing both current and historical  migration and 
-gene flow between them.

### 2024-10-30 tex/body-text-of-paper.tex
Use Table~ and Figure~ throughout.

@@
 three rivers of the California coast (Russian, Eel, and Smith)
-(Fig.~\ref{fig:map}, Table~\ref{tab:samples}).
+(Figure~\ref{fig:map}, Table~\ref{tab:samples}).
 %%%%%%%%%%%%%
@@
 \LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run.
-Names of each location are followed by the location codes (Tab.~\ref{tab:samples}).
+Names of each location are followed by the location codes (Table~\ref{tab:samples}).
 Abbreviations used in location names as follows:
@@
 a pictorial representation of the proportion of each collection sampled in each month.  Colors in
-this column correspond to run-timing groups as described in the caption of Fig.~\protect\ref{fig:map}.}
+this column correspond to run-timing groups as described in the caption of Figure~\protect\ref{fig:map}.}
 \label{tab:samples}
@@
 All samples were sequenced on a MiSeq using $2\times 75$-cycle paired-end sequencing protocols.
-Primer sequences and concentrations for making the primer pool are in Supplemental Table 1.
+Primer sequences and concentrations for making the primer pool are in Supplemental Data~1.
 
@@
 late-fall fish to fall-run fish in the Feather River Hatchery and of late-fall fish to the fall-run fish in
-the San Joaquin River (Fig.~\ref{fig:lfar-assoc}).
+the San Joaquin River (Figure~\ref{fig:lfar-assoc}).
 
-The Manhattan plot of Fig.~\ref{fig:lfar-assoc} shows sporadic SNPs with low association
+The Manhattan plot of Figure~\ref{fig:lfar-assoc} shows sporadic SNPs with low association
 $p$-values.  These are not shared between the two different late-fall to fall comparisons, and
@@
 
-The total number of alleles per locus varied between 1 and 10 (Fig.~\ref{fig:num-alle}), with the
+The total number of alleles per locus varied between 1 and 10 (Figure~\ref{fig:num-alle}), with the
 only locus with a single allele being the species-specific marker Ots\_coho001\_05\_32691399
@@
 {\em a priori} knowledge about which collections in the baseline could and should be
-grouped together into reporting units (Fig.~\ref{fig:struct}).  At every value of $K$, CLUMPAK discerned at least
-12 of 20 replicates in the major mode (Fig.~\ref{fig:struct}).   
+grouped together into reporting units (Figure~\ref{fig:struct}).  At every value of $K$, CLUMPAK discerned at least
+12 of 20 replicates in the major mode (Figure~\ref{fig:struct}).   
 Clustering solutions in the minor
-modes generally converged on a single alternative solution and appear in Fig.~\ref{fig:minor-modes}.
+modes generally converged on a single alternative solution and appear in Figure~\ref{fig:minor-modes}.
 %%%%%%%%%%%%%%%%%%%%
@@
 20 assignments are discarded, and half of those (10) were the incorrect assignments.  Thus,
-at the 0.8 threshold, the total fraction of correct assignments is 99.1\% (Fig.~\ref{fig:eighty}).
+at the 0.8 threshold, the total fraction of correct assignments is 99.1\% (Figure~\ref{fig:eighty}).
 Notably with the $>0.8$ criterion, the misassignment rates from the abundant CV-Fall
@@
 
-The results are summarized in Fig.~\ref{fig:fprs}.
+The results are summarized in Figure~\ref{fig:fprs}.
 %%%%%%%%%%%%%%%%%%%%%
@@
 rates (FNRs) for distinguishing between different relationships.
-Groupings for panels that are neither a previously defined reporting unit nor a single collection (Table \protect\ref{tab:samples}) are as follows: TRH = \{TRS, TRF\}; FRH = \{FRHS, FRHF\}; CVF = \{BCF, MDF, SJRF\}.
+Groupings for panels that are neither a previously defined reporting unit nor a single collection (Table~\protect\ref{tab:samples}) are as follows: TRH = \{TRS, TRF\}; FRH = \{FRHS, FRHF\}; CVF = \{BCF, MDF, SJRF\}.
 {\bf (a)} FPRs calculated
@@
 exceedingly low for all population groups, even when using stringent false negative thresholds as
-low as 0.05 (Fig.~\ref{fig:fprs}a).
+low as 0.05 (Figure~\ref{fig:fprs}a).
 For example, the false positive rate for FS identification is less than one error in 10 million
@@
 facilities for assessing error rates between different kin groups, and results for these are
-summarised in (Fig.~\ref{fig:fprs}b). Here, again, focusing on the results for the genetically
+summarised in (Figure~\ref{fig:fprs}b). Here, again, focusing on the results for the genetically
 depauperate SRW, and using an FNR of 0.1 we see that the chance of incorrectly
@@
 unprecedented power for identifying fish from all of the Chinook salmon ESUs in California,
-as well as individual populations within those ESUs (Fig.~\ref{fig:gsi}).
+as well as individual populations within those ESUs (Figure~\ref{fig:gsi}).
 The California Central Valley (CCV) is one of the largest river
@@
 near complete accuracy in distinguishing these 'stocks'. Moreover, the few fish that are
-apparently misidentified (Fig. 3) are likely to be primarily migrants and not true misidentifications.
+apparently misidentified (Figure~\ref{fig:gsi}) are likely to be primarily migrants and not true misidentifications.
 For example, the three fish that were field-characterized as spring-run from Butte Creek, but are
@@
 %\comment{Maybe discuss: A large fraction of the Feather River
-%Hatchery Spring-run (FRHS) collection into a cluster (colored orange in Fig.~\ref{fig:struct}).
+%Hatchery Spring-run (FRHS) collection into a cluster (colored orange in Figure~\ref{fig:struct}).
 %A large fraction of the samples from the fall-run hatchery
@@
 one parent is sampled, as well as identifying pairs of full siblings when no parents are sampled.
-Groups of full siblings larger than two are identified even more accurately than expressed in Fig.~\ref{fig:fprs}.
+Groups of full siblings larger than two are identified even more accurately than expressed in Figure~\ref{fig:fprs}.
 Although some additional errors might be

### 2024-10-31 tex/body-text-of-paper.tex

@@
 
-The results are summarized in Figure~\ref{fig:fprs}.
 %%%%%%%%%%%%%%%%%%%%%
@@
 
+Examination of the geographic patterns of allele frequencies for the RoSA-associated loci found a
+clear instance of the early-migrating haplotype present in the California Coastal Chinook salmon lineage.
+This is of note because this lineage does not currently have an early migrating component, although it
+has long been speculated that the Eel River, at least, historically harbored early-migrating Chinook salmon
+and early migrating steelhead ({\em O. mykiss}) persist in the basin \citep{bjorkstedt2005analysis}. Moreover, we
+discovered several recombinant RoSA haplotypes in both of the CCC populations that indicate a long-standing
+presence of both the early- and late-migrating haplotypes in these populations. This is consistent with the
+idea that early-migrating populations arise through migration of small numbers of individuals carrying 
+E-lineage haplotypes, in either heterozygous or homozygous form, which are then positively selected
+when appropriate habitat conditions exist. A similar pattern of recombination, due to long-standing
+co-occurence of the E and L haplotypes in heterozygotes, has been previously documented in the Klamath 
+River \citep{thompson2020complex}.
+
 As sequencing costs drop, whole genome sequence (WGS) data may become the preferred
@@
 Such an approach would be highly flexible and extensible, as it would allow for the use
-just about any marker type, including reduced representation DNA sequencing (e.g. RADseq,
-\citealt{meek2020identifying,thompson2024genomics}),
+just about any marker type, including reduced representation DNA sequencing
+\citep[e.g.,~RADseq,][]{meek2020identifying,thompson2024genomics},
 as regional monitoring datasets would not need to be replicable.  For the present, however,

### 2024-10-31 tex/body-text-of-paper.tex
fix the citealts

@@
 stock identification
-(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
+\citep[GSI:~][]{milner1982genetic,beacham2004dna,seeb2007development}
 and parentage based tagging
-(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).
+\citep[PBT:~][]{anderson2006power, garza2007large, abadia2013large, steele2013validation}.
 
@@
 in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring
-run, and 188 fall/late-fall run. We divided the fish identified as fall/late-fall (via the 125-locus genetic
-stock identification data set of \citealt{thompson2020complex} into fall run and late-fall run according
+run, and 188 fall/late-fall run. We divided the fish identified
+from the 125-locus genetic
+stock identification data set of \citet{thompson2020complex}  as fall/late-fall into fall run and late-fall run according
 to their sampling date at Keswick:  99 fish arriving before Dec 26 were classified as fall run and 89 fish arriving after Dec 26 were
@@
 
-We include a marker that targets the sex-determining region ({\em sdY} on the Y chromosome in Chinook salmon.
+We include a marker that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
 This marker characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}.
@@
 The paired-end sequence reads were joined together using Fast Length Adjustment
-of SHort reads (FLASH: \citealt{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
+of SHort reads \citep[FLASH:~][]{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
 Mapped reads were converted to BAM files using SAMtools (Li et al.
@@
 Each locus was assessed for the presence of null alleles using the R package `whoa'
-(\url{https://github.com/eriqande/whoa}; {\em see} \citealt{hendricks2018recent}).
+\citep[\url{https://github.com/eriqande/whoa};  see][]{hendricks2018recent}.
 
@@
 broken down according to the genotype of the fish at the markers within
-the RoSA.  For this we identified the RoSA genotype (EE, EL, or LL, see
-\citealt{thompson2020complex}) from the concordance of the genotypes
+the RoSA.  For this we identified the RoSA genotype \citep[EE, EL, or LL, see~][]{thompson2020complex} from the concordance of the genotypes
 at all 8 markers described in \citep{thompson2020complex}, requiring no missing

### 2024-11-05 tex/body-text-of-paper.tex
Stuff

@@
 
-We include a marker that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
-This marker characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}.
+We include a novel marker that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
+We designed this marker through examination of whole-genome sequence data in multiple populations of California Chinook salmon (Thompson et al., 2020.)
+This marker characterizes genetic sex in diverse California salmon populations more accurately than previously published marker assays \comment{\citep{von2015development}}.
 And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in

### 2024-11-06 tex/abstract.tex
Eric's near-final edits

@@
 pedigree reconstruction to identify relationships between individuals within populations
-have become common. We present, here, a novel set of
+have also become common. We present, here, a novel set of
 multi-allelic microhaplotype genetic markers for Chinook salmon which provide unprecedented
-resolution for population discrimination and relationship identification for a rapidly and
+resolution for population discrimination and relationship identification from a rapidly and
 economically assayed panel of markers. We show how this set of microhaplotypes

### 2024-11-06 tex/body-text-of-paper.tex
Eric's near-final edits

@@
 Stone National Fish Hatchery for supplementation (released as juveniles into the river) and as a captive broodstock;
-sampling of this population was performed by hatchery staff during spawning. In Central Valley Rivers with both
+sampling of this population was performed by hatchery staff during spawning. In Central Valley rivers with both
 spring-run and fall-run ecotypes (i.e.
@@
 early returning fish in May and April (expressing the spring-run life history) are visibly tagged and re-released
-to the river. Subsequently in the Fall, when hatchery raceways are reopened, only fish with the visible tag are included
+to the river. Subsequently in the fall, when hatchery raceways are reopened, only fish with the visible tag are included
 as spring-run broodstock, and the rest are included in the fall-run broodstock. We sampled putative spring-run and fall-run fish 
@@
 Similar to broodstock collections in the CCV, samples from spring- and fall-run fish in the Trinity River were selected on the basis of spawn timing at the Trinity River Hatchery with migration-timing ecotype then confirmed by RoSA genotype.
-Elsewhere, samples were taken from live broodstock at Cole Rivers (Rogue), Rowdy Creek (Smith), Iron Gate (Klamath) and Warm Springs (Russian River) Hatcheries, where only fall-run fish have been documented.
+Elsewhere, adult samples were taken from live broodstock at Rowdy Creek (Smith), Iron Gate (Klamath) and Warm Springs (Russian River) Hatcheries and juveniles were sampled from Blue Creek (Klamath); all locations where only fall-run fish have been documented.
+Spring run adult fish were sampled from the spring-run broodstock at Cole Rivers Hatchery on the Rogue River.
 Samples from the Eel River were taken from live fish ascending a ladder in the upper basin and were confirmed as fall-run ecotype by RoSA genotype.
@@
 \caption{\footnotesize Collections of fish in the reference baseline.  The Sampling Months column shows
-a pictorial representation of the proportion of each collection sampled in each month.  Colors in
+a pictorial representation of the proportion of each collection sampled in each month. BCkF were sampled as juveniles, all other collections were of adults. Colors in
 this column correspond to run-timing groups as described in the caption of Figure~\protect\ref{fig:map}.}
@@
 
-Stacks v1.45 \cite{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome
+Stacks v1.45 \citep{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome
 at the time, a {\em de novo} assembly was performed.  Initial filtering retained RAD loci with two or more SNPs present
@@
 or greater and a read depth of 10 or higher. Further filtering and analysis of the loci, including
-assessment of Hardy-Weinberg equilibrium \citep{hardy1908mendelian} was conducted in MICROHAPLOT (Ng et al., DOI: 10.5281/zenodo.820110).
+assessment of Hardy-Weinberg equilibrium \citep{hardy1908mendelian} was conducted in MICROHAPLOT (Ng {\em et al.}, DOI: 10.5281/zenodo.820110).
 Loci that produced more than two haplotypes within an individual, as well as those clearly violating Hardy-Weinberg equilibrium within populations were removed.
@@
 We include a novel marker that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
-We designed this marker through examination of whole-genome sequence data in multiple populations of California Chinook salmon (Thompson et al., 2020.)
-This marker characterizes genetic sex in diverse California salmon populations more accurately than previously published marker assays \comment{\citep{von2015development}}.
+We designed this marker through examination of whole-genome sequence data in multiple populations of California Chinook salmon \citep{thompson2020complex}.
+This marker characterizes genetic sex in diverse California salmon populations more accurately than previously published marker assays \citep{von2015development}.
 And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
@@
 
-Amplicon sequencing was performed with the Genotyping-in-Thousands by sequencing (GT-seq)
+Amplicon sequencing was performed with the Genotyping-in-Thousands-by-sequencing (GT-seq)
 method \citep{campbell2015genotyping}. All loci were
@@
 The paired-end sequence reads were joined together using Fast Length Adjustment
-of SHort reads \citep[FLASH:~][]{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci using BWA (Li and Durbin 2010).
-Mapped reads were converted to BAM files using SAMtools (Li et al.
-2009) and freebayes (Garrison and Marth 2012) was used to call variants.
+of SHort reads \citep[FLASH:~][]{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci (for the novel microhaplotypes, SNPlicons and RoSA markers) or from the entire Otsh\_v1.0 genome
+(for the LFAR, WRAP, VGLL3, and Six6 markers) using BWA mem\citep{bwa-mem2009}.
+Mapped reads were converted to BAM files using SAMtools \citep{li2009sequence} and freebayes \citep{garrison2012haplotype} was used to call variants.
 
@@
 the RoSA.  For this we identified the RoSA genotype \citep[EE, EL, or LL, see~][]{thompson2020complex} from the concordance of the genotypes
-at all 8 markers described in \citep{thompson2020complex}, requiring no missing
-data at any of those 8 markers, and discarding data from 12 individuals
+at all 8 markers described in \citep{thompson2020complex}, requiring that no genotypes be missing
+within and individual from any of those 8 markers, and discarding data from 12 individuals
 in which the E or L haplotypes had clearly recombined.
@@
 The total number of alleles per locus varied between 1 and 10 (Figure~\ref{fig:num-alle}), with the
-only locus with a single allele being the species-specific marker Ots\_coho001\_05\_32691399
-used to distinguish coho from Chinook salmon. Biallelic
-loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
+only locus bearing a single allele in Chinook salmon being the species-specific marker Ots\_coho001\_05\_32691399. 
+Biallelic loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
 showed more than two alleles, which has been shown to provide significantly more
@@
 
-At value of $K$ from 2 to 7, clusters identified by the program STRUCTURE
+At values of $K$ from 2 to 7, clusters identified by the program STRUCTURE
 generally corresponded to groupings of related populations, and confirmed
@@
 CV-Winter, CV-Spring, CV-Fall, and  CV-Late-Fall.  These groupings correspond
-largely to the divisions in the data found by STRUCTURE, they also correspond
+largely to the divisions in the data found by STRUCTURE, and they also correspond
 with divisions amongst the stocks as defined for management, and, as seen in the
@@
 SRW, were clearly fish from SRW that had been incorrectly sampled into a different
-collection, likely due to straying. (This is evident because SRW are so distinct that it is not possible
-to incorrectly identify them).  Of the remaining 20 misassignments,  8 involved fish from the
+collection, likely due to straying. (This is evident because SRW are so distinct that it is highly improbable they
+would be incorrectly identified).  Of the remaining 20 misassignments,  8 involved fish from the
 CHLF (late fall run) collection being misassigned to the CV-Fall reporting unit and 6 were
@@
 at the 0.8 threshold, the total fraction of correct assignments is 99.1\% (Figure~\ref{fig:eighty}).
-Notably with the $>0.8$ criterion, the misassignment rates from the abundant CV-Fall
+Notably with the $>$0.8 criterion, the misassignment rates from the abundant CV-Fall
 reporting unit to CV-Late-Fall or to CV-Spring both drop below 0.2\% (1/502).
@@
 
-While there is not a currently available method (like that of \citeauthor{anderson2006power}) for estimating error rates in
+While there is not a currently available method \citep[like that of][]{anderson2006power}) for estimating error rates in
 inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which 
@@
 relationship inference, gene regions identified in whole-genome sequence data
-for increased power for identification of specific populations and incorporates markers that have
-long been in use for both GSI and PBT in California (Clemento et al. 2014), by converting them
-into so-called 'SNPlicons'. This set of markers lays the foundation for a comprehensive 
+for increased power for identification of specific populations, and markers that have
+long been in use for both GSI and PBT in California \citep{clemento2014evaluation}, by converting them
+into so-called 'SNPlicons.' This set of markers lays the foundation for a comprehensive 
 genetic monitoring and evaluation effort that facilitates multiple types of inference 
@@
 ecotypes in the species range, with four named ecotypes, two of which are protected
-under the US Endangered Species Act. It is also the source of water for several major urban areas as well as the
-largest and most valuable agriculture area in North America. As such, accurate identification of the distinct ecotypes of CCV Chinook salmon is of utmost importance for monitoring and
+under the US Endangered Species Act. It is also the source of water for several major urban areas as well as, arguably,  
+the most productive and most valuable agriculture area in North America. As such, accurate identification of the distinct 
+ecotypes of CCV Chinook salmon is of utmost importance for monitoring and
 evaluation of individual ecotypes, as well as designing and implementing
@@
 Results from the model-based clustering analysis with the program STRUCTURE, revealed patterns 
-of population relationships that are coincident with previous work (Clemento et al. 2014; 
-Kinziger et al. 2013), emphasizing the distinction between Chinook salmon populations 
+of population relationships that are coincident with previous work 
+\citep{clemento2014evaluation,kinziger2013contemporary}, emphasizing
+the distinction between Chinook salmon populations 
 in Coastal California and the Central Valley. These analyses also uncover some additional patterns, 
@@
 This is of note because this lineage does not currently have an early migrating component, although it
-has long been speculated that the Eel River, at least, historically harbored early-migrating Chinook salmon
+has long been speculated that the Eel River, at least, historically harbored early-migrating Chinook salmon,
 and early migrating steelhead ({\em O. mykiss}) persist in the basin \citep{bjorkstedt2005analysis}. Moreover, we
@@
 data type for salmon genetics. If so, it may become possible to include
-all >200 genetically distinct Chinook salmon populations within a single standardized reference baseline
+all the >200 genetically distinct Chinook salmon populations in North America within a
+single standardized reference baseline
 constructed with WGS data
 that performs equally well at broad and regional scales \citep{desaixINPRESSpopulation}.
-Such an approach would be highly flexible and extensible, as it would allow for the use
+Such an approach would be highly flexible and extensible, as it would allow for the use of
 just about any marker type, including reduced representation DNA sequencing
@@
 one parent is sampled, as well as identifying pairs of full siblings when no parents are sampled.
-Groups of full siblings larger than two are identified even more accurately than expressed in Figure~\ref{fig:fprs}.
+Groups of full siblings larger than two can be identified even more accurately than expressed in Figure~\ref{fig:fprs},
+when using statistical approaches that account for the joint relationships between more than two individuals,
+such as COLONY \citep{wang2004sibship}.
 Although some additional errors might be
 expected when related but non-target kin pairs are sampled,  the false positive rates
-for these non-targe kin pairs are sufficiently small
+for these non-target kin pairs are sufficiently small
 for almost all realistic scenarios.

### 2024-11-06 tex/supplemental-figures.tex
Eric's near-final edits

@@
 \caption[Assignment table for fish with scaled likelihood $ > 0.8$]{\footnotesize Assignment table
-like that in Fig.~\ref{fig:gsi}b in the paper, but constrained so that only fish assigning
+like that in Figure~\ref{fig:gsi}b in the paper, but constrained so that only fish assigning
 to a reporting unit with scaled likelihood greater than 0.8 are included.}
@@
 \caption[Assignment table by RoSA genotype]{\footnotesize Assignment table
-like that in Fig.~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
+like that in Figure~\ref{fig:gsi}b in the paper, but with numbers according to genotypes
 at the RoSA. In each cell, the top left entry gives the number of EE (early run allele homozygotes)
@@
 	for the 61 candidate SNPs (in red) to design for winter-run-associated polymorphisms
-	(WRAPs).  Other sites are shown in black.  The $x$-axis shows position on
+	(WRAPs).  Other sites are shown in blue.  The $x$-axis shows position on
 	each chromsome in the Otsh\_v1.0 assembly.  The chromsome name is indicated in the facet headers by

### 2024-11-06 tex/supplemental-text.tex
Eric's near-final edits

@@
 analysis.  We first investigated the distribution throughout the genome of $|d|$, the absolute difference
-of the alternate allele frequency between the two groups  (Fig.~\ref{fig:wrap-absdiff}). This revealed
+of the alternate allele frequency between the two groups  (Figure~\ref{fig:wrap-absdiff}). This revealed
 that many loci had large values of $|d|$, but there were several
@@
 also adjacent to at least one window in which $>10\%$ of sites with $|d|>0.5$ also had $|d|>0.9$ 
-(Fig.~\ref{fig:wrap-slide-window}).  Within the windows found on four different chromosomes using
+(Figure~\ref{fig:wrap-slide-window}).  Within the windows found on four different chromosomes using
 the above criteria, we then attempted to design amplicons to type the SNPs at a subset of the sites
 within each window.  We chose all sites with $|d|>0.975$, as well as the 8 SNPs on each chromosome
-with the highest values of $|d|$, yielding 61 candidate SNPs (Fig.~\ref{fig:wrap-candi}).
+with the highest values of $|d|$, yielding 61 candidate SNPs (Figure~\ref{fig:wrap-candi}).
 
@@
 GATK.  The resulting VCF file was used to estimate allele frequency differences between
-winter run and feather river spring.  We also processed the sequence data using
-microhaplot (\url{https://github.com/ngthomas/microhaplot}) and visually inspected
-loci for consisten allele depth ratio and numbers of haplotypes.  We then chose 24 amplicons for further
+winter run and Feather River spring run.  We also processed the sequence data using
+the R package,
+'microhaplot' (\url{https://github.com/ngthomas/microhaplot}), and visually inspected
+loci for consistent allele depth ratio and numbers of haplotypes.  We then chose 24 amplicons for further
 testing on the basis of allele frequency differences between Feather River spring run and
@@
 8, 12, and 16.  The estimated frequencies of all the alleles present in the reference baseline
-in those three amplicons shows that there are not fixed differences between
+in those three amplicons shows that there are not fixed differences at these markers between
 winter run and all other reporting units (Table~\ref{tab:wrap-freqs}).

### 2024-11-06 002-GSI_and_Fst.Rmd
Change a footnote reference in Table 1

@@
     Run = run_timing_for_table,
-    `Reporting Unit$^4$` = repunit,
+    `Reporting Unit$^3$` = repunit,
     `ESU$^4$` = ESU,

### 2024-11-06 003-ckmr-sim-evaluation.Rmd
Fix ckmr-comp figure to have the right pop names

@@
   unnest(Qsims) %>%
-  select(condition, ckmr_sim_group, true_relat, logl_ratio)
+  select(condition, ckmr_sim_group, true_relat, logl_ratio) %>%
+  mutate(  # change the names around
+    group = str_replace(ckmr_sim_group, "^ck-", ""),
+    group = case_match(
+      group,
+      "LF" ~ "CHLF",
+      "IGH" ~ "IGHF",
+      "SONC" ~ "SO-NCal-Coast",
+      "CCC" ~ "Cent-Cal-Coast",
+      .default = group
+    ),
+    .before = ckmr_sim_group
+  ) %>%
+  mutate(group = factor(group, levels = c("SO-NCal-Coast", "IGHF", "TRH", "Cent-Cal-Coast", "SRW", "BCS", "MDS", "FRH", "CVF", "CHLF")))
 ```
@@
   geom_density(alpha = 0.5) +
-  facet_wrap(~ ckmr_sim_group + condition, ncol = 1) +
+  facet_wrap(~ group + condition, ncol = 1) +
   theme_bw() +

### 2024-11-06 tex/supplemental-figures.tex
Fix ckmr-comp figure to have the right pop names

@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{images/cmkr-comp-figure-crop.pdf}
+\includegraphics[width=0.9\textwidth]{images/ckmr-comp-figure-crop.pdf}
 \end{center}
@@
 amplicon ("top\_snps" in the panel headers).  Results shown for four hatchery collections in California 
-(ck-FRH: Feather River Hatchery [spring and fall]; ck-IGH: Iron Gate Hatchery; ck-SRW: Sacramento Winter Run;
-ck-TRH: Trinity River Hatchery [spring and fall]).  The density plots show
+(FRH: Feather River Hatchery [spring and fall]; IGH: Iron Gate Hatchery; SRW: Sacramento Winter Run;
+TRH: Trinity River Hatchery [spring and fall]).  The density plots show
 the distribution of log-likelihood ratios for Parent-Offspring vs Unrelated for four different relationships:

### 2024-11-06 extras/Reordering-forward-and-reverse-primers/primer-pair-reorder.R
Fix forward and reverse primer sequences
(they were reversed for the snplicons)

@@ -1,0 +1,47 @@
+
+
+# Anthony noted that the primer pairs were reversed in the supp data.
+# So, we just want to fix that here.
+
+# Here is what Anthony said:
+# I was just further perusing this table and it looks like the
+# forward and reverse primers are swapped for the SNPlicon markers.
+# It certainly doesn't really matter for the chemistry, but looking
+# back at the actual IDT order forms, the forward primers start
+# with CGA and the reverse with GTG.
+
+
+# So, let's put the current state of things into the directory where this
+# script is and fix it.
+library(tidyverse)
+x <- read_csv("extras/Reordering-forward-and-reverse-primers/OLD-Calif-Chinook-Amplicon-Panel-Information.csv")
+
+
+# this should be easy:
+check_it <- x %>%
+  mutate(
+    tmp_forward = case_when(
+      str_detect(ForwardPrimerSequence, "^CGA") ~ ForwardPrimerSequence,
+      str_detect(ReversePrimerSequence, "^CGA") ~ ReversePrimerSequence,
+      TRUE ~ "BAD NEWS FORWARD!!"
+    ),
+    tmp_reverse = case_when(
+      str_detect(ReversePrimerSequence, "^GTG") ~ ReversePrimerSequence,
+      str_detect(ForwardPrimerSequence, "^GTG") ~ ForwardPrimerSequence,
+      TRUE ~ "BAD NEWS REVERSE!!"
+    ),
+    .before = ForwardPrimerSequence
+  )
+
+
+# that all checks out, so now just copy those to the original columns and remove
+final <- check_it %>%
+  mutate(
+    ForwardPrimerSequence = tmp_forward,
+    ReversePrimerSequence = tmp_reverse
+  ) %>%
+  select(-tmp_forward, -tmp_reverse)
+
+
+write_csv(final, "inputs/Calif-Chinook-Amplicon-Panel-Information.csv")
+write_csv(final, "tex/supp_data/Supp-Data-1-Amplicon-info.csv")

### 2024-11-07 tex/body-text-of-paper.tex

@@
 
-As sequencing costs drop, whole genome sequence (WGS) data may become the preferred
-data type for salmon genetics. If so, it may become possible to include
-all the >200 genetically distinct Chinook salmon populations in North America within a
-single standardized reference baseline
-constructed with WGS data
-that performs equally well at broad and regional scales \citep{desaixINPRESSpopulation}.
-Such an approach would be highly flexible and extensible, as it would allow for the use of
-just about any marker type, including reduced representation DNA sequencing
-\citep[e.g.,~RADseq,][]{meek2020identifying,thompson2024genomics},
-as regional monitoring datasets would not need to be replicable.  For the present, however,
-baselines tailored to specific regions are essential for regional management questions and
-targeted sequencing approaches have proven to be the most practical for large-scale
-applications of either GSI, PBT or an integrated monitoring program \citep{beacham2021parentage}.
-
 %\comment{Maybe discuss: A large fraction of the Feather River
@@
 typically relies on many hundreds of microhaplotype markers \citep{baetscher2018microhaplotypes} or thousands of SNP markers \citep{hillary2018genetic}.
+
+As sequencing costs drop, whole genome sequence (WGS) data may become the preferred
+data type for salmon genetics. If so, it may become possible to include
+all the >200 genetically distinct Chinook salmon populations in North America within a
+single standardized reference baseline constructed with WGS data
+that performs equally well at broad and regional scales \citep{desaixINPRESSpopulation}.
+Such an approach would be highly flexible and extensible, as it would allow for the 
+assignment of unknown-origin fish using
+just about any marker type, including reduced representation DNA sequencing
+\citep[e.g.,~RADseq,][]{meek2020identifying,thompson2024genomics},
+as the variation used by such region-specific panels of markers should be contained within the WGS baseline dataset.
+For the present, however,
+baselines tailored to specific regions are essential for regional management questions and
+targeted sequencing approaches have proven to be the most practical for large-scale
+applications of either GSI, PBT or an integrated monitoring program \citep{beacham2021parentage}.
+
+\section*{Acknowledgements}
+Many people from various agencies and watershed groups contributed to the collection
+of the samples analyzed here, including staff from the California Department of Fish and Wildlife,
+National Marine Fisheries Service, US Fish and Wildlife Service, US Army Corps of Engineers and 
+Pacific States Marine Fisheries Commission. We also thank Elena Correa, Nicole Anderson and Libby
+Gilbert-Horvath for contributions to data collection. The manuscript was improved by comments from
+Jeff Rodzen.

### 2024-11-19 tex/body-text-of-paper.tex
Incorporate Cassie's edits

@@
 Using the whole genome sequencing data of \citet{thompson2020complex}, we also sought
-genomic regions that might provide diagnostic markers for the the Sacramento winter-run
+genomic regions that might provide diagnostic markers for the Sacramento winter-run
 Chinook salmon.  This ecotype is highly differentiated from all other Chinook salmon populations
@@
 Including all these markers in the panel allows them to be routinely typed, creating
- a data base for understanding the spatial and temporal
+ a database for understanding the spatial and temporal
 distribution of the alleles in this genomic region; however we have omitted these markers for calculating likelihoods for
@@
 And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
-coho salmon and Chinook salmon.  This is helpful when samplers have unwitting
+coho salmon and Chinook salmon.  This is helpful when samplers have unwittingly
 collected coho salmon from ocean catches.
@@
 and then purified using Agencourt AMPure XP magnetic beads.
-Libraries were quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New
-England BioLabs) before dilution and pooling for sequencing.
+Libraries were originally quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New
+England BioLabs); however more recently we have quanitifed our libraries
+with the Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific)
+before dilution and pooling for sequencing.
 All samples were sequenced on a MiSeq using $2\times 75$-cycle paired-end sequencing protocols.
@@
 at all 8 markers described in \citep{thompson2020complex}, requiring that no genotypes be missing
-within and individual from any of those 8 markers, and discarding data from 12 individuals
+within an individual from any of those 8 markers, and discarding data from 12 individuals
 in which the E or L haplotypes had clearly recombined.

### 2024-11-19 tex/supplemental-text.tex
Cassie's edits on supplement

@@
 large allele frequency differences between winter-run Chinook salmon and all the other
-Chinook salmon ecotypes in the Central Valley of California (CCV).  Because winter-run Chinook are
+Chinook salmon ecotypes in the Central Valley of California (CCV).  Because winter-run Chinook salmon are
 already highly differentiated from all others, we did not pursue an association study as used

### 2024-11-19 tex/body-text-of-paper.tex
Incorporate Mac's comments

@@
 Similar to broodstock collections in the CCV, samples from spring- and fall-run fish in the Trinity River were selected on the basis of spawn timing at the Trinity River Hatchery with migration-timing ecotype then confirmed by RoSA genotype.
-Elsewhere, adult samples were taken from live broodstock at Rowdy Creek (Smith), Iron Gate (Klamath) and Warm Springs (Russian River) Hatcheries and juveniles were sampled from Blue Creek (Klamath); all locations where only fall-run fish have been documented.
+Elsewhere, adult samples were taken from live broodstock at Rowdy Creek (Smith River), Iron Gate (Klamath River) and Warm Springs (Russian River) Hatcheries and juveniles were sampled from Blue Creek (Klamath River); all locations where only fall-run fish have been documented.
 Spring run adult fish were sampled from the spring-run broodstock at Cole Rivers Hatchery on the Rogue River.
@@
 The late-fall-run occurs only in the CCV and shares a genomic background with the more
-common CCV fall-run salmon ecotype, so has been refractory to previous GSI efforts with other marker types \citep{seeb2007development,clemento2014evaluation,thompson2024genomics}.
+common CCV fall-run salmon ecotype, so has been refractory to previous GSI efforts with other marker types \citep{seeb2007development,clemento2014evaluation,meek2016sequencing,thompson2024genomics}.
+
 Moreover, we show how the FRH spring-run lineage is

### 2024-11-19 tex/author-title-etc.tex
Eric's near-final edits
Incorporate Neil's comments

@@
 Clemento$^{1,2}$, Matthew A.
-Campell$^{1,2\dagger}$, Devon Pearse$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
-Thompson$^{1,2\ddag}$, John Carlos Garza$^{1,2}$}
+Campell$^{1,2\dagger}$, Devon Pearse$^{1,2}$, Anne K.
+Beulke$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
+Thompson$^{1,2\ddag}$, John Carlos Garza$^{1,2,\S}$}
 \author{Eric C.
@@
 Clemento$^{1,2}$, Matthew A.
-Campell$^{1,2,\dagger}$, Devon Pearse$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
-Thompson$^{1,2,\ddag}$, John Carlos Garza$^{1}$}
+Campell$^{1,2,\dagger}$, Devon Pearse$^{1,2}$, Anne K.
+Beulke$^{1,2}$, Cassie Columbus$^{1,2}$, Ellen Campbell$^{1,2}$, Neil F.
+Thompson$^{1,2,\ddag}$, John Carlos Garza$^{1,2,\S}$}
 
@@
 $^\dagger$Current address: Centre for Carbon, Water and Food, The University of Sydney, 380 Werombi Road, NSW 2570, Australia.
-$^\ddag$Current address: Pacific Shellfish Breeding Center, Agricultural Research Service, US Department of Agriculture, Newport, Oregon, USA.
+$^\ddag$Current address: Pacific Shellfish Research Unit, Agricultural Research Service, US Department of Agriculture, Newport, Oregon, USA.
 }

### 2024-11-19 tex/body-text-of-paper.tex
Incorporate Neil's comments

@@
 designation at the Feather River Hatchery is more complicated but it is managed by the hatchery practices in place:
-early returning fish in May and April (expressing the spring-run life history) are visibly tagged and re-released
+early returning fish in May and April (expressing the spring-run life history) are visibly tagged and returned
 to the river. Subsequently in the fall, when hatchery raceways are reopened, only fish with the visible tag are included
-as spring-run broodstock, and the rest are included in the fall-run broodstock. We sampled putative spring-run and fall-run fish 
+as spring-run broodstock, and the rest are included in the fall-run broodstock.
+This situation carries the potential for untagged spring run 
+phenotypes to be included as full-run spawners if they were not 
+encountered during the spring trapping season.
+ We sampled putative spring-run and fall-run fish 
 from these two broodstocks.
@@
 coho salmon and Chinook salmon.  This is helpful when samplers have unwittingly
-collected coho salmon from ocean catches.
+collected coho salmon.
 
@@
 largely to the divisions in the data found by STRUCTURE, and they also correspond
-with divisions amongst the stocks as defined for management, and, as seen in the
-figure, they correspond with groups of populations that can be reliably distinguished
+with divisions amongst the stocks as defined for management, and, as seen in
+Figure~\ref{fig:gsi}b, they correspond with groups of populations that can be reliably distinguished
 from one another for population assignment.
@@
 terms of RoSA genotypes (Figure~\ref{fig:rosa-gsi}). In this context, it is clear that the ``fall-run'' hatchery
-program in the Feather River both spawns and produces many spring-run fish and RoSA heterozygotes.
+program in the Feather River generates many spring-run fish and RoSA heterozygotes.
 Similarly, in other regions, such as the Trinity River, the RoSA markers are the only reliable means to distinguish fall-
@@
 distinction of the SWR population, which is likely at least partially due to 
-the extreme bottleneck that it passed through 15-20 generations prior. The inclusion 
+the extreme bottleneck that it passed through in the 1980's and 1990's \citep{hedrick1994effective}. The inclusion 
 of the LFAR markers also resulted in the clear distinction of the CHLF group, which 
@@
 Gilbert-Horvath for contributions to data collection. The manuscript was improved by comments from
-Jeff Rodzen.
+Jeff Rodzen. USDA is an equal opportunity provider and employer. Mention of trade names
+or commercial products in this publication is solely for the purpose of providing specific
+information and does not imply recommendation or endorsements by the U.S. Department of Agriculture.

### 2024-11-25 006-winter-run-vs-all-cv-alle-freqs.Rmd
Incorporate Devon's comments

@@
   xlab("Position along chromosome") +
-  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")
+  ylab("Absolute value of allele frequency difference, winter-run vs.\ non-winter-run")
 
@@
 ggsave(mh_plot + ylim(0.5, 1.0), filename = output_list$mh_plot,
-       height = , width = 12)
+       height = 7, width = 12)
 

### 2024-11-25 tex/body-text-of-paper.tex
Incorporate Devon's comments

@@
 
-Here, we present a reference baseline for Chinook salmon, focusing on the regional scale of
-rivers within the state of California, and particularly targeted to the complex population
-structure of Chinook salmon within California's Central Valley (CCV). Chinook salmon of the two main CCV river basins---the
-Sacramento and the San Joaquin---exhibit the greatest run-timing diversity
-within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
-(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California \citep{fisher1994past}.
-
 Chinook salmon are the largest of the Pacific salmonids and have historically been the target of extremely high-value and culturally important
@@
 
+Here, we present a reference baseline for Chinook salmon, focusing on the regional scale of
+rivers within the state of California, and particularly targeted to the complex population
+structure of Chinook salmon within California's Central Valley (CCV). Chinook salmon of the two main CCV river basins---the
+Sacramento and the San Joaquin---exhibit the greatest run-timing diversity
+within the species.  With four recognized ecotypes, delineated primarily on the basis of run timing
+(fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
+in freshwater every month of the year in California \citep{fisher1994past}.  Our previous work \citep{thompson2020complex} documented
+a strong association between these ecotypes and the alleles carried near the GREB1L and ROCK1 genes
+in a section of chromosome 28 that we termed the region of strongest association (RoSA), a pattern
+also observed in other river basins of the West Coast \citep[summarized in][]{waples2022implications}.
+
 In spite of their distinct phenotypes and life history patterns, the four ecotypes of Chinook salmon in the Central Valley are closely related
@@
 In the following, we present a new reference baseline that includes 1,636 Chinook salmon individuals
-from 17 collections, genotyped at 204 microhaplotypes \citep{baetscher2018microhaplotypes} from throughout the genome.
+from 17 collections, genotyped at 204 amplicon sequences \citep{baetscher2018microhaplotypes} from throughout the genome.
 This baseline is
@@
 This situation carries the potential for untagged spring run 
-phenotypes to be included as full-run spawners if they were not 
+phenotypes to be included as fall-run spawners if they were not 
 encountered during the spring trapping season.
@@
 with either fall-run or late-fall run being nearly fixed for one of the alleles, or
-\item An annotation from snpEff \citep{cingolani2012program} of “High” or “Moderate”, or
+\item An annotation from snpEff \citep{cingolani2012program} of ``High'' or ``Moderate,'' or
 \item $|d|>0.48$ and a lower confidence interval of $>$0.25 and a
@@
 try to find markers that could be diagnostic for the winter run, distinguishing the
-winter run from all other groups of Chinook salmon in the CCV.  We were unable to
-find a single diagnostic locus; however three markers from that effort are included in
+winter run from all other groups of Chinook salmon in the CCV.  We found
+a number of SNPs that were fixed for alternate alleles between the 16 winter-run fish
+and 84 non-winter run fish that we had sequenced; however after designing assays for many of these
+SNPs and testing them in a much larger sample, we did not find any truly diagnostic loci.
+Nonetheless three markers from that effort are included in
 the reference baseline.  The process of discovering them is described in the
@@
 In California, only the Sacramento and  Klamath basins have documented spring-run
-salmon and their historical occurrence in other basins has been unclear. We find no
-instance of spring-run fish in these coastal basins, but did identify one copy of the early-lineage
+salmon and their historical occurrence in other basins has been unclear. Although all fish
+sampled from coastal basins in our study were phenotypically fall run, we did identify one copy of the early-lineage
 RoSA haplotype in the Eel River. Moreover, we identified three additional copies of a
@@
 ecotypes in the species range, with four named ecotypes, two of which are protected
-under the US Endangered Species Act. It is also the source of water for several major urban areas as well as, arguably,  
-the most productive and most valuable agriculture area in North America. As such, accurate identification of the distinct 
+under the US Endangered Species Act. It is also the source of household water for
+tens of millions of California residents and millions of acres of arguably
+the most productive and valuable agriculture area in North America. As such, accurate identification of the distinct 
 ecotypes of CCV Chinook salmon is of utmost importance for monitoring and

### 2024-11-25 tex/GitLatexDiffIt.sh

@@ -1,0 +1,38 @@
+
+
+if [ $# -ne 1 ]; then
+    echo " Syntax: GitLatexDiffIt.sh  ref "; echo
+    echo "   where ref is a hash of tag for the starting commit you want.  "; echo
+    echo "  It will then be compared to the current working copy (even if uncommitted) "
+    echo
+    echo
+    echo " Example: GitLatexDiffIt.sh toCoauthors "
+    echo
+fi
+
+
+mkdir -p diffs
+
+M="diffs/main-diff-to-$1.pdf"
+S="diffs/supplement-diff-to-$1.pdf"
+
+# get the main document (note we have to latex supplement.tex first for xr)
+git-latexdiff \
+    --prepare 'pdflatex supplement.tex' \
+    --bibtex \
+    --main main.tex \
+    --output  $M \
+    $1  -- && \
+    open $M;
+
+
+
+# get the supplement
+git-latexdiff \
+    --prepare 'pdflatex main.tex' \
+    --bibtex \
+    --main supplement.tex \
+    --output  $S \
+    $1 -- &&
+    open $S
+

### 2024-11-27 tex/body-text-of-paper.tex
Incorporate Anne's comments

@@
 rivers within the state of California, and particularly targeted to the complex population
-structure of Chinook salmon within California's Central Valley (CCV). Chinook salmon of the two main CCV river basins---the
+structure of Chinook salmon within the California Central Valley (CCV). Chinook salmon of the two main CCV river basins---the
 Sacramento and the San Joaquin---exhibit the greatest run-timing diversity
@@
 in freshwater every month of the year in California \citep{fisher1994past}.  Our previous work \citep{thompson2020complex} documented
-a strong association between these ecotypes and the alleles carried near the GREB1L and ROCK1 genes
+a strong association between these ecotypes in the CCV and the alleles carried near the GREB1L and ROCK1 genes
 in a section of chromosome 28 that we termed the region of strongest association (RoSA), a pattern
@@
 among the reads on the sequencer.  We
-used freebayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
+used FreeBayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
 microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by freebayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
+that were considered multinucleotide or complex variants by FreeBayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
 
@@
 we first conducted a simple case-control GWAS using the low-coverage whole-genome sequencing
-data of \citet{thompson2020complex}, mapped to Otsh\_v1.0.  We conducted
+data of \citet{thompson2020complex}, mapped to the Otsh\_v1.0 Chinook salmon assembly
+(GenBank accession GCA\_002872995.1).  We conducted
 two separate association-study comparisons.
@@
 
-We include a novel marker that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
-We designed this marker through examination of whole-genome sequence data in multiple populations of California Chinook salmon \citep{thompson2020complex}.
-This marker characterizes genetic sex in diverse California salmon populations more accurately than previously published marker assays \citep{von2015development}.
+We include a novel amplicon that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
+We designed this amplicon to target the SNP described by \citet{bertho2022nonfunctional} that differentiates the functional {\em sdY} and a nonfunctional {\em sdY} pseudogene.
+Examination of whole-genome sequence data in multiple populations of California Chinook salmon \citep{thompson2020complex} shows that this
+marker characterizes genetic sex in diverse California salmon populations more accurately than previously published \citep{von2015development} marker assays.
 And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
@@
 consensus sequences (roughly 100--300 bp) around the markers
-to both the Chinook salmon version 1 (Otsh\_v1.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_002872995.1/}) and version 2  (Otsh\_v2.0; \url{https://www.ncbi.nlm.nih.gov/datasets/genome/GCF_018296145.1/}) genome assemblies,
+to both the Chinook salmon version 1 (Otsh\_v1.0, GenBank accession GCA\_002872995.1)
+and version 2  (Otsh\_v2.0, GenBank accession GCA\_018296145.1) genome assemblies,
 using both bwa-mem \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
@@
 (for the LFAR, WRAP, VGLL3, and Six6 markers) using BWA mem\citep{bwa-mem2009}.
-Mapped reads were converted to BAM files using SAMtools \citep{li2009sequence} and freebayes \citep{garrison2012haplotype} was used to call variants.
+Mapped reads were converted to BAM files using SAMtools \citep{li2009sequence} and FreeBayes \citep{garrison2012haplotype} was used to call variants.
 
@@
 recombinant haplotype that carries half of the RoSA SNPs from both the E and L lineages in the Eel and Russian rivers.
-No other clear recombinants were identified in the study populations, indicating that it arose in the California Coastal Chinook Salmon lineage, and providing further evidence of past presence of RoSA haplotypes associated with early migration.
+No other clear recombinants were identified in the study populations, indicating that it might have arisen in the California Coastal Chinook Salmon lineage, and providing further evidence of past presence of RoSA haplotypes associated with early migration.
 

### 2024-11-28 tex/body-text-of-paper.tex
Carlos and Eric finalize edits

@@
 (fall-, late-fall-, winter-, and spring-run), adult Chinook salmon can be found migrating or residing
-in freshwater every month of the year in California \citep{fisher1994past}.  Our previous work \citep{thompson2020complex} documented
-a strong association between these ecotypes in the CCV and the alleles carried near the GREB1L and ROCK1 genes
-in a section of chromosome 28 that we termed the region of strongest association (RoSA), a pattern
-also observed in other river basins of the West Coast \citep[summarized in][]{waples2022implications}.
+in freshwater every month of the year in California \citep{fisher1994past}.
 
@@
 In the following, we present a new reference baseline that includes 1,636 Chinook salmon individuals
-from 17 collections, genotyped at 204 amplicon sequences \citep{baetscher2018microhaplotypes} from throughout the genome.
+from 17 collections, genotyped at 204 loci, mostly microhaplotypes \citep{baetscher2018microhaplotypes}, from throughout the genome.
 This baseline is
@@
 as spring-run broodstock, and the rest are included in the fall-run broodstock.
-This situation carries the potential for untagged spring run 
-phenotypes to be included as fall-run spawners if they were not 
-encountered during the spring trapping season.
 We sampled putative spring-run and fall-run fish 
@@
 
-\subsubsection*{Run-timing associated markers}
+\subsubsection*{Adult-migration-timing associated markers}
 
 We include six amplicons that capture the eight  SNPs within the ``Region of Strongest Assocation'' (RoSA)
-between spring-run and fall-run fish on Chromosome 28 listed in Table~S3 of \citet{thompson2020complex}.
+between spring-run and fall-run fish, near the ROCK1 and GREB1L genes,  on Chromosome 28 listed in Table~S3 of \citet{thompson2020complex}.
 We add three more amplicons in the same RoSA region to the panel, as well as one more amplicon that
@@
 
-We include a novel amplicon that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
-We designed this amplicon to target the SNP described by \citet{bertho2022nonfunctional} that differentiates the functional {\em sdY} and a nonfunctional {\em sdY} pseudogene.
+We include an assay from a novel locus that targets the sex-determining region ({\em sdY}) on the Y chromosome in Chinook salmon.
+We designed this assay to target the SNP described by \citet{bertho2022nonfunctional} that differentiates the functional {\em sdY} and a nonfunctional {\em sdY} pseudogene.
 Examination of whole-genome sequence data in multiple populations of California Chinook salmon \citep{thompson2020complex} shows that this
-marker characterizes genetic sex in diverse California salmon populations more accurately than previously published \citep{von2015development} marker assays.
+marker characterizes genetic sex in diverse California salmon populations more accurately than previously published assays  \citep{von2015development}.
 And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
@@
 and then purified using Agencourt AMPure XP magnetic beads.
-Libraries were originally quantified by qPCR with the NEBNext Library Quant Kit for Illumina (New
-England BioLabs); however more recently we have quanitifed our libraries
-with the Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific)
+Libraries were quantified either by qPCR with the NEBNext Library Quant Kit for Illumina (New
+England BioLabs) or with the Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific)
 before dilution and pooling for sequencing.
@@
 try to find markers that could be diagnostic for the winter run, distinguishing the
-winter run from all other groups of Chinook salmon in the CCV.  We found
-a number of SNPs that were fixed for alternate alleles between the 16 winter-run fish
-and 84 non-winter run fish that we had sequenced; however after designing assays for many of these
-SNPs and testing them in a much larger sample, we did not find any truly diagnostic loci.
-Nonetheless three markers from that effort are included in
-the reference baseline.  The process of discovering them is described in the
-supplemental text section \ref{sec:wrap-methods}.
+winter run from all other groups of Chinook salmon in the CCV. Initial examination of sequence data 
+in the 16 winter-run fish and 84 non-winter run fish revealed
+a number of SNPs that were potentially diagnostic, but none were fixed for alternative alleles 
+in a larger sample.
+Nonetheless, three of these markers from that effort had large frequency differences
+between the groups and are included in the reference baseline baseline.  The process of 
+discovering them is described in the supplemental text section \ref{sec:wrap-methods}.
 
@@
 Feather River spring-migrating fish from the other CCV stocks with a predominantly fall-run
-genomic background. This is evident when the assignment matrix is enumerated in
+genomic background.
+
+FRHS fish are produced by selecting spring-tagged fish that again return in fall, 
+when the hatchery ladder reopens. This results in untagged spring run, fish not 
+encountered during the spring trapping season, being used as broodstock for the fall-run program.
+This is evident when the assignment matrix is enumerated in
 terms of RoSA genotypes (Figure~\ref{fig:rosa-gsi}). In this context, it is clear that the ``fall-run'' hatchery
@@
 In California, only the Sacramento and  Klamath basins have documented spring-run
-salmon and their historical occurrence in other basins has been unclear. Although all fish
-sampled from coastal basins in our study were phenotypically fall run, we did identify one copy of the early-lineage
-RoSA haplotype in the Eel River. Moreover, we identified three additional copies of a
+salmon and their historical occurrence in other basins has been unclear; however
+we identified three additional copies of a
 recombinant haplotype that carries half of the RoSA SNPs from both the E and L lineages in the Eel and Russian rivers.
-No other clear recombinants were identified in the study populations, indicating that it might have arisen in the California Coastal Chinook Salmon lineage, and providing further evidence of past presence of RoSA haplotypes associated with early migration.
+No other clear recombinants were identified in the study populations, suggesting that it arose in the California Coastal Chinook Salmon lineage, and providing further evidence of past presence of RoSA haplotypes associated with early migration.
 
@@
 distinction of the SWR population, which is likely at least partially due to 
-the extreme bottleneck that it passed through in the 1980's and 1990's \citep{hedrick1994effective}. The inclusion 
+the extreme bottleneck that it passed through in the 1980s and 1990s \citep{hedrick1994effective}. The inclusion 
 of the LFAR markers also resulted in the clear distinction of the CHLF group, which 

### 2025-03-14 008-lfar-gwas-and-markers.Rmd
Add lfar markers notebook

---
title: "Late Fall Associated Region: Genome Wide Association Study and Marker Selection"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

# Introduction

Here we do a GWAS to find a genomic region that is substantially differentiated between
the late-fall and fall runs in the Central Valley, and then we identify several
candidate sites to target for amplicon sequencing.

# Input and Output Paths

```{r}
start_time <- Sys.time()

# inputs:
input_list <- list(
  chr34_late_fall_mafs = "stored_results/5Mb_Chr34-late-fall.mafs.gz",
  chr34_cv_fall_mafs = "stored_results/5Mb_Chr34-cv-fall.mafs.gz",
  gwas_lf_v_sanjo = "stored_results/lfar-gwas/lf_v_sanjo-NC_037130.1.lrt0.gz",
  gwas_lf_v_frhf = "stored_results/lfar-gwas/lf_v_frhf-NC_037130.1.lrt0.gz",
  ots34_snpeff = "stored_results/thompson2020-vcf/ots34-5Mb-snpEff.vcf.gz",
  pvals = "stored_results/stored_p_vals.rds"
)
# outputs:
output_list <- list(
  late_fall_ids = "outputs/007/late-fall-ids.txt",
  cv_fall_ids = "outputs/007/cv-fall-ids.txt",
  candidates_for_assays = "outputs/007/candidates_for_assays.csv"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# GWAS

This is all executed with a Snakefile and associated files that reside in the
repository at: `SnakemakeSubDirs/cga-late-fall-bam-assoc`.  The contents of that
directory were originally part of another GitHub repo---simply `eriqande/cga-late-fall-bam-assoc`---but
have been inserted here for ease of archiving.

The figure created by that Snakefile is included in the paper.  We also have saved
the likelihood ratios for the various markers on Chromosome 34 (NC_037130.1)
for both the Late-fall vs San Joaquin fall
and the Late-Fall vs Feather River fall comparisons in `stored_results` so that we can have
those when visualizing allele frequencies.

Additionally, information for all markers with a p-value > 3 in both the SanJo-vs-LateFall and 
the FeatherRiverHatcheryFall-vs-LateFall comparisons was saved in `stored_results/stored_p_vals.rds`
for use now.

## Investigate GWAS pvalues

```{r}
library(tidyverse)
library(vcfR)

# read in the pvals and label the comparisons FRFH and SanJo and then pivot that 
# so we can compare values from the two analyses at each marker
pvals <- read_rds(input_list$pvals) %>%
  mutate(comparison = ifelse(comparison == "Late Fall versus Feather River Hatchery Fall", "FRHF", "SanJo")) %>%
  select(-(Major:Frequency), -(mh_label:xpos), -cgroup) %>%
  pivot_wider(
    names_from = comparison,
    values_from = c(LRT, neg_log10_p)
  )

# here are the total number of SNPs in each comparison:
FRH_snps <- 13477607
SJ_snps <- 12752546

# now, we can look for those that have have log10 pvals in both of the comparisons.
# maybe just plot them first:
ggplot(pvals, aes(x = neg_log10_p_FRHF, y = neg_log10_p_SanJo)) +
  geom_point() +
  geom_hline(yintercept = -log10(5e-8), linetype = "dashed") +
  geom_vline(xintercept = -log10(5e-8), linetype = "dashed")
```

So, that is only a single SNP in both:
```{r}
pvals %>%
  filter(neg_log10_p_FRHF > -log10(5e-8), neg_log10_p_SanJo > -log10(5e-8))
```

Now, let's consider a Bonferroni correction:
```{r}
pv2 <- pvals %>%
  mutate(
    bonFRHF = (10 ^ -neg_log10_p_FRHF) * FRH_snps,
    bonSJ = (10 ^ -neg_log10_p_SanJo) * SJ_snps,
  ) %>%
  arrange(bonFRHF)

pv2
```

So, that is only a single SNP that would be "significant" under a Bonferroni correction,
in the FRFH comparison.

Likewise only a single SNP is "significant" at a Bonferroni corrected 0.05 level
in the SJ comparison, hich we show by sorting the above table on the SanJo
values:
```{r}
pv2 %>%
  arrange(bonSJ)
```

# Marker Selection

This was originally done in
`/Users/eriq/Documents/git-repos/further-chinook-wgs/101-design-late-fall-assays.html`
and it has been transferred here, now.

Mac originally ran the association analysis between Late Fall and the two fall run groups in
the Central Valley (Feather and San Joaquin), finding significant differences
on Chromosome 34 (NC_037130.1).   We have seen the same thing with the GWAS just described.

Accordingly, we will investigate a 5 Mb chunck of Chromosome 34 to:

1. confirm those markers and explore patterns of differentiation in the region
2. See if the differentiation corresponds to annotated variants of large effect
3. identify candidate markers that we should be able to type in many more individuals.
We will provide sequences flanking these candidates so that
Anthony can design primers for them.

## Overview of allele frequency differences in the region

We will simply compare the allele frequencies estimated in the CV fall-run vs the late-fall
run using ANGSD.  I will do this in a 5 Mb chunk for a little context.

### Get the IDs.
```{r, message=FALSE}

# grab the meta data from the Science paper off of GitHub 
nooks <- read_csv("https://raw.githubusercontent.com/eriqande/thompson-et-al-2020-chinook-salmon-migration-timing/master/data/wgs-chinook-samples.csv")

# write out the names of the late-fall fish
nooks %>%
  filter(Population == "Coleman Hatchery Late Fall") %>%
  pull(vcf_name) %>%
  cat(., sep = "\n", file = output_list$late_fall_ids)

# same for the fall-run fish in the CV:
nooks %>%
  filter(Population %in% c("Feather River Hatchery Fall", "San Joaquin River Fall")) %>%
  pull(vcf_name) %>%
  cat(., sep = "\n", file = output_list$cv_fall_ids)
```

### Extract 5 Mb and the two sets of indivs

We show the code but don't actually do it, here.

```{sh, eval=FALSE}
bcftools view -S outputs/007/late-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/late-fall.vcf.gz
bcftools view -S outputs/007/cv-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/cv-fall.vcf.gz
```

### Run ANGSD to estimate allele frequencies

Here is the code used.  We don't do it here, because in and update from angsd 0.920
to 0.938 it broke.  But I have the estimated allele freqs in stored results.

```{sh, eval=FALSE}
cd outputs/007
angsd -vcf-gl late-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 16 -domaf 3 -out late-fall 2> late-fall.redirect.stderr
angsd -vcf-gl cv-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 32 -domaf 3 -out cv-fall 2> cv-fall.redirect.stderr
```

### Read the estimated allele freqs in and analyse them

```{r, message=FALSE}
af1 <- list(
  late_fall = read_tsv(input_list$chr34_late_fall_mafs),
  cv_fall = read_tsv(input_list$chr34_cv_fall_mafs)
) %>%
  bind_rows(.id = "run") %>%
  pivot_wider(names_from = run, values_from = c(knownEM, unknownEM, nInd)) %>%
  mutate(abs_diff = abs(knownEM_late_fall - knownEM_cv_fall))
```

Let's only bother with ones that have 8 or more indivs with reads in late fall and 16 or more in cv-fall, and
then we can plot them.
```{r}
af2 <- af1 %>%
  filter(nInd_late_fall >= 8 & nInd_cv_fall >= 16)

ggplot(af2, aes(x = position, y = abs_diff)) +
  geom_point(alpha = 0.4)
```

That is pretty cool.  It shows that characteristic pattern near 800 Kb that we saw around GREB1L,
where there a lots of SNPs with large allele frequency differences, and lots with very small
allele frequency differences, but very few with intermediate differences.

One thing I want to confirm is that all the SNPs with small allele freq differences
in that region are actually at low frequency in both groups.  We can do that by
coloring the points by the max frequency (between the two groups) of the minor
allele---that is the allele that is at lowest _average_ frequency in the two populations.

```{r}
af3 <- af2 %>%
  mutate(is_maf = (knownEM_cv_fall + knownEM_late_fall) < 0.5,
         max_maf = pmax(ifelse(is_maf, knownEM_late_fall, 1 - knownEM_late_fall),
                        ifelse(is_maf, knownEM_cv_fall, 1 - knownEM_cv_fall))
         )

mmp <- ggplot(af3, aes(x = position, y = abs_diff, color = max_maf)) +
  geom_point() +
  scale_color_viridis_c()
mmp
```

And, now, let us also do a version in which the high max_mafs are printed on top, so that we can see them if they exist...
```{r}
mmp_sort <- ggplot(af3 %>% arrange(max_maf), 
                   aes(x = position, y = abs_diff, color = max_maf)) +
  geom_point() +
  scale_color_viridis_c()
mmp_sort
```

Outstanding.  It is clear that there is a fairly large haplotype/region
that is differentiated between fall and late fall there in the first
2 Mb of the chromosome.

## Display some of the highly associated SNPs there

The GWAS turned up a number of highly associated SNPs.  Let's add circles for some
of the ones with the largest likelihood ratio statistic.

We will take the top 6 from each of the comparisons:
```{r}
lfvsj <- read_tsv(input_list$gwas_lf_v_sanjo) %>%
  arrange(desc(LRT))
lfvfrhf <- read_tsv(input_list$gwas_lf_v_frhf) %>%
  arrange(desc(LRT))

# now get the top 6 from each for illustration
top6_from_each <- list(
  lfvsj %>% slice(1:6),
  lfvfrhf %>% slice(1:6)
) %>%
  bind_rows() %>%
  arrange(desc(LRT)) %>%
  filter(!duplicated(Position))

# print those out
top6_from_each
```

So, add those to the plot
```{r}
top6freqs <- af3 %>% semi_join(top6_from_each, by = c("position" = "Position"))
mmp_sort +
  geom_point(
    data = top6freqs,
    shape = 21,
    size = 3, 
    color = "black"
    )
```

That shows us that there are a lot of other markers with substantial allele freq
diffs in this region, but the association test puts thew big circles up as the most significant, perhaps
because of the sample sizes.  It sort of reinforces my informal belief that many insights
can come from looking at absolute allele freq diffs, at least in these sorts of
cases where you have a haplotypic region with
very large allele freq differences).

## Annotate the variants

Before we start designing assays and things, I want to look for large-effect mutations.
We have the annotations for the positions in these 5 Mb of Ots34 in the stored results
(from the Science paper), so we will use them here:

First we read in the annotations:
```{r}
ann_vcf <- read.vcfR(input_list$ots34_snpeff)

avt <- vcfR2tidy(ann_vcf, info_only = TRUE)

# get column heading for the snpEff output
ann_column_names <- avt$meta %>%
  filter(Tag == "INFO", ID == "ANN") %>%
  pull(Description) %>%
  str_replace_all(., pattern = "^.* '| +|'", "") %>%
  str_split(., "\\|") %>%
  pluck(1)

anns <- avt$fix %>%
  select(CHROM, POS, REF, ALT, ANN) %>%
  separate(ANN, into = ann_column_names, sep = "\\|")

```
I don't think we need to worry about the discarded pieces there.

Now, we can join those annotations onto the allele frequencies and plot them,
coloring by predicted effect:

```{r}
af3_ann <- left_join(af3, anns, by = c("position" = "POS")) %>%
  mutate(Annotation_Impact_f = factor(Annotation_Impact,
                                      levels = c("HIGH", "LOW", "MODERATE", "MODIFIER"))) %>%
  arrange(desc(Annotation_Impact_f))

imp_col <- ggplot(af3_ann, aes(x = position, y = abs_diff, colour = Annotation_Impact)) +
  geom_point() +
  scale_colour_manual(values = c(HIGH = "red", LOW = "green", MODERATE = "orange", MODIFIER = "gray90")) +
  theme_bw()
imp_col
```

So, there are a few that we would want to be sure to hit.

## Finding candidates for assay design

When I originally did this, I also used a hack to include indels in here.  I spare
the reader from the gory details of that.

We could rank things by allele frequency difference, but I would also like
to incorporate an estimate of the variance in these absolute frequency
differences.  The variance in that difference will be the sum of the
two variances, so it is easy to calculate.  We don't really know what the
sample size is, but I am going to throw out as an approximation that each
individual in the `nInd` field represents a sample of about 1.2 gene copies.

We will estimate the variances from each different estimate and then calculate
two standard deviations below the actual estimate and have that as a "lower
bound" on the absolute value of the allele frequency difference.

```{r}
# here is a function for computing the lower bound on the absolute alle freq diff
lub <- function(f1, f2, n1, n2, s = 1.2) {
  var <- (f1 * (1 - f1) / (n1 * s)) + (f2 * (1 - f2) / (n2 * s))
  sd <- sqrt(var)

  abs(f1 - f2) - 2 * sd
}

saind2 <- af3_ann %>%
  mutate(abs_diff_lb = lub(knownEM_late_fall, knownEM_cv_fall, nInd_late_fall, nInd_cv_fall))
```

Plot those to make sure they look right, and denote the most highly associated ones
so that we can see those. 
```{r}
top6again <- saind2 %>%
  semi_join(top6_from_each, by = c("position" = "Position"))

ggplot(saind2, aes(x = abs_diff, y = abs_diff_lb, colour = max_maf)) +
  geom_point() +
  scale_colour_viridis_c() +
  geom_point(data = top6again, fill = NA, colour = "black", size = 3, shape = 21)
```

All right.  When I look at that, I think to myself, The ones I would want to design
assays for (for distinguishing fall-run from late-fall), would be:

1. (1) The ones with the highest association values
2. (2) The ones with an abs_diff > 0.75 and an abs_diff "lower bound" above 0.5.  
3. (4) Then, for any remaining ones we might want to do, I would choose things with an
abs_diff > 0.5, and abs_diff_lb > 0.25, and a max_maf as close to 1 as possible.  
4. (8) Annotation_Impact %in% c("Moderate", "High")

We will designate those with a bit mask, with bits being 2 raised to their number 
minus 1 (as given in the parentheses, above).

**Another Category**  
I will also add another category of candidate, which is out in the flanks and might
be interesting for assessing LD out there.

* (16) Position > 2.2 Mb and abs_diff > 0.48 and abs_diff_lb > 0.25

Mark them.
```{r}
assay_candi <- saind2 %>%
  mutate(candi_category = 1 * position %in% top6_from_each$Position +
           2 * (abs_diff > 0.75 & abs_diff_lb > 0.5) +
           4 * (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98) + 
           8 * (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) +
           16 * (abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6)) %>%
  select(chromo:max_maf, Annotation_Impact_f:candi_category, everything())

# how many?
assay_candi %>%
  count(candi_category)
```

Just to be explicit about what these candi_categories mean:

* 0 : not selected for any reason
* 1 : Hi-assoc (from Mac)
* 2 : abs_diff > 0.75 & abs_diff_lb > 0.5
* 3 : 1 and 2
* 4 : (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98)
* 7 : 1 and 2 and 4
* 8 : (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) )
* 16 : abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6
* 20 : 4 and 16

Plot them on a plot like the above so we can see them.
```{r}
ggplot(mapping = aes(x = abs_diff, y = abs_diff_lb, colour = factor(candi_category))) +
  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")

```

And see where those candidates live:
```{r}
ggplot(mapping = aes(x = position, y = abs_diff, colour = factor(candi_category))) +
  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")
```

So, we have good candidates across a couple Mb.  That is great.

Let's make a plot of that for the supplement:
```{r}
assay_candi4fig <- assay_candi %>%
  mutate(
    candi_cat = case_match(
      candi_category,
      0 ~ NA_character_,
      1 ~ "1: One of the 10 SNPs with lowest p-values",
      2 ~ "2: |d| > 0.75 with lower confidence interval >0.5",
      3 ~ "3: 1 and 2",
      4 ~ "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
      7 ~ "7: 1 and 2 and 4",
      8 ~ "8: snpEff annotation of High or Moderate",
      16 ~ "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
      20 ~ "20: 4 and 16"
    ),
    `Candidate Marker Category` = factor(
      candi_cat,
      levels = c(
        "1: One of the 10 SNPs with lowest p-values",
        "2: |d| > 0.75 with lower confidence interval >0.5",
        "3: 1 and 2",
        "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
        "7: 1 and 2 and 4",
        "8: snpEff annotation of High or Moderate",
        "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
        "20: 4 and 16"
      ))
  )

g <- ggplot(mapping = aes(x = position, y = abs_diff, colour = `Candidate Marker Category`)) +
  geom_point(data = assay_candi4fig %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi4fig %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2") +
  xlab("Position on Chromosome 34 (Mb)") +
  ylab("|d|, Absolute Difference in Allele Frequency")

ggsave(g, filename = "tex/images/lfar-candidates.pdf", width = 8, height = 5)
```

## Preparing an assay order

We prepare assay order information for these 49 good candidates, and handed that
to Anthony, along with the sequences, to design primers.

That process involved some custom scripts and Primer3.  Once the candidates
came back, they were tested in a variety of Chinook samples and it was found that
only 8 gave reliable amplicons that mapped reliably, etc.  And those
are used in the ensuing work.

Here, print out the actual candidates:
```{r}
actual_candidates <- assay_candi %>%
  filter(candi_category > 0) %>% 
  arrange(position)

# write that out for fun:
write_csv(actual_candidates, file = output_list$candidates_for_assays)
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```

### 2025-03-14 tex/body-text-of-paper.tex
Got the One-column version typesetting well
Tweaks.  Mostly anonymizing for review
Edits responding to referee comments

@@
 revealed that these allozyme markers
-exhibited different allele frequencies among major lineages, or stocks, of Chinook salmon on the West Coast.
+exhibited different allele frequencies among major lineages, or stocks, of Chinook salmon on the West Coast of the USA.
 These allele frequency differences make GSI possible, and
@@
 
-In the following, we present a new reference baseline that includes 1,636 Chinook salmon individuals
-from 17 collections, genotyped at 204 loci, mostly microhaplotypes \citep{baetscher2018microhaplotypes}, from throughout the genome.
-This baseline is
-highly effective for GSI within California and the markers within it provide ample power
-for PBT, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program 
+We present a new reference baseline that includes 1,636 Chinook salmon individuals from 17 collections that were genotyped at 204 loci distributed throughout the genome.
+This baseline is highly effective for GSI within California and provides ample power for PBT, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program 
 \citep{garza2007large,beacham2021parentage}.
@@
 California's Central Valley, and the two main tributaries (Klamath and Trinity Rivers) in the Klamath basin in Northern California, and
-three rivers of the California coast (Russian, Eel, and Smith)
+three rivers of the California coast (Russian, Eel, and Smith).
+Of the 1594 total fish samples from 17 different collections, 724 fish from 10 collections were of the fall run ecotype,
+498 fish from 5 collections were spring run, and 261 and 111 fish were of the winter run and late-fall run ecotypes, 
+respectively,  each represented by a single collection.
 (Figure~\ref{fig:map}, Table~\ref{tab:samples}).
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{images/map-crop.pdf}
+\includegraphics[width=0.48\textwidth]{images/map-crop.pdf}
 \end{center}
@@
 In the San Joaquin River, in 2016, at the time of sampling, only
-fall-run Chinook salmon adults were found, so the samples there were
-easily categorized. Run-timing
+fall-run Chinook salmon adults were found, therefore the samples could be categorized.  Run-timing
 designation at the Feather River Hatchery is more complicated but it is managed by the hatchery practices in place:
@@
 These gene regions were PCR-amplified and sequenced on a MiSeq sequencer using a $2\times 75$-cycle kit.
-The sequences from each locus were mapped to the de novo reference sequences (produced by Stacks) with
+The sequences from each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
 the Burrows-Wheeler aligner using the mem algorithm \citep{bwa-mem2009}. 
@@
 read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus
-Specific Primers respectively.  We then dropped any loci that produced genotypes that were
-frequently inconsistent with genotypes produced with the SNPtype assays and any loci that
-did not amplify well or those that amplified so well that they were grossly over-represented
-among the reads on the sequencer.  We
-used FreeBayes \citep{garrison2012haplotype} to identify all variants on the fragment (not just the SNPs targeted by the original TaqMan assays).  This allows us to score each locus as a
-microhaplotype \citep{baetscher2018microhaplotypes} when multiple variants occur in the sequence.  We filtered out variants
-that were considered multinucleotide or complex variants by FreeBayes. The resulting genetic markers are hereafter referred to as 'SNPlicons'.
+Specific Primers respectively.  Loci were removed that were frequently inconsistent with SNPtype assays,
+had poor amplification, or had excessive amplification and were subsequently over-represented in the sequencing library.
+FreeBayes \citep{garrison2012haplotype} was used to identify all SNPs (additional to the TaqMan-assay targets) along the fragment, which facilitated scoring fragment microhaplotypes when multiple SNPs were present along the read \citep{baetscher2018microhaplotypes}.
+The resulting genetic markers are hereafter referred to as `SNPlicons' to record their origin as TaqMan assays in a previous
+baseline.
 
@@
 To identify loci with large allele frequency differences between late-fall and fall-run Chinook salmon,
-we first conducted a simple case-control GWAS using the low-coverage whole-genome sequencing
+we first conducted a simple case-control genome-wide association study (GWAS) using the low-coverage whole-genome sequencing
 data of \citet{thompson2020complex}, mapped to the Otsh\_v1.0 Chinook salmon assembly
@@
 
-There was only one peak in association scores that was shared between both comparisons (see
+We surveyed the GWAS $p$-values relative to the commonly used significance threshold for
+GWAS of $5\times 10^{-8}$ \citep{chen2021revisiting}.  Only a single site exceeded that threshold in both comparisons and it occurred
+atop a peak in association scores on chromosome 34 (see
 Results). From this peak we identified candidates for follow-up genotyping by including the 10
-SNPs with the lowest association $p$-values, and supplementing those with additional SNPs that
+SNPs in that peak with the lowest association $p$-values, and supplementing those with additional SNPs that
 showed:
@@
 \end{itemize}
-The final condition was implemented to gather several SNPs within the flanking region to the right of
-the main peak of association.
+The final condition was implemented in order to gather several SNPs with large allele frequency differences
+that were adjacent to, but not directly within, the main peak of association, as we had hoped that those SNPs
+might enable us to learn about recombination in the region.
 
@@
 and version 2  (Otsh\_v2.0, GenBank accession GCA\_018296145.1) genome assemblies,
-using both bwa-mem \citep{bwa-mem2009} and the BLAST-like Alignment tool \citep{kent2002blat}.
+using both bwa mem \citep{bwa-mem2009} and the BLAST-like Alignment tool, BLAT \citep{kent2002blat}.
 A similar mapping exercise was performed for the remaining markers so that we can
@@
 The paired-end sequence reads were joined together using Fast Length Adjustment
-of SHort reads \citep[FLASH:~][]{magovc2011flash}) and mapped to an indexed sequence reference created from the Stacks target loci (for the novel microhaplotypes, SNPlicons and RoSA markers) or from the entire Otsh\_v1.0 genome
-(for the LFAR, WRAP, VGLL3, and Six6 markers) using BWA mem\citep{bwa-mem2009}.
-Mapped reads were converted to BAM files using SAMtools \citep{li2009sequence} and FreeBayes \citep{garrison2012haplotype} was used to call variants.
+of SHort reads \citep[FLASH:~][]{magovc2011flash} and mapped to an indexed sequence reference created from the Stacks target loci (for the novel microhaplotypes, SNPlicons and RoSA markers) or from the entire Otsh\_v1.0 genome
+(for the LFAR, WRAP, VGLL3, and Six6 markers) using bwa mem. Mapped reads were converted to BAM files using SAMtools and FreeBayes was used to call variants.
 
@@
 
-For errors involving misidentifying
+For errors of misidentifying
 unrelated (U) pairs as parent-offspring (PO), full-sibling (FS), or half-sibling (HS)
 pairs we used importance sampling to estimate the very small FPRs.  The
-FNRs were estimated in each case using simple Monte Carlo.  For the FS and
-HS cases the FNRs were calculated from Monte Carlo sampling while taking account
+FNRs were estimated in each case using simple Monte Carlo taking account
 of the physical linkage of the markers by inputting their genomic positions into
@@
 software to simulate genotypes of related pairs
-in the presence of physical linkage.  For PO pairs (and U pairs),
-FNRs were calculated without reference to the genomic locations of the markers
-because physical linkage does not
-affect the distribution of log-likelihood ratios in those two relationship types,
-because there is no variation, anywhere
-in the genome, in the number of gene copies shared identical by descent between
-a parent and its child or between two unrelated individuals.
-
-For errors involving misidentifications between the relationship
+in the presence of physical linkage.  
+For errors of misidentifications between the relationship
 types of avuncular (AN: aunt-niece, uncle-nephew, etc.), PO, FS, and HS,
@@
 of all the amplicons used in the baseline is available
-in Data Supplement~1.
+in Data Supplement~1.  An overview of genomic coordinates is found in Figure~\ref{fig:genome-locations}
 
@@
 
-The case-control association test revealed a single peak on chromosome 34 with large allele frequency differences
+The case-control association identified only a single SNP with $p < 5\times 10^{-8}$ in the comparison
+between late-fall run and both the FRH fall-run and the San Joaquin fall run.  This SNP was within
+a single evident peak on chromosome 34 with large allele frequency differences
 between Central Valley late-fall and fall-run fish.  This peak was evident both in the comparison of
@@
 10 SNPs with the lowest association $p$-values, and including additional ones from the filtering criteria given in the
-Methods, yielded 49 candidate markers for which we designed primers for amplication.  Only 8
+Methods, yielded 49 candidate markers for which we designed primers for amplication (Figure~\ref{fig:lfar-candi}).  Only 8
 of these 49 primer pairs yielded reliable amplification and mapping.   Of these 8, two of the SNPs
@@
 {\footnotesize
-\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrr}
+\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrrr}
 \hline\hline
@@
 Nonetheless, three of these markers from that effort had large frequency differences
-between the groups and are included in the reference baseline baseline.  The process of 
+between the groups and are included in the reference baseline.  The process of 
 discovering them is described in the supplemental text section \ref{sec:wrap-methods}.
@@
 Of the 184 loci that originated from our novel
-microhaplotype discovery or from SNPtype assays, 179 of them mapped identically to a single
-location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
+microhaplotype discovery or from SNPtype assays, 179 of them were mapped by both bwa mem and
+BLAT to exactly the same location
+on an assembled chromosome in the Otsh\_v2.0 genome.
 Of the remaining 5 loci, one of them was mapped to the same location by BLAT and bwa mem, but
@@
 \caption{\footnotesize Simple population genetic summaries. All quantities are means
-over loci.  $\bar{N}$ is the number of fish with observed genotypes; $\bar{N}_\mathrm{A,ss}$ is the
+over loci.  $\bar{N}$ is the average over loci of the number of fish in each collection with observed genotypes
+at a locus; $\bar{N}_\mathrm{A,ss}$ is the
 average number of alleles and $\bar{P}_\mathrm{poly,ss}$ the fraction of polymorphic loci,
@@
 {\footnotesize
-\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrrrr}
+\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrrrr}
 \hline\hline
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{images/gsi_and_fst_fig-crop.pdf}
+\includegraphics[width=0.48\textwidth]{images/gsi_and_fst_fig-crop.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{images/clumpak-crop.pdf}
+\includegraphics[width=0.48\textwidth]{images/clumpak-crop.pdf}
 \end{center}
@@
 Another noteworthy feature appears at $K=7$, with STRUCTURE separating the fall-run reporting unit
-into two clusters (orange and red) that do not correspond exactly with the individual
+into two clusters that do not correspond exactly with the individual
 collections within the reference baseline. For example, fish from the
-orange cluster that predominates in the FRHS collection are also found in the FRHF and MDS collections.
+cluster that predominates in the FRHS collection are also found in the FRHF and MDS collections.
 
@@
 While fish from CHLF were found to misassign at a rate of 2.7\% (8/300) and MDS at a rate
-of 7.1\% (6/84) to CV-Fall, the
-converse is not evident---that is, fish from the CV-Fall are misassigned to CHLF at a rate
+of 7.1\% (6/84) to CV-Fall, misassignment in the other direction occurs at a much lower
+rate---that is, fish from the CV-Fall are misassigned to CHLF at a rate
 of only 0.6\% (3/508) and to the CV-Spring reporting unit (to which MDS belongs) at a rate
 of only 0.4\% (2/508).
-
 When assignments are only accepted with a scaled likelihood greater than 0.8,
-20 assignments are discarded, and half of those (10) were the incorrect assignments.  Thus,
+10 incorrect assignments and 10 correct assigments are discarded.  Thus,
 at the 0.8 threshold, the total fraction of correct assignments is 99.1\% (Figure~\ref{fig:eighty}).
@@
 
+The `CKMRsim' power analyses for this set of markers demonstrates that they have ample variation
+for accurate identification of parent-offspring (PO) and full-sibling (FS) pairs in almost all
+realistic situations. The predicted false positive rates from unrelated pairs for both the
+parent-offspring (PO) and full-sibling (FS) relationships are
+exceedingly low for all population groups, even when using stringent false negative thresholds as
+low as 0.05 (Figure~\ref{fig:fprs}a).
+For example, the false positive rate for FS identification is less than one error in 10 million
+comparisons of potential siblings in the California Chinook salmon population with the lowest
+heterozygosity (SRW) at a false negative rate of 0.05.
+
 %%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%
-
-The `CKMRsim' power analyses for this set of markers demonstrates that they have ample variation
-for accurate identification of parent-offspring (PO) and full-sibling (FS) pairs in almost all
-realistic situations. The predicted false positive rates from unrelated pairs for both the
-parent-offspring (PO) and full-sibling (FS) relationships are
-exceedingly low for all population groups, even when using stringent false negative thresholds as
-low as 0.05 (Figure~\ref{fig:fprs}a).
-For example, the false positive rate for FS identification is less than one error in 10 million
-comparisons of potential siblings in the California Chinook salmon population with the lowest
-heterozygosity (SRW) at a false negative rate of 0.05.
 
@@
 While there is not a currently available method \citep[like that of][]{anderson2006power}) for estimating error rates in
-inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which 
-require much more statistical power, it is clear that PO-trio inference with these markers will be essentially error free.
+inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying
+PO and FS pairs, which 
+require much more statistical power, we hypothesize that PO-trio inference with these markers will be very accurate.
 
@@
 This marker set and baseline reference dataset provide
-unprecedented power for identifying fish from all of the Chinook salmon ESUs in California,
+excellent power for identifying fish from all of the Chinook salmon ESUs in California,
 as well as individual populations within those ESUs (Figure~\ref{fig:gsi}).

### 2025-03-14 tex/responses-to-referees/responses-to-reviewers-1.Rmd
supp figs

---
title: "EVA-2025-003-RA Responses to Reviewers"
output: pdf_document
date: "2025-03-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reviewer 1

Authors present on the development of a new set of markers, with many in close proximity to additional SNPs that can be used to assemble microhaplotypes. The marker set shows high discrimination power between the multiple ecotypes present with a large data set and fits well within the scope of the journal. Overall, the study is of excellent quality however, I have the following comments below.

### Major Comments

* I’m confused by the amount of sequencing and a clear purpose of each sequencing run that should be clarified. There was ddRAD to identify novel polymorphisms/microhaplotypes, the second run using a using PCR-GBS type sequencing has an unclear purpose. Was this on the full population? Another “amplicon sequencing” run was performed to convert SNPtype polymorphisms, which sounds similar to the second sequencing run of ddRAD polymorphisms but uses different terminology. Was this part of the previous run? Was this run a few samples or the full population? I presume these were different runs but maybe they were the same. The GT-seq method, does this include all samples and the full set of markers to this point?

* Authors provide considerable detail on sample location in the methods but only provide the sample number in the introduction. I believe it would be useful to have summary sentences of Table 1 that include number sites for spring, fall, etc run, and the number of samples within each ecotype, total samples within the methods, etc.

    - _Thank you for this suggestion.  We added the following to the end of the first paragraph in the Population Sampling portion of the
Methods section:_
"Of the 1594 total fish samples from 17 different collections, 724 fish from 10 collections were of the Fall run ecotype,
498 fish from 5 collections were Spring run, and 261 and 111 fish were of the Winter run and Late-fall run ecotypes, 
respectively,  each represented by a single collection."

* Is there any LOD threshold for your GWAS that allowed identifying the chr34 peak? 
    - _We updated our treatment of this topic to note that we relied upon the commonly used
    GWAS $p$-value threshold of $5 \times 10^{-8}$.  We also added an additional
    supplemental figure that shows all the candidates chosen for followup within and
    near that peak._

* In addition, L239-240 implies you were you only interested in right side of the peak. I would assume you were interested in up and downstream of the peak unless I missed something. In addition, please clarify why 2.2 Mb selected. Is this an LD block? Does this mean 2.2 Mb in each direction or total?
    - _This is a good question.  It turns out that the main peak of association is at the
    beginning of chromosome 34, so there is no flanking region to the left of the peak. To
    clarify why those candidate SNPs in the flanking region were chosen we now state:_ 
    "The final condition was implemented in order to gather several SNPs with
    large allele frequency differences that were adjacent to, but not directly
    within, the main peak of association, as we had hoped that those SNPs might
    enable us to learn about recombination in the region."

* The writing of the manuscript is largely well written but makes high use of words such as “we” or “us”, which although stylistic, makes the manuscript longer than needed in some cases such as “we then dropped” or “this allows us”. In many cases these can be removed (examples in minor comments).

* There is also use of informal language e.g. “amplified so well” or unnecessary words such as “then” that could be removed.

* I commend the use of the distribution of log-likelihood ratio for relationship to show the power of haplotypes vs SNPs and would argue this could be elevated to a full figure, potentially in a condensed form, maybe less panels somehow?

* If possible, I think it would also be good to show how much of an improvement the discrimination power improved from the old marker data. In theory, this data has already been generated for the full population if you strip out the novel SNPs of the current study.

* It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.

### Minor Comments

- L117-119: There are a lot of clauses in this sentence, consider: “We present a new reference baseline that includes 1,636 Chinook salmon individuals from 17 collections that were genotyped at 204 loci distributed throughout the genome. The Baetscher reference is also inappropriate in this sentence. If you want to expand with additional sentences that these are predominately fragments that can be scored as haplotypes and the reason behind this (the power) the reference would be appropriate. It may also be worth including how haplotypes are being used in other organisms.

    - _We whittled that sentence down._

- L119-122: Similar long sentence, consider: “This baseline is highly effective for GSI within California and provides ample power for BST, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program.
    - _We used your edit._

- L147-148: Consider: “therefore the samples could be categorized” instead of “so the samples there were easily categorized”
    - _Done._

- L191: Italicize de novo
    - _Fixed._

- L207-210: Informal sentence, consider: “Loci were removed that were frequently inconsistent with SNPtype assays, had poor amplification, or over amplified and were subsequently over-represented in the sequencing library”.
    - _We changed that sentence to:_ "Loci were removed that were frequently inconsistent with SNPtype assays,
had poor amplification, or had excessive amplification and were subsequently over-represented in the sequencing library."

- L210-212: Unnecessary words, consider: “FreeBayes was used to identify all SNPs (additional to TaqMan assays) along the fragment which facilitated scoring fragment microhaplotypes when multiple SNPs were present along the read (Baetscher et al. 2018)”. Since MNP and complex variants were not included, consider just using SNPs where appropriate to avoid confusion with boarder term of variants.
    - _We used your suggested edit and removed discussion of MNPs and complex variants._

- L217: Define GWAS since you defined SNP
    - _Done_

- L296 & L317: I’m not sure of the difference between the microhaplotype markers and SNPlicon targets. I presumed they are the same thing after it was stated that microhaplotypes would “hereafter referred to as SNPlicons” but this term is rarely used again, please decide whether you want to keep this term and avoid using both names or clarify if these are different.
    - _We updated the introduction of the term SNPlicon to specifically state that it refers to the microhaplotypes
    that were obatined from sequencing around the sites of the Taq-Man assays that were previously used.
    We feel it is quite important to retain information about the provenance of these markers as it could
    facilitate merging amplicon-sequenced data with previously collected Taq-Man data, so we have retained
    the terminology._

- L316: Remove the double bracket/parentheses
    - _Done_

- L319: Package references only need to used on the first mention of the package
    - _Eliminated._

- L386-390: I had a hard time following this sentence, please clarify.
    - _This was just a technical statement about the way that CKMRsim works because
    of the fact that physical linkage in the absence of LD has no effect on the likelihoods
    for parent-offspring pairs and unrelated pairs.  However since simulation with
    physical linkage in PO and U relationships yields identical results to those without
    physical linkage, we just simplify the manuscript by not mentioning that technical
    detail that can be confusing._

- L402 & 416: Inconsistent use of “Figure” or “Fig.”. Supplemental figures are also mentioned out of order
    - _We changed that single occurrence of Fig. to Figure. Also, thanks for catching that
    mis-ordered supplemental figure.  We have renumbered them correctly._

- L447: What do you mean by mapped identically? This would imply there is no polymorphism along the locus. You can either remove the word or say “uniquely mapped” if I understand what you mean correctly.
    - _This was meant to say that they were mapped by both bwa mem and blat
    to exactly the same location in the genome. We updated the text to be more clear._

- L448: Use of BLAT, you did not define this at L299
    - _Fixed._

- L486: Not sure how easily this it to recolour, but “red” looks more bright pink and then you also have another pink for BCS. Maybe easier to just say it is a different colour.
    - _We chose not to try to name the colors from the default CLUMPAK palette. Rather we simply
    refer to the collections where the changes are apparent._

- L511-L514: I do not agree. Although though it is a much lower rate, they still misassign
    - _We rephrased that to be correct._

- L576-577: strong statement when this has not been tested. Consider altering the strength of the statement such “hypothesize this will be error free”
    - _We edited this as suggested._

- Figure 4. Avoid method and interpretation sentences in caption and use in main text instead

## Reviewer 2

This study develops a microhaplotype panel (based on the GT-seq method) to create a training set to identify different populations of genetic stock and relationships between individuals in America Salmon. My comments relating to this manuscript are outlined below:

### Major Comments

* Lines 399-444: My biggest compliant is the paper states that it creates a “baseline” for GSI and PBT analysis but it is not always clear what this final baseline is and only at the end of the introduction does it state what this baseline contains (in terms of number of animals and number of loci). I feel this needs to also be stated in the abstract. In addition, in the results section it is not easy for work out how many markers made into the final reference baseline, and I think this information should be included in a table that also shows how markers from each source of information. I’m assuming that this is what is being used in the subsequent analysis.

* Table 3: The values in second column (the average number of fish with observed genotypes) seem odd to me. I would have thought these values would be less than seventh column in Table 1 (column N). For example, CHLF has 379.6 in Table 3 > 261 in Table 1. Can the authors explain this? Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.

### Minor Comments

- Line 48: “… West Coast” should be “West Coast of the USA”?
    - _Fixed_

- Line 66: The term “reference baselines” is used through the manuscript but it is quite vague what this means. Reading the manuscript, it would seem to me that the reference baseline has two parts (1) “a multihaplotype marker/SNP set” and (2) “a reference dataset for training the GSI and PBT methods”. I think the authors define the term “reference baselines” here in the introduction.

- Introduction – The paper seems to rely heavily on the work by Thompson et al (2020) in the methods, but this is not mentioned in the introduction. The authors should make a comment towards of the introduction that the work depends on some of the data and results from this paper

- Line 181: “a de novo assembly was performed”. How was this performed? More details are needed.

- Line 214: The use of the term “SNPlicons” seems a little inappropriate. My understanding is the authors are using more than just the information at the SNP and are scoring microhaplotype, so I feel the name is a little misleading.
    - _We provide more explanation of which markers/microhaplotypes are referred to as the SNPlicons.  Since
    this terminology is already used in the distributed version of the baseline that is being
    used by several labs in California, we felt it best to leave the terminology as is._

- Line 443: The word “baseline” is duplicated.

    - _Fixed_

- Line 516: “… half of those (10) were the incorrect assignments”. I found this a little confusing. Are you saying that 10 of the discarded assignments were from the 24 incorrect assignments reported in line 501? This sentence could possibly do with a rephasing.
    - _Edited for clarity_

- Caption of Table 3: “N is the number of fish …” should be “N is the average number of fish …”
    - _Fixed_

- Figure 1 and 2: The colours for “spring run” and “Late-fall run” are very difficult to separate for a colour-blind person. Since the groups are not in a specific order on the graph, it is very hard work to out what is what. I’d recommend the authors use a selection of colours that will work for colour-blind people.

- Line 589: “unprecedented” is an over-statement as there is no independent validation set. The authors should use a word that less forceful.
    - _Changed to "excellent"_

- Sentence at lines 645-647. Can the authors provide a reference to back up this statement.

- The online data and scripts (thank you for providing these) needs a proper readme file describing what the folders and files relate to.

### 2025-03-14 tex/supplemental-figures.tex
Tweaks.  Mostly anonymizing for review
supp figs

@@
 %!TEX root = supplement.tex
-
-%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-\includegraphics[width=\textwidth]{images/lfar-assoc-faceted.jpg}
-\caption[
-	Negative log base 10 of association $p$-values for
-	individual SNPs for late-fall versus fall run
-]{
-	\footnotesize Negative log base 10 of association $p$-values for
-	individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
-	with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
-	to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
-	upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
-	panel is the comparison of late-fall to San Joaquin River fall. 
-}
-\label{fig:lfar-assoc}
-\end{figure}
 
@@
 named chromosome/linkage group in the assembly.}
-\label{fig:num-alle}
+\label{fig:genome-locations}
 \end{figure}
 %%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=\textwidth]{images/lfar-assoc-faceted-min-0.png}
+\caption[
+	Negative log base 10 of association $p$-values for
+	individual SNPs for late-fall versus fall run
+]{
+	\footnotesize Negative log base 10 of association $p$-values for
+	individual SNPs for late-fall versus fall run.  $x$ axis shows position in genome (in megabases),
+	with color alternating by chromosome, as indicated by numbers above the $x$-axis. ``Unk'' refers
+	to unplaced scaffolds in the Otsh\_v1.0 genome assembly \citep{christensen2018chinook}. The 
+	upper panel is the comparison between late-fall and Feather River Hatchery fall, while the lower 
+	panel is the comparison of late-fall to San Joaquin River fall.  The dashed line corresponds to a
+	significance threshold of $5 \times 10^{-8}$.  The only SNPs exceeding that threshold in both
+	comparisons are the ones atop the prominent peaks on chromosome 34.
+}
+\label{fig:lfar-assoc}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\includegraphics[width=\textwidth]{images/lfar-candidates.pdf}
+\caption[
+	Candidates chosen for assay design for the late-fall associated region
+]{
+	\footnotesize Positions (on $x$ axis) and absolute allele frequency difference
+	between whole-genome sequenced fall-run and late-fall run fish (on $y$ axis)
+	of 49 candidate SNPs for which amplicons were designed.  Gray points are SNPs that
+	were not candidates.  Colors of points denote the reason each SNP was chosen
+	as a candidate.  See main text ``Methods" for further explanation of categories.
+}
+\label{fig:lfar-candi}
+\end{figure}
 

### 2025-03-14 002-GSI_and_Fst.Rmd
Get some counts

@@
 
+```
+
+## Get a quick summary of number of collections and samples by ecotype
+
+One of the reviewers wanted summaries in the text.  That's a great idea.
+Here we get the numbers that we need.  
+```{r}
+# total fish:
+sum(fbmet4$N)
+```
+
+```{r}
+# numbers of samples and collection by ecotype
+fbmet4 %>%
+  group_by(`Run Timing`) %>%
+  summarise(
+    num_collections = n(),
+    num_samples = sum(N)
+  )
 ```

### 2025-03-17 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Pushing the revisions further along

@@
 
+_Thank you for your comments and suggestions.  We have addressed all of them, as
+detailed below, and we are pleased with how this has improved the manuscript._
+
 ### Major Comments
 
-* I’m confused by the amount of sequencing and a clear purpose of each sequencing run that should be clarified.
+* **I’m confused by the amount of sequencing and a clear purpose of each sequencing run that should be clarified.
 There was ddRAD to identify novel polymorphisms/microhaplotypes, the second run using a using PCR-GBS type sequencing has an unclear purpose.
@@
 I presume these were different runs but maybe they were the same.
-The GT-seq method, does this include all samples and the full set of markers to this point?
+The GT-seq method, does this include all samples and the full set of markers to this point?**
+
+    - **Should be able to just add a few sentences to make this more clear.**
 
@@
 I believe it would be useful to have summary sentences of Table 1 that include number sites for spring, fall, etc run, and the number of samples within each ecotype, total samples within the methods, etc.
-
     - _Thank you for this suggestion.  We added the following to the end of the first paragraph in the Population Sampling portion of the
 Methods section:_
-"Of the 1594 total fish samples from 17 different collections, 724 fish from 10 collections were of the Fall run ecotype,
-498 fish from 5 collections were Spring run, and 261 and 111 fish were of the Winter run and Late-fall run ecotypes, 
+"Of the 1594 total fish samples from 17 different populations, 724 fish from 10 collections were of the fall run ecotype,
+498 fish from 5 collections were spring run, and 261 and 111 fish were of the winter run and late-fall run ecotypes, 
 respectively,  each represented by a single collection."
@@
 Does this mean 2.2 Mb in each direction or total?
-    - _This is a good question.  It turns out that the main peak of association is at the
+    - _It turns out that the main peak of association is at the
     beginning of chromosome 34, so there is no flanking region to the left of the peak. To
     clarify why those candidate SNPs in the flanking region were chosen we now state:_ 
-    "The final condition was implemented in order to gather several SNPs with
-    large allele frequency differences that were adjacent to, but not directly
-    within, the main peak of association, as we had hoped that those SNPs might
-    enable us to learn about recombination in the region."
+    "The final condition was implemented in order to gather several SNPs with large allele frequency differences
+    that were adjacent to, but not directly within, the main peak of association, so as to possibly learn about
+    recombination in the region."
 
@@
 In many cases these can be removed (examples in minor comments).
+    - _We addressed all the examples in the minor comments and also removed additional instances of the
+    use of such language where it was practical._
 
 * There is also use of informal language e.g. “amplified so well” or unnecessary words such as “then” that could be removed.
+    - _Following this suggestion, we removed many occurrences of "then" and "as well as"_
 
-* I commend the use of the distribution of log-likelihood ratio for relationship to show the power of haplotypes vs SNPs and would argue this could be elevated to a full figure, potentially in a condensed form, maybe less panels somehow?
+* **I commend the use of the distribution of log-likelihood ratio for relationship to show the power of haplotypes vs SNPs and would argue this could be elevated to a full figure, potentially in a condensed form, maybe less panels somehow?**
 
-* If possible, I think it would also be good to show how much of an improvement the discrimination power improved from the old marker data.
-In theory, this data has already been generated for the full population if you strip out the novel SNPs of the current study.
+    - **I am planning to break one population out into a single panel and show different logl-ratios
+    for different relationships.  Probably a 2x2 or 2x2 grid. Intend to leave the others in the
+    supplement as is.**
 
-* It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.
+* **If possible, I think it would also be good to show how much of an improvement the discrimination power improved from the old marker data.
+In theory, this data has already been generated for the full population if you strip out the novel SNPs of the current study.**
+    - **We could do this for the 78 SNPlicons plus 16 other single-SNPs, but it would be a little hard. We might just
+    say no.**
+
+* **It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.**
+    - **Should be easy to throw a few citations in probably in a paragraph in the discussion.**
 
@@
 - Figure 4. Avoid method and interpretation sentences in caption and use in main text instead
+    - _We removed the final few sentences from Fig. 4's caption and edited them appropriately
+    to include them in the text in the Results section._
 
@@
 I feel this needs to also be stated in the abstract.
-In addition, in the results section it is not easy for work out how many markers made into the final reference baseline, and I think this information should be included in a table that also shows how markers from each source of information.
+    - _As per reviewer \#1's suggestion, the methods now includes textual comment on the
+    number of samples from each collection. Additionally, we have added several sentences
+    in the third paragraph of the introduction to define the terms reference data set and
+    baseline, and we have also added a summary in the abstract
+    on the total number of amplicons, individuals and populations in the reference data set._
+
+* In addition, in the results section it is not easy for work out how many markers made into the final reference baseline, and I think this information should be included in a table that also shows how markers from each source of information.
 I’m assuming that this is what is being used in the subsequent analysis.
+    - _Excellent suggestion.  This information now appears in the new Table 2 which shows the total number of
+    amplicons and the total number of variant sites within each amplicon for each of several
+    groups of markers that correspond to the different sources of information._
 
-* Table 3: The values in second column (the average number of fish with observed genotypes) seem odd to me.
+* **Table 3: The values in second column (the average number of fish with observed genotypes) seem odd to me.
 I would have thought these values would be less than seventh column in Table 1 (column N).
@@
 Can the authors explain this?
-Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.
+Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.**
+    - **Gotta still see what he is talking about here.**
 
@@
 I think the authors define the term “reference baselines” here in the introduction.
+    - _Thank you.  The paper really needed this.  We now have a paragraph in the 
+    introduction that defines these terms_
 
 - Introduction – The paper seems to rely heavily on the work by Thompson et al (2020) in the methods, but this is not mentioned in the introduction.
-The authors should make a comment towards of the introduction that the work depends on some of the data and results from this paper
+The authors should make a comment towards of the introduction that the work depends on some of the data and results from this paper.
+    - _Added a clause in the last paragraph of the introduction about this._
 
-- Line 181: “a de novo assembly was performed”. How was this performed? More details are needed.
+- Line 181: "a de novo assembly was performed". How was this performed? More details are needed.
+    - _We clarified that that we used ustacks from the Stacks program to assemble the reads into
+    RAD loci._
 
@@
 
-- Figure 1 and 2: The colours for “spring run” and “Late-fall run” are very difficult to separate for a colour-blind person.
+- **Figure 1 and 2: The colours for “spring run” and “Late-fall run” are very difficult to separate for a colour-blind person.
 Since the groups are not in a specific order on the graph, it is very hard work to out what is what.
-I’d recommend the authors use a selection of colours that will work for colour-blind people.
+I’d recommend the authors use a selection of colours that will work for colour-blind people.**
+
+    - **Planning to just cross-hatch the light blue of the late fall**
 
@@
 
-- Sentence at lines 645-647. Can the authors provide a reference to back up this statement.
+- **Sentence at lines 645-647. Can the authors provide a reference to back up this statement.**
 
-- The online data and scripts (thank you for providing these) needs a proper readme file describing what the folders and files relate to.
+- **The online data and scripts (thank you for providing these) needs a proper readme file describing what the folders and files relate to.**
+
+    - **Gonna make a README and store the htmls in the docs/ folder like
+    we did for the Science paper.**

### 2025-03-17 002-GSI_and_Fst.Rmd
Pushing the revisions further along

@@
   )
+```
+
+# Let's also just do a summary of the number of markers of different kinds
+
+```{r}
+# get the distinct loci in it
+loci_tib <- read_csv("tex/supp_data/Supp-Data-1-Amplicon-info.csv")
+
+# now, pull out their specifiers and count
+loc_table <- loci_tib %>%
+  mutate(
+    spec = str_sub(AmpliconName, 5, 8),
+    num_snps = str_count(SNPLocationsInAmpliconSequence, ",") + 1L
+  ) %>%
+  group_by(spec) %>%
+  summarise(num_amplicons = n(), num_variants = sum(num_snps)) %>%
+  arrange(desc(num_amplicons)) %>%
+  rename(
+    Type = spec,
+    `N (amplicons)` = num_amplicons,
+    `N (variants)` = num_variants
+  )
+
+# just write this out and then stick it in the doc
+write.table(loc_table, sep = "&", na = "", quote = FALSE, row.names = FALSE)
 ```

### 2025-03-17 tex/abstract.tex
Pushing the revisions further along

@@
 resolution for population discrimination and relationship identification from a rapidly and
-economically assayed panel of markers. We show how this set of microhaplotypes
+economically assayed panel of markers. We show how this set of genetic markers assayed
+by sequencing of 204 amplicons, in tandem with a reference data set of 1594 individual
+samples from 17 populations, 
 provides definitive power to identify all known lineages of Chinook salmon in

### 2025-03-17 tex/body-text-of-paper.tex
Pushing the revisions further along

@@
 has expanded considerably.
+
+A critical ingredient for GSI is the {\em reference data set}, which consists of individuals
+sampled from across a range of populations and with each individual genotyped using the same
+panel of genetic markers.  When individuals of unknown origin are genotyped
+at the same panel of genetic markers, the availability of the reference data set
+makes it possible to do inference regarding the population of origin of the
+unknown individuals.  Such reference data sets have seen widespread use and
+development in salmon management, where they are often called
+``genetic baselines," and defined as, ``databases of genotypes from
+breeding populations"\citep{seeb2007development}.  Throughout, we will
+use ``reference" or ``reference data set'' and ``baseline'' interchangeably; however,
+a key point to keep in mind is that as genetic technologies
+evolve over time, reference data sets continue to be developed and
+refined to exploit the new technologies.
 
@@
 We describe how we identified and compiled the markers, including several new markers
+identified through re-analysis of the whole genome sequencing data of \citet{thompson2020complex},
 that are particularly effective at distinguishing between closely related groups of fish,
@@
 
-We compiled samples of fish for the reference baselines from 13 locations within California, and one
+Samples of fish for the reference baseline were compiled from 13 locations within California, and one
 within Oregon.
@@
 three rivers of the California coast (Russian, Eel, and Smith).
-Of the 1594 total fish samples from 17 different collections, 724 fish from 10 collections were of the fall run ecotype,
+Of the 1594 total fish samples from 17 different populations, 724 fish from 10 collections were of the fall run ecotype,
 498 fish from 5 collections were spring run, and 261 and 111 fish were of the winter run and late-fall run ecotypes, 
@@
 As such, sampling targeted carcasses found in each of these
-rivers at times representative of their run-timing designation, which we were able to further confirm using their genotypes
+rivers at times representative of their run-timing designation, which were further confirmed using their genotypes
 within the region of strongest association (RoSA) on chromosome 28 \citep{thompson2020complex}.
@@
 as spring-run broodstock, and the rest are included in the fall-run broodstock.
-We sampled putative spring-run and fall-run fish 
-from these two broodstocks.
-Similar to broodstock collections in the CCV, samples from spring- and fall-run fish in the Trinity River were selected on the basis of spawn timing at the Trinity River Hatchery with migration-timing ecotype then confirmed by RoSA genotype.
+Putative spring-run and fall-run fish 
+from these two broodstocks were sampled.
+Similar to broodstock collections in the CCV, samples from spring- and fall-run fish in the Trinity River were selected on the basis of spawn timing at the Trinity River Hatchery with migration-timing ecotype confirmed by RoSA genotype.
 Elsewhere, adult samples were taken from live broodstock at Rowdy Creek (Smith River), Iron Gate (Klamath River) and Warm Springs (Russian River) Hatcheries and juveniles were sampled from Blue Creek (Klamath River); all locations where only fall-run fish have been documented.
@@
 The genetic markers in the baseline include newly discovered microhaplotypes, SNP assays
-that were translated into amplicons, markers associated with phenotypes, as well
-as species- and sex-specific markers.
+that were translated into amplicons, markers associated with phenotypes, and
+species- and sex-specific markers.
 These genetic markers were compiled from several different
@@
 
-We identified candidate loci with multiple single-nucleotide polymorphisms (SNPs) among multiple
-populations using reduced-representation genome sequencing of 2--4 Chinook salmon, each, from a number of
+Candidate loci with multiple single-nucleotide polymorphisms (SNPs) among multiple
+populations were identified using reduced-representation genome sequencing of 2--4 Chinook salmon, each, from a number of
 populations ranging from California to British
@@
 Genomic DNA was extracted using Qiagen DNeasy 96 Blood and Tissue Kits on a BioRobot 3000
-(Qiagen, Inc). We used double-digest restriction-site associated DNA sequencing (ddRAD) with different size selections (400 bp and 500 bp
-fragments) to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
+(Qiagen, Inc). Double-digest restriction-site associated DNA sequencing (ddRAD) with different size selections (400 bp and 500 bp
+fragments) were used to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
 \citet{peterson2012double} with the modifications described in \citep{baetscher2018microhaplotypes}.
@@
 Stacks v1.45 \citep{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome
-at the time, a {\em de novo} assembly was performed.  Initial filtering retained RAD loci with two or more SNPs present
+at the time, the ustacks module of Stacks were used to assemble reads {\em de novo} into RAD loci.  Initial filtering retained RAD loci with two or more SNPs present
 in at least 10 samples. After this initial filtering, there were 3,931 RAD loci retained in the ddRAD data with 400 bp size selection and
 4,794 retained in the data with 500 bp size selection.
-Gene regions were then selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation amongst the winter run and at least one other population.
+Gene regions were selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation amongst the winter run and at least one other population.
 A 100 bp window was used because the success of multiplex PCR and MiSeq sequencing is diminished in amplicons larger than 100 bp. Primer3
@@
 assessment of Hardy-Weinberg equilibrium \citep{hardy1908mendelian} was conducted in MICROHAPLOT (Ng {\em et al.}, DOI: 10.5281/zenodo.820110).
-Loci that produced more than two haplotypes within an individual, as well as those clearly violating Hardy-Weinberg equilibrium within populations were removed.
+Loci that produced more than two haplotypes within an individual, and those clearly violating Hardy-Weinberg equilibrium within populations were removed.
 
@@
 using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (Standard BioTools Inc., South San Francisco, CA).
-In order to type these markers using amplicon sequencing, we followed the methods in
-\citet{campbell2015genotyping} and added the Illumina small RNA sequencing primer and
+These markers were typed using amplicon sequencing, following the methods in
+\citet{campbell2015genotyping} with the addition of the Illumina small RNA sequencing primer and
 read 2 sequencing primer to the SNPtype Specific Target Amplification Primer and Locus
@@
 
-To identify loci with large allele frequency differences between late-fall and fall-run Chinook salmon,
-we first conducted a simple case-control genome-wide association study (GWAS) using the low-coverage whole-genome sequencing
+Loci with large allele frequency differences between late-fall and fall-run Chinook salmon,
+were identified with a simple case-control genome-wide association study (GWAS) using the low-coverage whole-genome sequencing
 data of \citet{thompson2020complex}, mapped to the Otsh\_v1.0 Chinook salmon assembly
@@
 
-We surveyed the GWAS $p$-values relative to the commonly used significance threshold for
+The GWAS $p$-values were compared relative to the commonly used significance threshold for
 GWAS of $5\times 10^{-8}$ \citep{chen2021revisiting}.  Only a single site exceeded that threshold in both comparisons and it occurred
 atop a peak in association scores on chromosome 34 (see
-Results). From this peak we identified candidates for follow-up genotyping by including the 10
+Results). Candidates for follow-up genotyping were identified in the peak by including the 10
 SNPs in that peak with the lowest association $p$-values, and supplementing those with additional SNPs that
@@
 The final condition was implemented in order to gather several SNPs with large allele frequency differences
-that were adjacent to, but not directly within, the main peak of association, as we had hoped that those SNPs
-might enable us to learn about recombination in the region.
+that were adjacent to, but not directly within, the main peak of association, so as to possibly learn about
+recombination in the region.
 
-For the candidate markers found by the above process, we designed amplification primers using
+We designed amplification primers for the candidates using
 Primer3 \citep{koressaar2007enhancements,untergasser2012primer3} and tested amplification in a
 subset of Chinook salmon
-samples. The primer pairs that amplified sequence that reliably mapped back to the association-peak region were then typed
+samples. The primer pairs that amplified sequence that reliably mapped back to the association-peak region were typed
 in a separate collection of 1,638 Sacramento River Chinook salmon sampled from the fish trap below Keswick Dam, including 1,169 winter run, 281 spring
@@
 were
-estimated for each of these four groups: winter, spring, fall, and late-fall, and we retained
-amplicons showing large allele frequency differences between late-fall fish and all other runs.
+estimated for each of these four groups: winter, spring, fall, and late-fall, and
+amplicons showing large allele frequency differences between late-fall fish and all other runs were retained.
 
@@
 a simple association study, with only 16 samples, as done to identify the LFAR.  Consequently we pursued a bespoke
-analysis to find regions we could expect to have large allele frequency differences between
+analysis to find regions expected to have large allele frequency differences between
 the winter-run and all other ecotypes of Chinook salmon in California's Central Valley.   This analysis
@@
  a database for understanding the spatial and temporal
-distribution of the alleles in this genomic region; however we have omitted these markers for calculating likelihoods for
+distribution of the alleles in this genomic region; however these markers are omitted from likelihood calculations for
 GSI, since their distribution across populations in the Feather River has been clearly and evidently
@@
 in two genes, VGLL3 and SIX6,
-that have been found to be associated with phenotypes related to reproductive maturity in both
-Atlantic \citep{barson2015sex} and Pacific \citep{waters2021heterogeneous} salmonids. We include
-them both to provide additional variation for population genetic analyses and to
+found to be associated with phenotypes related to reproductive maturity in both
+Atlantic \citep{barson2015sex} and Pacific \citep{waters2021heterogeneous} salmonids. They
+provide additional variation for population genetic analyses and to
 monitor populations for potential phenotypic patterns.
@@
 marker characterizes genetic sex in diverse California salmon populations more accurately than previously published assays  \citep{von2015development}.
-And, finally, we include marker Ots\_coho001\_05\_32691399 which is fixed for alternate alleles in
-coho salmon and Chinook salmon.  This is helpful when samplers have unwittingly
+Finally, marker Ots\_coho001\_05\_32691399, which is fixed for alternate alleles in
+coho salmon and Chinook salmon, is included to diagnose when samplers have unwittingly
 collected coho salmon.
@@
 using both bwa mem \citep{bwa-mem2009} and the BLAST-like Alignment tool, BLAT \citep{kent2002blat}.
-A similar mapping exercise was performed for the remaining markers so that we can
+A similar mapping exercise was performed for the remaining markers to
 report their locations in both Otsh\_v1.0 and Otsh\_v2.0.
@@
 to the manufacturer's instructions. Samples were pooled by plate
-and then purified using Agencourt AMPure XP magnetic beads.
+and purified using Agencourt AMPure XP magnetic beads.
 Libraries were quantified either by qPCR with the NEBNext Library Quant Kit for Illumina (New
@@
 using simple operations in the R programming language, Version 4.3
-\citep{rcore}. We started with a simple tally of the number of alleles
-found at each locus.  To provide a summary of the genetic variation and
-missing data we further summarized:
+\citep{rcore}. Summaries provided are:
 \begin {itemize}
@@
 When subsampling each collection to the minimum number of genotyped
-fish across all collections at a locus, we subsampled individuals without
+fish across all collections at a locus, individuals were subsampled without
 replacement, taking the average of 1,000 different random subsamples.
@@
 all collections, using the {\footnotesize\tt pp.fst} function from the R
-package `hierfstat' \citep{hierfstat}. Finally, we evaluated population
-structure in the data by performing unsupervised clustering 
+package `hierfstat' \citep{hierfstat}. Population
+structure in the data was evaluated
 using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
@@
 using default settings of the program.  STRUCTURE output was
-a summarized and visualized using CLUMPAK
+summarized and visualized using CLUMPAK
 \citep{kopelman2015clumpak}.
@@
 To assess the power of the marker panel for GSI, we used the R package
-`rubias' \citep{moran2019bayesian}.  We used the function {\tt self\_assign()}
-to perform assignment of each individual in the reference baseline to one of the
+`rubias' \citep{moran2019bayesian}.  The function {\tt self\_assign()} was used
+to assign each individual in the reference baseline to one of the
 collections within the baseline.  Each fish was removed from the reference baseline
@@
 broken down according to the genotype of the fish at the markers within
-the RoSA.  For this we identified the RoSA genotype \citep[EE, EL, or LL, see~][]{thompson2020complex} from the concordance of the genotypes
+the RoSA.  RoSA genotype \citep[EE, EL, or LL, see~][]{thompson2020complex}
+was assessed from the concordance of the genotypes
 at all 8 markers described in \citep{thompson2020complex}, requiring that no genotypes be missing
@@
 different values of the false negative rate (FNR) for a variety of different pairwise
-relationships in different populations.  For these purposes, we divided the
-collections into 10 different groups.  The members of each group are
+relationships in different populations.  For these purposes, collections were divided into
+10 different groups.  The members of each group are
 relatively, genetically similar, but the groups do not necessarily correspond to reporting
-units, because, in some cases, we desired FPRs and FNRs for a single
-collection.   For all cases, we assumed that the genotyping error rate
+units, because, in some cases, FPRs and FNRs are desired for a single
+collection.   The assumed genotyping error rate
 was 1\% per locus and the genotyping error model was the default
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+\caption{\footnotesize  Summary of amplicons in the reference data set.  Source denotes
+which discovery effort yielded the amplicon. N (amplicon) is the number of amplicons
+and N (variant) is the total number of variant sites routinely assayed within the sequence of each
+amplicon.  Source abbreviations are as follows: mhap = novel microhaplotype discovery, scon = conversion of existing SNPtype
+assays, rosa = adult-migration-timing associated markers, wrap = winter-run associated
+polymorphisms, lfar = late-fall-run associated polymorphisms, sixx and vgll = amplicons
+targeting the SIX6 and VGLL3 genes, respectively, sexy = sexID marker, coho =
+distinguishes coho from Chinook salmon. The $^*$ denotes markers routinely
+excluded from likelihood calculations for GSI.}
+\label{tab:loci-summ}
+{
+\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrr}
+\hline\hline
+Type&N (amplicon)&N (variant)\tabularnewline
+\hline
+\input{inputs/amplicon-summary.tex}
+\end{tabular*}
+}
+\vspace*{-2.3ex}\hrule width0.48\textwidth
+\vspace*{0.3ex}\hrule width0.48\textwidth
+\end{table}
+%%%%%%%%%%%%%%%
+
 A summary, which includes genomic location, consensus sequence, and primer sequences,
@@
 
-We tested the 229 candidates from our novel microhaplotype discovery process and
-initially retained the 125 markers described in \citep{thompson2020complex}
+Of 229 tested candidates from our novel microhaplotype discovery process
+125 markers were retained \citep{thompson2020complex}
 for use in stock identification within
 the Klamath Basin.  However, further evaluation of those markers in other
-populations led us to remove some of the markers.
-Twelve loci appeared to have null alleles in the winter-run population and were removed.
-Additionally, two loci were monomorphic in the winter-run, even though they contained
-multiple alleles in other populations, and an additional 5 amplicons were deemed
-difficult to score in routine application of the baseline, and were removed, leaving 106
+populations led the removal of
+twelve loci with apparent null alleles in the winter-run population, two that were monomorphic
+in the winter-run, despite having
+multiple alleles in other populations, and an additional 5 loci that were
+difficult to score in routine application of the baseline. This left 106
 novel microhaplotype amplicons in the current panel.
@@
 $p$-values.  These are not shared between the two different late-fall to fall comparisons, and
-sites that we investigated were concentrated in areas of poor mapping, suggesting that their
+were concentrated in areas of poor mapping, suggesting that their
 $p$-values were artifactual.  By contrast, a large number of sites near the prominent peak on
@@
 10 SNPs with the lowest association $p$-values, and including additional ones from the filtering criteria given in the
-Methods, yielded 49 candidate markers for which we designed primers for amplication (Figure~\ref{fig:lfar-candi}).  Only 8
+Methods, yielded 49 candidate markers for which we designed primers for amplification (Figure~\ref{fig:lfar-candi}).  Only 8
 of these 49 primer pairs yielded reliable amplification and mapping.   Of these 8, two of the SNPs
@@
 }
-\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\vspace*{-2.3ex}\hrule width0.48\textwidth
+\vspace*{0.3ex}\hrule width0.48\textwidth
 \end{table}
@@
 
-Though the winter run fish are easily identified, we undertook an additional effort to
-try to find markers that could be diagnostic for the winter run, distinguishing the
-winter run from all other groups of Chinook salmon in the CCV. Initial examination of sequence data 
+Initial examination of sequence data 
 in the 16 winter-run fish and 84 non-winter run fish revealed
-a number of SNPs that were potentially diagnostic, but none were fixed for alternative alleles 
+a number of SNPs that were potentially diagnostic, but none were fixed for alternate alleles 
 in a larger sample.
-Nonetheless, three of these markers from that effort had large frequency differences
+Nonetheless, three of these markers had large frequency differences
 between the groups and are included in the reference baseline.  The process of 
@@
 All individuals were assigned here, using a maximum-scaled-likelihood rule to reporting unit and
-then to collection within reporting unit as described in the text. For the assignment matrix resulting
+to collection within reporting unit as described in the text. For the assignment matrix resulting
 when only fish with a scaled likelihood $>0.80$ to reporting unit are assigned, see Figure~\ref{fig:eighty}}
@@
 incorrectly categorized as parent-offspring (PO); as well as FPRs of parent-offspring (PO) pairs or half-sibling (HS)
-pairs being incorrectly inferred as full siblings (FS).  In both {\em a} and {\em b}, 100,000 Monte Carlo replicates
-were simulated.  Importance sampling allows estimation of very small false positive rates for unrelated pairs, but
-cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities
-($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$
-is the estimated standard error of the mean of the Monte Carlo sample.  This also is the reason that many of the lines terminate.   }
+pairs being incorrectly inferred as full siblings (FS).  In both {\bf a} and {\bf b}, 100,000 Monte Carlo replicates
+were simulated.}
 \begin{center}
@@
 summarised in (Figure~\ref{fig:fprs}b). Here, again, focusing on the results for the genetically
-depauperate SRW, and using an FNR of 0.1 we see that the chance of incorrectly
+depauperate SRW, and using an FNR of 0.1, the chance of incorrectly
 categorizing an HS as an FS,  or an FS as a PO is less than 1 in 100. Likewise, 
@@
 at later life stages,  so a FPR of  $<$0.001 is comfortably low in that case.
+Estimation of
+error rates between different kin-categories cannot be done in `CKMRsim' using
+importance sampling. Consequently, it is difficult to accurately estimate the small
+probabilities ($10^{-4}$ or less) in Figure~\ref{fig:fprs}b, as indicated by the error bars,
+which represent the estimate $\pm2s$ where s is the estimated standard error of the
+mean of the Monte Carlo sample.  However, as noted above, because there are
+far fewer related pairs than unrelated pairs, it is not typically essential to
+have accurate estimates of very low FPRs for related pairs.
 
@@
 ecotypes of CCV Chinook salmon is of utmost importance for monitoring and
-evaluation of individual ecotypes, as well as designing and implementing
+evaluation of individual ecotypes, and for designing and implementing
 effective conservation and management actions. This has been challenging,
@@
 idea that early-migrating populations arise through migration of small numbers of individuals carrying 
-E-lineage haplotypes, in either heterozygous or homozygous form, which are then positively selected
+E-lineage haplotypes, in either heterozygous or homozygous form, which are subsequently positively selected
 when appropriate habitat conditions exist. A similar pattern of recombination, due to long-standing
@@
 salmon, by providing sufficient power for parent-offspring relationship inference, when only
-one parent is sampled, as well as identifying pairs of full siblings when no parents are sampled.
+one parent is sampled, and identifying pairs of full siblings when no parents are sampled.
 Groups of full siblings larger than two can be identified even more accurately than expressed in Figure~\ref{fig:fprs},
@@
 Gilbert-Horvath for contributions to data collection. The manuscript was improved by comments from
-Jeff Rodzen. USDA is an equal opportunity provider and employer. Mention of trade names
+Jeff Rodzen and from two anonymous referees. USDA is an equal opportunity provider and employer. Mention of trade names
 or commercial products in this publication is solely for the purpose of providing specific

### 2025-03-24 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Getting close to finishing reponses to referees

@@
 
-* **I commend the use of the distribution of log-likelihood ratio for relationship to show the power of haplotypes vs SNPs and would argue this could be elevated to a full figure, potentially in a condensed form, maybe less panels somehow?**
+* I commend the use of the distribution of log-likelihood ratio for relationship to show the power of haplotypes vs SNPs and would argue this could be elevated to a full figure, potentially in a condensed form, maybe less panels somehow?
+    - _We appreciate the referee's feedback on this.  In response, we have created a
+    new figure in the paper itself that examines the distribution of log-likelihood
+    ratios only in the Feather River Hatchery.  This figure distinguishes between
+    different log-likelihood ratios (PO/U, FS/U, and HS/U) and nicely shows the
+    advantages of using microhaplotypes for relationship inference. We have also left
+    the original figure in the supplement for any readers curious about the effect
+    in different collections._
 
-    - **I am planning to break one population out into a single panel and show different logl-ratios
-    for different relationships.  Probably a 2x2 or 2x2 grid. Intend to leave the others in the
-    supplement as is.**
-
-* **If possible, I think it would also be good to show how much of an improvement the discrimination power improved from the old marker data.
-In theory, this data has already been generated for the full population if you strip out the novel SNPs of the current study.**
-    - **We could do this for the 78 SNPlicons plus 16 other single-SNPs, but it would be a little hard. We might just
-    say no.**
+* If possible, I think it would also be good to show how much of an improvement the discrimination power improved from the old marker data.
+In theory, this data has already been generated for the full population if you strip out the novel SNPs of the current study.
+    - _After long and careful consideration of this, we determined that it would not
+    be possible to make an accurate comparison between the old and the new marker
+    data in this regard.  Unfortunately, only 78 of the original 94 TaqMan assays
+    were able to be converted into amplicon-sequenced markers, so we are not able
+    to reproduce the old marker set with the same reference samples._
 
@@
 
-- **Sentence at lines 645-647. Can the authors provide a reference to back up this statement.**
+- Sentence at lines 645-647. Can the authors provide a reference to back up this statement.
+    - _We added a citation here that extensively describes the use of a Fluidigm chip-based panel for
+    parent pair allocation in Chinook salmon in California._
 

### 2025-03-24 003-ckmr-sim-evaluation.Rmd
Getting close to finishing reponses to referees

@@
     ckmr_comp_crop = "results/ckmr-sim-eval/cmkr-comp-figure-crop.pdf",
-    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf"
+    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf",
+    frh_comp = "results/ckmr-sim-eval/frh-comp-figure.pdf",
+    frh_comp_crop = "results/ckmr-sim-eval/frh-comp-figure-crop.pdf",
+    frh_comp_tex = "tex/images/frh-comp-figure-crop.pdf"
 
@@
   rename(Locus = AmpliconName, Chrom = Otsh_v2.0_chromosome, Pos = AmpliconStartPos) %>%
-  select(Locus, Chrom, Pos)
+  select(Locus, Chrom, Pos) %>%
+  filter(!str_detect(Locus, "sexy"))
 
@@
     ungroup() %>%
+    filter(!is.na(Locus)) %>%  # this gets rid of the rosa stuff
     left_join(GC2, by = join_by(Locus)) %>%
@@
 
-source("R/shift_legend.R")
-
 # first do the importance sampling ones
@@
 
-a <- grid.arrange(shift_legend(g))
+library(lemon) 
+a <- reposition_legend(g, panel = "panel-2-3",  x = 0, y = 0, just = c(0,0))
 ```
@@
 
-a2 <- grid.arrange(shift_legend(g2))
-
+a2 <- reposition_legend(g2, panel = "panel-2-3", x = 0, y = 0, just = c(0,0))
 ```
@@
 
-Neil suggested, and Carlos and I agreed, that one way to see they improvement in
+Neil suggested, and Carlos and I agreed, that one way to see the improvement in
 relationship inference obtained by using multiallelic microhaps is to look at the
@@
 
-  # now extract the logl_ratios for PO vs U and return those
+  # now extract the logl_ratios
   extract_logls(Q_link, numer = c(PO = 1), denom = c(U = 1))
@@
 ```{r}
-# collections / groups we want to do
+# collections / groups we want to do.  I originally did many, but
+# just whittled it down to include in main text following a reviewer request.
 pops <- c("ck-FRH", "ck-IGH", "ck-SRW", "ck-TRH")
@@
 ```
+
+## Making a comparison figure for the main text
+
+Reviewer one wanted to see this in the main text.  I will show just a single
+figure for FRH.  I also want to look at PO, FS, and FS versus U specifically.
+```{r}
+pop <- c("ck-FRH")
+
+# let's put the microhaps and the best SNPs together in a long format.
+input_tibbles2 <- bind_rows(
+  ckmr_sim_nests %>%
+    filter(ckmr_sim_group %in% pop) %>%
+    mutate(condition = "microhaps", .before = ckmr_sim_group),
+  ckmr_sim_nests %>%
+    filter(ckmr_sim_group %in% pop) %>%
+    mutate(tops = map(data, retain_top_snp)) %>%
+    mutate(data = tops) %>%
+    select(-tops) %>%
+    mutate(condition = "top_snps", .before = ckmr_sim_group)
+)
+
+# a new function that gets multiple logls out of it
+get_logl_rats_multi <- function(X, GC2 = gen_coords) {
+
+  Y <- ckmr_afreqs_ready(X, GC2)
+
+  C <- ckmr_create(Y)
+
+  # throttle text output to console
+  options(CKMRsim.discard_stdout = TRUE)
+  options(CKMRsim.linkage_verbosity = 0)
+
+  # Simulate some Q_ij's. Suppress output messages that the CKMRsim package
+  # produces
+  Q_link <- ckmr_simQ(C) #%>% suppressMessages()
+
+  # now extract the logl_ratios for PO, FS, or HS vs U and return those
+  list(
+    POU = extract_logls(Q_link, numer = c(PO = 1), denom = c(U = 1)),
+    FSU = extract_logls(Q_link, numer = c(FS = 1), denom = c(U = 1)),
+    HSU = extract_logls(Q_link, numer = c(HS = 1), denom = c(U = 1))
+  ) %>%
+    bind_rows(.id = "ratio_type")
+}
+
+Qsims2 <- lapply(input_tibbles2$data, get_logl_rats_multi)
+
+AllSims2 <- tibble(
+  condition = input_tibbles2$condition,
+  ckmr_sim_group = input_tibbles2$ckmr_sim_group,
+  Qsims = Qsims2
+) %>% 
+  unnest(Qsims) %>%
+  filter(  # only keep the relevant ones
+    (ratio_type == "POU" & true_relat %in% c("U", "PO")) |
+      (ratio_type == "FSU" & true_relat %in% c("U", "FS")) |
+      (ratio_type == "HSU" & true_relat %in% c("U", "HS"))
+  )
+```
+
+Now, we make a picture of that.
+```{r}
+library(ggpattern)
+AS3 <- AllSims2 %>%
+  mutate(
+    ratio_type = case_match(
+      ratio_type,
+      "POU" ~ "Parent-offspring versus Unrelated log-likelihood ratio",
+      "FSU" ~ "Full-sibling versus Unrelated log-likelihood ratio",
+      "HSU" ~ "Half-sibling versus Unrelated log-likelihood ratio"
+    ),
+    ratio_type = factor(
+      ratio_type,
+      levels = c(
+        "Parent-offspring versus Unrelated log-likelihood ratio",
+        "Full-sibling versus Unrelated log-likelihood ratio",
+        "Half-sibling versus Unrelated log-likelihood ratio"
+      )
+    )
+  ) %>%
+  rename(`True\nRelationship` = true_relat)
+
+g8 <- ggplot(AS3, aes(x = logl_ratio, fill = `True\nRelationship`, alpha = condition)) +
+  geom_density() +
+  scale_alpha_manual(values = c(microhaps = 0.7, top_snps = 0.3)) +
+  facet_wrap(~ratio_type, ncol = 1, scales = "free") +
+  xlab("Log-likelihood Ratio") +
+  ylab("Density") +
+  theme_bw() +
+  theme(legend.position = "top") +
+  guides(alpha = "none")
+
+ggsave(g8, filename = output_list$frh_comp, width = 4, height = 6)
+
+# let's crop that down for use, too
+CALL <- paste("pdfcrop ", output_list$frh_comp, collapse = " ")
+system(CALL)
+
+file.rename(from = output_list$frh_comp_crop, to = output_list$frh_comp_tex)
+
+```

### 2025-03-24 tex/body-text-of-paper.tex
Getting close to finishing reponses to referees

@@
 4,794 retained in the data with 500 bp size selection.
-Gene regions were selected that (1) contained multiple SNPs within a 100 bp window, and (2) showed haplotype variation amongst the winter run and at least one other population.
-A 100 bp window was used because the success of multiplex PCR and MiSeq sequencing is diminished in amplicons larger than 100 bp. Primer3
+
+The 8,725 discovered RAD loci were filtered down to a set of markers suitable for conversion
+to amplicon sequencing assays. Gene regions were selected from the RAD loci that (1) contained multiple SNPs within a 100 bp window,
+and (2) showed haplotype variation amongst the winter run and at least one other population.
+A 100 bp window was used because the success of multiplex PCR and MiSeq sequencing is
+diminished in amplicons larger than 100 bp. Primer3
 software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.
-
-These gene regions were PCR-amplified and sequenced on a MiSeq sequencer using a $2\times 75$-cycle kit.
+These 229 gene regions were then tested for PCR-amplification and consistency.  This step included
+amplification in thousands of samples from a number of different collections of Chinook salmon for sampled
+for several projects, with
+subsequent sequencing on a
+MiSeq sequencer using a $2\times 75$-cycle kit.
 The sequences from each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
 the Burrows-Wheeler aligner using the mem algorithm \citep{bwa-mem2009}. 
-Subsequently, alignmets were sorted with SAMtools \citep{li2009sequence} and provided to
+Subsequently, alignments were sorted with SAMtools \citep{li2009sequence} and provided to
 FreeBayes \citep{garrison2012haplotype} for variant calling.
@@
 amplicons, versus using the most heterozygous single SNP from each
-amplicon (\ref{fig:ckmr-comp}).  The figures show a considerable increase
-in separation between the relationships categories from calling genotypes
+amplicon for all fish from the FRH (Fig.~\ref{fig:ckmr-comp}), with results for the PO vs Unrelated
+log likelihood ratio shown for more collections in Fig.~\ref{fig:ckmr-comp-supp}.  The figures show a considerable increase
+in separation between the relationship categories from calling genotypes
 as microhaplotypes in these amplicons.
+%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\newcommand{\fprcap}{\footnotesize The distribution of log-likelihood ratios for
+fish from Feather River Hatchery when scoring markers as microhaplotypes versus as
+single SNPs.  Three different likelihood ratios are shown in the three panels.  The opacity
+of the distributions indicates the marker type: the more opaque setting indicates markers
+scored as microhaplotypes; the more transparent setting indicates the use of only the most
+heterozygous SNP from amongst all SNPs in the microhaplotype.  Evidently, scoring the markers
+as microhaplotypes produces more separation between the distributions for the relationships of PO, FS, or HS, against
+the relationship of unrelated (U).}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{images/frh-comp-figure-crop.pdf}
+\end{center}
+\caption[\fprcap]{\fprcap}
+\label{fig:ckmr-comp}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
 
@@
 Previous genetic marker sets for California Chinook salmon had sufficient power for inferring
-parent-offspring relationships, but only when both parents were sampled and genotyped.
+parent-offspring relationships, but only when both parents were sampled and genotyped \citep{clemento2013creation}.
 This marker set considerably increases the capacity for relationship inference in California

### 2025-03-24 001-map-of-samples.Rmd
Make run types different shapes on the map

@@
 
+# define the shapes
+point_shapes <- c(
+  `Fall run` = 21,
+  `Spring run` = 22,
+  `Winter run` = 23,
+  `Late-fall run` = 24
+)
+
 mm <- base_map + 
   geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", linewidth = 0.4) +
-  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
+  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing, shape = run_timing), size = 3.5) +
   geom_label(
@@
   scale_fill_manual(values = run_time_colors) +
+  scale_shape_manual(values = point_shapes) +
   theme_bw() +
@@
 
-file.copy(from = output_list$cropmap, to = output_list$texmap)
+file.copy(from = output_list$cropmap, to = output_list$texmap, overwrite = TRUE)
 
@@
   g <- ggplot(tmp, aes(x = 1, y = 1)) +
-    geom_point(shape = 21, fill = tmp$value, size = 3.5) +
+    geom_point(shape = point_shapes[tmp$name], fill = tmp$value, size = 3.5) +
     theme_void()

### 2025-03-24 R/table-as-figure.R
Add stripes to Late-fall color blocks

@@ -103,17 +103,25 @@ table_as_figure <- function(
     summarise(val = last(val)) %>%
     filter(val != max(val))
 
+  require(ggpattern)
+  patterns = c(`Fall_run` = "none", `Late-fall run` = "stripe", `Spring run` = "none", `Winter run` = "none")
+
   # then put these together into a big ggplot
   all_but_cell_labels <- ggplot() +
-    geom_rect(  # outer edge filled rectangles
+    geom_rect_pattern(  # outer edge filled rectangles
       data = Z,
       mapping = aes(
         xmin = xmin,
         xmax = xmax,
         ymin = ymin,
         ymax = ymax,
-        fill = external_group
-      )
+        fill = external_group,
+        pattern = external_group
+      ),
+      pattern_fill    = 'black',
+      pattern_colour  = 'black',
+      pattern_size = 0.05,
+      pattern_spacing = 0.005
     ) +
     geom_label( # outer edge labels
       data = Z,
@@ -147,6 +155,7 @@ table_as_figure <- function(
     geom_hline(yintercept = N - internal_lines_tib$val, linewidth = thick_line_width) +
     coord_cartesian(xlim = c(-1, N+1), ylim = c(-1, N+1), expand = FALSE) +
     scale_fill_manual(values = c(external_colors, internal_colors), na.value = "white") +
+    scale_pattern_manual(values = patterns) +
     theme_void() +
     theme(
       legend.position = "none",
@@ -202,12 +211,28 @@ table_as_figure <- function(
   # the legends with cowplot
   for_external_legend <- RC_groups %>%
     mutate(egf = factor(external_group, levels = unique(external_group))) %>%
-    ggplot(aes(x = egf, fill = egf)) +
-    geom_bar(color = "black") +
+    ggplot() +
+    geom_bar_pattern(
+      aes(
+        x = egf,
+        fill = egf,
+        pattern = egf
+      ),
+      color = "black",
+      pattern_fill    = 'black',
+      pattern_colour  = 'black',
+      pattern_size = 0.05,
+      pattern_spacing = 0.01
+    ) +
     scale_fill_manual(
       values = external_colors,
       name = external_color_legend_name
+    ) +
+    scale_pattern_manual(
+      values = patterns,
+      name = external_color_legend_name
     )
+
 
   for_internal_legend <- RC_groups %>%
     mutate(igf = factor(internal_group, levels = unique(internal_group))) %>%

### 2025-03-25 002-GSI_and_Fst.Rmd
small tweaks

@@
   select(-old_name) %>%
-  select(indiv, repunit, collection, sample_type, everything()) %>%
-  filter(!duplicated(indiv))  # necessary cuz clemento stuck mill and deer together into a single code
+  select(indiv, repunit, collection, sample_type, everything())
 

### 2025-03-25 tex/responses-to-referees/responses-to-reviewers-1.Rmd
more responses

@@
 Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.**
-    - **Gotta still see what he is talking about here.**
+    - **Oh crap! We were missing 42 fish from the meta-data.  When Anthony gets
+    these for me, I will include and regenerate Table 1.  I will also just do
+    proportion of missing data, instead of average num indivs.**
 
@@
 
-- **Figure 1 and 2: The colours for “spring run” and “Late-fall run” are very difficult to separate for a colour-blind person.
+- Figure 1 and 2: The colours for “spring run” and “Late-fall run” are very difficult to separate for a colour-blind person.
 Since the groups are not in a specific order on the graph, it is very hard work to out what is what.
-I’d recommend the authors use a selection of colours that will work for colour-blind people.**
+I’d recommend the authors use a selection of colours that will work for colour-blind people.
 
-    - **Planning to just cross-hatch the light blue of the late fall**
+    - _We greatly appreciate this feedback and apologize for the difficulty in
+    discerning those two colors. We have introduced several changes to the figures
+    that we hope will make the distinction between those colors much easier.
+    In order to keep the color scheme consistent, as much as possible,
+    with previous publications we have employed different shapes and striping to
+    make the different run types more distinct for colour-blind people. Specifically,
+    Figure 1 (the map) now uses different colors and shapes for the different run
+    timing groups.  Figure 2 using striping across the Late-fall run group sectors
+    to clearly distinguish that color from the spring-run color._
 

### 2025-03-25 002-GSI_and_Fst.Rmd
Use fraction missing instead of average fish non-missing

@@
     rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf",
-    pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv"
+    pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv",
+    popgen_summ = "tex/inputs/popgen-summary.tex"
   )
@@
   nest()
-
-# Remove this once I get the full meta data
-fbmet <- fbmet %>% filter(!is.na(WATERSHED))
 
@@
 ```{r}
-fbl_no_miss <- full_base2 %>%
+# first pivot and get whether each locus is Hz or not
+fbl_10 <- full_base2 %>%
   select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
@@
   mutate(isHz = value[1] != value[2]) %>%
-  ungroup() %>%
+  ungroup()
+
+# then get the proportion of missing data for each population
+miss_props <- fbl_10 %>% 
+  group_by(repunit, collection) %>%
+  summarise(fractMiss = mean(is.na(value))) %>%
+  ungroup()
+
+fbl_no_miss <- fbl_10 %>%
   filter(!is.na(value) & !is.na(isHz))
@@
 
-# here, we get the minimum number of diploids at each locus
+# here, we get the minimum number of diploids at each locus, which we
+# use for rarefaction sampling.
 min_gene_copies <- simple_summaries %>%
@@
 
-# and now we can summarise there into means and SDs
+# and now we can summarise these into means and SDs
 mean_Hzs_etc <- locus_summaries %>%
@@
   left_join(mean_Hzs_etc, by = join_by(repunit, collection)) %>%
+  left_join(miss_props, by = join_by(repunit, collection)) %>%
   mutate(collection = factor(collection, levels = unique(pop_labels$collection))) %>%
@@
   ungroup() %>%
-  select(collection, mean_num_dip, mean_num_alle, mean_fract_poly, mean_expHz, mean_obsHz) %>%
+  select(collection, fractMiss, mean_num_alle, mean_fract_poly, mean_expHz, mean_obsHz) %>%
   mutate(
     collection = as.character(collection),
-    mean_num_dip = sprintf("%.1f", mean_num_dip),
+    fractMiss = sprintf("%.3f", fractMiss),
     mean_num_alle = sprintf("%.2f", mean_num_alle),
@@
     Code = collection,
-    `$\\bar{N}$` = mean_num_dip,
+    `$M$` = fractMiss,
     `$\\bar{N}_\\mathrm{A,ss}$` = mean_num_alle,
@@
   delim = "&",
-  file = "tex/inputs/popgen-summary.tex", 
+  file = output_list$popgen_summ, 
   eol = "\\tabularnewline\n", 

### 2025-03-25 tex/body-text-of-paper.tex
Make run types different shapes on the map
Use fraction missing instead of average fish non-missing

@@
 three rivers of the California coast (Russian, Eel, and Smith).
-Of the 1594 total fish samples from 17 different populations, 724 fish from 10 collections were of the fall run ecotype,
-498 fish from 5 collections were spring run, and 261 and 111 fish were of the winter run and late-fall run ecotypes, 
+Of the 1636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall run ecotype,
+498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter run and late-fall run ecotypes, 
 respectively,  each represented by a single collection.
@@
 each location name represent the different run-timing groups of Chinook salmon
-included in the collections from the location.  Color codes are as follows: \Wball~=~Winter run,
+included in the collections from the location.  Color and shape codes are as follows: \Wball~=~Winter run,
 \LFball{}~=~Late-fall run, \Fball~=~Fall run, \Sball~=~Spring run.
@@
 which discovery effort yielded the amplicon. N (amplicon) is the number of amplicons
-and N (variant) is the total number of variant sites routinely assayed within the sequence of each
-amplicon.  Source abbreviations are as follows: mhap = novel microhaplotype discovery, scon = conversion of existing SNPtype
-assays, rosa = adult-migration-timing associated markers, wrap = winter-run associated
+and N (variant) is the total number of variant sites assayed within the amplified sequence of all the
+amplicons.  Source abbreviations are as follows: mhap = novel microhaplotype discovery, scon = conversion of existing SNPtype
+assays (the `SNPlicons'), rosa = adult-migration-timing associated markers, wrap = winter-run associated
 polymorphisms, lfar = late-fall-run associated polymorphisms, sixx and vgll = amplicons
@@
 \caption{\footnotesize Simple population genetic summaries. All quantities are means
-over loci.  $\bar{N}$ is the average over loci of the number of fish in each collection with observed genotypes
-at a locus; $\bar{N}_\mathrm{A,ss}$ is the
+over loci.   $M$ is the total fraction of missing genotypes; $\bar{N}_\mathrm{A,ss}$ is the
 average number of alleles and $\bar{P}_\mathrm{poly,ss}$ the fraction of polymorphic loci,

### 2025-03-25 tex/inputs/samples-table.tex
Eric's near-final edits
Use fraction missing instead of average fish non-missing

@@
-Mainstem River$^1$&Code&Location Name$^2$&Run&Reporting Unit$^4$&ESU$^4$&$N$&$N_\mathrm{day}^5$&Sampling Months&Year Range\tabularnewline
+Mainstem River$^1$&Code&Location Name$^2$&Run&Reporting Unit$^3$&ESU$^4$&$N$&$N_\mathrm{day}^5$&Sampling Months&Year Range\tabularnewline
 \hline Rogue R.&CRHS&Cole Rivers H.&Spring&SO-NCal-Coast&SONCC&50&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CRHS_Cole-Rivers-Hatchery.pdf}}&2004--2010\tabularnewline
@@
 Sacramento R.&FRHS&Feather R. H.&Spring&CV-Fall&CVSR&202&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
-Sacramento R.&FRHF&Feather R. H.&Fall&CV-Fall&CVFR&106&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
-Sacramento R.&BCF&Butte Ck.&Fall&CV-Fall&CVFR&45&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
-Sacramento R.&MDF&Deer Ck.&Fall&CV-Fall&CVFR&24&24&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
+Sacramento R.&FRHF&Feather R. H.&Fall&CV-Fall&CVFR&107&107&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
+Sacramento R.&BCF&Butte Ck.&Fall&CV-Fall&CVFR&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
+Sacramento R.&MDF&Deer Ck.&Fall&CV-Fall&CVFR&25&25&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
 Sacramento R.&MDF&Mill Ck.&Fall&CV-Fall&CVFR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
 San Joaquin R.&SJRF&San Joaquin R.&Fall&CV-Fall&CVFR&79&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
-Sacramento R.&CHLF&Coleman NFH&Late-fall&CV-Late-Fall&CVFR&261&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
+Sacramento R.&CHLF&Coleman NFH&Late-fall&CV-Late-Fall&CVFR&300&299&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2017\tabularnewline

### 2025-03-25 tex/body-text-of-paper.tex
fix text for pg-summ table

@@
 only locus bearing a single allele in Chinook salmon being the species-specific marker Ots\_coho001\_05\_32691399. 
-Biallelic loci were most common, at 66 of the total, but two-thirds of the microhaplotype markers
+Biallelic loci were most common, at 66, but two-thirds of the microhaplotype markers
 showed more than two alleles, which has been shown to provide significantly more
@@
 %%%%%%%%%%%%%%%
-The average number of fish with non-missing data in each population varied from
-a low of 40.2 (TRF) to a high of 279.6 (CHLF).  The average number of alleles
+The total fraction of missing data by collection varied from a low of
+0.008  in CRHS to a high of 0.338 in SRW, the latter due to a tranche of
+samples from degraded carcasses. The average number of alleles
 per locus, standardized to the smallest sample size, ranged from 2.32 (SRW) to

### 2025-03-25 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Done addressing comments (except Repo Readme)

@@
 
-* **I’m confused by the amount of sequencing and a clear purpose of each sequencing run that should be clarified.
+* I’m confused by the amount of sequencing and a clear purpose of each sequencing run that should be clarified.
 There was ddRAD to identify novel polymorphisms/microhaplotypes, the second run using a using PCR-GBS type sequencing has an unclear purpose.
@@
 I presume these were different runs but maybe they were the same.
-The GT-seq method, does this include all samples and the full set of markers to this point?**
-
-    - **Should be able to just add a few sentences to make this more clear.**
+The GT-seq method, does this include all samples and the full set of markers to this point?
+    - _We added to this section to make it more clear that the ddRAD one was to find gene regions, and the
+    subsequent sequencing runs were made on amplicons from those gene regions, and we are also explicit about
+    how many fish were used in those GT-seq optimization runs._
 
@@
 Methods section:_
-"Of the 1594 total fish samples from 17 different populations, 724 fish from 10 collections were of the fall run ecotype,
-498 fish from 5 collections were spring run, and 261 and 111 fish were of the winter run and late-fall run ecotypes, 
+"Of the 1636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall run ecotype,
+498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter run and late-fall run ecotypes, 
 respectively,  each represented by a single collection."
@@
 
-* **It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.**
-    - **Should be easy to throw a few citations in probably in a paragraph in the discussion.**
+* It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.
+    - _We have included four additional citation in the first paragraph of the discussion
+    to place this work in the context of those other investigations._
 
@@
 
-* **Table 3: The values in second column (the average number of fish with observed genotypes) seem odd to me.
+* Table 3: The values in second column (the average number of fish with observed genotypes) seem odd to me.
 I would have thought these values would be less than seventh column in Table 1 (column N).
@@
 Can the authors explain this?
-Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.**
-    - **Oh crap! We were missing 42 fish from the meta-data.  When Anthony gets
-    these for me, I will include and regenerate Table 1.  I will also just do
-    proportion of missing data, instead of average num indivs.**
+Furthermore, the proportion of individuals with observed genotypes is a more useful metric than the average number of non-missing genotypes, since there the number of fish between sampling sites is very different.
+    - _Thank you very much for noting this.  We have now corrected this problem. The table was
+    created from an old version of our metadata file that was missing entried from 42 fish.  These
+    fish are now properly included in the metadata file and the figures in Table 1 have been
+    regenerated. Also, we agree with the referee that the average number of non-missing genotypes
+    is a less useful metric than one quantified as a proportion.  Therefore we have replaced
+    the average number of non-missing genotypes with the total proportion of missing data in Table 1._
 

### 2025-03-25 tex/body-text-of-paper.tex
Done addressing comments (except Repo Readme)

@@
 (Qiagen, Inc). Double-digest restriction-site associated DNA sequencing (ddRAD) with different size selections (400 bp and 500 bp
-fragments) were used to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
-\citet{peterson2012double} with the modifications described in \citep{baetscher2018microhaplotypes}.
+fragments) was used to produce a broad array of locus targets. Library preparation and sequencing methods followed those of
+\citet{peterson2012double} with the modifications described in \citet{baetscher2018microhaplotypes}.
 Sequencing was performed on a MiSeq (Illumina, Inc) using $2 \times 300$-cycle paired-end sequencing.
-
 Stacks v1.45 \citep{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome
@@
 software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.
-These 229 gene regions were then tested for PCR-amplification and consistency.  This step included
-amplification in thousands of samples from a number of different collections of Chinook salmon for sampled
-for several projects, with
-subsequent sequencing on a
-MiSeq sequencer using a $2\times 75$-cycle kit.
-The sequences from each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
+These 229 gene regions were then tested for PCR-amplification and consistency in a separate sample
+of 288 Chinook salmon: 96 SRW fish and 48 each from the Entiat River (a Columbia River
+tributary), TRH, FRHS, and CRLF\@.  Eight GT-seq amplifications were made and these were sequenced
+on eight runs of a MiSeq sequencer
+using a $2\times 75$-cycle kit to identify problematic loci and to optimize primer dilutions.
+The sequences from each run at each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
 the Burrows-Wheeler aligner using the mem algorithm \citep{bwa-mem2009}. 
@@
 \begin {itemize}
-\item the average over loci of the number of fish
-in which each locus was successfully genotyped, $\bar{N}$;
+\item the total fraction of missing genotypes, $M$;
 \item the average number of alleles per locus in each collection with all collections
@@
 \subsection*{Genetic Markers}
+
+The number of markers and variants obtained from each of our discovery efforts appears in
+Table~\ref{tab:loci-summ}.  Further information, including genomic location, consensus sequence, and primer sequences,
+of all the amplicons used in the reference data set is available
+in Data Supplement~1.  A graphical overview of genomic coordinates is found in Figure~\ref{fig:genome-locations}.
 
@@
 %%%%%%%%%%%%%%%
-
-A summary, which includes genomic location, consensus sequence, and primer sequences,
-of all the amplicons used in the baseline is available
-in Data Supplement~1.  An overview of genomic coordinates is found in Figure~\ref{fig:genome-locations}
 
@@
 
-While there is not a currently available method \citep[like that of][]{anderson2006power}) for estimating error rates in
+While there is not a currently available method \citep[like that of][]{anderson2006power} for estimating error rates in
 inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying
@@
 long been in use for both GSI and PBT in California \citep{clemento2014evaluation}, by converting them
-into so-called 'SNPlicons.' This set of markers lays the foundation for a comprehensive 
+into so-called 'SNPlicons,' which are treated as multiallelic haplotypes in the presence
+of additional variation surrounding the originally assayed SNP. Our work adds to a growing
+chorus of recent studies finding that microhaplotype markers can provide additional
+power over single SNPs for both GSI \citep{mckinney2017managing,mckinney2022new,hargrove2024efficient} and
+relationship inference \citep{baetscher2018microhaplotypes,bootsma2020gt,delomas2024microhaplotypes}.
+This set of markers lays the foundation for a comprehensive 
 genetic monitoring and evaluation effort that facilitates multiple types of inference 

### 2025-03-25 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Editing the replies to referees

@@
 that were adjacent to, but not directly within, the main peak of association, so as to possibly learn about
-    recombination in the region."
+recombination in the region. "
 
@@
     ratios only in the Feather River Hatchery.  This figure distinguishes between
-    different log-likelihood ratios (PO/U, FS/U, and HS/U) and nicely shows the
+    different log-likelihood ratios (PO/U, FS/U, and HS/U) and nicely shows, in a graphical way, the
     advantages of using microhaplotypes for relationship inference. We have also left
@@
 * It may also be worth discussing other studies on other organisms that either support or do not support the utilization and increased power achieved through haplotypes over SNPs.
-    - _We have included four additional citation in the first paragraph of the discussion
+    - _We have included four additional citations in the first paragraph of the discussion
     to place this work in the context of those other investigations._
@@
     physical linkage, we just simplify the manuscript by not mentioning that technical
-    detail that can be confusing._
+    detail._
 
@@
     number of samples from each collection. Additionally, we have added several sentences
-    in the third paragraph of the introduction to define the terms reference data set and
-    baseline, and we have also added a summary in the abstract
+    in the third paragraph of the introduction to define the terms "reference data set" and
+    "baseline," and we have also added a summary in the abstract
     on the total number of amplicons, individuals and populations in the reference data set._
@@
     - _Thank you very much for noting this.  We have now corrected this problem. The table was
-    created from an old version of our metadata file that was missing entried from 42 fish.  These
-    fish are now properly included in the metadata file and the figures in Table 1 have been
+    created from an old version of our metadata file that was missing entries from 42 fish.  These
+    fish are now properly included in the metadata file and the quantities in Table 1 have been
     regenerated. Also, we agree with the referee that the average number of non-missing genotypes
@@
 - Line 181: "a de novo assembly was performed". How was this performed? More details are needed.
-    - _We clarified that that we used ustacks from the Stacks program to assemble the reads into
+    - _We clarified that we used ustacks from the Stacks program to assemble the reads into
     RAD loci._
@@
     Figure 1 (the map) now uses different colors and shapes for the different run
-    timing groups.  Figure 2 using striping across the Late-fall run group sectors
+    timing groups.  Figure 2 uses striping across the Late-fall run group sectors
     to clearly distinguish that color from the spring-run color._

### 2025-03-25 tex/abstract.tex
Edits ID'd looking at the latexdiff

--- 2024-11-06 tex/abstract.tex
@@
 resolution for population discrimination and relationship identification from a rapidly and
-economically assayed panel of markers. We show how this set of microhaplotypes
+economically assayed panel of markers. We show how this set of genetic markers assayed
+by sequencing of 204 amplicons, in tandem with a reference data set of 1636 individual
+samples from 17 populations, 
 provides definitive power to identify all known lineages of Chinook salmon in

### 2025-03-25 tex/body-text-of-paper.tex
Edits ID'd looking at the latexdiff

@@
 We present a new reference baseline that includes 1,636 Chinook salmon individuals from 17 collections that were genotyped at 204 loci distributed throughout the genome.
-This baseline is highly effective for GSI within California and provides ample power for PBT, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program 
+This baseline is highly effective for GSI within California and provides ample power for PBT inference of other close relatives, thus enabling a highly effective and efficient integrated GSI/PBT monitoring and evaluation program
 \citep{garza2007large,beacham2021parentage}.
-We describe how we identified and compiled the markers, including several new markers
+We describe how we developed and compiled the markers, including several new markers
 identified through re-analysis of the whole genome sequencing data of \citet{thompson2020complex},
@@
 Stacks v1.45 \citep{catchen2013stacks} was used to analyze the ddRAD sequence data.  As there was not a well-assembled Chinook salmon genome
-at the time, the ustacks module of Stacks were used to assemble reads {\em de novo} into RAD loci.  Initial filtering retained RAD loci with two or more SNPs present
+at the time, the ustacks module of Stacks was used to assemble reads {\em de novo} into RAD loci.  Initial filtering retained RAD loci with two or more SNPs present
 in at least 10 samples. After this initial filtering, there were 3,931 RAD loci retained in the ddRAD data with 400 bp size selection and
@@
 software \citep{untergasser2012primer3} was used to design primers for 229 gene regions that met these criteria.
-These 229 gene regions were then tested for PCR-amplification and consistency in a separate sample
+These 229 gene regions were tested for PCR-amplification and consistency in a separate sample
 of 288 Chinook salmon: 96 SRW fish and 48 each from the Entiat River (a Columbia River
@@
 The sequences from each run at each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
-the Burrows-Wheeler aligner using the mem algorithm \citep{bwa-mem2009}. 
+the Burrows-Wheeler (bwa) aligner using the mem algorithm \citep{bwa-mem2009}. 
 Subsequently, alignments were sorted with SAMtools \citep{li2009sequence} and provided to
@@
 had poor amplification, or had excessive amplification and were subsequently over-represented in the sequencing library.
-FreeBayes \citep{garrison2012haplotype} was used to identify all SNPs (additional to the TaqMan-assay targets) along the fragment, which facilitated scoring fragment microhaplotypes when multiple SNPs were present along the read \citep{baetscher2018microhaplotypes}.
+FreeBayes was used to identify all SNPs (additional to the TaqMan-assay targets) along the fragment, which facilitated scoring fragment microhaplotypes when multiple SNPs were present along the read \citep{baetscher2018microhaplotypes}.
 The resulting genetic markers are hereafter referred to as `SNPlicons' to record their origin as TaqMan assays in a previous
@@
 and version 2  (Otsh\_v2.0, GenBank accession GCA\_018296145.1) genome assemblies,
-using both bwa mem \citep{bwa-mem2009} and the BLAST-like Alignment tool, BLAT \citep{kent2002blat}.
+using both bwa mem and the BLAST-like Alignment tool, BLAT \citep{kent2002blat}.
 A similar mapping exercise was performed for the remaining markers to
@@
 broken down according to the genotype of the fish at the markers within
-the RoSA.  RoSA genotype \citep[EE, EL, or LL, see~][]{thompson2020complex}
+the RoSA.  RoSA genotype \citep[EE, EL, or LL, see][]{thompson2020complex}
 was assessed from the concordance of the genotypes
-at all 8 markers described in \citep{thompson2020complex}, requiring that no genotypes be missing
+at all 8 markers described in \citet{thompson2020complex}, requiring that no genotypes be missing
 within an individual from any of those 8 markers, and discarding data from 12 individuals
@@
 \label{tab:loci-summ}
-{
 \begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrr}
@@
 \end{tabular*}
-}
-\vspace*{-2.3ex}\hrule width0.48\textwidth
-\vspace*{0.3ex}\hrule width0.48\textwidth
+\par\vspace*{-3.2ex}\noindent\rule{0.48\textwidth}{0.5pt}
+\par\vspace*{-2.4ex}\noindent\rule{0.48\textwidth}{0.5pt}
 \end{table}
@@
 the Klamath Basin.  However, further evaluation of those markers in other
-populations led the removal of
+populations led to the removal of
 twelve loci with apparent null alleles in the winter-run population, two that were monomorphic
@@
 }
-\vspace*{-2.3ex}\hrule width0.48\textwidth
-\vspace*{0.3ex}\hrule width0.48\textwidth
+\par\vspace*{-2.2ex}\noindent\rule{0.48\textwidth}{0.5pt}
+\par\vspace*{-2.4ex}\noindent\rule{0.48\textwidth}{0.5pt}
 \end{table}
@@
 into so-called 'SNPlicons,' which are treated as multiallelic haplotypes in the presence
-of additional variation surrounding the originally assayed SNP. Our work adds to a growing
+of additional variation surrounding the originally assayed SNP\@. Our work adds to a growing
 chorus of recent studies finding that microhaplotype markers can provide additional

### 2025-03-26 001-map-of-samples.Rmd
First pass at cleaning up output_lists

@@
   output_list <- list(
-    map = "results/map_of_samples/map.pdf",
-    cropmap = "results/map_of_samples/map-crop.pdf",
     texmap = "tex/images/map-crop.pdf"
@@
 }
+
+# add the other entries based on the final texmap
+output_list$map = "results/map_of_samples/map.pdf"
+output_list$cropmap = "results/map_of_samples/map-crop.pdf"
 

### 2025-03-26 002-GSI_and_Fst.Rmd
First pass at cleaning up output_lists

@@
   output_list <- list(
+    tex_table = "tex/inputs/samples-table.tex",
+    gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf",
+    ass_table80_tex = "tex/images/ass-table-80-crop.pdf",
+    rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf",
+    pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv",
+    popgen_summ = "tex/inputs/popgen-summary.tex"
+  )
+}
+
+# then add the intermediates to it:
+# then add the intermediate things to it
+output_list <- c(
+  output_list,
   assig_table = "results/GSI_and_Fst/assignment-table-full-baseline.pdf",
@@
   rosa_gsi_table = "results/GSI_and_Fst/rosa-gsi-table.pdf",
-    tex_table = "tex/inputs/samples-table.tex",
   gsi_fst_fig = "results/GSI_and_Fst/gsi_and_fst_fig.pdf",
   gsi_fst_fig_crop = "results/GSI_and_Fst/gsi_and_fst_fig-crop.pdf",
-    gsi_fst_fig_tex = "tex/images/gsi_and_fst_fig-crop.pdf",
   ass_table80_crop = "results/GSI_and_Fst/ass-table-80-crop.pdf",
-    ass_table80_tex = "tex/images/ass-table-80-crop.pdf",
-    rosa_gsi_table_crop = "results/GSI_and_Fst/rosa-gsi-table-crop.pdf",
-    rosa_gsi_table_tex = "tex/images/rosa-gsi-table-crop.pdf",
-    pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv",
-    popgen_summ = "tex/inputs/popgen-summary.tex"
-  )
-}
+  rosa_gsi_table_crop = "results/GSI_and_Fst/rosa-gsi-table-crop.pdf"
+)
 

### 2025-03-26 003-ckmr-sim-evaluation.Rmd
First pass at cleaning up output_lists

@@
   output_list <- list(
+    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
+    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf",
+    frh_comp_tex = "tex/images/frh-comp-figure-crop.pdf"
+  )
+}
+
+# add the intermediates that are not really outputs
+output_list <- c(
+  output_list,
   ckmr_figure = "results/ckmr-sim-eval/fpr-fnr-figure.pdf",
   ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
-    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
   ckmr_comp = "results/ckmr-sim-eval/cmkr-comp-figure.pdf",
   ckmr_comp_crop = "results/ckmr-sim-eval/cmkr-comp-figure-crop.pdf",
-    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf",
   frh_comp = "results/ckmr-sim-eval/frh-comp-figure.pdf",
-    frh_comp_crop = "results/ckmr-sim-eval/frh-comp-figure-crop.pdf",
-    frh_comp_tex = "tex/images/frh-comp-figure-crop.pdf"
-
+  frh_comp_crop = "results/ckmr-sim-eval/frh-comp-figure-crop.pdf"
 )
-}
 

### 2025-03-26 005-plotting-genome-locations-of-markers.Rmd
count up marker types in 005
First pass at cleaning up output_lists

@@
   output_list <- list(
-    pdf = "results/genomic-locations-plot/gen-loc.pdf",
+    pdf = "results/genomic-locations-plot/gen-loc.pdf"
+  )
+}
+
+# add the intermediate results/paths we need
+output_list <- c(
+  output_list,
   pdfcrop = "results/genomic-locations-plot/gen-loc-crop.pdf",
@@
 )
-}
+
 # you can create the necessary output directories like this:
@@
   mutate(chrom_dbl = as.numeric(chrom))
+```
+
+Let's make a quick tally of what we have here.
+```{r}
+amplicons %>%
+  count(LocusOrigin)
 ```

### 2025-03-26 001-map-of-samples.Rmd
Notebooks all rendering.

@@
     toc_float: true
-bibliography: "`r here::here('tex/citation.bib')`"
 ---
@@
 knitr::opts_chunk$set(message = FALSE)
-knitr::opts_knit$set(root.dir = here::here())
 start_time <- Sys.time()
@@
 library(plotly)
+library(maps) # needed for map_data()
+library(mapproj)  # making it explicit for renv
 ```
@@
 
-# Citations
-
-::: {#refs}
-:::
-
 # Session Info

### 2025-03-26 002-GSI_and_Fst.Rmd
Notebooks all rendering.

@@
 knitr::opts_chunk$set(message = FALSE)
-knitr::opts_knit$set(root.dir = here::here())
 start_time <- Sys.time()
-
 ```
@@
 ```
+
+# Session Info
+
+```{r}
+sessioninfo::session_info()
+```
+
+# Running Time
+
+Running the code and rendering this notebook required approximately this much time:
+
+```{r}
+Sys.time() - start_time
+```

### 2025-03-26 005-plotting-genome-locations-of-markers.Rmd
Notebooks all rendering.

@@
   output_list <- list(
-    pdf = "results/genomic-locations-plot/gen-loc.pdf"
+    genome_locations_tex = "tex/images/genomic-locations-plot.pdf"
   )
@@
   output_list,
-  pdfcrop = "results/genomic-locations-plot/gen-loc-crop.pdf",
-  pdftex = "tex/images/genomic-locations-plot.pdf"
+  genome_locations_pdf = "results/genomic-locations-plot/gen-loc.pdf",
+  genome_locations_crop = "results/genomic-locations-plot/gen-loc-crop.pdf"
 )
@@
 ```{r}
-ggsave(final, file = output_list$pdf, width = 7, height = 8.5)
+ggsave(final, file = output_list$genome_locations_pdf, width = 7, height = 8.5)
 ```
@@
 # let's crop that down for use, too
-CALL <- paste("pdfcrop ", output_list$pdf, collapse = " ")
+CALL <- paste("pdfcrop ", output_list$genome_locations_pdf, collapse = " ")
 system(CALL)
 
-file.copy(from = output_list$pdfcrop, to = output_list$pdftex, overwrite = TRUE)
+file.copy(from = output_list$genome_locations_crop, to = output_list$genome_locations_tex, overwrite = TRUE)
 ```
-
-# Citations
-
-::: {#refs}
-:::
 

### 2025-03-26 006-winter-run-vs-all-cv-alle-freqs.Rmd
Notebooks all rendering.

@@
   output_list <- list(
-    mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
-    mh4tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
-    pdf = "results/genomic-locations-plot/gen-loc.pdf",
-    hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
+    winter_v_nonwinter_mh_tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
     hundy_kb4tex = "tex/images/wrap-slide-window.pdf",
-    fuv2 = "results/wrap-discover/follow_up_variants2.rds",
-    wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
     wrap61_4tex = "tex/images/wrap-candi.pdf"
@@
 }
+
+# now add the intermediates
+output_list <- c(
+  output_list,
+  mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
+  hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
+  wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
+  fuv2 = "results/wrap-discover/follow_up_variants2.rds"
+)
+
 # you can create the necessary output directories like this:
@@
 
-file.copy(from = output_list$mh_plot, to = output_list$mh4tex, overwrite = TRUE)
+file.copy(from = output_list$mh_plot, to = output_list$winter_v_nonwinter_mh_tex, overwrite = TRUE)
 ```

### 2025-03-26 007-whoa-on-chinook.Rmd
Notebooks all rendering.

@@
 ---
-title: "Running the California Microhap Baseline through `whoa` and Deciding We Should Drop Two of the Amplicons"
+title: "Running the California Microhap Baseline through `whoa` and Deciding We Could Drop Two of the Amplicons in a Future Revision of the Baseline"
 author: "Eric C. Anderson"
@@
   output_list <- list(
-    zs = "results/whoa/heterozygote-z-scores.pdf"
+    whoa_zs = "results/whoa/heterozygote-z-scores.pdf"
   )
@@
 
-  ggsave(zs, filename = output_list$zs, width = 20, height = 30)
+  ggsave(zs, filename = output_list$whoa_zs, width = 20, height = 30)
 ```
@@
 populations there are a few that look a little wonky, but those are only
-in a few pops.  So, I don't think we really need to toss any others.
+in a few pops.  So, I don't think we really would be called upon to toss any others
+in a future revision of the baseline.
+
+# Session Info
+
+```{r}
+sessioninfo::session_info()
+```
+
+# Running Time
+
+Running the code and rendering this notebook required approximately this much time:
+
+```{r}
+Sys.time() - start_time
+```

### 2025-03-26 008-lfar-gwas-and-markers.Rmd
Notebooks all rendering.

@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(message = FALSE)
+start_time <- Sys.time()
+```
+
 # Introduction
@@
 ```{r}
-start_time <- Sys.time()
-
 # inputs:
@@
 output_list <- list(
-  late_fall_ids = "outputs/007/late-fall-ids.txt",
-  cv_fall_ids = "outputs/007/cv-fall-ids.txt",
-  candidates_for_assays = "outputs/007/candidates_for_assays.csv"
+  lfar_candidates_tex = "tex/images/lfar-candidates.pdf"
 )
+
+# additional outputs not specified in Snakefile:
+output_list <-  c(
+  output_list,
+  late_fall_ids = "results/lfar/late-fall-ids.txt",
+  cv_fall_ids = "results/lfar/cv-fall-ids.txt",
+  candidates_for_assays = "results/lfar/candidates_for_assays.csv"
+)
+
 # you can create the necessary output directories like this:
@@
 
-Accordingly, we will investigate a 5 Mb chunck of Chromosome 34 to:
+Accordingly, we will investigate a 5 Mb chunk of Chromosome 34 to:
 
@@
 
-### Get the IDs.
+### Get the IDs
 ```{r, message=FALSE}
@@
 
-ggsave(g, filename = "tex/images/lfar-candidates.pdf", width = 8, height = 5)
+ggsave(g, filename = output_list$lfar_candidates, width = 8, height = 5)
 ```

### 2025-03-26 003-ckmr-sim-evaluation.Rmd
Notebooks all rendering.
Fix threads iun 003

--- 2025-03-24 003-ckmr-sim-evaluation.Rmd
@@
   output_list <- snakemake@output
+  threads <- as.numeric(snakemake@threads)
 } else {
@@
   output_list <- list(
+    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
+    frh_comp_tex = "tex/images/frh-comp-figure-crop.pdf"
+  )
+  threads <- 8
+}
+
+# add the intermediates that are not really outputs
+output_list <- c(
+  output_list,
   ckmr_figure = "results/ckmr-sim-eval/fpr-fnr-figure.pdf",
   ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
-    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
   ckmr_comp = "results/ckmr-sim-eval/cmkr-comp-figure.pdf",
   ckmr_comp_crop = "results/ckmr-sim-eval/cmkr-comp-figure-crop.pdf",
-    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf",
   frh_comp = "results/ckmr-sim-eval/frh-comp-figure.pdf",
   frh_comp_crop = "results/ckmr-sim-eval/frh-comp-figure-crop.pdf",
-    frh_comp_tex = "tex/images/frh-comp-figure-crop.pdf"
-
+  ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf" # ggsave doesn't save this, so see code comments below
 )
-}
 
@@
 library(gridExtra)
+
+# ensure that Mendel is installed into the CKMRsim package tree:
+install_mendel(Dir = system.file(package = "CKMRsim"))
 
@@
   FUN = function(x) get_FPRs_and_FNRs(x),
-  mc.cores = 8
+  mc.cores = threads
 )
@@
 
-Qsims <- mclapply(input_tibbles$data, get_logl_rats, mc.cores = 8)
+Qsims <- mclapply(input_tibbles$data, get_logl_rats, mc.cores = threads)
 
@@
 # at size 13" high by 10" wide. Then I moved it to where it belongs in the tex directory.
+# at tex/images/cmkr-comp-figure-crop.pdf
 asg
@@
 
+g8
+
 ```
+
+# Session Info
+
+```{r}
+sessioninfo::session_info()
+```
+
+# Running Time
+
+Running the code and rendering this notebook required approximately this much time:
+
+```{r}
+Sys.time() - start_time
+```

### 2025-03-26 scripts/R_input_and_output_lists_to_snakemake_skeleton.sh
Snakemake stuff

@@ -1,0 +1,47 @@
+
+
+# this is a simple script to pull the output_list and
+# input_list names from the Rmd files and create a skeleton
+# Snakefile to render them all via Snakemake.
+
+# Basically, for each Rmd file it makes a snakemake rule
+# to render it using my script in scripts/render-rmd-for-snakemake.R
+
+
+
+make_rule () {
+    filename=$1
+
+    # first make a rule name by stripping the first four characters (numbers and dash)
+    rulename=$(echo $filename | awk '{a = $1; print substr(a, 5)}' | sed 's/.Rmd$//g; s/-/_/g;')
+    htmlname=$(echo $filename | sed 's/.Rmd$/.html/g;')
+
+
+    # now, get the input list elements
+
+    echo rule $rulename:
+    echo "    input:"
+    echo "        rmd=\"$filename\","
+    awk '/input_list <- list\(/ {go = 1; next} go==1 && $1== ")" {exit} go==1 {printf("    %s\n", $0)}' $filename
+    echo "    output:"
+    echo "        html=\"docs/$htmlname\","
+    awk '/output_list <- list\(/ {go = 1; next} go==1 && $1== ")" {exit} go==1 {printf("    %s\n", $0)}' $filename
+    echo "    log:"
+    echo "        \"results/logs/$rulename.log\""
+    echo "    conda:"
+    echo "        \"envs/pandoc.yaml\""
+    echo "    script:"
+    echo "        \"scripts/render-rmd-for-snakemake.R\""
+    echo
+    echo
+    echo
+
+    # then we also put all the output files there for rule all
+    awk '/output_list <- list\(/ {go = 1; next} go==1 && $1== ")" {exit} go==1 {printf("    %s\n", $0)}' $filename >> xxx_temp_outs
+}
+
+
+echo "" > xxx_temp_outs
+for i in 00*.Rmd; do
+    make_rule $i
+done

### 2025-03-26 scripts/render-rmd-for-snakemake.R

@@ -1,0 +1,51 @@
+
+# this is a simple R script lets let me sanely
+# render an Rmd with Snakemake, and put the html output where I
+# want to, and also to have other output files that are specified
+# in the snakefile rule block.  Also, chunks get evaluated in the
+# current working directory.
+
+# Assumes:
+#   - the Rmd input file is named "rmd" in the snakemake input object
+#   - the output html file is named html in the output object
+#   - that renv is being used for R package management
+
+# NOTE:
+#   - This script sets the working directory to be the top level of the
+#     directory, which is typically the RStudio directory if the thing is
+#     also an RStudio project.  To get good behavior when just knitting the
+#     file, you should set the individual Rmd directory to Project Directory
+#     using the drop-down on RStudio's Knit button
+#
+#
+
+# uncomment the two lines below for troubleshooting the
+# snakemake object
+#readr::write_rds(snakemake, file = "snakeobj.rds")
+#quit(save = FALSE)
+
+# this loads the package library managed by renv
+source("renv/activate.R");
+
+# redirect output and messages/errors to the log
+log <- file(snakemake@log[[1]], open="wt")
+sink(log, type = "output")
+sink(log, type = "message")
+
+# snakemake doesn't seem to make the necessary directories
+# for output when running R?  Not sure what is up there...
+# dump <- lapply(snakemake@output, function(f) {
+#   dir.create(dirname(f), recursive = TRUE, showWarnings = FALSE)
+#   print(dirname(f))
+# })
+
+
+outfile <- basename(snakemake@output$html)
+outdir <- dirname(snakemake@output$html)
+
+rmarkdown::render(
+  input = snakemake@input$rmd,
+  output_file = outfile,
+  output_dir = outdir,
+  knit_root_dir = getwd()
+)

### 2025-03-26 README.Rmd
Add README.Rmd

---
title: "california-chinook-microhaps: Data and code repository for \"A multipurpose microhaplotype panel for genetic analysis of California Chinook salmon\""
author: "Eric C. Anderson (also repo maintainer/owner), Anthony J. Clemento, Matthew A. Campbell, Devon E. Pearse, Anne K. Beulke, Cassie Columbus, Ellen Campbell, Neil F. Thompson, John Carlos Garza"
output: 
  github_document:
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

**Last Updated:** `r Sys.Date()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# use this for the the README.md
GPbase <- "https://eriqande.github.io/california-chinook-microhaps/"

# use this for making the README.html that we will hard link to docs/index.html
Dbase <- "./"

# choose the one that you want here depending on if you are on the
# GitHub README or are making an index.html for the docs directory
hdir <- Dbase

make_link <- function(Rmd, base = hdir) {
  nb <- stringr::str_replace(Rmd, "Rmd$", "html")
  sprintf("%s\n\n(_Compiled RMarkdown HTML document on GitHub Pages:_ [%s](%s%s))\n\n", Rmd, nb, base, nb)
}
```

This repository includes code, data, and some intermediate results to
reproduce the results in Anderson et al. ("A multipurpose microhaplotype
panel for genetic analysis of California Chinook salmon").
You can get the whole thing by cloning or downloading the repo from 
[https://github.com/eriqande/california-chinook-microhaps](https://github.com/eriqande/california-chinook-microhaps).

If you are viewing this as the README of a GitHub repository, note that
you can read it in a somewhat friendlier format on GitHub pages at:
[https://eriqande.github.io/california-chinook-microhaps/](https://eriqande.github.io/california-chinook-microhaps/)

Most of this repository concerns analyses done using R and that are fully contained
within this repo and intended to run automatically to create the figures and
tables used in the paper.

In most cases, products from cluster computing are placed in this repository within
the `stored_results` directory.  This means that all the RMarkdown documents can be
evaluated successfully without first running the cluster-based analyses.

Analyses requiring cluster computing are documented, but do not run
automatically.

All the RMarkdown documents can be evaluated en masse using the included
Snakefile which also shows the different output created by each notebook.
Or, if desired, they can be run one at a time in, for example,
RStudio.  Each RMarkdown document creates an HTML file in the docs/
directory that is served up publicly on GitHub pages, so that the
analyses and codes can be read in that context.

On my Mac, it takes about 10 minutes to run all the notebooks, with
002 and 003 being the ones that are most computationally intensive.

# Preliminaries and Dependencies

To the extent possible, we have tried to provide automatic installs of R package
dependencies and some public data sets.  Accordingly, you should be connected
to the internet when you run any of this.

Running the code here requires the R programming language. It was built most
recently using R 4.4.3.  Presumably any R 4.4.x version will work.  Required R
packages (and their corresponding versions) are managed by `renv`.  If you use
RStudio then you should be prompted to install all the packages upon
opening the project.

The R code creates publication-ready figures and makes heavy use of the `pdfcrop`
utility.  This utility should be available on your path.  It comes with most TeX
distributions, I believe.  You will need LaTeX to typeset the paper if desired.
Alternatively, you can obtain `pdfcrop` from: 
[https://ctan.org/pkg/pdfcrop?lang=en](https://ctan.org/pkg/pdfcrop?lang=en).

# RMarkdown Documents

The following RMarkdown documents should be evaluated in order.
The script `render-numbered-Rmds.R` will do that (except for `002-allele-frequencies.Rmd`)
when run in the top level
of this repository.
Some RMarkdown documents rely on
outputs from previous ones.  Some of these RMarkdown documents include
calls to Unix utilities, so might not run on non-Unix or non-Linux architectures.

Outputs (figures, tables, R data objects, etc) from each RMarkdown document
are written to the `outputs/XXX` directories.  Intermediate files
written during evaluation of each RMarkdown document are written
to the `intermediates/XXX` directories. To facilitate working between
the cluster and a desktop/laptop, some outputs are written to
the `stored_results/XXX` directories which are version controlled and
included in this repo.

Thumbnails of the figures and tables generated by each RMarkdown document appear
below.  Additionally, at the top of each section is a link to the
compiled (HTML-version) of the RMarkdown document on GitHub Pages.

## `r make_link("001-align-coho-genome.Rmd")`

The products of this analysis go into creation of some of the trees (`009`) and also
in the designation of the alleles as ancestral or not in some of the haplotype rasters (`006`).

## `r make_link("002-allele-frequencies.Rmd")`

Calculation of allele frequency differences between early- and late-running forms
across the genome, and around GREB1L, and the generation of plots.

[![](./docs/images_and_thumbs/allele-freqs-figure_thumb.png)][alle_freqs]

## `r make_link("003-extract-johnson-creek-variants.Rmd")`

Using the previously assembled Chinook salmon genome from @narumGenomicVariationUnderlying2018
to ascertain variation from that Upper Columbia River summer/spring-run fish
around the GREB1L region.  The products of this analysis go into some later trees and
haplotype rasters.

## `r make_link("004-prepare-haplotypes.Rmd")`

Using BEAGLE to impute genotypes and infer haplotypes in 5 Mb around GREB1L.
The outputs of this analysis are used for a number of downstream analyses.

## `r make_link("005-annotating-variants-near-greb1l.Rmd")`

Using snpEff to annotate variants and then analyzing a few of them.  This
provides data that goes into Table S1 ("Two near-perfectly associated, non-synonymous variants in the GREB1L gene"),
and is also used later for the haplotype raster plots.

[![](./docs/images_and_thumbs/table-s2-2-candidate-snps_thumb.png)][snpeff_res]

## `r make_link("006-haplo-raster-plots.Rmd")`

Compiling some auxiliary information (gene/exon inclusion, etc.) and
then making haplotype raster plots using functions from the
'ecaRbioinf' package.

[![](./docs/images_and_thumbs/moderate-zoom-haplo-raster-with-JC-fish_thumb.png)][haplo_raster]
[![](./docs/images_and_thumbs/read-depth-raster_thumb.png)][read_depth_raster]

### 2025-03-26 SnakemakeSubDirs/cga-late-fall-bam-assoc/scripts/manhattan-plot.R
Add snakemmake Subdirs

@@ -1,0 +1,118 @@
+if(exists("snakemake")) {
+  
+  # redirect output and messages/errors to the log
+  log <- file(snakemake@log[[1]], open="wt")
+  sink(log, type = "output")
+  sink(log, type = "message")
+  
+  sgfile <- snakemake@input$sg
+  lf_v_frhf_lrts <- snakemake@input$lf_v_frhf_lrts
+  lf_v_sanjo_lrts <- snakemake@input$lf_v_sanjo_lrts
+  outfile <- snakemake@output$mh_plot
+  neg10 <- as.numeric(snakemake@params$neg_log10_cutoff)
+}
+
+require(tidyverse)
+require(vroom)
+
+
+
+
+
+#' function to process the data for a manhattan plot
+#' @param sgfile the scaffold groups file
+#' @param lrt_files vector of paths to lrt0.gz files (one for each chrom, etc.)
+#' @param outfile path to the output file.  Extension (.pdf/.jpg/etc) determines type
+#' @param neg_log10_cutoff  value below which you want to filter out points to make the
+#' plot smaller.
+mh_plot_data <- function(
+    sgfile = "inputs/scaffold_groups.tsv",
+    lrt_files = c(list.files(path="results/doAsso/lf_v_frhf", pattern = ".*lrt0.gz", full.names = TRUE),
+                  list.files(path="results/doAsso_unplaced/lf_v_frhf", pattern = ".*lrt0.gz", full.names = TRUE)),
+    neg_log10_cutoff=3
+) {
+  # first, get the cumulative position of the start of each chromosome
+  sg <- read_tsv(sgfile) %>%
+    mutate(
+      cstart = 1 + lag(cumsum(stop), default = 0),
+      cend = cstart + stop - 1
+    )
+  
+  # read in the LRTs and process them
+  r <- vroom(lrt_files) %>%
+    filter(LRT != -999) %>% # remove missing ones
+    mutate(neg_log10_p = -log10(pchisq(LRT, df = 1, lower.tail = FALSE))) %>%
+    filter(neg_log10_p > neg_log10_cutoff) %>% 
+    left_join(
+      .,
+      sg %>% select(chrom, mh_label, angsd_chrom, cstart),
+      by = c("Chromosome" = "chrom")
+    ) %>%
+    mutate(xpos = Position + cstart - 1) %>%
[…]

### 2025-03-26 SnakemakeSubDirs/cga-late-fall-bam-assoc/scripts/prep/make-scaff-beds.R

@@ -1,0 +1,20 @@
+scaffs <- read_tsv("inputs/scaffold_groups.tsv") %>% 
+  mutate(num = as.integer(str_replace_all(id, "[a-z_]*", ""))) %>% 
+  filter(num >= 35) %>%
+  mutate(grp = rep(seq(1:28), each = 500)[1:n()] ) %>%
+  mutate(id = sprintf("unplaced-%03d", grp)) %>%
+  mutate(rstr = str_c(chrom, ":", start, "-", stop))
+
+
+dir.create("inputs/region_files")
+
+scaffs %>%
+  select(id, rstr) %>%
+  group_by(id) %>%
+  nest() %>%
+  ungroup() %>%
+  mutate( what = walk2(
+    .x = id,
+    .y = data,
+    .f = function(x, y) write_tsv(y, col_names = FALSE, file = str_c("inputs/region_files/", x, ".txt")))
+)

### 2025-03-26 README.Rmd
Thumbnails and links in README

@@
 # GitHub README or are making an index.html for the docs directory
-hdir <- Dbase
+hdir <- GPbase
 
@@
 
+# Repo Contents
+
+The most important data in this repository are in:
+
+- `data/SWFSC-chinook-reference-baseline.csv.gz`: The actual reference data set---genotypes
+of all the fish in the baseline.
+- `data/baseline_repository_meta_complete.csv.gz`: Metadata from all those reference
+data set fish
+
+The rest of the repo is involved in analyses of those data sets.  
+Here is a an annotated graphical overview of the contents of the
+repo made using `tree`.   Some file-series are omitted and replaced with
+ellipses.
+
+```sh
+.
+├── .Rprofile                                             # For initiating renv
+├── .gitignore                                            # standard git ignore file
+├── 001-map-of-samples.Rmd                                # 8 Rmarkdown documents (where the analyses happen)
+├── 002-GSI_and_Fst.Rmd                                  
+├── 003-ckmr-sim-evaluation.Rmd
+├── 004-lfar-and-wrap-allele-freqs.Rmd
+├── 005-plotting-genome-locations-of-markers.Rmd
+├── 006-winter-run-vs-all-cv-alle-freqs.Rmd
+├── 007-whoa-on-chinook.Rmd
+├── 008-lfar-gwas-and-markers.Rmd
+├── R                                                     # Directory of R code that gets sourced and reused in the notebooks
+│   ├── colors.R
+│   ├── get_top_assignments.R
+│   ├── manhattan_plot_funcs.R
+│   ├── month_sample_sparkline.R
+│   ├── plot_self_assigment_table.R
+│   ├── shift_legend.R
+│   ├── table-as-figure.R
+│   └── taf-prep.R
+├── README.Rmd                                            # Source for README and docs/index
+├── README.md                                             # Produced from README.Rmd
+├── Snakefile                                             # Snakefile to run all R notebooks
+├── SnakemakeSubDirs                                      # Copies of other repos used for cluster-based analyses
+│   └── cga-late-fall-bam-assoc                           # LFAR GWAS using angsd code uses Snakemake
+│       ├── .gitignore
+│       ├── README.Rmd
+│       ├── Snakefile                                     # This has all the info on how to run the LFAR/GWAS analysis
+│       ├── cga-late-fall-bam-assoc.Rproj
+│       ├── envs
+│       │   ├── angsd.yaml
+│       │   └── samtools.yaml
+│       ├── inputs
+│       │   ├── lf_v_both.bamlist
+│       │   ├── lf_v_both.ybin
+│       │   ├── lf_v_frhf.bamlist
+│       │   ├── lf_v_frhf.ybin
+│       │   ├── lf_v_sanjo.bamlist
+│       │   ├── lf_v_sanjo.ybin
+│       │   ├── region_files
+│       │   │   ├── unplaced-001.txt
+│       │   │   ├── unplaced-002.txt
+│       │   │   ├── ...
+│       │   │   ├── unplaced-027.txt
+│       │   │   └── unplaced-028.txt
+│       │   └── scaffold_groups.tsv
+│       ├── scripts
+│       │   ├── manhattan-plot.R                          # Functions for making manhattan plots
+│       │   └── prep
+│       │       └── make-scaff-beds.R
+│       └── sedna                                         # cluster profile for Snakemake
+│           ├── config.yaml
+│           └── status-sacct.sh
+├── california-chinook-microhaps.Rproj                    # Rstudio project file
+├── data                                                  # Main data files
+│   ├── SWFSC-chinook-reference-baseline.csv.gz           # The actual reference data set
+│   └── baseline_repository_meta_complete.csv.gz          # Meta data for all fish in reference data set
+├── development                                           # Folder of various things that we ended up not using
+│   ├── genepop-conversion.Rmd
+│   └── genepop-conversion.nb.html
+├── docs                                                  # Directory of html files served up on GitHub Pages
+│   ├── 001-map-of-samples.html
+│   ├── 002-GSI_and_Fst.html
+│   ├── 003-ckmr-sim-evaluation.html
+│   ├── 004-lfar-and-wrap-allele-freqs.html
+│   ├── 005-plotting-genome-locations-of-markers.html
+│   ├── 006-winter-run-vs-all-cv-alle-freqs.html
+│   ├── 007-whoa-on-chinook.html
+│   └── 008-lfar-gwas-and-markers.html
+├── envs                                                  # YAMLS specifying conda environments for snakemake rules
+│   └── pandoc.yaml
+├── extras                                                # Scripts and code used in cleaning/standardizing the baseline prior to distribution 
+│   ├── .gitignore
+│   ├── Applying-New-Names-To-Baseline
+│   │   ├── .gitignore
+│   │   ├── applying-new-names-to-baseline.Rmd
+│   │   ├── subRoSA_baseline_rubias_JUN2024_w_mixture_complete-rosa_lfar_vgll3six6_added.csv
+│   │   └── subRoSA_baseline_rubias_with_rosa_wrap_lfar_vgll3six6_and_mixture_28JUN2024.csv
+│   ├── Renaming-amplicons
+│   │   ├── .gitignore
+│   │   ├── Amplicon-renaming.Rmd
+│   │   ├── ChinookFullPanelLociTable_May2024Updated.xlsx
+│   │   ├── Otsh_v2.0_chroms.tsv
+│   │   └── add-species-id.Rmd
+│   └── Reordering-forward-and-reverse-primers
+│       ├── OLD-Calif-Chinook-Amplicon-Panel-Information.csv
+│       └── primer-pair-reorder.R
+├── inputs                                                 # Small inputs used in most R notebooks
+│   ├── Calif-Chinook-Amplicon-Panel-Information.csv
+│   ├── Otsh_v1.0_chrom_lengths.tsv
+│   ├── Otsh_v2.0_chrom_lengths.tsv
+│   ├── Rivers_OR
+│   │   ├── Rivers_OR.dbf
+│   │   ├── Rivers_OR.prj
+│   │   ├── Rivers_OR.sbn
+│   │   ├── Rivers_OR.sbx
+│   │   ├── Rivers_OR.shp
+│   │   └── Rivers_OR.shx
+│   ├── map-notations.tsv
+│   ├── reference-collection-names.csv
+│   └── wrap
+│       └── chrom_lengths.txt
+├── renv                                                   # Folder of renv R package management stuff 
+│   ├── .gitignore
+│   ├── activate.R
+│   └── settings.json
+├── renv.lock                                              # renv R package management lockfile
+├── scripts                                                # Scripts used when running Snakemake or prepping the Snakefile
+│   ├── R_input_and_output_lists_to_snakemake_skeleton.sh
+│   └── render-rmd-for-snakemake.R
+├── stored_results                                         # Results produced on the HPCC stored here for further analysis
+│   ├── 5Mb_Chr34-cv-fall.mafs.gz
+│   ├── 5Mb_Chr34-late-fall.mafs.gz
+│   ├── lfar-gwas
+│   │   ├── lf_v_frhf-NC_037130.1.lrt0.gz
+│   │   └── lf_v_sanjo-NC_037130.1.lrt0.gz
+│   ├── stored_p_vals.rds
+│   ├── thompson2020-vcf
+│   │   ├── ots34-5Mb-snpEff.vcf.gz
+│   │   └── ots34-5Mb-snpEff.vcf.gz.csi
+│   └── win-non-win-abs-diffs-gt0.5.rds
+└── tex                                                    # Directory of files for typesetting the manuscript
+    ├── .gitignore
+    ├── GitLatexDiffIt.sh
+    ├── abstract.tex
+    ├── author-title-etc.tex
+    ├── body-text-of-paper.tex
+    ├── citation.bib
+    ├── data-accessibility.tex
+    ├── eca_molecolres.sty
+    ├── images                                             # Figures/images included in manuscript.  Many (but not all) are produced when running the notebooks
+    │   ├── Fall-run-ball-crop.pdf
+    │   ├── Late-fall-run-ball-crop.pdf
+    │   ├── Spring-run-ball-crop.pdf
+    │   ├── Winter-run-ball-crop.pdf
+    │   ├── ass-table-80-crop.pdf
+    │   ├── banner.pdf
+    │   ├── ckmr-comp-figure-crop.pdf
+    │   ├── clumpak-crop.pdf
+    │   ├── fpr-fnr-figure-crop.pdf
+    │   ├── frh-comp-figure-crop.pdf
+    │   ├── genomic-locations-plot.pdf
+    │   ├── gsi_and_fst_fig-crop.pdf
+    │   ├── lfar-assoc-faceted-min-0.png
+    │   ├── lfar-candidates.pdf
+    │   ├── major-modes-crop.pdf
+    │   ├── map-crop.pdf
+    │   ├── minor-modes-crop.pdf
+    │   ├── months-BCF_Butte-Creek.pdf
+    │   ├── months-BCS_Butte-Creek.pdf
+    │   ├── months-BCkF_Blue-Creek.pdf
+    │   ├── months-CHLF_Coleman-National-Fish-Hatchery.pdf
+    │   ├── months-CRHS_Cole-Rivers-Hatchery.pdf
+    │   ├── months-CoRHS_Cole-Rivers-Hatchery.pdf
+    │   ├── months-ERF_Van-Arsdale-Fisheries-Station.pdf
+    │   ├── months-FRHF_Feather-River-Hatchery.pdf
+    │   ├── months-FRHS_Feather-River-Hatchery.pdf
+    │   ├── months-IGHF_Iron-Gate-Hatchery.pdf
+    │   ├── months-MDF_Deer-Creek.pdf
+    │   ├── months-MDF_Mill-Creek.pdf
+    │   ├── months-MDS_Deer-Creek.pdf
+    │   ├── months-MDS_Mill-Creek.pdf
+    │   ├── months-RRF_Warm-Springs-Hatchery.pdf
+    │   ├── months-SJRF_San-Joaquin-River.pdf
+    │   ├── months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf
+    │   ├── months-SmRF_Smith-River.pdf
+    │   ├── months-TRF_Trinity-River-Hatchery.pdf
+    │   ├── months-TRS_Trinity-River-Hatchery.pdf
+    │   ├── num-alle-barplot.pdf
+    │   ├── rosa-gsi-table-crop.pdf
+    │   ├── winter-v-non-winter-mh-plot_nind_ge10.pdf
+    │   ├── wrap-candi.pdf
+    │   └── wrap-slide-window.pdf
+    ├── inputs                                              # LaTeX input for tables in manuscripts. Many (but not all) are produced when running the notebooks
+    │   ├── amplicon-summary.tex
+    │   ├── ascertainment-samples-supp-table.tex
+    │   ├── lfar-freqs.tex
+    │   ├── popgen-summary.tex
+    │   ├── samples-table.tex
+    │   └── wrap-freqs.tex
+    ├── main-one-col.tex                                   # Main TeX document for one-column format typesetting (for submission)
+    ├── main.bbl
+    ├── main.tex                                           # Main TeX document for two-column format typesetting
+    ├── men.bst
+    ├── newcomms.tex
+    ├── responses-to-referees
+    │   ├── .gitignore
+    │   └── responses-to-reviewers-1.Rmd
+    ├── supp_data                                          # Supplemental data
+    │   ├── Supp-Data-1-Amplicon-info.csv
+    │   └── Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv
+    ├── supplement.bbl
+    ├── supplement.tex                                    # Main TeX document for typesetting the supplement
+    ├── supplemental-figures.tex
+    ├── supplemental-tables.tex
+    └── supplemental-text.tex
+
+```
+
 # RMarkdown Documents
 
-The following RMarkdown documents should be evaluated in order.
-The script `render-numbered-Rmds.R` will do that (except for `002-allele-frequencies.Rmd`)
-when run in the top level
-of this repository.
-Some RMarkdown documents rely on
-outputs from previous ones.  Some of these RMarkdown documents include
-calls to Unix utilities, so might not run on non-Unix or non-Linux architectures.
+The following RMarkdown documents can be evaluated in any order.
 
-Outputs (figures, tables, R data objects, etc) from each RMarkdown document
-are written to the `outputs/XXX` directories.  Intermediate files
+Outputs (figures, tables, etc) from each RMarkdown document
+are written to the `tex/images` or `tex/inputs` directories.  Intermediate files
 written during evaluation of each RMarkdown document are written
-to the `intermediates/XXX` directories. To facilitate working between
-the cluster and a desktop/laptop, some outputs are written to
-the `stored_results/XXX` directories which are version controlled and
-included in this repo.
+into the `results` directories.
 
 Thumbnails of the figures and tables generated by each RMarkdown document appear
-below.  Additionally, at the top of each section is a link to the
+below each notebook that generates them.
+
+Click the thumbnail to see a full size PDF.
+
+Additionally, at the top of each section is a link to the
 compiled (HTML-version) of the RMarkdown document on GitHub Pages.
 
-## `r make_link("001-align-coho-genome.Rmd")`
+## `r make_link("001-map-of-samples.Rmd")`
 
-The products of this analysis go into creation of some of the trees (`009`) and also
-in the designation of the alleles as ancestral or not in some of the haplotype rasters (`006`).
+Makes a map of where the samples were drawn.
 
-## `r make_link("002-allele-frequencies.Rmd")`
+[![](./docs/images_and_thumbs/figure1_thumb.png)][fig1]
 
-Calculation of allele frequency differences between early- and late-running forms
-across the genome, and around GREB1L, and the generation of plots.
+## `r make_link("002-GSI_and_Fst.Rmd")`
 
-[![](./docs/images_and_thumbs/allele-freqs-figure_thumb.png)][alle_freqs]
+Summaries of the samples and population genetic statistics,
+as well as GSI power analysis and Fst calculation.
 
-## `r make_link("003-extract-johnson-creek-variants.Rmd")`
+[![](./docs/images_and_thumbs/table1_thumb.png)][tab1]
+[![](./docs/images_and_thumbs/figure2_thumb.png)][fig2]
+[![](./docs/images_and_thumbs/figureS6_thumb.png)][figS6]
+[![](./docs/images_and_thumbs/figureS7_thumb.png)][figS7]
 
-Using the previously assembled Chinook salmon genome from @narumGenomicVariationUnderlying2018
-to ascertain variation from that Upper Columbia River summer/spring-run fish
-around the GREB1L region.  The products of this analysis go into some later trees and
-haplotype rasters.
+[![](./docs/images_and_thumbs/supp_data2_thumb.png)][supp_data2]
+[![](./docs/images_and_thumbs/table4_thumb.png)][tab4]
 
-## `r make_link("004-prepare-haplotypes.Rmd")`
+## `r make_link("003-ckmr-sim-evaluation.Rmd")`
 
-Using BEAGLE to impute genotypes and infer haplotypes in 5 Mb around GREB1L.
-The outputs of this analysis are used for a number of downstream analyses.
+Using CKMRsim to estimate false positive rates and false negative rates
+of relationship inference.  Also comparisons of power for inferring relatives
+when using microhaplotypes versus single SNPs from each amplicon.
 
-## `r make_link("005-annotating-variants-near-greb1l.Rmd")`
+[![](./docs/images_and_thumbs/figure4_thumb.png)][fig4]
+[![](./docs/images_and_thumbs/figureS8_thumb.png)][figS8]
+[![](./docs/images_and_thumbs/figure5_thumb.png)][fig5]
 
-Using snpEff to annotate variants and then analyzing a few of them.  This
-provides data that goes into Table S1 ("Two near-perfectly associated, non-synonymous variants in the GREB1L gene"),
-and is also used later for the haplotype raster plots.
+## `r make_link("004-lfar-and-wrap-allele-freqs.Rmd")`
 
-[![](./docs/images_and_thumbs/table-s2-2-candidate-snps_thumb.png)][snpeff_res]
+Assessing allele frequency differences between groups of fish at the
+LFAR and WRAP amplicons.
 
-## `r make_link("006-haplo-raster-plots.Rmd")`
+[![](./docs/images_and_thumbs/table3_thumb.png)][tab3]
+[![](./docs/images_and_thumbs/tableS2_thumb.png)][tabS2]
 
-Compiling some auxiliary information (gene/exon inclusion, etc.) and
-then making haplotype raster plots using functions from the
-'ecaRbioinf' package.
+## `r make_link("005-plotting-genome-locations-of-markers.Rmd")`
 
-[![](./docs/images_and_thumbs/moderate-zoom-haplo-raster-with-JC-fish_thumb.png)][haplo_raster]
-[![](./docs/images_and_thumbs/read-depth-raster_thumb.png)][read_depth_raster]
+Plotting the positions of different amplicons in the genome.
+
+[![](./docs/images_and_thumbs/figureS1_thumb.png)][figS1]
+
+## `r make_link("006-winter-run-vs-all-cv-alle-freqs.Rmd")`
+
+Analyzing the allele frequencies from WGS data to identify regions
+that show large allele frequency differences between winter run
+and everyone else.  (Finding the WRAP markers...)
+
+[![](./docs/images_and_thumbs/figureS9_thumb.png)][figS9]
+[![](./docs/images_and_thumbs/figureS10_thumb.png)][figS10]
+[![](./docs/images_and_thumbs/figureS11_thumb.png)][figS11]
+
+## `r make_link("007-whoa-on-chinook.Rmd")`
+
+This was not included in the paper, but is an interesting analysis of
+the markers suggesting that two of them might have some genotyping
+issues and would be good candidates for removal in a further update
+to the baseline.
+
+[![](./docs/images_and_thumbs/whoa_thumb.png)][whoa]
+
+## `r make_link("008-lfar-gwas-and-markers.Rmd")`
+
+Analyzing the GWAS and allele frequency data from the LFAR search
+
+[![](./docs/images_and_thumbs/figureS3_thumb.png)][figS3]
+
+[fig1]: `r hdir`images_and_thumbs/figure1.pdf
+[tab1]: `r hdir`images_and_thumbs/table1.pdf
+[fig2]: `r hdir`images_and_thumbs/figure2.pdf
+[figS6]: `r hdir`images_and_thumbs/figureS6.pdf
+[figS7]: `r hdir`images_and_thumbs/figureS7.pdf
+[supp_data2]: `r hdir`images_and_thumbs/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv
+[tab4]: `r hdir`images_and_thumbs/table4.pdf
+[fig4]: `r hdir`images_and_thumbs/figure4.pdf
+[figS8]: `r hdir`images_and_thumbs/figureS8.pdf
+[fig5]: `r hdir`images_and_thumbs/figure5.pdf
+[tab3]: `r hdir`images_and_thumbs/table3.pdf
+[tabS2]: `r hdir`images_and_thumbs/tableS2.pdf
+[figS1]: `r hdir`images_and_thumbs/figureS1.pdf
+[figS9]: `r hdir`images_and_thumbs/figureS9.pdf
+[figS10]: `r hdir`images_and_thumbs/figureS10.pdf
+[figS11]: `r hdir`images_and_thumbs/figureS11.pdf
+[whoa]: `r hdir`images_and_thumbs/whoa.pdf
+[figS3]: `r hdir`images_and_thumbs/figureS3.pdf

### 2025-03-27 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Done

@@
 
-- **The online data and scripts (thank you for providing these) needs a proper readme file describing what the folders and files relate to.**
+- The online data and scripts (thank you for providing these) needs a proper readme file describing what the folders and files relate to.
 
-    - **Gonna make a README and store the htmls in the docs/ folder like
-    we did for the Science paper.**
+    - A comprehensive README has been added.  The repo and its README can be viewed at:
+    [https://github.com/eriqande/california-chinook-microhaps/](https://github.com/eriqande/california-chinook-microhaps/).  This will be archived at Zenodo upon acceptance.

### 2025-03-27 002-GSI_and_Fst.Rmd
Done

@@
     pop_gen_by_loc_coll = "tex/supp_data/Supp-Data-2-pop-gen-summaries-by-locus-and-collection.csv",
-    popgen_summ = "tex/inputs/popgen-summary.tex"
+    popgen_summ = "tex/inputs/popgen-summary.tex",
+    num_alle_barplot = "tex/images/num-alle-barplot.pdf"
   )
@@
 
-ggsave(gbp, filename = "tex/images/num-alle-barplot.pdf", width = 4, height = 3)
+ggsave(gbp, filename = output_list$num_alle_barplot, width = 4, height = 3)
 

### 2025-03-27 008-lfar-gwas-and-markers.Rmd
Done

@@
 have been inserted here for ease of archiving.
+
+To run that analysis it is necessary to first map the sequences of the salmon
+(publicly archived at [https://www.ncbi.nlm.nih.gov/sra/PRJNA667732](https://www.ncbi.nlm.nih.gov/sra/PRJNA667732)) to the RefSeq version of Otsh_v1.0, and name the resulting BAM files as indicated in
+the bamlists.  If you would like to reproduce the analysis, let us know, as we
+can provide those BAMS.
+
+The snakefile can can be launched something like this:
+```sh
+snakemake -np --use-envmodules --use-conda  --cores 24
+```
 

### 2025-03-27 009-structure.Rmd
Done

--- 2025-03-26 008-lfar-gwas-and-markers.Rmd
@@
 ---
-title: "Late Fall Associated Region: Genome Wide Association Study and Marker Selection"
+title: "Running Structure"
 author: "Eric C. Anderson"
@@
 
-Here we do a GWAS to find a genomic region that is substantially differentiated between
-the late-fall and fall runs in the Central Valley, and then we identify several
-candidate sites to target for amplicon sequencing.
+Anthony ran these analyses by using the reference data set and running them
+in the program Structure using the our lab's [slg_pipe](https://github.com/eriqande/slg_pipe)
+framework for parallelization.
 
-# Input and Output Paths
-
-```{r}
-# inputs:
-input_list <- list(
-  chr34_late_fall_mafs = "stored_results/5Mb_Chr34-late-fall.mafs.gz",
-  chr34_cv_fall_mafs = "stored_results/5Mb_Chr34-cv-fall.mafs.gz",
-  gwas_lf_v_sanjo = "stored_results/lfar-gwas/lf_v_sanjo-NC_037130.1.lrt0.gz",
-  gwas_lf_v_frhf = "stored_results/lfar-gwas/lf_v_frhf-NC_037130.1.lrt0.gz",
-  ots34_snpeff = "stored_results/thompson2020-vcf/ots34-5Mb-snpEff.vcf.gz",
-  pvals = "stored_results/stored_p_vals.rds"
-)
-# outputs:
-output_list <- list(
-  lfar_candidates_tex = "tex/images/lfar-candidates.pdf"
-)
-
-# additional outputs not specified in Snakefile:
-output_list <-  c(
-  output_list,
-  late_fall_ids = "results/lfar/late-fall-ids.txt",
-  cv_fall_ids = "results/lfar/cv-fall-ids.txt",
-  candidates_for_assays = "results/lfar/candidates_for_assays.csv"
-)
-
-# you can create the necessary output directories like this:
-dump <- lapply(output_list, function(x)
-  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
-)
-```
-
-# GWAS
-
-This is all executed with a Snakefile and associated files that reside in the
-repository at: `SnakemakeSubDirs/cga-late-fall-bam-assoc`.  The contents of that
-directory were originally part of another GitHub repo---simply `eriqande/cga-late-fall-bam-assoc`---but
-have been inserted here for ease of archiving.
-
-The figure created by that Snakefile is included in the paper.  We also have saved
-the likelihood ratios for the various markers on Chromosome 34 (NC_037130.1)
-for both the Late-fall vs San Joaquin fall
-and the Late-Fall vs Feather River fall comparisons in `stored_results` so that we can have
-those when visualizing allele frequencies.
-
-Additionally, information for all markers with a p-value > 3 in both the SanJo-vs-LateFall and 
-the FeatherRiverHatcheryFall-vs-LateFall comparisons was saved in `stored_results/stored_p_vals.rds`
-for use now.
-
-## Investigate GWAS pvalues
-
-```{r}
-library(tidyverse)
-library(vcfR)
-
-# read in the pvals and label the comparisons FRFH and SanJo and then pivot that 
-# so we can compare values from the two analyses at each marker
-pvals <- read_rds(input_list$pvals) %>%
-  mutate(comparison = ifelse(comparison == "Late Fall versus Feather River Hatchery Fall", "FRHF", "SanJo")) %>%
-  select(-(Major:Frequency), -(mh_label:xpos), -cgroup) %>%
-  pivot_wider(
-    names_from = comparison,
-    values_from = c(LRT, neg_log10_p)
-  )
-
-# here are the total number of SNPs in each comparison:
-FRH_snps <- 13477607
-SJ_snps <- 12752546
-
-# now, we can look for those that have have log10 pvals in both of the comparisons.
-# maybe just plot them first:
-ggplot(pvals, aes(x = neg_log10_p_FRHF, y = neg_log10_p_SanJo)) +
-  geom_point() +
-  geom_hline(yintercept = -log10(5e-8), linetype = "dashed") +
-  geom_vline(xintercept = -log10(5e-8), linetype = "dashed")
-```
-
-So, that is only a single SNP in both:
-```{r}
-pvals %>%
-  filter(neg_log10_p_FRHF > -log10(5e-8), neg_log10_p_SanJo > -log10(5e-8))
-```
-
-Now, let's consider a Bonferroni correction:
-```{r}
-pv2 <- pvals %>%
-  mutate(
-    bonFRHF = (10 ^ -neg_log10_p_FRHF) * FRH_snps,
-    bonSJ = (10 ^ -neg_log10_p_SanJo) * SJ_snps,
-  ) %>%
-  arrange(bonFRHF)
-
-pv2
-```
-
-So, that is only a single SNP that would be "significant" under a Bonferroni correction,
-in the FRFH comparison.
-
-Likewise only a single SNP is "significant" at a Bonferroni corrected 0.05 level
-in the SJ comparison, hich we show by sorting the above table on the SanJo
-values:
-```{r}
-pv2 %>%
-  arrange(bonSJ)
-```
-
-# Marker Selection
-
-This was originally done in
-`/Users/eriq/Documents/git-repos/further-chinook-wgs/101-design-late-fall-assays.html`
-and it has been transferred here, now.
-
-Mac originally ran the association analysis between Late Fall and the two fall run groups in
-the Central Valley (Feather and San Joaquin), finding significant differences
-on Chromosome 34 (NC_037130.1).   We have seen the same thing with the GWAS just described.
-
-Accordingly, we will investigate a 5 Mb chunk of Chromosome 34 to:
-
-1. confirm those markers and explore patterns of differentiation in the region
-2. See if the differentiation corresponds to annotated variants of large effect
-3. identify candidate markers that we should be able to type in many more individuals.
-We will provide sequences flanking these candidates so that
-Anthony can design primers for them.
-
-## Overview of allele frequency differences in the region
-
-We will simply compare the allele frequencies estimated in the CV fall-run vs the late-fall
-run using ANGSD.  I will do this in a 5 Mb chunk for a little context.
-
-### Get the IDs
-```{r, message=FALSE}
-
-# grab the meta data from the Science paper off of GitHub 
-nooks <- read_csv("https://raw.githubusercontent.com/eriqande/thompson-et-al-2020-chinook-salmon-migration-timing/master/data/wgs-chinook-samples.csv")
-
-# write out the names of the late-fall fish
-nooks %>%
-  filter(Population == "Coleman Hatchery Late Fall") %>%
-  pull(vcf_name) %>%
-  cat(., sep = "\n", file = output_list$late_fall_ids)
-
-# same for the fall-run fish in the CV:
-nooks %>%
-  filter(Population %in% c("Feather River Hatchery Fall", "San Joaquin River Fall")) %>%
-  pull(vcf_name) %>%
-  cat(., sep = "\n", file = output_list$cv_fall_ids)
-```
-
-### Extract 5 Mb and the two sets of indivs
-
-We show the code but don't actually do it, here.
-
-```{sh, eval=FALSE}
-bcftools view -S outputs/007/late-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/late-fall.vcf.gz
-bcftools view -S outputs/007/cv-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/cv-fall.vcf.gz
-```
-
-### Run ANGSD to estimate allele frequencies
-
-Here is the code used.  We don't do it here, because in and update from angsd 0.920
-to 0.938 it broke.  But I have the estimated allele freqs in stored results.
-
-```{sh, eval=FALSE}
-cd outputs/007
-angsd -vcf-gl late-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 16 -domaf 3 -out late-fall 2> late-fall.redirect.stderr
-angsd -vcf-gl cv-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 32 -domaf 3 -out cv-fall 2> cv-fall.redirect.stderr
-```
-
-### Read the estimated allele freqs in and analyse them
-
-```{r, message=FALSE}
-af1 <- list(
-  late_fall = read_tsv(input_list$chr34_late_fall_mafs),
-  cv_fall = read_tsv(input_list$chr34_cv_fall_mafs)
-) %>%
-  bind_rows(.id = "run") %>%
-  pivot_wider(names_from = run, values_from = c(knownEM, unknownEM, nInd)) %>%
-  mutate(abs_diff = abs(knownEM_late_fall - knownEM_cv_fall))
-```
-
-Let's only bother with ones that have 8 or more indivs with reads in late fall and 16 or more in cv-fall, and
-then we can plot them.
-```{r}
-af2 <- af1 %>%
-  filter(nInd_late_fall >= 8 & nInd_cv_fall >= 16)
-
-ggplot(af2, aes(x = position, y = abs_diff)) +
-  geom_point(alpha = 0.4)
-```
-
-That is pretty cool.  It shows that characteristic pattern near 800 Kb that we saw around GREB1L,
-where there a lots of SNPs with large allele frequency differences, and lots with very small
-allele frequency differences, but very few with intermediate differences.
-
-One thing I want to confirm is that all the SNPs with small allele freq differences
-in that region are actually at low frequency in both groups.  We can do that by
-coloring the points by the max frequency (between the two groups) of the minor
-allele---that is the allele that is at lowest _average_ frequency in the two populations.
-
-```{r}
-af3 <- af2 %>%
-  mutate(is_maf = (knownEM_cv_fall + knownEM_late_fall) < 0.5,
-         max_maf = pmax(ifelse(is_maf, knownEM_late_fall, 1 - knownEM_late_fall),
-                        ifelse(is_maf, knownEM_cv_fall, 1 - knownEM_cv_fall))
-         )
-
-mmp <- ggplot(af3, aes(x = position, y = abs_diff, color = max_maf)) +
-  geom_point() +
-  scale_color_viridis_c()
-mmp
-```
-
-And, now, let us also do a version in which the high max_mafs are printed on top, so that we can see them if they exist...
-```{r}
-mmp_sort <- ggplot(af3 %>% arrange(max_maf), 
-                   aes(x = position, y = abs_diff, color = max_maf)) +
-  geom_point() +
-  scale_color_viridis_c()
-mmp_sort
-```
-
-Outstanding.  It is clear that there is a fairly large haplotype/region
-that is differentiated between fall and late fall there in the first
-2 Mb of the chromosome.
-
-## Display some of the highly associated SNPs there
-
-The GWAS turned up a number of highly associated SNPs.  Let's add circles for some
-of the ones with the largest likelihood ratio statistic.
-
-We will take the top 6 from each of the comparisons:
-```{r}
-lfvsj <- read_tsv(input_list$gwas_lf_v_sanjo) %>%
-  arrange(desc(LRT))
-lfvfrhf <- read_tsv(input_list$gwas_lf_v_frhf) %>%
-  arrange(desc(LRT))
-
-# now get the top 6 from each for illustration
-top6_from_each <- list(
-  lfvsj %>% slice(1:6),
-  lfvfrhf %>% slice(1:6)
-) %>%
-  bind_rows() %>%
-  arrange(desc(LRT)) %>%
-  filter(!duplicated(Position))
-
-# print those out
-top6_from_each
-```
-
-So, add those to the plot
-```{r}
-top6freqs <- af3 %>% semi_join(top6_from_each, by = c("position" = "Position"))
-mmp_sort +
-  geom_point(
-    data = top6freqs,
-    shape = 21,
-    size = 3, 
-    color = "black"
-    )
-```
-
-That shows us that there are a lot of other markers with substantial allele freq
-diffs in this region, but the association test puts thew big circles up as the most significant, perhaps
-because of the sample sizes.  It sort of reinforces my informal belief that many insights
-can come from looking at absolute allele freq diffs, at least in these sorts of
-cases where you have a haplotypic region with
-very large allele freq differences).
-
-## Annotate the variants
-
-Before we start designing assays and things, I want to look for large-effect mutations.
-We have the annotations for the positions in these 5 Mb of Ots34 in the stored results
-(from the Science paper), so we will use them here:
-
-First we read in the annotations:
-```{r}
-ann_vcf <- read.vcfR(input_list$ots34_snpeff)
-
-avt <- vcfR2tidy(ann_vcf, info_only = TRUE)
-
-# get column heading for the snpEff output
-ann_column_names <- avt$meta %>%
-  filter(Tag == "INFO", ID == "ANN") %>%
-  pull(Description) %>%
-  str_replace_all(., pattern = "^.* '| +|'", "") %>%
-  str_split(., "\\|") %>%
-  pluck(1)
-
-anns <- avt$fix %>%
-  select(CHROM, POS, REF, ALT, ANN) %>%
-  separate(ANN, into = ann_column_names, sep = "\\|")
-
-```
-I don't think we need to worry about the discarded pieces there.
-
-Now, we can join those annotations onto the allele frequencies and plot them,
-coloring by predicted effect:
-
-```{r}
-af3_ann <- left_join(af3, anns, by = c("position" = "POS")) %>%
-  mutate(Annotation_Impact_f = factor(Annotation_Impact,
-                                      levels = c("HIGH", "LOW", "MODERATE", "MODIFIER"))) %>%
-  arrange(desc(Annotation_Impact_f))
-
-imp_col <- ggplot(af3_ann, aes(x = position, y = abs_diff, colour = Annotation_Impact)) +
-  geom_point() +
-  scale_colour_manual(values = c(HIGH = "red", LOW = "green", MODERATE = "orange", MODIFIER = "gray90")) +
-  theme_bw()
-imp_col
-```
-
-So, there are a few that we would want to be sure to hit.
-
-## Finding candidates for assay design
-
-When I originally did this, I also used a hack to include indels in here.  I spare
-the reader from the gory details of that.
-
-We could rank things by allele frequency difference, but I would also like
-to incorporate an estimate of the variance in these absolute frequency
-differences.  The variance in that difference will be the sum of the
-two variances, so it is easy to calculate.  We don't really know what the
-sample size is, but I am going to throw out as an approximation that each
-individual in the `nInd` field represents a sample of about 1.2 gene copies.
-
-We will estimate the variances from each different estimate and then calculate
-two standard deviations below the actual estimate and have that as a "lower
-bound" on the absolute value of the allele frequency difference.
-
-```{r}
-# here is a function for computing the lower bound on the absolute alle freq diff
-lub <- function(f1, f2, n1, n2, s = 1.2) {
-  var <- (f1 * (1 - f1) / (n1 * s)) + (f2 * (1 - f2) / (n2 * s))
-  sd <- sqrt(var)
-
-  abs(f1 - f2) - 2 * sd
-}
-
-saind2 <- af3_ann %>%
-  mutate(abs_diff_lb = lub(knownEM_late_fall, knownEM_cv_fall, nInd_late_fall, nInd_cv_fall))
-```
-
-Plot those to make sure they look right, and denote the most highly associated ones
-so that we can see those. 
-```{r}
-top6again <- saind2 %>%
-  semi_join(top6_from_each, by = c("position" = "Position"))
-
-ggplot(saind2, aes(x = abs_diff, y = abs_diff_lb, colour = max_maf)) +
-  geom_point() +
-  scale_colour_viridis_c() +
-  geom_point(data = top6again, fill = NA, colour = "black", size = 3, shape = 21)
-```
-
-All right.  When I look at that, I think to myself, The ones I would want to design
-assays for (for distinguishing fall-run from late-fall), would be:
-
-1. (1) The ones with the highest association values
-2. (2) The ones with an abs_diff > 0.75 and an abs_diff "lower bound" above 0.5.  
-3. (4) Then, for any remaining ones we might want to do, I would choose things with an
-abs_diff > 0.5, and abs_diff_lb > 0.25, and a max_maf as close to 1 as possible.  
-4. (8) Annotation_Impact %in% c("Moderate", "High")
-
-We will designate those with a bit mask, with bits being 2 raised to their number 
-minus 1 (as given in the parentheses, above).
-
-**Another Category**  
-I will also add another category of candidate, which is out in the flanks and might
-be interesting for assessing LD out there.
-
-* (16) Position > 2.2 Mb and abs_diff > 0.48 and abs_diff_lb > 0.25
-
-Mark them.
-```{r}
-assay_candi <- saind2 %>%
-  mutate(candi_category = 1 * position %in% top6_from_each$Position +
-           2 * (abs_diff > 0.75 & abs_diff_lb > 0.5) +
-           4 * (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98) + 
-           8 * (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) +
-           16 * (abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6)) %>%
-  select(chromo:max_maf, Annotation_Impact_f:candi_category, everything())
-
-# how many?
-assay_candi %>%
-  count(candi_category)
-```
-
-Just to be explicit about what these candi_categories mean:
-
-* 0 : not selected for any reason
-* 1 : Hi-assoc (from Mac)
-* 2 : abs_diff > 0.75 & abs_diff_lb > 0.5
-* 3 : 1 and 2
-* 4 : (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98)
-* 7 : 1 and 2 and 4
-* 8 : (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) )
-* 16 : abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6
-* 20 : 4 and 16
-
-Plot them on a plot like the above so we can see them.
-```{r}
-ggplot(mapping = aes(x = abs_diff, y = abs_diff_lb, colour = factor(candi_category))) +
-  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
-  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
-  theme_bw() +
-  scale_color_brewer(palette = "Dark2")
-
-```
-
-And see where those candidates live:
-```{r}
-ggplot(mapping = aes(x = position, y = abs_diff, colour = factor(candi_category))) +
-  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
-  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
-  theme_bw() +
-  scale_color_brewer(palette = "Dark2")
-```
-
-So, we have good candidates across a couple Mb.  That is great.
-
-Let's make a plot of that for the supplement:
-```{r}
-assay_candi4fig <- assay_candi %>%
-  mutate(
-    candi_cat = case_match(
-      candi_category,
-      0 ~ NA_character_,
-      1 ~ "1: One of the 10 SNPs with lowest p-values",
-      2 ~ "2: |d| > 0.75 with lower confidence interval >0.5",
-      3 ~ "3: 1 and 2",
-      4 ~ "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
-      7 ~ "7: 1 and 2 and 4",
-      8 ~ "8: snpEff annotation of High or Moderate",
-      16 ~ "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
-      20 ~ "20: 4 and 16"
-    ),
-    `Candidate Marker Category` = factor(
-      candi_cat,
-      levels = c(
-        "1: One of the 10 SNPs with lowest p-values",
-        "2: |d| > 0.75 with lower confidence interval >0.5",
-        "3: 1 and 2",
-        "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
-        "7: 1 and 2 and 4",
-        "8: snpEff annotation of High or Moderate",
-        "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
-        "20: 4 and 16"
-      ))
-  )
-
-g <- ggplot(mapping = aes(x = position, y = abs_diff, colour = `Candidate Marker Category`)) +
-  geom_point(data = assay_candi4fig %>% filter(candi_category == 0), colour = "gray90") + 
-  geom_point(data = assay_candi4fig %>% filter(candi_category > 0)) +
-  theme_bw() +
-  scale_color_brewer(palette = "Dark2") +
-  xlab("Position on Chromosome 34 (Mb)") +
-  ylab("|d|, Absolute Difference in Allele Frequency")
-
-ggsave(g, filename = output_list$lfar_candidates, width = 8, height = 5)
-```
-
-## Preparing an assay order
-
-We prepare assay order information for these 49 good candidates, and handed that
-to Anthony, along with the sequences, to design primers.
-
-That process involved some custom scripts and Primer3.  Once the candidates
-came back, they were tested in a variety of Chinook samples and it was found that
-only 8 gave reliable amplicons that mapped reliably, etc.  And those
-are used in the ensuing work.
-
-Here, print out the actual candidates:
-```{r}
-actual_candidates <- assay_candi %>%
-  filter(candi_category > 0) %>% 
-  arrange(position)
-
-# write that out for fun:
-write_csv(actual_candidates, file = output_list$candidates_for_assays)
-```
+Subsequently, outputs were uploaded to the [CLUMPAK server](https://clumpak.tau.ac.il/) for
+visualization.
 

### 2025-03-27 README.Rmd
Done

@@
 title: "california-chinook-microhaps: Data and code repository for \"A multipurpose microhaplotype panel for genetic analysis of California Chinook salmon\""
-author: "Eric C. Anderson (also repo maintainer/owner), Anthony J. Clemento, Matthew A. Campbell, Devon E. Pearse, Anne K. Beulke, Cassie Columbus, Ellen Campbell, Neil F. Thompson, John Carlos Garza"
+author: "Eric C. Anderson (also repo maintainer), Anthony J. Clemento, Matthew A. Campbell, Devon E. Pearse, Anne K. Beulke, Cassie Columbus, Ellen Campbell, Neil F. Thompson, John Carlos Garza"
 output: 
-  github_document:
-    toc: true
-    toc_depth: 3
   html_document:
@@
     toc_float: true
+    toc_depth: 3
+  github_document:
+    toc: true
     toc_depth: 3
@@
 
+# Evaluating all the R Markdown documents using Snakemake
+
+The `Snakefile` contains a recipe for evaluating all the R Markdown documents.
+Within a conda environment that has Snakemake, you can run them all like this:L
+```sh
+snakemake -np --use-conda --forceall  --cores 8
+```
+If you are not familiar with Snakemake, it will probably be better to
+simply evaluate each one separately within RStudio if you want to 
+reproduce the results.
+
+# RMarkdown Documents
+
+The following RMarkdown documents can be evaluated in any order.
+
+Outputs (figures, tables, etc) from each RMarkdown document
+are written to the `tex/images` or `tex/inputs` directories.  Intermediate files
+written during evaluation of each RMarkdown document are written
+into the `results` directories.
+
+Thumbnails of the figures and tables generated by each RMarkdown document appear
+below each notebook that generates them.
+
+Click the thumbnail to see a full size PDF.
+
+Additionally, at the top of each section is a link to the
+compiled (HTML-version) of the RMarkdown document on GitHub Pages.
+
+## `r make_link("001-map-of-samples.Rmd")`
+
+Makes a map of where the samples were drawn. (Figure 1)
+
+[![](./docs/images_and_thumbs/figure1_thumb.png)][fig1]
+
+## `r make_link("002-GSI_and_Fst.Rmd")`
+
+Summaries of the samples and population genetic statistics,
+as well as GSI power analysis and Fst calculation.
+(Tables 1, 4;  Figures 2, S4 S6, S7; SuppData 2)
+
+[![](./docs/images_and_thumbs/table1_thumb.png)][tab1]
+[![](./docs/images_and_thumbs/table4_thumb.png)][tab4]
+
+[![](./docs/images_and_thumbs/figure2_thumb.png)][fig2]
+[![](./docs/images_and_thumbs/figureS4_thumb.png)][figS4]
+[![](./docs/images_and_thumbs/figureS6_thumb.png)][figS6]
+[![](./docs/images_and_thumbs/figureS7_thumb.png)][figS7]
+
+[![](./docs/images_and_thumbs/supp_data2_thumb.png)][supp_data2]
+
+## `r make_link("003-ckmr-sim-evaluation.Rmd")`
+
+Using CKMRsim to estimate false positive rates and false negative rates
+of relationship inference.  Also comparisons of power for inferring relatives
+when using microhaplotypes versus single SNPs from each amplicon.
+(Figures 4, 5, S8)
+
+[![](./docs/images_and_thumbs/figure4_thumb.png)][fig4]
+[![](./docs/images_and_thumbs/figureS8_thumb.png)][figS8]
+[![](./docs/images_and_thumbs/figure5_thumb.png)][fig5]
+
+## `r make_link("004-lfar-and-wrap-allele-freqs.Rmd")`
+
+Assessing allele frequency differences between groups of fish at the
+LFAR and WRAP amplicons. (Tables 3, S2)
+
+[![](./docs/images_and_thumbs/table3_thumb.png)][tab3]
+[![](./docs/images_and_thumbs/tableS2_thumb.png)][tabS2]
+
+## `r make_link("005-plotting-genome-locations-of-markers.Rmd")`
+
+Plotting the positions of different amplicons in the genome.
+(Figure S1)
+
+[![](./docs/images_and_thumbs/figureS1_thumb.png)][figS1]
+
+## `r make_link("006-winter-run-vs-all-cv-alle-freqs.Rmd")`
+
+Analyzing the allele frequencies from WGS data to identify regions
+that show large allele frequency differences between winter run
+and everyone else to discover the WRAP markers. (Figures S9, S10, S11)
+
+[![](./docs/images_and_thumbs/figureS9_thumb.png)][figS9]
+[![](./docs/images_and_thumbs/figureS10_thumb.png)][figS10]
+[![](./docs/images_and_thumbs/figureS11_thumb.png)][figS11]
+
+## `r make_link("007-whoa-on-chinook.Rmd")`
+
+This was not included in the paper, but is an interesting analysis of
+the markers suggesting that two of them might have some genotyping
+issues and would be good candidates for removal in a further update
+to the baseline.
+
+[![](./docs/images_and_thumbs/whoa_thumb.png)][whoa]
+
+## `r make_link("008-lfar-gwas-and-markers.Rmd")`
+
+Analyzing the GWAS and allele frequency data from the LFAR search. (Figures S2, S3).
+
+[![](./docs/images_and_thumbs/figureS2_thumb.png)][figS2]
+[![](./docs/images_and_thumbs/figureS3_thumb.png)][figS3]
+
+## `r make_link("009-structure.Rmd")`
+
+Running _structure_ analyses on the data set. (Figures 3, S5).
+
+[![](./docs/images_and_thumbs/figure3_thumb.png)][fig3]
+[![](./docs/images_and_thumbs/figureS5_thumb.png)][figS5]
+
 # Repo Contents
@@
 repo made using `tree`.   Some file-series are omitted and replaced with
-ellipses.
+ellipses, and this is not guaranteed to be current, but it gives
+a clear picture of the main parts of the repo.
 
@@
 
-# RMarkdown Documents
-
-The following RMarkdown documents can be evaluated in any order.
-
-Outputs (figures, tables, etc) from each RMarkdown document
-are written to the `tex/images` or `tex/inputs` directories.  Intermediate files
-written during evaluation of each RMarkdown document are written
-into the `results` directories.
-
-Thumbnails of the figures and tables generated by each RMarkdown document appear
-below each notebook that generates them.
-
-Click the thumbnail to see a full size PDF.
-
-Additionally, at the top of each section is a link to the
-compiled (HTML-version) of the RMarkdown document on GitHub Pages.
-
-## `r make_link("001-map-of-samples.Rmd")`
-
-Makes a map of where the samples were drawn.
-
-[![](./docs/images_and_thumbs/figure1_thumb.png)][fig1]
-
-## `r make_link("002-GSI_and_Fst.Rmd")`
-
-Summaries of the samples and population genetic statistics,
-as well as GSI power analysis and Fst calculation.
-
-[![](./docs/images_and_thumbs/table1_thumb.png)][tab1]
-[![](./docs/images_and_thumbs/figure2_thumb.png)][fig2]
-[![](./docs/images_and_thumbs/figureS6_thumb.png)][figS6]
-[![](./docs/images_and_thumbs/figureS7_thumb.png)][figS7]
-
-[![](./docs/images_and_thumbs/supp_data2_thumb.png)][supp_data2]
-[![](./docs/images_and_thumbs/table4_thumb.png)][tab4]
-
-## `r make_link("003-ckmr-sim-evaluation.Rmd")`
-
-Using CKMRsim to estimate false positive rates and false negative rates
-of relationship inference.  Also comparisons of power for inferring relatives
-when using microhaplotypes versus single SNPs from each amplicon.
-
-[![](./docs/images_and_thumbs/figure4_thumb.png)][fig4]
-[![](./docs/images_and_thumbs/figureS8_thumb.png)][figS8]
-[![](./docs/images_and_thumbs/figure5_thumb.png)][fig5]
-
-## `r make_link("004-lfar-and-wrap-allele-freqs.Rmd")`
-
-Assessing allele frequency differences between groups of fish at the
-LFAR and WRAP amplicons.
-
-[![](./docs/images_and_thumbs/table3_thumb.png)][tab3]
-[![](./docs/images_and_thumbs/tableS2_thumb.png)][tabS2]
-
-## `r make_link("005-plotting-genome-locations-of-markers.Rmd")`
-
-Plotting the positions of different amplicons in the genome.
-
-[![](./docs/images_and_thumbs/figureS1_thumb.png)][figS1]
-
-## `r make_link("006-winter-run-vs-all-cv-alle-freqs.Rmd")`
-
-Analyzing the allele frequencies from WGS data to identify regions
-that show large allele frequency differences between winter run
-and everyone else.  (Finding the WRAP markers...)
-
-[![](./docs/images_and_thumbs/figureS9_thumb.png)][figS9]
-[![](./docs/images_and_thumbs/figureS10_thumb.png)][figS10]
-[![](./docs/images_and_thumbs/figureS11_thumb.png)][figS11]
-
-## `r make_link("007-whoa-on-chinook.Rmd")`
-
-This was not included in the paper, but is an interesting analysis of
-the markers suggesting that two of them might have some genotyping
-issues and would be good candidates for removal in a further update
-to the baseline.
-
-[![](./docs/images_and_thumbs/whoa_thumb.png)][whoa]
-
-## `r make_link("008-lfar-gwas-and-markers.Rmd")`
-
-Analyzing the GWAS and allele frequency data from the LFAR search
-
-[![](./docs/images_and_thumbs/figureS3_thumb.png)][figS3]
-
 [fig1]: `r hdir`images_and_thumbs/figure1.pdf
@@
 [fig2]: `r hdir`images_and_thumbs/figure2.pdf
+[figS4]: `r hdir`images_and_thumbs/figureS4.pdf
 [figS6]: `r hdir`images_and_thumbs/figureS6.pdf
@@
 [whoa]: `r hdir`images_and_thumbs/whoa.pdf
+[figS2]: `r hdir`images_and_thumbs/figureS2.png
 [figS3]: `r hdir`images_and_thumbs/figureS3.pdf
+[fig3]: `r hdir`images_and_thumbs/figure3.pdf
+[figS5]: `r hdir`images_and_thumbs/figureS5.pdf

### 2025-03-27 tex/responses-to-referees/responses-to-reviewers-1.Rmd
make response to refs more cover-lettery

@@
 ---
-title: "EVA-2025-003-RA Responses to Reviewers"
+title: "EVA-2025-003-RA Cover Letter and Responses to Reviewers"
 output: pdf_document
-date: "2025-03-07"
+date: "2025-03-26"
 ---
@@
 ```
+
+Thank you for the opportunity to revise our manuscript for _Evolutionary Applications_.
+We have made a number of changes as recommended by the referees and feel that
+these have greatly improved the manuscript.
+
+In addition to the responses to referees (below) we are uploading a version of
+the manuscript created with `latexdiff` that shows the changes we have made in
+the document since the previously submitted version.
+
+It has been a pleasure working with the journal's editorial staff.
 
@@
     - A comprehensive README has been added.  The repo and its README can be viewed at:
-    [https://github.com/eriqande/california-chinook-microhaps/](https://github.com/eriqande/california-chinook-microhaps/).  This will be archived at Zenodo upon acceptance.
+    [https://anonymous.4open.science/r/california-chinook-microhaps-A661/README.md](https://anonymous.4open.science/r/california-chinook-microhaps-A661/README.md).
+This will be archived at Zenodo upon acceptance.

### 2025-03-31 tex/responses-to-referees/responses-to-reviewers-1.Rmd
Updates/edits from Carlos

@@
 Methods section:_
-"Of the 1636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall run ecotype,
-498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter run and late-fall run ecotypes, 
+"Of the 1,636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall-run ecotype,
+498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter-run and late-fall-run ecotypes, 
 respectively,  each represented by a single collection."

### 2025-03-31 tex/abstract.tex
Updates/edits from Carlos

--- 2024-11-06 tex/abstract.tex
@@
 have also become common. We present, here, a novel set of
-multi-allelic microhaplotype genetic markers for Chinook salmon which provide unprecedented
+multi-allelic microhaplotype genetic markers for Chinook salmon which provide excellent
 resolution for population discrimination and relationship identification from a rapidly and
-economically assayed panel of markers. We show how this set of microhaplotypes
+economically assayed panel of markers. We show how this set of genetic markers assayed
+by sequencing of 204 amplicons, in tandem with a reference data set of 1,636 individual
+samples from 17 populations,
 provides definitive power to identify all known lineages of Chinook salmon in

### 2025-03-31 tex/body-text-of-paper.tex
Updates/edits from Carlos

@@
 three rivers of the California coast (Russian, Eel, and Smith).
-Of the 1636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall run ecotype,
-498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter run and late-fall run ecotypes, 
+Of the 1,636 total fish samples from 17 different populations, 727 fish from 10 collections were of the fall-run ecotype,
+498 fish from 5 collections were spring run, and 111 and 300 fish were of the winter-run and late-fall-run ecotypes, 
 respectively,  each represented by a single collection.
@@
 Elsewhere, adult samples were taken from live broodstock at Rowdy Creek (Smith River), Iron Gate (Klamath River) and Warm Springs (Russian River) Hatcheries and juveniles were sampled from Blue Creek (Klamath River); all locations where only fall-run fish have been documented.
-Spring run adult fish were sampled from the spring-run broodstock at Cole Rivers Hatchery on the Rogue River.
+Spring-run adult fish were sampled from the spring-run broodstock at Cole Rivers Hatchery on the Rogue River.
 Samples from the Eel River were taken from live fish ascending a ladder in the upper basin and were confirmed as fall-run ecotype by RoSA genotype.
@@
 The sequences from each run at each locus were mapped to the {\em de novo} reference sequences (produced by Stacks) with
-the Burrows-Wheeler (bwa) aligner using the mem algorithm \citep{bwa-mem2009}. 
+bwa mem \citep{bwa-mem2009}. 
 Subsequently, alignments were sorted with SAMtools \citep{li2009sequence} and provided to
@@
 and PBT in California \citep{clemento2014evaluation}. These markers were originally assayed
-using TaqMan (Life Technologies Corp., Carlsbad, CA) probes or SNPtype assays (Standard BioTools Inc., South San Francisco, CA).
+using TaqMan (Life Technologies Corp.) probes or SNPtype assays (Standard BioTools Inc.).
 These markers were typed using amplicon sequencing, following the methods in
@@
 Initial examination of sequence data 
-in the 16 winter-run fish and 84 non-winter run fish revealed
+in the 16 winter-run fish and 84 non-winter-run fish revealed
 a number of SNPs that were potentially diagnostic, but none were fixed for alternate alleles 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-The late-fall run collection (CHLF) appears as a separate cluster at $K$ values
+The late-fall-run collection (CHLF) appears as a separate cluster at $K$ values
 as low as 4, with the inclusion of the LFAR markers.  This is notable, since accurately discriminating between the late-fall and
@@
 
-Overall, of the 1636 fish in the reference baseline, 1612 (98.5\%) were correctly
+Overall, of the 1,636 fish in the reference baseline, 1,612 (98.5\%) were correctly
 assigned to their reporting unit of origin, while 24 (1.5\%) were incorrectly assigned.
@@
 would be incorrectly identified).  Of the remaining 20 misassignments,  8 involved fish from the
-CHLF (late fall run) collection being misassigned to the CV-Fall reporting unit and 6 were
+CHLF (late-fall run) collection being misassigned to the CV-Fall reporting unit and 6 were
 fish from the MDS (Mill-Deer spring run) collection being assigned to the CV-Fall reporting unit.
@@
 of only 0.6\% (3/508) and to the CV-Spring reporting unit (to which MDS belongs) at a rate
-of only 0.4\% (2/508).
-When assignments are only accepted with a scaled likelihood greater than 0.8,
+of only 0.4\% (2/508). When assignments are only accepted with a scaled likelihood greater than 0.8,
 10 incorrect assignments and 10 correct assigments are discarded.  Thus,
@@
 amplicons, versus using the most heterozygous single SNP from each
-amplicon for all fish from the FRH (Fig.~\ref{fig:ckmr-comp}), with results for the PO vs Unrelated
-log likelihood ratio shown for more collections in Fig.~\ref{fig:ckmr-comp-supp}.  The figures show a considerable increase
+amplicon for all fish from the FRH (Figure~\ref{fig:ckmr-comp}), with results for the PO vs Unrelated
+log likelihood ratio shown for more collections in Figure~\ref{fig:ckmr-comp-supp}.  The figures show a considerable increase
 in separation between the relationship categories from calling genotypes
@@
 of additional variation surrounding the originally assayed SNP\@. Our work adds to a growing
-chorus of recent studies finding that microhaplotype markers can provide additional
+number of recent studies finding that microhaplotype markers can provide additional
 power over single SNPs for both GSI \citep{mckinney2017managing,mckinney2022new,hargrove2024efficient} and
@@
 near complete accuracy in distinguishing these 'stocks'. Moreover, the few fish that are
-apparently misidentified (Figure~\ref{fig:gsi}) are likely to be primarily migrants and not true misidentifications.
+apparently misidentified (Figure~\ref{fig:gsi},~\ref{fig:eighty}) are likely to be primarily migrants and not true misidentifications.
 For example, the three fish that were field-characterized as spring-run from Butte Creek, but are
